### 2021-10-16 MultipleRoundSims/MultipleRoundSims.tex
initial commit

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
\author{ }
\date{ }

\begin{document}

\maketitle

\begin{abstract}
    Risk-Limiting Audits (RLAs) guarantee with known probability 
    that if the outcome of an 
    election is incorrectly announced, it will be detected, 
    and a full hand recount will be performed. 
    In Ballot-polling RLAs, samples of ballots are drawn and tallied.
    In this paper we present an audit simulation framework
    that can be used to confirm theoretical results of audits.
    We present such experimental results for multiple round 
    ballot-polling RLAs.
    BRAVO [citation] has long been the standard ballot-polling RLA,
    while Minerva was recently introduced with the claim
    that fewer ballots on average are necessary to conclude 
    an audit.
    [Minerva paper] presented experimental
    results for one round audits, and here
    we present results
    from simulations of multiple round audits using Minerva 
    as well as BRAVO 
    (both Selection-Ordered 
    BRAVO and End-of-Round BRAVO).
    The simulation results agree within reasonable error with
    the mathematical properties claimed in [Minerva paper].
    On average, BRAVO audits are unnecessarily conservative 
    while Minerva audits stop with fewer ballots. We also
    present details on software implementing Minerva and
    the simulations themselves.
\end{abstract}

\section{Introduction: B2 and R2 Ballot Polling RLAs}
    In ballot-polling RLAs, samples of ballots are drawn and tallied
    in rounds
    after each of which a statistical measure determines whether to
    continue.

\section{Simulations to Confirm Theoretical Results}
The outcomes of RLAs depend on random chance; some random samples
support the alternative hypothesis more than expected, resulting in 
quick low-risk conclusions, while other unlucky samples require subsequent
rounds to confirm the announced result.
We can simulate random samples for various true underlying ballot 
distributions by computing pseudorandom samples. 
By applying an audit's stopping condition to thousands of such
simulated samples, the stopping behavior will tend towards the true
frequencies for the audit. 
In this way, we can examine whether theoretical claims for an audit are
actually correct.
For this paper, we simulated audits for all US states with at least a $5\%$ 
margin in the 2020 Presidential election. 
For each of these states, we simulated 
$10,000=10^4$ audits with the announced
underyling ballot distribution
and an additional $10,000=10^4$ audits with a tie
as the underlying ballot distribution.
\subsection{End-of-Round BRAVO}
\subsection{Selection-Ordered BRAVO}
\subsection{Minerva}

\end{document}

### 2021-10-18 MultipleRoundSims/MultipleRoundSims.tex
proprogress

@@
 \usepackage[utf8]{inputenc}
+\usepackage{graphicx}
+\graphicspath{{./imgs/}}
 
@@
 continue. 
+Some ballot-polling audits (B2 audits) 
+are designed to make a stopping decision
+after each individual ballot is sampled.
+Other ballot-polling audits (R2 audits) assume many ballots are drawn
+at once and make a stopping decision at the conclusion of each 
+so-called round.
+In practice, election officials draw many ballots at once.
+For this reason, R2 audits are used or B2 audits are applied to 
+rounds.
+BRAVO is a B2 audit. 
+When an audit is performed in rounds, BRAVO can have its
+stopping condition applied once at the end of each round
+(End-of Round (EoR) BRAVO).
+This method ignores the possibility of stopping earlier in the 
+sample and thus requires more ballots on average to sample.
+Alternatively, the order of ballots in the sample can be tracked
+by election officials and the B2 BRAVO stopping condition can 
+be applied retroactively to all subsamples 
+(Selection-Ordered (SO) BRAVO).
+SO BRAVO requires fewer ballots on average than EoR BRAVO but
+requies the work of tracking the order of ballots rather than
+just their tally.
+Minerva was designed for R2 audits and applies its stopping rule
+once for each round.
+Minerva, by design then, requires fewer ballots to be sampled when 
+an audit is performed in rounds.
 
 \section{Simulations to Confirm Theoretical Results}
+
 The outcomes of RLAs depend on random chance; some random samples
@@
 quick low-risk conclusions, while other unlucky samples require subsequent
-rounds to confirm the announced result.
-We can simulate random samples for various true underlying ballot 
+rounds to confirm the announced results.
+We can simulate random samples for various underlying ballot 
 distributions by computing pseudorandom samples. 
 By applying an audit's stopping condition to thousands of such
-simulated samples, the stopping behavior will tend towards the true
-frequencies for the audit. 
-In this way, we can examine whether theoretical claims for an audit are
+simulated samples, the average behavior of the simulated
+audits will tend towards the true behavior of the audit.
+In this way, we can examine whether theoretical claims about an audit are
 actually correct.
-For this paper, we simulated audits for all US states with at least a $5\%$ 
-margin in the 2020 Presidential election. 
+
+For this paper, we simulated audits for 
+the 2020 Presidential election
+in all US states whose margin was at least $5\%$.
+Round sizes increase roughly proportional to the inverse
+square of the margin, so 
+smaller margins are computationally more expensive to simulate.
 For each of these states, we simulated 
@@
 as the underlying ballot distribution.
+These are reasonable choices for inital simulation experiments
+because most audits frame the stopping decision as a binary
+hypothesis test where the null hypothesis assumes an underyling tie
+and the alternative hypothesis assumes the announced distribution.
+A standard first round size in ballot-polling audits
+has been one which achieves a $90\%$ probability
+of stopping, and we chose round sizes to reflect this standard
+in our simulations.
+We ran our simulations for up to five rounds.
+With a $90\%$ stopping probability, 
+an audit only has a $a$ probablity of not stopping
+by the end of the fifth round, assuming the outcome was correctly
+announced.
+
 \subsection{End-of-Round BRAVO}
-\subsection{Selection-Ordered BRAVO}
+For the EoR BRAVO simulations, our software estimated for each round
+the round size that would achieve a $90\%$ probability of stopping.
+For the simulations with the announced outcome as the underlying
+ballot distribution, for each round $j$ we computed the probability
+of stopping given that an audit had continued through round $j$.
+That is, among the audits remaining at the beginning of round $j$, 
+we found the proportion which then stopped in round $j$.
+%This metric shows whether our software accurately estimated the minimum number of ballots that achieve the $90\%$ probability of stopping.
+
+%\includegraphics[scale=.6]{eor_bravo_90perc_10^4_corrected/cond_sprob/r1_cond_sprob.png}
+\begin{center}
+\includegraphics[scale=.41]{eor_bravo_90perc_10^4_corrected/cond_sprob_first_3_rounds.png}
+\end{center}
+[Figure 1] shows the proportion of audits that stopped in the $j$th round
+to all audits which had not already stopped in a preceding round.
+This proportion is shown for each state whose margin is given on the 
+horizontal axis.
+Here we display only the first three rounds
+since very few audits, $(.1)^{j-1}$ on average, 
+make it to the $jth$ round, 
+and as a result the data becomes exponentially less representative
+of the true probabilities.
+
+The proportion shown in [Fig 1] gives an estimate
+of the true probability of an EOR BRAVO audit stopping in the $j$th round,
+given that it has not already stopped in a previous round. 
+In [Figure 1] we see that, especially in earlier rounds for which 
+the values are more representative of true audit behavior, 
+our predictions are accurate.
+In particular the average across all margins is just above $.9=90\%$ for
+all three rounds.
+
+The total risk across all $5$ rounds for this audit is arguably
+the most important metric to consider.
+The risk of an audit is the probability that the audit
+stops despite an underlying tie. 
+%That is, the risk $R$ of audit $\mathcal{A}$ is 
+%$$R(\mathcal{A})=Pr[\mathcal{A}=\text{Correct}\mid H_0].$$
+Therefore, we now consider the proportion of audits with an underlying
+tie.
+
+%\includegraphics[scale=.5]{eor_bravo_90perc_10^4_corrected/total_risk.png}
+\includegraphics[scale=.7]{eor_bravo_90perc_10^4_corrected/total_risk_with_alpha.png}
+
+[Fig 2] shows, for each state margin,
+the proportion of EOR BRAVO audits with an underlying
+tie which stopped before the conclusion of the $5$ rounds to the 
+total number of audits, $10^4$.
+
+[Fig 2] makes it clear that the risk of EOR BRAVO is roughly
+an order of magnitude less than the risk limit. 
+That is, on average, roughly $10$ times more audits could have stopped 
+with the audit still meeting the risk limit.
+These simulations support the claim made in [Minerva paper]
+that EOR BRAVO is unnecessarily conservative and thus requires
+more ballots on average.
+
+% \subsection{Selection-Ordered BRAVO}
+% next_sample_size code is slow... tie simulations running still
+
 \subsection{Minerva}
+For Minerva, it has not been shown that round sizes can be chosen
+during the audit. That is, an adversary with knowledge of the history
+of the audit may be able to choose round sizes which cause the 
+risk of the audit to exceed the risk limit.
+For this reason, we have to choose the round sizes of a Minerva 
+audit a priori.
+For this paper, we consider two choices of round sizes.
+For both, we estimate and then use the minimum first round size 
+which achieves
+a $90\%$ probability of stopping.
+Then, for subsequent rounds, we either (i) multiply the previous
+round size by a factor of $1.5$ and samples this many new ballots 
+or (ii) multiply by a factor of $1$ and sample this many.
+We choose $1.5$ because it gives a very rough approximation of the
+round sizes which give a $90\%$ probability of stopping.
+We choose $1$ because it may reflect a practical way to continue an
+audit; if election officials have selected some first round size within
+reasonable logistical bounds, drawing the same sample in subsequent rounds
+could be practical.
+
+\subsubsection{Roundsizes with Multiple of $1.5$}
+
+\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
+
+\subsubsection{Roundsizes with Multiple of $1.0$}
+
+\includegraphics[scale=.7]{minerva_multiround_1x_10^4/total_risk.png}
 

### 2021-10-19 MultipleRoundSims/MultipleRoundSims.tex
minerva explanation

@@
 a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) multiply the previous
-round size by a factor of $1.5$ and samples this many new ballots 
-or (ii) multiply by a factor of $1$ and sample this many.
-We choose $1.5$ because it gives a very rough approximation of the
-round sizes which give a $90\%$ probability of stopping.
-We choose $1$ because it may reflect a practical way to continue an
+Then, for subsequent rounds, we either (i) 
+draw the same number of ballots in each round or (ii)
+multiply the previous round size by a factor of $1.5$ and 
+sample this many new ballots.
+We consider the case of drawing samples of the same size
+because it may reflect a practical way to continue an
 audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same sample in subsequent rounds
-could be practical.
+reasonable logistical bounds, drawing the same number of 
+ballots in subsequent rounds may be practical.
+We also consider round sizes with samples increasing by a multiple
+of $1.5$ because this multiple gives a very rough approximation of 
+round sizes with a $90\%$ probability of stopping.
 
-\subsubsection{Roundsizes with Multiple of $1.5$}
+\subsubsection{Round Sizes with Multiple of $1.0$}
 
-\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
-
-\subsubsection{Roundsizes with Multiple of $1.0$}
+As with the preceding simulations, we ran $10,000=10^4$ trials
+per state for both the underlying tie and underlying announced
+outcome.
 
@@
 
+[Fig above] shows that all simulations had a risk of less than 
+$.1$, the risk limit. 
+Unlike EOR BRAVO, the risks here are much closer to the risk limit,
+showing that Minerva stops on average with sharper risk.
+
+\subsubsection{Round Sizes with Multiple of $1.5$}
+
+%For the Minerva simulations with a round size multiple of $1.5$,
+%we increased the number of simulations to $10^6$ per state for 
+%both an underlying tie and underlying announced outcome.
+
+%\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
+
 \end{document}

### 2021-10-20 MultipleRoundSims/MultipleRoundSims.tex
new comp

@@
 \graphicspath{{./imgs/}}
+\usepackage{amsthm}
+\newtheorem{definition}{Definition}
 
@@
 an audit is performed in rounds.
+
+Now we present relevant definitions.
+We consider a two candidate plurality contest.
+To begin we define an audit.
+\begin{definition}
+An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
+following decisions
+\begin{enumerate}
+\item
+$Correct$: the audit is complete
+\item
+$Uncertain$: continue the audit
+\end{enumerate}
+\end{definition}
+Audits in this paper are all formed as a binary hypothesis test.
+Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
+That is, the true underlying ballot distribution is given by the announced ballot tallies.
+Under the null hypothesis, $H_0$, the true outcome is a tie 
+(or a the announced winner losing by one vote if there is an odd number of total ballots).
+Now we define key attributes of an audit that we will consider while analyzing our simulations.
+The risk of an audit is the probability that an audit stops when a tie has occurred.
+\begin{definition}[Risk]
+The risk $R$ of an audit $\mathcal{A}$ is
+$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
+\end{definition}
+This leads us to the following simple definition of an $\alpha$-RLA.
+\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+risk limit $\alpha$ iff 
+$$R(\mathcal{A}) \le \alpha.$$
+\end{definition}
+
+It is useful to discuss the probability of an audit stopping in
+some round, if the outcome is correctly announced.
+\begin{definition}[Stopping Probability]
+The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
+$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
+\end{definition}
+The notion of stopping probability can be useful when selecting round sizes.
 

### 2021-10-21 MultipleRoundSims/MultipleRoundSims.tex
eorbravo and minerva plots all in; add simulator discussion

@@
     We present such experimental results for multiple round 
-    ballot-polling RLAs.
-    BRAVO [citation] has long been the standard ballot-polling RLA,
+    ballot polling RLAs.
+    BRAVO [citation] has long been the standard ballot polling RLA,
     while Minerva was recently introduced with the claim
@@
 \section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
+In ballot polling RLAs, samples of ballots are drawn and tallied
 in rounds
@@
 continue. 
-Some ballot-polling audits (B2 audits) 
+Some ballot polling audits (B2 audits) 
 are designed to make a stopping decision
 after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
+Other ballot polling audits (R2 audits) assume many ballots are drawn
 at once and make a stopping decision at the conclusion of each 
@@
 
-[Fig above] shows that all simulations had a risk of less than 
-$.1$, the risk limit. 
-Unlike EOR BRAVO, the risks here are much closer to the risk limit,
-showing that Minerva stops on average with sharper risk.
+[Fig above] shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.
+
+Notice in [fig above] that all simulations had a proportion of audits that stopped less than $.1$, the risk limit, supporting
+the claim that Minerva is a risk limiting audit. 
+Unlike EOR BRAVO, the experimental risks here are much closer to the risk limit,
+showing that Minerva stops on average with a less conservative risk; Minerva is sharper.
 
@@
 %\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
+\includegraphics[scale=.7]{minerva_multiround_1p5x_10^4/total_risk.png}
+
+[Fig above] shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.
+
+\includegraphics[scale=.7]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
 

### 2021-10-21 MultipleRoundSims/samplepaper.tex
eorbravo and minerva plots all in; add simulator discussion

--- 2021-10-20 MultipleRoundSims/MultipleRoundSims.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
+\usepackage{float}
 \usepackage{graphicx}
 \graphicspath{{./imgs/}}
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
-\author{ }
-\date{ }
+% nice looking audit titles
+\newcommand{\Minerva}{\textsc{Minerva}\xspace}
+\newcommand{\B}{{{B2}}\xspace}
+\newcommand{\R}{{{R2}}\xspace}
+\newcommand{\Bravo}{\textsc{BRAVO}\xspace}
 
 \begin{document}
-
-\maketitle
-
+%
+\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
+%add \thanks{} within the title brackets if want to refer to supporting organization ^^
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
+Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
+Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+%
+\authorrunning{F. Author et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Princeton University, Princeton NJ 08544, USA \and
+Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
+\email{lncs@springer.com}\\
+\url{http://www.springer.com/gp/computer-science/lncs} \and
+ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
+\email{\{abc,lncs\}@uni-heidelberg.de}}
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
@@
     In this paper we present an audit simulation framework
-    that can be used to confirm theoretical results of audits.
+    that can be used to support theoretical results of audits.
     We present such experimental results for multiple round 
@@
     the simulations themselves.
+
+\keywords{Risk-limiting audit  \and Ballot polling audit}
 \end{abstract}
+%
+%
+%
 
@@
 \end{definition}
-Audits in this paper are all formed as a binary hypothesis test.
+All of the audits discussed in this paper are modeled as binary hypothesis tests.
 Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
@@
 \end{definition}
-The notion of stopping probability can be useful when selecting round sizes.
+The notion of stopping probability can be useful for selecting round sizes.
 
-\section{Simulations to Confirm Theoretical Results}
+\section{Simulator}
+\subsection{Simulations to Support Theoretical Audit Properties}
 
@@
 support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other unlucky samples require subsequent
-rounds to confirm the announced results.
+quick low-risk conclusions, while other samples require subsequent
+rounds in order to confirm the announced results.
 We can simulate random samples for various underlying ballot 
@@
 actually correct.
+
+\subsection{Software for Simulations}
+Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
+all written in Python.
+For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
+of the minimum round size to achieve a desired stopping probability. 
+For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
+Ballots can be sampled from any distribution of the users choosing. 
+It is often useful to consider the distribution of ballots corresponding to a tie and the 
+distribution of ballots corresponding with the announced results; these are the distributions represented
+by the null and alternative hypotheses.
+After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
+If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
+The abstract simulator class does not prescribe any one method for choosing round sizes. 
+We implement several classes to support various round size choices: 
+round sizes from an estimate to achieve a desired probability of stopping, 
+predetermined round sizes, and random round sizes.
+
+\section{Simulation Results}
 
@@
 \subsection{End-of-Round BRAVO}
-For the EoR BRAVO simulations, our software estimated for each round
-the round size that would achieve a $90\%$ probability of stopping.
+For the EoR BRAVO simulations, our software estimated and used for each round
+the minimum round size that would achieve a $90\%$ probability of stopping.
 For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the probability
-of stopping given that an audit had continued through round $j$.
-That is, among the audits remaining at the beginning of round $j$, 
-we found the proportion which then stopped in round $j$.
-%This metric shows whether our software accurately estimated the minimum number of ballots that achieve the $90\%$ probability of stopping.
+ballot distribution, for each round $j$ we computed the proportion of audits 
+that stopped in round $j$ among those that had not stopped before round $j$.
 
-%\includegraphics[scale=.6]{eor_bravo_90perc_10^4_corrected/cond_sprob/r1_cond_sprob.png}
-\begin{center}
-\includegraphics[scale=.41]{eor_bravo_90perc_10^4_corrected/cond_sprob_first_3_rounds.png}
-\end{center}
-[Figure 1] shows the proportion of audits that stopped in the $j$th round
-to all audits which had not already stopped in a preceding round.
-This proportion is shown for each state whose margin is given on the 
-horizontal axis.
-Here we display only the first three rounds
-since very few audits, $(.1)^{j-1}$ on average, 
-make it to the $jth$ round, 
-and as a result the data becomes exponentially less representative
-of the true probabilities.
+\begin{figure}
+\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
+For the first three rounds of the EoR BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
+to all audits which had not yet stopped before the $j$th round.}
+\label{fig:eor_bravo_sprob}
+\end{figure}
 
-The proportion shown in [Fig 1] gives an estimate
+In Figure \ref{fig:eor_bravo_sprob}, we display proportions for only the first three rounds
+since very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
+make it to the $jth$ round.
+As  a result, the proportions are based on an exponentially smaller dataset in each round.
+The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
 of the true probability of an EOR BRAVO audit stopping in the $j$th round,
 given that it has not already stopped in a previous round. 
-In [Figure 1] we see that, especially in earlier rounds for which 
+In \ref{fig:eor_bravo_sprob}, we see that, especially in earlier rounds for which 
 the values are more representative of true audit behavior, 
 our predictions are accurate.
-In particular the average across all margins is just above $.9=90\%$ for
+In particular, the average across all margins is just above $.9=90\%$ for
 all three rounds.
 
-The total risk across all $5$ rounds for this audit is arguably
-the most important metric to consider.
-The risk of an audit is the probability that the audit
-stops despite an underlying tie. 
-%That is, the risk $R$ of audit $\mathcal{A}$ is 
-%$$R(\mathcal{A})=Pr[\mathcal{A}=\text{Correct}\mid H_0].$$
-Therefore, we now consider the proportion of audits with an underlying
-tie.
+The risk of this audit, across all $5$ rounds, is an important metric since it determines whether an audit is risk-limiting.
+Therefore, we now consider the proportion of audits that stopped with an underlying tie.
+This proportion, for a risk-limiting audit, should approach a value less than the risk limit, $0.1$, as more audits are performed.
 
-%\includegraphics[scale=.5]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\includegraphics[scale=.7]{eor_bravo_90perc_10^4_corrected/total_risk_with_alpha.png}
+\begin{figure}
+\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
+\caption{For each state margin, this plot shows
+the proportion of EOR BRAVO audits with an underlying
+tie that stopped.}
+\label{fig:eor_bravo_risk}
+\end{figure}
 
-[Fig 2] shows, for each state margin,
-the proportion of EOR BRAVO audits with an underlying
-tie which stopped before the conclusion of the $5$ rounds to the 
-total number of audits, $10^4$.
-
-[Fig 2] makes it clear that the risk of EOR BRAVO is roughly
+Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR BRAVO is roughly
 an order of magnitude less than the risk limit. 
@@
 % \subsection{Selection-Ordered BRAVO}
-% next_sample_size code is slow... tie simulations running still
+% next_sample_size code is slow... tie simulations running still!
 
@@
 As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying announced
+per state for both the underlying tie and underlying reported
 outcome.
 
-\includegraphics[scale=.7]{minerva_multiround_1x_10^4/total_risk.png}
+\begin{figure}
+\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
+\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.}
+\label{fig:minerva1_risk}
+\end{figure}
 
-[Fig above] shows that all simulations had a risk of less than 
-$.1$, the risk limit. 
-Unlike EOR BRAVO, the risks here are much closer to the risk limit,
-showing that Minerva stops on average with sharper risk.
+Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
+the claim that Minerva is a risk limiting audit. 
+Unlike EOR BRAVO, the experimental risks here are much closer to the risk limit,
+showing that Minerva stops on average with a less conservative risk; Minerva is sharper.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with the announced results as the underlying distribution
+in the Minerva simulations with a round size multiple of $1.0$.}
+\label{fig:minerva1_sprob}
+\end{figure}
+
+Figure~\ref{fig:minerva1_sprob} shows that the round size estimate for the first round achieved the desired
+stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
+probability, but it was within  roughly $10\%$.
 
@@
 
-%\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
+%\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^6/total_risk.png}
 
+\begin{figure}
+\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
+\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.}
+\label{fig:minerva1p5_risk}
+\end{figure}
+
+Figure~\ref{fig:minerva1p5_risk} shows, for each state's margin, the proportion of audits that stopped of
+all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state's margin, the proportion of audits that stopped in each round
+with the announced results as the underlying distribution
+in the Minerva simulations with a round size multiple of $1.5$.}
+\label{fig:minerva1p5_sprob}
+\end{figure}
+
+Figure~\ref{fig:minerva1p5_sprob} shows that the round size estimate for the first round achieved the desired
+stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
+probability, but it was within  roughly $10\%$. 
+% we could suggest that predicting accurate multiples is possible (same curve rather than line)
+
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+% \bibliographystyle{splncs04}
+% \bibliography{mybibliography}
+%
+\begin{thebibliography}{8}
+\bibitem{ref_article1}
+Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+
+\bibitem{ref_lncs1}
+Author, F., Author, S.: Title of a proceedings paper. In: Editor,
+F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
+Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+
+\bibitem{ref_book1}
+Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
+Location (1999)
+
+\bibitem{ref_proc1}
+Author, A.-B.: Contribution title. In: 9th International Proceedings
+on Proceedings, pp. 1--2. Publisher, Location (2010)
+
+\bibitem{ref_url1}
+LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
+Oct 2017
+\end{thebibliography}
 \end{document}

### 2021-10-21 MultipleRoundSims/draft.tex
replace audit names with fancy text

--- 2021-10-21 MultipleRoundSims/samplepaper.tex
@@
 \usepackage{float}
+\usepackage{xspace}
 \usepackage{graphicx}
 \graphicspath{{./imgs/}}
+\usepackage{comment}
 % Used for displaying a sample figure. If possible, figure files should
@@
 \newcommand{\R}{{{R2}}\xspace}
-\newcommand{\Bravo}{\textsc{BRAVO}\xspace}
+\newcommand{\BRAVO}{\textsc{Bravo}\xspace}
 
@@
 %
+\begin{comment}
 \author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
@@
 %
+\end{comment}
 \maketitle              % typeset the header of the contribution
@@
     ballot-polling RLAs.
-    BRAVO [citation] has long been the standard ballot-polling RLA,
-    while Minerva was recently introduced with the claim
+    \BRAVO [citation] has long been the standard ballot-polling RLA,
+    while \Minerva was recently introduced with the claim
     that fewer ballots on average are necessary to conclude 
@@
     we present results
-    from simulations of multiple round audits using Minerva 
-    as well as BRAVO 
+    from simulations of multiple round audits using \Minerva 
+    as well as \BRAVO 
     (both Selection-Ordered 
-    BRAVO and End-of-Round BRAVO).
+    \BRAVO and End-of-Round \BRAVO).
     The simulation results agree within reasonable error with
     the mathematical properties claimed in [Minerva paper].
-    On average, BRAVO audits are unnecessarily conservative 
-    while Minerva audits stop with fewer ballots. We also
-    present details on software implementing Minerva and
+    On average, \BRAVO audits are unnecessarily conservative 
+    while \Minerva audits stop with fewer ballots. We also
+    present details on software implementing \Minerva and
     the simulations themselves.
@@
 rounds.
-BRAVO is a B2 audit. 
-When an audit is performed in rounds, BRAVO can have its
-stopping condition applied once at the end of each round
-(End-of Round (EoR) BRAVO).
+\BRAVO is a B2 audit. 
+When an audit is performed in rounds, \BRAVO can have its
+stopping condition applied once at the end of each round: End-of Round (EoR) \BRAVO.
 This method ignores the possibility of stopping earlier in the 
@@
 Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 BRAVO stopping condition can 
-be applied retroactively to all subsamples 
-(Selection-Ordered (SO) BRAVO).
-SO BRAVO requires fewer ballots on average than EoR BRAVO but
+by election officials and the B2 \BRAVO stopping condition can 
+be applied retroactively to all subsamples: Selection-Ordered (SO) \BRAVO.
+SO \BRAVO requires fewer ballots on average than EoR \BRAVO but
 requies the work of tracking the order of ballots rather than
 just their tally.
-Minerva was designed for R2 audits and applies its stopping rule
+\Minerva was designed for R2 audits and applies its stopping rule
 once for each round.
-Minerva, by design then, requires fewer ballots to be sampled when 
+By design then, \Minerva requires fewer ballots to be sampled when 
 an audit is performed in rounds.
@@
 
-\subsection{End-of-Round BRAVO}
-For the EoR BRAVO simulations, our software estimated and used for each round
+\subsection{End-of-Round \BRAVO}
+For the EoR \BRAVO simulations, our software estimated and used for each round
 the minimum round size that would achieve a $90\%$ probability of stopping.
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
-For the first three rounds of the EoR BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
+For the first three rounds of the EoR \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
 to all audits which had not yet stopped before the $j$th round.}
@@
 The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
-of the true probability of an EOR BRAVO audit stopping in the $j$th round,
+of the true probability of an EOR \BRAVO audit stopping in the $j$th round,
 given that it has not already stopped in a previous round. 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
 \caption{For each state margin, this plot shows
-the proportion of EOR BRAVO audits with an underlying
+the proportion of EOR \BRAVO audits with an underlying
 tie that stopped.}
@@
 
-Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR BRAVO is roughly
+Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR \BRAVO is roughly
 an order of magnitude less than the risk limit. 
@@
 with the audit still meeting the risk limit.
-These simulations support the claim made in [Minerva paper]
-that EOR BRAVO is unnecessarily conservative and thus requires
+These simulations support the claim made in [\Minerva paper]
+that EOR \BRAVO is unnecessarily conservative and thus requires
 more ballots on average.
@@
 
-\subsection{Minerva}
-For Minerva, it has not been shown that round sizes can be chosen
+\subsection{\Minerva Simulations}
+For \Minerva, it has not been shown that round sizes can be chosen
 during the audit. That is, an adversary with knowledge of the history
@@
 risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a Minerva 
+For this reason, we have to choose the round sizes of a \Minerva 
 audit a priori.
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
@@
 Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
-the claim that Minerva is a risk limiting audit. 
-Unlike EOR BRAVO, the experimental risks here are much closer to the risk limit,
-showing that Minerva stops on average with a less conservative risk; Minerva is sharper.
+the claim that \Minerva is a risk limiting audit. 
+Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
+showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
@@
 all $10^4$ audits with the announced results as the underlying distribution
-in the Minerva simulations with a round size multiple of $1.0$.}
+in the \Minerva simulations with a round size multiple of $1.0$.}
 \label{fig:minerva1_sprob}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
@@
 with the announced results as the underlying distribution
-in the Minerva simulations with a round size multiple of $1.5$.}
+in the \Minerva simulations with a round size multiple of $1.5$.}
 \label{fig:minerva1p5_sprob}

### 2021-10-22 MultipleRoundSims/draft.tex
wording
comments
One more sentence about ballot-polling RLAs

@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
+% This is a draft of the coming paper "Multiple Round Ballot Polling Risk-Limiting Audit Simulations".
+% 2021.
 %
@@
 \usepackage{comment}
-% Used for displaying a sample figure. If possible, figure files should
+
+% \usepackage{graphicx}  --> Used for displaying a sample figure. If possible, figure files should
 % be included in EPS format.
@@
 \newcommand{\BRAVO}{\textsc{Bravo}\xspace}
+
+\usepackage{color}
+\definecolor{green}{rgb}{0.31, 0.78, 0.47}
+\newcommand{\fpo}[1]{\textcolor{green}{#1}}
 
@@
     In Ballot-polling RLAs, samples of ballots are drawn and tallied.
+    \fpo{Then the sample tally is (statistically) compared with announced tally.}
+
     In this paper we present an audit simulation framework
@@
 square of the margin, so 
-smaller margins are computationally more expensive to simulate.
+smaller margins are computationally much more expensive to simulate.
 For each of these states, we simulated 

### 2021-10-24 MultipleRoundSims/Extras.tex
Rewrote abstract, changed title, added authors and affiliations

--- 2021-10-21 MultipleRoundSims/draft.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{float}
-\usepackage{xspace}
-\usepackage{graphicx}
-\graphicspath{{./imgs/}}
-\usepackage{comment}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-% nice looking audit titles
-\newcommand{\Minerva}{\textsc{Minerva}\xspace}
-\newcommand{\B}{{{B2}}\xspace}
-\newcommand{\R}{{{R2}}\xspace}
-\newcommand{\BRAVO}{\textsc{Bravo}\xspace}
-
-\begin{document}
-%
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
-%add \thanks{} within the title brackets if want to refer to supporting organization ^^
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\begin{comment}
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\end{comment}
-\maketitle              % typeset the header of the contribution
-%
 \begin{abstract}
@@
 \end{abstract}
-%
-%
-%
-
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-\BRAVO is a B2 audit. 
-When an audit is performed in rounds, \BRAVO can have its
-stopping condition applied once at the end of each round: End-of Round (EoR) \BRAVO.
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 \BRAVO stopping condition can 
-be applied retroactively to all subsamples: Selection-Ordered (SO) \BRAVO.
-SO \BRAVO requires fewer ballots on average than EoR \BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-\Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-By design then, \Minerva requires fewer ballots to be sampled when 
-an audit is performed in rounds.
-
-Now we present relevant definitions.
-We consider a two candidate plurality contest.
-To begin we define an audit.
-\begin{definition}
-An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
-following decisions
-\begin{enumerate}
-\item
-$Correct$: the audit is complete
-\item
-$Uncertain$: continue the audit
-\end{enumerate}
-\end{definition}
-All of the audits discussed in this paper are modeled as binary hypothesis tests.
-Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
-That is, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or a the announced winner losing by one vote if there is an odd number of total ballots).
-Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The risk of an audit is the probability that an audit stops when a tie has occurred.
-\begin{definition}[Risk]
-The risk $R$ of an audit $\mathcal{A}$ is
-$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
-\end{definition}
-This leads us to the following simple definition of an $\alpha$-RLA.
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff 
-$$R(\mathcal{A}) \le \alpha.$$
-\end{definition}
-
-It is useful to discuss the probability of an audit stopping in
-some round, if the outcome is correctly announced.
-\begin{definition}[Stopping Probability]
-The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
-\end{definition}
-The notion of stopping probability can be useful for selecting round sizes.
-
-\section{Simulator}
-\subsection{Simulations to Support Theoretical Audit Properties}
-
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other samples require subsequent
-rounds in order to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
-
-\subsection{Software for Simulations}
-Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
-all written in Python.
-For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
-of the minimum round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
-Ballots can be sampled from any distribution of the users choosing. 
-It is often useful to consider the distribution of ballots corresponding to a tie and the 
-distribution of ballots corresponding with the announced results; these are the distributions represented
-by the null and alternative hypotheses.
-After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
-If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
-The abstract simulator class does not prescribe any one method for choosing round sizes. 
-We implement several classes to support various round size choices: 
-round sizes from an estimate to achieve a desired probability of stopping, 
-predetermined round sizes, and random round sizes.
-
-\section{Simulation Results}
-
-For this paper, we simulated audits for 
-the 2020 Presidential election
-in all US states whose margin was at least $5\%$.
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits with the announced
-underyling ballot distribution
-and an additional $10,000=10^4$ audits with a tie
-as the underlying ballot distribution.
-These are reasonable choices for inital simulation experiments
-because most audits frame the stopping decision as a binary
-hypothesis test where the null hypothesis assumes an underyling tie
-and the alternative hypothesis assumes the announced distribution.
-A standard first round size in ballot-polling audits
-has been one which achieves a $90\%$ probability
-of stopping, and we chose round sizes to reflect this standard
-in our simulations.
-We ran our simulations for up to five rounds.
-With a $90\%$ stopping probability, 
-an audit only has a $a$ probablity of not stopping
-by the end of the fifth round, assuming the outcome was correctly
-announced.
-
-\subsection{End-of-Round \BRAVO}
-For the EoR \BRAVO simulations, our software estimated and used for each round
-the minimum round size that would achieve a $90\%$ probability of stopping.
-For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the proportion of audits 
-that stopped in round $j$ among those that had not stopped before round $j$.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
-For the first three rounds of the EoR \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
-to all audits which had not yet stopped before the $j$th round.}
-\label{fig:eor_bravo_sprob}
-\end{figure}
-
-In Figure \ref{fig:eor_bravo_sprob}, we display proportions for only the first three rounds
-since very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
-make it to the $jth$ round.
-As  a result, the proportions are based on an exponentially smaller dataset in each round.
-The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
-of the true probability of an EOR \BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round. 
-In \ref{fig:eor_bravo_sprob}, we see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our predictions are accurate.
-In particular, the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-The risk of this audit, across all $5$ rounds, is an important metric since it determines whether an audit is risk-limiting.
-Therefore, we now consider the proportion of audits that stopped with an underlying tie.
-This proportion, for a risk-limiting audit, should approach a value less than the risk limit, $0.1$, as more audits are performed.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\caption{For each state margin, this plot shows
-the proportion of EOR \BRAVO audits with an underlying
-tie that stopped.}
-\label{fig:eor_bravo_risk}
-\end{figure}
-
-Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR \BRAVO is roughly
-an order of magnitude less than the risk limit. 
-That is, on average, roughly $10$ times more audits could have stopped 
-with the audit still meeting the risk limit.
-These simulations support the claim made in [\Minerva paper]
-that EOR \BRAVO is unnecessarily conservative and thus requires
-more ballots on average.
-
-% \subsection{Selection-Ordered BRAVO}
-% next_sample_size code is slow... tie simulations running still!
-
-\subsection{\Minerva Simulations}
-For \Minerva, it has not been shown that round sizes can be chosen
-during the audit. That is, an adversary with knowledge of the history
-of the audit may be able to choose round sizes which cause the 
-risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a \Minerva 
-audit a priori.
-For this paper, we consider two choices of round sizes.
-For both, we estimate and then use the minimum first round size 
-which achieves
-a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) 
-draw the same number of ballots in each round or (ii)
-multiply the previous round size by a factor of $1.5$ and 
-sample this many new ballots.
-We consider the case of drawing samples of the same size
-because it may reflect a practical way to continue an
-audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same number of 
-ballots in subsequent rounds may be practical.
-We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping.
-
-\subsubsection{Round Sizes with Multiple of $1.0$}
-
-As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying reported
-outcome.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_risk}
-\end{figure}
-
-Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
-the claim that \Minerva is a risk limiting audit. 
-Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
-showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_sprob}
-\end{figure}
-
-Figure~\ref{fig:minerva1_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$.
-
-\subsubsection{Round Sizes with Multiple of $1.5$}
-
-%For the Minerva simulations with a round size multiple of $1.5$,
-%we increased the number of simulations to $10^6$ per state for 
-%both an underlying tie and underlying announced outcome.
-
-%\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^6/total_risk.png}
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.}
-\label{fig:minerva1p5_risk}
-\end{figure}
-
-Figure~\ref{fig:minerva1p5_risk} shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped in each round
-with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.5$.}
-\label{fig:minerva1p5_sprob}
-\end{figure}
-
-Figure~\ref{fig:minerva1p5_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$. 
-% we could suggest that predicting accurate multiples is possible (same curve rather than line)
-
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}

### 2021-10-24 MultipleRoundSims/draft.tex
Rewrote abstract, changed title, added authors and affiliations

@@
 %
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
+\title{Simulations of Ballot Polling Risk-Limiting Audits}
 %add \thanks{} within the title brackets if want to refer to supporting organization ^^
@@
 %
-\begin{comment}
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+%\begin{comment}
+\author{Oliver Broadrick\inst{1} \and
+Sarah Morin\inst{1} \and
+Grant McClearn\inst{2} \and
+Filip Zagorski\inst{3} \and Neal McBurnett \and Poorvi L. Vora\inst{1}}
 %
-\authorrunning{F. Author et al.}
+\authorrunning{Broadrick, Morin et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\institute{Department of Computer Science, The George Washington University 
+\email{obroadrick@gwmail.gwu.edu}\\
+\and
+Department of Computer Science, Stanford University 
+\email{obroadrick@gwmail.gwu.edu}\\
+\and Wroclaw University of Science and Technology\\
+\email{filip.zagorski@gmail.com}}
 %
-\end{comment}
+%\end{comment}
 \maketitle              % typeset the header of the contribution
@@
 \begin{abstract}
-    Risk-Limiting Audits (RLAs) guarantee with known probability 
-    that if the outcome of an 
-    election is incorrectly announced, it will be detected, 
-    and a full hand recount will be performed. 
-    In Ballot-polling RLAs, samples of ballots are drawn and tallied.
-    \fpo{Then the sample tally is (statistically) compared with announced tally.}
+    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling RLAs \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO.
+We also
+    present details on the R2B2 open source software library and the related simulation software.
 
-    In this paper we present an audit simulation framework
-    that can be used to support theoretical results of audits.
-    We present such experimental results for multiple round 
-    ballot-polling RLAs.
-    \BRAVO [citation] has long been the standard ballot-polling RLA,
-    while \Minerva was recently introduced with the claim
-    that fewer ballots on average are necessary to conclude 
-    an audit.
-    [Minerva paper] presented experimental
-    results for one round audits, and here
-    we present results
-    from simulations of multiple round audits using \Minerva 
-    as well as \BRAVO 
-    (both Selection-Ordered 
-    \BRAVO and End-of-Round \BRAVO).
-    The simulation results agree within reasonable error with
-    the mathematical properties claimed in [Minerva paper].
-    On average, \BRAVO audits are unnecessarily conservative 
-    while \Minerva audits stop with fewer ballots. We also
-    present details on software implementing \Minerva and
-    the simulations themselves.
+    \BRAVO is the most commonly used ballot-polling RLA.
+\Minerva was recently proposed, and requires fewer first-round ballots, on average, than both Selection-Ordered \BRAVO and End-of-Round \BRAVO when the first-round stopping probability is large.
+
+    An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.
+In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $0.9$, a common choice among election officials.
+Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).
+
+    We show that the simulation results are consistent with predictions of the R2B2 open-source library for ballot polling audits.
+We also observe that both \BRAVO audits are unnecessarily conservative
+    while \Minerva audits stop with fewer ballots.
 

### 2021-10-29 MultipleRoundSims/draft.tex
organizing intro to motivate problem

@@
 %
-%\begin{comment}
+\begin{comment}
 \author{Oliver Broadrick\inst{1} \and
 Sarah Morin\inst{1} \and
-Grant McClearn\inst{2} \and
-Filip Zagorski\inst{3} \and Neal McBurnett \and Poorvi L. Vora\inst{1}}
+Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
+Filip Zagorski\inst{3} }
 %
@@
 Department of Computer Science, Stanford University 
-\email{obroadrick@gwmail.gwu.edu}\\
+\email{grantmcc@stanford.edu}\\
 \and Wroclaw University of Science and Technology\\
@@
 %
-%\end{comment}
+\end{comment}
 \maketitle              % typeset the header of the contribution
@@
 
-    \BRAVO is the most commonly used ballot-polling RLA.
+    \BRAVO is the most commonly used ballot-polling risk limiting audit (RLA).
 \Minerva was recently proposed, and requires fewer first-round ballots, on average, than both Selection-Ordered \BRAVO and End-of-Round \BRAVO when the first-round stopping probability is large.
@@
 
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-\BRAVO is a B2 audit. 
-When an audit is performed in rounds, \BRAVO can have its
-stopping condition applied once at the end of each round: End-of Round (EoR) \BRAVO.
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 \BRAVO stopping condition can 
-be applied retroactively to all subsamples: Selection-Ordered (SO) \BRAVO.
-SO \BRAVO requires fewer ballots on average than EoR \BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-\Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-By design then, \Minerva requires fewer ballots to be sampled when 
-an audit is performed in rounds.
+\section{Introduction}
+\subsection{Background}
+In the general ballot-polling risk limiting audit (RLA), a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
+A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit.
+Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
+The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.
 
+The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced, and stopping decisions are made after each ballot draw.
+In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2.
+There are two obvious approaches.
+The B2 stopping condition can be applied once at the end of each round: End-of Round (EoR) \BRAVO.
+Alternatively, the order of ballots in the sample can be tracked by election officials and the B2 \BRAVO stopping condition can be applied retroactively after each ballot drawn: Selection-Ordered (SO) \BRAVO.
+SO \BRAVO requires fewer ballots on average than EoR \BRAVO but requires the work of tracking the order of ballots rather than just their tally.
+
+\Minerva was designed for R2 audits and applies its stopping rule once for each round.
+Thus it does not require the tracking of ballots that SO \BRAVO does.
+It has been proven that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds.
+Computations of first-round size for a 0.9 stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO.
+First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability 0.9.
+There are no results, either theoretical or based on simulations, regarding the expected number of ballots drawn in a \Minerva audit.
+Further, there is no literature comparing simulations of \Minerva and EoR or SO \BRAVO, or studying the sizes of multiple-round \Minerva audits.
+
+\subsection{Our Results}
+
+\subsection{Organization}
+
+\section{Definitions}
 Now we present relevant definitions.

### 2021-10-29 MultipleRoundSims/draft.tex
Re organized into section files

--- 2021-10-24 MultipleRoundSims/draft.tex
@@
 \begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling RLAs \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO.
+    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO.
 We also
@@
     We show that the simulation results are consistent with predictions of the R2B2 open-source library for ballot polling audits.
-We also observe that both \BRAVO audits are unnecessarily conservative
+We also observe that both \BRAVO audits are more conservative  than needed,
     while \Minerva audits stop with fewer ballots.
@@
 %
+\section{Introduction}
+\label{sec:intro}
+\input{sections/01-intro}
 
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-\BRAVO is a B2 audit. 
-When an audit is performed in rounds, \BRAVO can have its
-stopping condition applied once at the end of each round: End-of Round (EoR) \BRAVO.
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 \BRAVO stopping condition can 
-be applied retroactively to all subsamples: Selection-Ordered (SO) \BRAVO.
-SO \BRAVO requires fewer ballots on average than EoR \BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-\Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-By design then, \Minerva requires fewer ballots to be sampled when 
-an audit is performed in rounds.
+\section{Definitions}
+\label{sec:defs}
+\input{sections/02-defs}
 
-Now we present relevant definitions.
-We consider a two candidate plurality contest.
-To begin we define an audit.
-\begin{definition}
-An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
-following decisions
-\begin{enumerate}
-\item
-$Correct$: the audit is complete
-\item
-$Uncertain$: continue the audit
-\end{enumerate}
-\end{definition}
-All of the audits discussed in this paper are modeled as binary hypothesis tests.
-Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
-That is, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or a the announced winner losing by one vote if there is an odd number of total ballots).
-Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The risk of an audit is the probability that an audit stops when a tie has occurred.
-\begin{definition}[Risk]
-The risk $R$ of an audit $\mathcal{A}$ is
-$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
-\end{definition}
-This leads us to the following simple definition of an $\alpha$-RLA.
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff 
-$$R(\mathcal{A}) \le \alpha.$$
-\end{definition}
-
-It is useful to discuss the probability of an audit stopping in
-some round, if the outcome is correctly announced.
-\begin{definition}[Stopping Probability]
-The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
-\end{definition}
-The notion of stopping probability can be useful for selecting round sizes.
+\section{Related work}
+\label{sec:related}
+\input{sections/03-related}
 
 \section{Simulator}
-\subsection{Simulations to Support Theoretical Audit Properties}
-
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other samples require subsequent
-rounds in order to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
-
-\subsection{Software for Simulations}
-Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
-all written in Python.
-For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
-of the minimum round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
-Ballots can be sampled from any distribution of the users choosing. 
-It is often useful to consider the distribution of ballots corresponding to a tie and the 
-distribution of ballots corresponding with the announced results; these are the distributions represented
-by the null and alternative hypotheses.
-After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
-If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
-The abstract simulator class does not prescribe any one method for choosing round sizes. 
-We implement several classes to support various round size choices: 
-round sizes from an estimate to achieve a desired probability of stopping, 
-predetermined round sizes, and random round sizes.
+\label{sec:simulator}
+\input{sections/04-simulator}
 
 \section{Simulation Results}
+\label{sec:results}
+\input{sections/05-results}
 
-For this paper, we simulated audits for 
-the 2020 Presidential election
-in all US states whose margin was at least $5\%$.
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally much more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits with the announced
-underyling ballot distribution
-and an additional $10,000=10^4$ audits with a tie
-as the underlying ballot distribution.
-These are reasonable choices for inital simulation experiments
-because most audits frame the stopping decision as a binary
-hypothesis test where the null hypothesis assumes an underyling tie
-and the alternative hypothesis assumes the announced distribution.
-A standard first round size in ballot-polling audits
-has been one which achieves a $90\%$ probability
-of stopping, and we chose round sizes to reflect this standard
-in our simulations.
-We ran our simulations for up to five rounds.
-With a $90\%$ stopping probability, 
-an audit only has a $a$ probablity of not stopping
-by the end of the fifth round, assuming the outcome was correctly
-announced.
-
-\subsection{End-of-Round \BRAVO}
-For the EoR \BRAVO simulations, our software estimated and used for each round
-the minimum round size that would achieve a $90\%$ probability of stopping.
-For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the proportion of audits 
-that stopped in round $j$ among those that had not stopped before round $j$.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
-For the first three rounds of the EoR \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
-to all audits which had not yet stopped before the $j$th round.}
-\label{fig:eor_bravo_sprob}
-\end{figure}
-
-In Figure \ref{fig:eor_bravo_sprob}, we display proportions for only the first three rounds
-since very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
-make it to the $jth$ round.
-As  a result, the proportions are based on an exponentially smaller dataset in each round.
-The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
-of the true probability of an EOR \BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round. 
-In \ref{fig:eor_bravo_sprob}, we see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our predictions are accurate.
-In particular, the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-The risk of this audit, across all $5$ rounds, is an important metric since it determines whether an audit is risk-limiting.
-Therefore, we now consider the proportion of audits that stopped with an underlying tie.
-This proportion, for a risk-limiting audit, should approach a value less than the risk limit, $0.1$, as more audits are performed.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\caption{For each state margin, this plot shows
-the proportion of EOR \BRAVO audits with an underlying
-tie that stopped.}
-\label{fig:eor_bravo_risk}
-\end{figure}
-
-Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR \BRAVO is roughly
-an order of magnitude less than the risk limit. 
-That is, on average, roughly $10$ times more audits could have stopped 
-with the audit still meeting the risk limit.
-These simulations support the claim made in [\Minerva paper]
-that EOR \BRAVO is unnecessarily conservative and thus requires
-more ballots on average.
-
-% \subsection{Selection-Ordered BRAVO}
-% next_sample_size code is slow... tie simulations running still!
-
-\subsection{\Minerva Simulations}
-For \Minerva, it has not been shown that round sizes can be chosen
-during the audit. That is, an adversary with knowledge of the history
-of the audit may be able to choose round sizes which cause the 
-risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a \Minerva 
-audit a priori.
-For this paper, we consider two choices of round sizes.
-For both, we estimate and then use the minimum first round size 
-which achieves
-a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) 
-draw the same number of ballots in each round or (ii)
-multiply the previous round size by a factor of $1.5$ and 
-sample this many new ballots.
-We consider the case of drawing samples of the same size
-because it may reflect a practical way to continue an
-audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same number of 
-ballots in subsequent rounds may be practical.
-We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping.
-
-\subsubsection{Round Sizes with Multiple of $1.0$}
-
-As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying reported
-outcome.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_risk}
-\end{figure}
-
-Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
-the claim that \Minerva is a risk limiting audit. 
-Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
-showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_sprob}
-\end{figure}
-
-Figure~\ref{fig:minerva1_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$.
-
-\subsubsection{Round Sizes with Multiple of $1.5$}
-
-%For the Minerva simulations with a round size multiple of $1.5$,
-%we increased the number of simulations to $10^6$ per state for 
-%both an underlying tie and underlying announced outcome.
-
-%\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^6/total_risk.png}
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.}
-\label{fig:minerva1p5_risk}
-\end{figure}
-
-Figure~\ref{fig:minerva1p5_risk} shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped in each round
-with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.5$.}
-\label{fig:minerva1p5_sprob}
-\end{figure}
-
-Figure~\ref{fig:minerva1p5_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$. 
-% we could suggest that predicting accurate multiples is possible (same curve rather than line)
+\section{Conclusions and Future Work}
+\label{sec:conc}
+\input{sections/06-conc}
 

### 2021-10-29 MultipleRoundSims/sections/01-intro.tex
Re organized into section files

%Intro
Need motivating paragraph and two sentence summary of results here.

\subsection{Background}
In the general ballot-polling risk limiting audit (RLA), a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}. A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit. Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit. The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.

The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced, and stopping decisions are made after each ballot draw. In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2. There are two obvious approaches. The B2 stopping condition can be applied once at the end of each round: End-of Round (EoR) \BRAVO.  Alternatively, the order of ballots in the sample can be tracked by election officials and the B2 \BRAVO stopping condition can be applied retroactively after each ballot drawn: Selection-Ordered (SO) \BRAVO. SO \BRAVO requires fewer ballots on average than EoR \BRAVO but requires the work of tracking the order of ballots rather than just their tally.

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. It has been proven that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds. Computations of first-round size for a 0.9 stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO. First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability 0.9. There are no results, either theoretical or based on simulations, regarding the expected number of ballots drawn in a \Minerva audit. Further, there is no literature comparing simulations of \Minerva and EoR or SO \BRAVO, or studying the sizes of multiple-round \Minerva audits.

\subsection{Our Results}

\subsection{Organization}

### 2021-10-29 MultipleRoundSims/sections/02-defs.tex
Re organized into section files

--- 2021-10-20 MultipleRoundSims/MultipleRoundSims.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\graphicspath{{./imgs/}}
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
-\author{ }
-\date{ }
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-    Risk-Limiting Audits (RLAs) guarantee with known probability 
-    that if the outcome of an 
-    election is incorrectly announced, it will be detected, 
-    and a full hand recount will be performed. 
-    In Ballot-polling RLAs, samples of ballots are drawn and tallied.
-    In this paper we present an audit simulation framework
-    that can be used to confirm theoretical results of audits.
-    We present such experimental results for multiple round 
-    ballot-polling RLAs.
-    BRAVO [citation] has long been the standard ballot-polling RLA,
-    while Minerva was recently introduced with the claim
-    that fewer ballots on average are necessary to conclude 
-    an audit.
-    [Minerva paper] presented experimental
-    results for one round audits, and here
-    we present results
-    from simulations of multiple round audits using Minerva 
-    as well as BRAVO 
-    (both Selection-Ordered 
-    BRAVO and End-of-Round BRAVO).
-    The simulation results agree within reasonable error with
-    the mathematical properties claimed in [Minerva paper].
-    On average, BRAVO audits are unnecessarily conservative 
-    while Minerva audits stop with fewer ballots. We also
-    present details on software implementing Minerva and
-    the simulations themselves.
-\end{abstract}
-
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-BRAVO is a B2 audit. 
-When an audit is performed in rounds, BRAVO can have its
-stopping condition applied once at the end of each round
-(End-of Round (EoR) BRAVO).
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 BRAVO stopping condition can 
-be applied retroactively to all subsamples 
-(Selection-Ordered (SO) BRAVO).
-SO BRAVO requires fewer ballots on average than EoR BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-Minerva, by design then, requires fewer ballots to be sampled when 
-an audit is performed in rounds.
-
+%Definitions
 Now we present relevant definitions.
@@
 \end{definition}
-The notion of stopping probability can be useful when selecting round sizes.
-
-\section{Simulations to Confirm Theoretical Results}
-
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other unlucky samples require subsequent
-rounds to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
-
-For this paper, we simulated audits for 
-the 2020 Presidential election
-in all US states whose margin was at least $5\%$.
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits with the announced
-underyling ballot distribution
-and an additional $10,000=10^4$ audits with a tie
-as the underlying ballot distribution.
-These are reasonable choices for inital simulation experiments
-because most audits frame the stopping decision as a binary
-hypothesis test where the null hypothesis assumes an underyling tie
-and the alternative hypothesis assumes the announced distribution.
-A standard first round size in ballot-polling audits
-has been one which achieves a $90\%$ probability
-of stopping, and we chose round sizes to reflect this standard
-in our simulations.
-We ran our simulations for up to five rounds.
-With a $90\%$ stopping probability, 
-an audit only has a $a$ probablity of not stopping
-by the end of the fifth round, assuming the outcome was correctly
-announced.
-
-\subsection{End-of-Round BRAVO}
-For the EoR BRAVO simulations, our software estimated for each round
-the round size that would achieve a $90\%$ probability of stopping.
-For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the probability
-of stopping given that an audit had continued through round $j$.
-That is, among the audits remaining at the beginning of round $j$, 
-we found the proportion which then stopped in round $j$.
-%This metric shows whether our software accurately estimated the minimum number of ballots that achieve the $90\%$ probability of stopping.
-
-%\includegraphics[scale=.6]{eor_bravo_90perc_10^4_corrected/cond_sprob/r1_cond_sprob.png}
-\begin{center}
-\includegraphics[scale=.41]{eor_bravo_90perc_10^4_corrected/cond_sprob_first_3_rounds.png}
-\end{center}
-[Figure 1] shows the proportion of audits that stopped in the $j$th round
-to all audits which had not already stopped in a preceding round.
-This proportion is shown for each state whose margin is given on the 
-horizontal axis.
-Here we display only the first three rounds
-since very few audits, $(.1)^{j-1}$ on average, 
-make it to the $jth$ round, 
-and as a result the data becomes exponentially less representative
-of the true probabilities.
-
-The proportion shown in [Fig 1] gives an estimate
-of the true probability of an EOR BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round. 
-In [Figure 1] we see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our predictions are accurate.
-In particular the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-The total risk across all $5$ rounds for this audit is arguably
-the most important metric to consider.
-The risk of an audit is the probability that the audit
-stops despite an underlying tie. 
-%That is, the risk $R$ of audit $\mathcal{A}$ is 
-%$$R(\mathcal{A})=Pr[\mathcal{A}=\text{Correct}\mid H_0].$$
-Therefore, we now consider the proportion of audits with an underlying
-tie.
-
-%\includegraphics[scale=.5]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\includegraphics[scale=.7]{eor_bravo_90perc_10^4_corrected/total_risk_with_alpha.png}
-
-[Fig 2] shows, for each state margin,
-the proportion of EOR BRAVO audits with an underlying
-tie which stopped before the conclusion of the $5$ rounds to the 
-total number of audits, $10^4$.
-
-[Fig 2] makes it clear that the risk of EOR BRAVO is roughly
-an order of magnitude less than the risk limit. 
-That is, on average, roughly $10$ times more audits could have stopped 
-with the audit still meeting the risk limit.
-These simulations support the claim made in [Minerva paper]
-that EOR BRAVO is unnecessarily conservative and thus requires
-more ballots on average.
-
-% \subsection{Selection-Ordered BRAVO}
-% next_sample_size code is slow... tie simulations running still
-
-\subsection{Minerva}
-For Minerva, it has not been shown that round sizes can be chosen
-during the audit. That is, an adversary with knowledge of the history
-of the audit may be able to choose round sizes which cause the 
-risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a Minerva 
-audit a priori.
-For this paper, we consider two choices of round sizes.
-For both, we estimate and then use the minimum first round size 
-which achieves
-a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) 
-draw the same number of ballots in each round or (ii)
-multiply the previous round size by a factor of $1.5$ and 
-sample this many new ballots.
-We consider the case of drawing samples of the same size
-because it may reflect a practical way to continue an
-audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same number of 
-ballots in subsequent rounds may be practical.
-We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping.
-
-\subsubsection{Round Sizes with Multiple of $1.0$}
-
-As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying announced
-outcome.
-
-\includegraphics[scale=.7]{minerva_multiround_1x_10^4/total_risk.png}
-
-[Fig above] shows that all simulations had a risk of less than 
-$.1$, the risk limit. 
-Unlike EOR BRAVO, the risks here are much closer to the risk limit,
-showing that Minerva stops on average with sharper risk.
-
-\subsubsection{Round Sizes with Multiple of $1.5$}
-
-%For the Minerva simulations with a round size multiple of $1.5$,
-%we increased the number of simulations to $10^6$ per state for 
-%both an underlying tie and underlying announced outcome.
-
-%\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
-
-\end{document}
+The notion of stopping probability can be useful for selecting round sizes.

### 2021-10-29 MultipleRoundSims/sections/04-simulator.tex
Re organized into section files

--- 2021-10-20 MultipleRoundSims/MultipleRoundSims.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\graphicspath{{./imgs/}}
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
-\author{ }
-\date{ }
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-    Risk-Limiting Audits (RLAs) guarantee with known probability 
-    that if the outcome of an 
-    election is incorrectly announced, it will be detected, 
-    and a full hand recount will be performed. 
-    In Ballot-polling RLAs, samples of ballots are drawn and tallied.
-    In this paper we present an audit simulation framework
-    that can be used to confirm theoretical results of audits.
-    We present such experimental results for multiple round 
-    ballot-polling RLAs.
-    BRAVO [citation] has long been the standard ballot-polling RLA,
-    while Minerva was recently introduced with the claim
-    that fewer ballots on average are necessary to conclude 
-    an audit.
-    [Minerva paper] presented experimental
-    results for one round audits, and here
-    we present results
-    from simulations of multiple round audits using Minerva 
-    as well as BRAVO 
-    (both Selection-Ordered 
-    BRAVO and End-of-Round BRAVO).
-    The simulation results agree within reasonable error with
-    the mathematical properties claimed in [Minerva paper].
-    On average, BRAVO audits are unnecessarily conservative 
-    while Minerva audits stop with fewer ballots. We also
-    present details on software implementing Minerva and
-    the simulations themselves.
-\end{abstract}
-
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-BRAVO is a B2 audit. 
-When an audit is performed in rounds, BRAVO can have its
-stopping condition applied once at the end of each round
-(End-of Round (EoR) BRAVO).
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 BRAVO stopping condition can 
-be applied retroactively to all subsamples 
-(Selection-Ordered (SO) BRAVO).
-SO BRAVO requires fewer ballots on average than EoR BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-Minerva, by design then, requires fewer ballots to be sampled when 
-an audit is performed in rounds.
-
-Now we present relevant definitions.
-We consider a two candidate plurality contest.
-To begin we define an audit.
-\begin{definition}
-An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
-following decisions
-\begin{enumerate}
-\item
-$Correct$: the audit is complete
-\item
-$Uncertain$: continue the audit
-\end{enumerate}
-\end{definition}
-Audits in this paper are all formed as a binary hypothesis test.
-Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
-That is, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or a the announced winner losing by one vote if there is an odd number of total ballots).
-Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The risk of an audit is the probability that an audit stops when a tie has occurred.
-\begin{definition}[Risk]
-The risk $R$ of an audit $\mathcal{A}$ is
-$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
-\end{definition}
-This leads us to the following simple definition of an $\alpha$-RLA.
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff 
-$$R(\mathcal{A}) \le \alpha.$$
-\end{definition}
-
-It is useful to discuss the probability of an audit stopping in
-some round, if the outcome is correctly announced.
-\begin{definition}[Stopping Probability]
-The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
-\end{definition}
-The notion of stopping probability can be useful when selecting round sizes.
-
-\section{Simulations to Confirm Theoretical Results}
+%Simulator
+\subsection{Simulations to Support Theoretical Audit Properties}
 
@@
 
-For this paper, we simulated audits for 
-the 2020 Presidential election
-in all US states whose margin was at least $5\%$.
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits with the announced
-underyling ballot distribution
-and an additional $10,000=10^4$ audits with a tie
-as the underlying ballot distribution.
-These are reasonable choices for inital simulation experiments
-because most audits frame the stopping decision as a binary
-hypothesis test where the null hypothesis assumes an underyling tie
-and the alternative hypothesis assumes the announced distribution.
-A standard first round size in ballot-polling audits
-has been one which achieves a $90\%$ probability
-of stopping, and we chose round sizes to reflect this standard
-in our simulations.
-We ran our simulations for up to five rounds.
-With a $90\%$ stopping probability, 
-an audit only has a $a$ probablity of not stopping
-by the end of the fifth round, assuming the outcome was correctly
-announced.
-
-\subsection{End-of-Round BRAVO}
-For the EoR BRAVO simulations, our software estimated for each round
-the round size that would achieve a $90\%$ probability of stopping.
-For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the probability
-of stopping given that an audit had continued through round $j$.
-That is, among the audits remaining at the beginning of round $j$, 
-we found the proportion which then stopped in round $j$.
-%This metric shows whether our software accurately estimated the minimum number of ballots that achieve the $90\%$ probability of stopping.
-
-%\includegraphics[scale=.6]{eor_bravo_90perc_10^4_corrected/cond_sprob/r1_cond_sprob.png}
-\begin{center}
-\includegraphics[scale=.41]{eor_bravo_90perc_10^4_corrected/cond_sprob_first_3_rounds.png}
-\end{center}
-[Figure 1] shows the proportion of audits that stopped in the $j$th round
-to all audits which had not already stopped in a preceding round.
-This proportion is shown for each state whose margin is given on the 
-horizontal axis.
-Here we display only the first three rounds
-since very few audits, $(.1)^{j-1}$ on average, 
-make it to the $jth$ round, 
-and as a result the data becomes exponentially less representative
-of the true probabilities.
-
-The proportion shown in [Fig 1] gives an estimate
-of the true probability of an EOR BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round. 
-In [Figure 1] we see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our predictions are accurate.
-In particular the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-The total risk across all $5$ rounds for this audit is arguably
-the most important metric to consider.
-The risk of an audit is the probability that the audit
-stops despite an underlying tie. 
-%That is, the risk $R$ of audit $\mathcal{A}$ is 
-%$$R(\mathcal{A})=Pr[\mathcal{A}=\text{Correct}\mid H_0].$$
-Therefore, we now consider the proportion of audits with an underlying
-tie.
-
-%\includegraphics[scale=.5]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\includegraphics[scale=.7]{eor_bravo_90perc_10^4_corrected/total_risk_with_alpha.png}
-
-[Fig 2] shows, for each state margin,
-the proportion of EOR BRAVO audits with an underlying
-tie which stopped before the conclusion of the $5$ rounds to the 
-total number of audits, $10^4$.
-
-[Fig 2] makes it clear that the risk of EOR BRAVO is roughly
-an order of magnitude less than the risk limit. 
-That is, on average, roughly $10$ times more audits could have stopped 
-with the audit still meeting the risk limit.
-These simulations support the claim made in [Minerva paper]
-that EOR BRAVO is unnecessarily conservative and thus requires
-more ballots on average.
-
-% \subsection{Selection-Ordered BRAVO}
-% next_sample_size code is slow... tie simulations running still
-
-\subsection{Minerva}
-For Minerva, it has not been shown that round sizes can be chosen
-during the audit. That is, an adversary with knowledge of the history
-of the audit may be able to choose round sizes which cause the 
-risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a Minerva 
-audit a priori.
-For this paper, we consider two choices of round sizes.
-For both, we estimate and then use the minimum first round size 
-which achieves
-a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) 
-draw the same number of ballots in each round or (ii)
-multiply the previous round size by a factor of $1.5$ and 
-sample this many new ballots.
-We consider the case of drawing samples of the same size
-because it may reflect a practical way to continue an
-audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same number of 
-ballots in subsequent rounds may be practical.
-We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping.
-
-\subsubsection{Round Sizes with Multiple of $1.0$}
-
-As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying announced
-outcome.
-
-\includegraphics[scale=.7]{minerva_multiround_1x_10^4/total_risk.png}
-
-[Fig above] shows that all simulations had a risk of less than 
-$.1$, the risk limit. 
-Unlike EOR BRAVO, the risks here are much closer to the risk limit,
-showing that Minerva stops on average with sharper risk.
-
-\subsubsection{Round Sizes with Multiple of $1.5$}
-
-%For the Minerva simulations with a round size multiple of $1.5$,
-%we increased the number of simulations to $10^6$ per state for 
-%both an underlying tie and underlying announced outcome.
-
-%\includegraphics[scale=.7]{minerva_multiround_1p5x_10^6/total_risk.png}
-
-\end{document}
+\subsection{Software for Simulations}
+Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
+all written in Python.
+For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
+of the minimum round size to achieve a desired stopping probability. 
+For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
+Ballots can be sampled from any distribution of the users choosing. 
+It is often useful to consider the distribution of ballots corresponding to a tie and the 
+distribution of ballots corresponding with the announced results; these are the distributions represented
+by the null and alternative hypotheses.
+After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
+If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
+The abstract simulator class does not prescribe any one method for choosing round sizes. 
+We implement several classes to support various round size choices: 
+round sizes from an estimate to achieve a desired probability of stopping, 
+predetermined round sizes, and random round sizes.

### 2021-10-29 MultipleRoundSims/sections/05-results.tex
Re organized into section files

--- 2021-10-21 MultipleRoundSims/samplepaper.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{float}
-\usepackage{graphicx}
-\graphicspath{{./imgs/}}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-% nice looking audit titles
-\newcommand{\Minerva}{\textsc{Minerva}\xspace}
-\newcommand{\B}{{{B2}}\xspace}
-\newcommand{\R}{{{R2}}\xspace}
-\newcommand{\Bravo}{\textsc{BRAVO}\xspace}
-
-\begin{document}
-%
-\title{Multiple Round Ballot Polling Risk-Limiting Audit Simulations}
-%add \thanks{} within the title brackets if want to refer to supporting organization ^^
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-    Risk-Limiting Audits (RLAs) guarantee with known probability 
-    that if the outcome of an 
-    election is incorrectly announced, it will be detected, 
-    and a full hand recount will be performed. 
-    In Ballot-polling RLAs, samples of ballots are drawn and tallied.
-    In this paper we present an audit simulation framework
-    that can be used to support theoretical results of audits.
-    We present such experimental results for multiple round 
-    ballot-polling RLAs.
-    BRAVO [citation] has long been the standard ballot-polling RLA,
-    while Minerva was recently introduced with the claim
-    that fewer ballots on average are necessary to conclude 
-    an audit.
-    [Minerva paper] presented experimental
-    results for one round audits, and here
-    we present results
-    from simulations of multiple round audits using Minerva 
-    as well as BRAVO 
-    (both Selection-Ordered 
-    BRAVO and End-of-Round BRAVO).
-    The simulation results agree within reasonable error with
-    the mathematical properties claimed in [Minerva paper].
-    On average, BRAVO audits are unnecessarily conservative 
-    while Minerva audits stop with fewer ballots. We also
-    present details on software implementing Minerva and
-    the simulations themselves.
-
-\keywords{Risk-limiting audit  \and Ballot polling audit}
-\end{abstract}
-%
-%
-%
-
-\section{Introduction: B2 and R2 Ballot Polling RLAs}
-In ballot-polling RLAs, samples of ballots are drawn and tallied
-in rounds
-after each of which a statistical measure determines whether to
-continue. 
-Some ballot-polling audits (B2 audits) 
-are designed to make a stopping decision
-after each individual ballot is sampled.
-Other ballot-polling audits (R2 audits) assume many ballots are drawn
-at once and make a stopping decision at the conclusion of each 
-so-called round.
-In practice, election officials draw many ballots at once.
-For this reason, R2 audits are used or B2 audits are applied to 
-rounds.
-BRAVO is a B2 audit. 
-When an audit is performed in rounds, BRAVO can have its
-stopping condition applied once at the end of each round
-(End-of Round (EoR) BRAVO).
-This method ignores the possibility of stopping earlier in the 
-sample and thus requires more ballots on average to sample.
-Alternatively, the order of ballots in the sample can be tracked
-by election officials and the B2 BRAVO stopping condition can 
-be applied retroactively to all subsamples 
-(Selection-Ordered (SO) BRAVO).
-SO BRAVO requires fewer ballots on average than EoR BRAVO but
-requies the work of tracking the order of ballots rather than
-just their tally.
-Minerva was designed for R2 audits and applies its stopping rule
-once for each round.
-Minerva, by design then, requires fewer ballots to be sampled when 
-an audit is performed in rounds.
-
-Now we present relevant definitions.
-We consider a two candidate plurality contest.
-To begin we define an audit.
-\begin{definition}
-An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
-following decisions
-\begin{enumerate}
-\item
-$Correct$: the audit is complete
-\item
-$Uncertain$: continue the audit
-\end{enumerate}
-\end{definition}
-All of the audits discussed in this paper are modeled as binary hypothesis tests.
-Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
-That is, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or a the announced winner losing by one vote if there is an odd number of total ballots).
-Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The risk of an audit is the probability that an audit stops when a tie has occurred.
-\begin{definition}[Risk]
-The risk $R$ of an audit $\mathcal{A}$ is
-$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
-\end{definition}
-This leads us to the following simple definition of an $\alpha$-RLA.
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff 
-$$R(\mathcal{A}) \le \alpha.$$
-\end{definition}
-
-It is useful to discuss the probability of an audit stopping in
-some round, if the outcome is correctly announced.
-\begin{definition}[Stopping Probability]
-The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
-\end{definition}
-The notion of stopping probability can be useful for selecting round sizes.
-
-\section{Simulator}
-\subsection{Simulations to Support Theoretical Audit Properties}
-
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other samples require subsequent
-rounds in order to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
-
-\subsection{Software for Simulations}
-Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
-all written in Python.
-For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
-of the minimum round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
-Ballots can be sampled from any distribution of the users choosing. 
-It is often useful to consider the distribution of ballots corresponding to a tie and the 
-distribution of ballots corresponding with the announced results; these are the distributions represented
-by the null and alternative hypotheses.
-After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
-If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
-The abstract simulator class does not prescribe any one method for choosing round sizes. 
-We implement several classes to support various round size choices: 
-round sizes from an estimate to achieve a desired probability of stopping, 
-predetermined round sizes, and random round sizes.
-
-\section{Simulation Results}
-
+%Results
 For this paper, we simulated audits for 
@@
 % we could suggest that predicting accurate multiples is possible (same curve rather than line)
-
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}

### 2021-10-29 MultipleRoundSims/sections/04-software.tex
added material on R2B2
changed definition of stopping prob

%Software
In this section we describe the software implementing ballot polling audits and the simulator, termed the R2B2 library. All the software is released as open source under the MIT License.
\subsection{R2B2 Library}

The R2B2 Python library provides a framework for the exploration of round-by-round
and ballot-by-ballot Risk-Limiting Audits (RLAs). The goal in designing R2B2 is two fold:
\begin{enumerate}
    \item Provide an elegant Python library which can be easily imported and used
    in any other code base.
    \item Provide an interactive set of tools which can be utilized `out-of-the-box'
    for experimenting with and learning about risk-limiting audits.
\end{enumerate}
The following describes the design, usage and possible next steps for the R2B2 library.

\subsection{Design}

The high-level design of R2B2 is an object-oriented view of election audits.
The three main object classes, \code{Election}, \code{Contest}, and \code{Audit},
-serve to group data into logically independent structures.

\subsubsection{Election}
The \code{Election} contains the information that comprises an entire election,
most importantly, the total number of ballots cast in the election and the list
of \code{Contest}s from the election. At the moment \code{Election} does not offer
functionality beyond grouping \code{Contest}s; potential functionality discussed
in 1.3 Future Work.

\subsubsection{Contest}
The \code{Contest} contains the information related to a single
contest such as the ballots cast in that contest, the candidates, the type of contest,
and the reported tally. Providing a structure to hold this information independent of
any particular audit is especially useful for exploratory work.

\subsubsection{Audit}

The \code{Audit} contains information related to a single contests audit parameters
such as the risk limit, sampling method, and \code{Contest} to audit. It is important
to note the \code{Audit} is an Abstract Base Class upon which specific RLAs are built.
It only contains the parameters and attributes common to all RLAs and provides a set
of methods that can be called by any audit implementation. The functionality of
\code{Audit} can be divided into two basic groups: \textit{interactive}
and \textit{bulk}.\\
\\
The interactive implementation allows users to execute an audit step-by-step as it
might progress during a live election audit through the following:

\begin{itemize}
    \item The \code{run()} method begins an interactive audit executing where users
    are prompted for round sizes and the counts of winner ballots found in the sample
    and in return are given information about the current state of the audit and whether
    the stopping condition(s) have been met.
    \item Two distributions representing the null and alternative hypotheses are maintained
    and allow for computation of the audits per-round risk and stopping probability
    schedules.
    \item Before each round, the audit will recommend possible next round sizes given
    different criteria, such as a set of desired stopping probabilities.
\end{itemize}
The bulk implementations allows users to generate a larger set of data from an audit
such as:

\begin{itemize}
    \item A set of stopping conditions given a set of round sizes.
    \item A set of risk levels given a set of round size and winner ballots pairs.
    \item A list of all stopping conditions from the minimum to the maximum round size.
\end{itemize}

\subsection{Usage}

R2B2 makes understanding and exploring election audits simple for the user with no
Python knowledge while simultaneously providing a comprehensive set of tools for
the experienced Python developer.

\subsubsection{Python Library}

Using R2B2 is as simple as using any other Python library: simply import the library
and all of the functionality is at your finger tips. Not only does this allow users
to write their own Python scripts for exploring RLAs, it also allows R2B2 to be plugged
in to any other Python library. See the following Jupyter Notebooks for information on
the usage of R2B2:

\begin{itemize}
    \item \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/R2B2%20Basics.ipynb}{Basic Usage}
    \item \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/Generating%20Graphs.ipynb}{Generating Graphs}
\end{itemize}

\subsubsection{Stand-Alone Tool}

R2B2 provides a significant amount of functionality `out-of-the-box' for educational
or exploratory use. For those who wish learn about RLAs without having to write any
code themselves, R2B2 provides a command line tool for both interactive auditing and
generating audit results and statistics for larger data sets.

\subsection{Simulations to Support Theoretical Audit Properties}

The outcomes of RLAs depend on random chance; some random samples
support the alternative hypothesis more than expected, resulting in 
quick low-risk conclusions, while other samples require subsequent
rounds in order to confirm the announced results.
We can simulate random samples for various underlying ballot 
distributions by computing pseudorandom samples. 
By applying an audit's stopping condition to thousands of such
simulated samples, the average behavior of the simulated
audits will tend towards the true behavior of the audit.
In this way, we can examine whether theoretical claims about an audit are
actually correct.

\subsection{Software for Simulations}
Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
all written in Python.
For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
of the minimum round size to achieve a desired stopping probability. 
For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
Ballots can be sampled from any distribution of the users choosing. 
It is often useful to consider the distribution of ballots corresponding to a tie and the 
distribution of ballots corresponding with the announced results; these are the distributions represented
by the null and alternative hypotheses.
After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
The abstract simulator class does not prescribe any one method for choosing round sizes. 
We implement several classes to support various round size choices: 
round sizes from an estimate to achieve a desired probability of stopping, 
predetermined round sizes, and random round sizes.

### 2021-10-29 MultipleRoundSims/sections/R2B2.tex
added material on R2B2
changed definition of stopping prob

--- 2021-10-29 MultipleRoundSims/sections/04-software.tex
@@
-%Software
-In this section we describe the software implementing ballot polling audits and the simulator, termed the R2B2 library. All the software is released as open source under the MIT License.
-\subsection{R2B2 Library}
+\section{R2B2 Library}
 
@@
 generating audit results and statistics for larger data sets.
-
-\subsection{Simulations to Support Theoretical Audit Properties}
-
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other samples require subsequent
-rounds in order to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
-
-\subsection{Software for Simulations}
-Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
-all written in Python.
-For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
-of the minimum round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
-Ballots can be sampled from any distribution of the users choosing. 
-It is often useful to consider the distribution of ballots corresponding to a tie and the 
-distribution of ballots corresponding with the announced results; these are the distributions represented
-by the null and alternative hypotheses.
-After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
-If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
-The abstract simulator class does not prescribe any one method for choosing round sizes. 
-We implement several classes to support various round size choices: 
-round sizes from an estimate to achieve a desired probability of stopping, 
-predetermined round sizes, and random round sizes.

### 2021-10-30 MultipleRoundSims/sections/05-results.tex
Added subsection on Selection-Ordered BRAVO
- plots
- figures, captions
- explanation

@@
 
-% \subsection{Selection-Ordered BRAVO}
+\subsection{Selection-Ordered \BRAVO}
 % next_sample_size code is slow... tie simulations running still!
+For the SO \BRAVO simulations, our software estimated and used for each round
+the minimum round size that would achieve a $90\%$ probability of stopping.
+For the simulations with the announced outcome as the underlying
+ballot distribution, for each round $j$ we computed the proportion of audits 
+that stopped in round $j$ among those that had not stopped before round $j$.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
+For the first three rounds of the SO \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
+to all audits which had not yet stopped before the $j$th round.}
+\label{fig:so_bravo_sprob}
+\end{figure}
+
+In Figure \ref{fig:so_bravo_sprob}, we again display proportions for only the first three rounds.
+The proportions shown in Figure \ref{fig:so_bravo_sprob}, like those in \ref{fig:eor_bravo_sprob}, give an estimate
+of the true probability of an SO \BRAVO audit stopping in the $j$th round,
+given that it has not already stopped in a previous round. 
+In \ref{fig:so_bravo_sprob}, we see that our round size predictions are relatively accurate,
+all three rounds being near $90\%$.
+
+Again, we will next consider the proportion of audits that stopped with an underlying tie over all five rounds.
+This proportion, for a risk-limiting audit, should approach a value less than the risk limit, 
+$0.1$, as more audits are performed.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{so_bravo_90perc_10^4/total_risk.png}
+\caption{For each state margin, this plot shows
+the proportion of SO \BRAVO audits with an underlying
+tie that stopped in any of the five rounds.}
+\label{fig:so_bravo_risk}
+\end{figure}
+
+In Figure~\ref{fig:so_bravo_risk} we show only the results for the $13$
+states whose simulations with an underlying tie have finished running.
+To estimate the minimum next round size that achieves a desired stopping probability,
+the SO \BRAVO software generates the probability distribution ballot by ballot since
+the stopping condition needs to be evaluated for each individual ballot drawn.
+This means, when an underyling tie causes audits to quickly move on to larger rounds, 
+the simulations are computationally expensive. 
+Regardless, the claim that these simulations support, that SO \BRAVO is a Risk-Limiting Audit,
+is a proven and well known result.
+Notice in Figure~\ref{fig:so_bravo_risk},
+that the risk of SO \BRAVO is much
+nearer the risk limit than that of EoR \BRAVO.
+
+%TODO SO and EoR macros
 

### 2021-10-30 MultipleRoundSims/sections/04-software.tex
edited software file

@@
 %Software
-In this section we describe the software implementing ballot polling audits and the simulator, termed the R2B2 library. All the software is released as open source under the MIT License.
+In this section we describe the software implementing ballot polling audits, termed the R2B2 library, and the simulator software. All the software is released as open source under the MIT License.
 \subsection{R2B2 Library}
@@
 The R2B2 Python library provides a framework for the exploration of round-by-round
-and ballot-by-ballot Risk-Limiting Audits (RLAs). The goal in designing R2B2 is two fold:
+and ballot-by-ballot RLAs. The goal in designing R2B2 is two fold:
 \begin{enumerate}
@@
 \end{enumerate}
-The following describes the design, usage and possible next steps for the R2B2 library.
 
-\subsection{Design}
+\subsubsection{Design}
 
@@
 
-\subsubsection{Election}
 The \code{Election} contains the information that comprises an entire election,
@@
 of \code{Contest}s from the election. At the moment \code{Election} does not offer
-functionality beyond grouping \code{Contest}s; potential functionality discussed
-in 1.3 Future Work.
+functionality beyond grouping \code{Contest}s.
 
-\subsubsection{Contest}
 The \code{Contest} contains the information related to a single
@@
 
-\subsubsection{Audit}
-
-The \code{Audit} contains information related to a single contests audit parameters
+The \code{Audit} contains information related to the audit parameters for a single contest, 
 such as the risk limit, sampling method, and \code{Contest} to audit. It is important
@@
 
-\subsection{Usage}
+\subsubsection{Usage}
 
@@
 the experienced Python developer.
-
-\subsubsection{Python Library}
 
@@
 
-\subsubsection{Stand-Alone Tool}
-
-R2B2 provides a significant amount of functionality `out-of-the-box' for educational
+R2B2 also provides a significant amount of functionality `out-of-the-box' for educational
 or exploratory use. For those who wish learn about RLAs without having to write any

### 2021-10-30 MultipleRoundSims/sections/01-intro.tex
bib file, changes to intro, reordered related work and defs

@@
 %Intro
-Need motivating paragraph and two sentence summary of results here.
+The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances.
+For this reason,
+{\em evidence-based elections} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability.
+One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA).
+This paper provides insight into the main approach to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
+While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
+This paper examines the number of ballots drawn over multiple rounds of both audits, and also examines the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.
 
 \subsection{Background}
+This paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology.
 In the general ballot-polling risk limiting audit (RLA), a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
@@
 It has been proven that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds.
-Computations of first-round size for a 0.9 stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO.
-First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability 0.9.
-There are no results, either theoretical or based on simulations, regarding the expected number of ballots drawn in a \Minerva audit.
-Further, there is no literature comparing simulations of \Minerva and EoR or SO \BRAVO, or studying the sizes of multiple-round \Minerva audits.
+Simulations have also been presented to understand better its first round properties; these show that \Minerva also draws fewer ballots than SO \BRAVO for large first rounds.
+There are no results, however, either theoretical or based on simulations, regarding the expected number of ballots drawn over multiple rounds in a \Minerva audit.
+Further, there is no literature comparing the number of ballots drawn by \Minerva and EoR or SO \BRAVO over multiple rounds.
+
+Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, available for use in real election audits.
+Both have been used in real election audits.
+For this reason, it is very important to understand their properties over multiple rounds.
 

### 2021-10-30 MultipleRoundSims/sections/02-related.tex
bib file, changes to intro, reordered related work and defs

%Related Work
Computations of first-round size for a 0.9 stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO. First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability 0.9.

### 2021-10-30 MultipleRoundSims/sections/03-defs.tex
bib file, changes to intro, reordered related work and defs

--- 2021-10-29 MultipleRoundSims/sections/02-defs.tex
@@
 We consider a two candidate plurality contest.
-To begin we define an audit.
+To begin we define an audit. We assume that ballots are sampled with replacement. 
 \begin{definition}
@@
 Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or a the announced winner losing by one vote if there is an odd number of total ballots).
+(or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
 Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The risk of an audit is the probability that an audit stops when a tie has occurred.
+The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. 
 \begin{definition}[Risk]
-The risk $R$ of an audit $\mathcal{A}$ is
+The maximum risk $R$ of an audit $\mathcal{A}$ is
 $$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
@@
 It is useful to discuss the probability of an audit stopping in
-some round, if the outcome is correctly announced.
+the $j^{th}$ round, given that the underlying election is as announced.
 \begin{definition}[Stopping Probability]
 The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X_j)=Correct \mid H_a]$$
+$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
 \end{definition}

### 2021-10-30 MultipleRoundSims/sections/04-software.tex
bib file, changes to intro, reordered related work and defs

@@
 %Software
-In this section we describe the software implementing ballot polling audits, termed the R2B2 library, and the simulator software. All the software is released as open source under the MIT License.
+In this section we describe the software implementing ballot polling audits, termed the R2B2 library, and the simulator software used for this research.
+All the software is released as open source under the MIT License.
 \subsection{R2B2 Library}
 
-The R2B2 Python library provides a framework for the exploration of round-by-round
+The R2B2 Python library \cite{r2b2} provides a framework for the exploration of round-by-round
 and ballot-by-ballot RLAs. The goal in designing R2B2 is two fold:
@@
 The three main object classes, \code{Election}, \code{Contest}, and \code{Audit},
--serve to group data into logically independent structures.
+serve to group data into logically independent structures.
 
@@
 to note the \code{Audit} is an Abstract Base Class upon which specific RLAs are built.
-It only contains the parameters and attributes common to all RLAs and provides a set
+It only contains the parameters and attributes common to the RLAs of this paper and provides a set
 of methods that can be called by any audit implementation. The functionality of
@@
 
-\subsection{Simulations to Support Theoretical Audit Properties}
+\subsection{Simulation Software}
+As described above, our open source audit software library R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
+all written in Python.
+For each of these audits, the software can compute the stopping condition for a given sample and estimates
+of the next round size to achieve a desired stopping probability. 
+For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, [need to check].
+given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).
 
-The outcomes of RLAs depend on random chance; some random samples
-support the alternative hypothesis more than expected, resulting in 
-quick low-risk conclusions, while other samples require subsequent
-rounds in order to confirm the announced results.
-We can simulate random samples for various underlying ballot 
-distributions by computing pseudorandom samples. 
-By applying an audit's stopping condition to thousands of such
-simulated samples, the average behavior of the simulated
-audits will tend towards the true behavior of the audit.
-In this way, we can examine whether theoretical claims about an audit are
-actually correct.
+When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes. 
+It is often useful to consider two underlying elections: one corresponding to a tie between the pair, and the other corresponding to the results as announced.
+These are the probability distributions represented
+by the null and alternative hypotheses.
 
-\subsection{Software for Simulations}
-Our open source audit software library r2b2 [link] has implementations of several ballot polling risk-limiting audits as well as a simulator, 
-all written in Python.
-For each of these audits, the software can evaluate the stopping condition for a given sample and can give estimates
-of the minimum round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples using the pseudorandom number generator, [need to check].
-Ballots can be sampled from any distribution of the users choosing. 
-It is often useful to consider the distribution of ballots corresponding to a tie and the 
-distribution of ballots corresponding with the announced results; these are the distributions represented
-by the null and alternative hypotheses.
-After drawing a sample, the simulator then evaluates the given audit's stopping condition for this simulated sample.
+After drawing a simulated sample of ballots, the simulator evaluates the given audit's stopping condition for this sample.
 If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
@@
 round sizes from an estimate to achieve a desired probability of stopping, 
-predetermined round sizes, and random round sizes.
+predetermined round sizes, and pseudorandomly-generated round sizes.
+
+\subsection{Testing}
+
+The R2B2 (python) software is used to compute stopping conditions and next round estimates, it is intended for use by us and other researchers, and designed for this purpose.
+We have also independently implemented all the functionality in matlab (the two codebases are written by different individuals) and have extensively checked the results of the two codebases.
+Additionally, for use in regular election audits by election officials, we have written an add-on to the {\em Arlo} risk-limiting audit software, which has also been extensively checked against these two codebases.

### 2021-10-30 MultipleRoundSims/sections/02-related.tex
Related work

@@
 %Related Work
-Computations of first-round size for a 0.9 stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO.
-First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability 0.9.
+The \BRAVO audit \cite{bravo} is a well-known ballot polling audit which has been used in numerous pilot and real audits.
+When used to audit a two-candidate election, it is an instance of Wald's sequential probability ratio test (SPRT) \cite{wald}, and inherits the SPRT property of being the most efficient test (requiring the smallest expected number of ballots) if the election is as announced.
+The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken.
+Real election audits invest in drawing a large number of ballots before making the decision.
+It is possible to apply \BRAVO to the sequence of ballots if the sequential order is retained.
+This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to a set of ballots drawn earlier.
+
+We do know a great deal about the properties of \BRAVO.
+The risk limiting property of \BRAVO follows from the similar property of the SPRT.
+Stopping probabilities for \BRAVO may be computed as described by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena}.
+The results of the computations match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
+
+The \Minerva audit \cite{usenix_minerva,arxiv_athena} was developed for large first round sizes which enable election officials to be done in one round with large probability.
+It uses information from the entire sample, and has been proven to be risk limiting when the round schedule for the audit is determined before the audit begins.
+That is, information about the actual ballots drawn in the first round cannot be used to determine future round sizes.
+First-round sizes for a $0.9$ stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO.
+First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability $0.9$.
+
+Ballot polling audit simulations have been used to familiarize election officials and the public with the approach \cite{dice}.
+McLaughlin and Stark \cite{mclaughlin_thesis,simulations_house} compare the workload for the Canvass Audits by Sampling and Testing (CAST) and Kaplan-Markov (KM) audits using simulations.
+Blom {\em et al.} demonstrate the efficiency of their ballot polling approach to audit instant runoff voting (IRV) using simulations \cite{blom_IRV}.
+Huang {\em et al.} present a framework generalizing a number of ballot polling audits and compare their performance (round sizes and stopping probabilities) using simulations \cite{DBLP:conf/evoteid/HuangRSTV20}.
+This work was prior to the development of \Minerva, and focuses on the comparison between Bayesian audits \cite{bayesian-audits} and \BRAVO; essentially studying the impact of the prior of the Bayesian RLA.

### 2021-10-31 MultipleRoundSims/draft.tex
added material on R2B2
changed definition of stopping prob
bib file, changes to intro, reordered related work and defs
Related work
Done with a first edit

@@
 \usepackage{comment}
+\usepackage{listings,hyperref}
 
@@
 
+\def\code#1{\texttt{#1}}
+
 \begin{document}
+\lstset{language=Python}
 %
@@
 Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
-Filip Zagorski\inst{3} }
+Filip Zag{\'o}rski\inst{3} }
 %
@@
     An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.
-In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $0.9$, a common choice among election officials.
+In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $90\%$, a common choice among election officials.
 Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).
@@
 
-\section{Definitions}
-\label{sec:defs}
-\input{sections/02-defs}
-
 \section{Related work}
 \label{sec:related}
-\input{sections/03-related}
+\input{sections/02-related}
 
-\section{Simulator}
-\label{sec:simulator}
-\input{sections/04-simulator}
+\section{Software}
+\label{sec:software}
+\input{sections/03-software}
+
+\section{Experiments}
+\label{sec:expts}
+\input{sections/04-expts}
 
@@
 %
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
+\bibliographystyle{splncs04}
+\bibliography{audits}
 %
+\end{document}
+
 \begin{thebibliography}{8}
@@
 \end{thebibliography}
-\end{document}

### 2021-10-31 MultipleRoundSims/sections/03-software.tex
Done with a first edit

--- 2021-10-30 MultipleRoundSims/sections/04-software.tex
@@
 
-The R2B2 Python library \cite{r2b2} provides a framework for the exploration of round-by-round
+The R2B2 Python library \cite{r2b2_anon} provides a framework for the exploration of round-by-round
 and ballot-by-ballot RLAs. The goal in designing R2B2 is two fold:
@@
 \subsection{Simulation Software}
-As described above, our open source audit software library R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
+As described above R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
 all written in Python.
@@
 When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes.
-It is often useful to consider two underlying elections: one corresponding to a tie between the pair, and the other corresponding to the results as announced.
-These are the probability distributions represented
-by the null and alternative hypotheses.
 
@@
 
-The R2B2 (python) software is used to compute stopping conditions and next round estimates, it is intended for use by us and other researchers, and designed for this purpose.
-We have also independently implemented all the functionality in matlab (the two codebases are written by different individuals) and have extensively checked the results of the two codebases.
-Additionally, for use in regular election audits by election officials, we have written an add-on to the {\em Arlo} risk-limiting audit software, which has also been extensively checked against these two codebases.
+The R2B2 software is used to compute stopping conditions and next round estimates.
+It is intended for use by us and other researchers, and designed for this purpose.
+We have also independently implemented all the functionality in matlab \cite{brla_explore_anon} (the two codebases are written by different individuals) and have extensively checked the results of the two codebases.
+Additionally, for use in regular election audits by election officials, we have written an add-on \cite{athena_anon} to the {\em Arlo} risk-limiting audit software, the results of which have also been extensively checked against the other two codebases.

### 2021-10-31 MultipleRoundSims/sections/04-expts.tex
Done with a first edit

--- 2021-10-30 MultipleRoundSims/sections/03-defs.tex
@@
 %Definitions
-Now we present relevant definitions.
-We consider a two candidate plurality contest.
-To begin we define an audit. We assume that ballots are sampled with replacement. 
+In this section, we motivate and describe the experiments. We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
+
+We first present relevant definitions.
+
 \begin{definition}
-An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
+An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives one of the 
 following decisions
@@
 \end{definition}
+
 All of the audits discussed in this paper are modeled as binary hypothesis tests.
+
 Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
 That is, the true underlying ballot distribution is given by the announced ballot tallies.
+
 Under the null hypothesis, $H_0$, the true outcome is a tie 
 (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
-Now we define key attributes of an audit that we will consider while analyzing our simulations.
-The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. 
+
+The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora show that this is the maximum risk \cite{Bayesian-RLA}.) 
 \begin{definition}[Risk]
@@
 \end{definition}
-This leads us to the following simple definition of an $\alpha$-RLA.
+
+This leads us to the following definition of an $\alpha$-RLA.
 \begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
@@
 The notion of stopping probability can be useful for selecting round sizes.
+
+We performed simulations to study $S_j$ and obtain estimates of the expected number of ballots drawn for the various audits.
+We know that \BRAVO uses the smallest expected number of ballots when drawn ballot-by-ballot.
+However, this may not be the case when \BRAVO is used as an R2 audit for a large first round size.
+We also studied the risk values.
+While all the audits we studied have been proven to be RLAs, it is useful to observe how close to the risk limit each audit gets over a small, finite number of rounds as would be the case in a real election.

### 2021-10-31 MultipleRoundSims/sections/05-results.tex
Done with a first edit

--- 2021-10-29 MultipleRoundSims/sections/05-results.tex
@@
 the 2020 Presidential election
-in all US states whose margin was at least $5\%$.
+in all US states where the pairwise margin for the two main candidates was at least $5\%$. 
 Round sizes increase roughly proportional to the inverse
@@
 For each of these states, we simulated 
-$10,000=10^4$ audits with the announced
-underyling ballot distribution
-and an additional $10,000=10^4$ audits with a tie
-as the underlying ballot distribution.
-These are reasonable choices for inital simulation experiments
-because most audits frame the stopping decision as a binary
-hypothesis test where the null hypothesis assumes an underyling tie
-and the alternative hypothesis assumes the announced distribution.
-A standard first round size in ballot-polling audits
-has been one which achieves a $90\%$ probability
-of stopping, and we chose round sizes to reflect this standard
-in our simulations.
-We ran our simulations for up to five rounds.
-With a $90\%$ stopping probability, 
-an audit only has a $a$ probablity of not stopping
-by the end of the fifth round, assuming the outcome was correctly
-announced.
+$10,000=10^4$ audits assuming the underlying election was as announced,  
+and an additional $10,000=10^4$ audits assuming the underlying election was a tie.
+
+We ran simulations for a $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced.
+We ran our simulations for up to five rounds. We assume that the audit results in a hand count if it does not stop in any of $5$ rounds.
 
@@
 For the EoR \BRAVO simulations, our software estimated and used for each round
-the minimum round size that would achieve a $90\%$ probability of stopping.
-For the simulations with the announced outcome as the underlying
-ballot distribution, for each round $j$ we computed the proportion of audits 
-that stopped in round $j$ among those that had not stopped before round $j$.
+the round size that would achieve a $90\%$ probability of stopping if the true election outcome were as announced.
+In Figure \ref{fig:eor_bravo_sprob}, we display the proportion of EoR \BRAVO audits that stopped in the $j^{th}$ round
+to all audits which had not yet stopped before the $j^{th}$ round, for only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
+make it to the $j^{th}$ round. The proportions give an estimate
+of the true probability of an EoR \BRAVO audit stopping in the $j$th round,
+given that it has not already stopped in a previous round, when the underlying election is as announced. 
+We see that, especially in earlier rounds for which 
+the values are more representative of true audit behavior, 
+our round size predictions are accurate.
+In particular, the average across all margins is just above $.9=90\%$ for
+all three rounds.
 
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
-For the first three rounds of the EoR \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
-to all audits which had not yet stopped before the $j$th round.}
+We also study the proportion of audits that stopped when the underlying election was a tie.
+This proportion should approach a value less than the risk limit, $0.1$, as more audits are performed.
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
+This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
+as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
 \label{fig:eor_bravo_sprob}
@@
 
-In Figure \ref{fig:eor_bravo_sprob}, we display proportions for only the first three rounds
-since very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
-make it to the $jth$ round.
-As  a result, the proportions are based on an exponentially smaller dataset in each round.
-The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
-of the true probability of an EOR \BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round. 
-In \ref{fig:eor_bravo_sprob}, we see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our predictions are accurate.
-In particular, the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-The risk of this audit, across all $5$ rounds, is an important metric since it determines whether an audit is risk-limiting.
-Therefore, we now consider the proportion of audits that stopped with an underlying tie.
-This proportion, for a risk-limiting audit, should approach a value less than the risk limit, $0.1$, as more audits are performed.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\caption{For each state margin, this plot shows
-the proportion of EOR \BRAVO audits with an underlying
-tie that stopped.}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
+\caption{This plot shows, for each state margin,
+the fraction of EoR \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
 \label{fig:eor_bravo_risk}
@@
 
-Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR \BRAVO is roughly
+We observe (see Figure~\ref{fig:eor_bravo_risk}) that the risk of EoR \BRAVO is roughly
 an order of magnitude less than the risk limit. 
-That is, on average, roughly $10$ times more audits could have stopped 
-with the audit still meeting the risk limit.
-These simulations support the claim made in [\Minerva paper]
-that EOR \BRAVO is unnecessarily conservative and thus requires
-more ballots on average.
+These results are as expected, because EoR \BRAVO is known to be too conservative \cite{usenix_minerva}.
 
-% \subsection{Selection-Ordered BRAVO}
+\subsection{Selection-Ordered \BRAVO}
 % next_sample_size code is slow... tie simulations running still!
+For the SO \BRAVO simulations, our software estimated and used for each round
+the round size that would achieve a $90\%$ probability of stopping.
+
+In Figure \ref{fig:so_bravo_sprob}, we again display proportions of SO \BRAVO audits that stopped in the $j^{th}$ round
+to all audits which had not yet stopped before the $j^{th}$ round for only the first three rounds.
+The proportions shown in Figure \ref{fig:so_bravo_sprob}, like those in \ref{fig:eor_bravo_sprob}, give an estimate
+of the true probability of an SO \BRAVO audit stopping in the $j^{th}$ round,
+given that it has not already stopped in a previous round. 
+In \ref{fig:so_bravo_sprob}, we see that our round size predictions are relatively accurate,
+all three rounds being near $90\%$.
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
+This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
+as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
+\label{fig:so_bravo_sprob}
+\end{figure}
+
+As before, we will next consider the proportion of audits that stopped with an underlying tie over all five rounds.
+This proportion, for a risk-limiting audit, should approach a value less than the risk limit, 
+$0.1$, as more audits are performed.
+
+In Figure~\ref{fig:so_bravo_risk} we show only the results for the $13$
+states whose simulations with an underlying tie have finished running.
+To estimate the next round size that achieves a desired stopping probability,
+the SO \BRAVO software generates the probability distribution on the number of ballots in the sample ballot by ballot (see \cite{usenix_minerva}) since
+the stopping condition needs to be evaluated for each individual ballot drawn.
+because the underlying tied election causes audits to move on to larger rounds, 
+the simulations are computationally expensive. SO \BRAVO is proven to be a Risk-Limiting Audit,
+and we observe in Figure~\ref{fig:so_bravo_risk},
+that the risk of SO \BRAVO is much
+nearer the risk limit than that of EoR \BRAVO, as expected.
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/total_risk.png}
+\caption{This plot shows, for each state margin,
+the fraction of SO \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:so_bravo_risk}
+\end{figure}
+
+%TODO SO and EoR macros
 
 \subsection{\Minerva Simulations}
-For \Minerva, it has not been shown that round sizes can be chosen
-during the audit. That is, an adversary with knowledge of the history
-of the audit may be able to choose round sizes which cause the 
-risk of the audit to exceed the risk limit.
-For this reason, we have to choose the round sizes of a \Minerva 
-audit a priori.
+The proof that \Minerva is an RLA \cite{usenix_minerva} assumes that the round schedule is pre-determined. For this reason, we have to choose the round sizes of a \Minerva 
+audit {\em a priori}.
 For this paper, we consider two choices of round sizes.
-For both, we estimate and then use the minimum first round size 
-which achieves
+For both, we estimate and use a first round size 
+for
 a $90\%$ probability of stopping.
@@
 of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping.
-
-\subsubsection{Round Sizes with Multiple of $1.0$}
+round sizes with a $90\%$ probability of stopping for \Minerva.
 
 As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both the underlying tie and underlying reported
-outcome.
+per state for both an underlying election as reported and an underlying tied election.
 
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_risk}
+Figure~\ref{fig:minerva1_sprob} and Figure~\ref{fig:minerva1p5_sprob} show that the first round size estimates were fairly accurate for multipliers of $10.$ and $1.5$ respectively, with first round stopping probabilities being very close to $90\%$.
+For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively did not consistently achieve $90\%$ stopping
+probability, as they were not accurate estimates of $90\%$ stopping probabilities.
+Note that we chose a simple multiplier for future rounds, but one could make more accurate round size estimates before the audit begins.
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.0$.}
+\label{fig:minerva1_sprob}
 \end{figure}
 
-Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
-the claim that \Minerva is a risk limiting audit. 
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.5$.}
+\label{fig:minerva1p5_sprob}
+\end{figure}
+
+Figures \ref{fig:minerva1_risk} and Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples of $1.0$ and $1.5$ respectively, as would be expected for an RLA with risk limit $0.1$.
 Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
@@
 
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.0$.}
-\label{fig:minerva1_sprob}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
+\caption{This plot shows, for each state margin,
+the fraction of \Minerva audits with a round size multiple of $1.0$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:minerva1_risk}
 \end{figure}
-
-Figure~\ref{fig:minerva1_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$.
-
-\subsubsection{Round Sizes with Multiple of $1.5$}
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.}
+% we could suggest that predicting accurate multiples is possible (same curve rather than line)
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
+\caption{This plot shows, for each state margin,
+the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
 \label{fig:minerva1p5_risk}
 \end{figure}
-
-Figure~\ref{fig:minerva1p5_risk} shows, for each state's margin, the proportion of audits that stopped of
-all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.
-
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state's margin, the proportion of audits that stopped in each round
-with the announced results as the underlying distribution
-in the \Minerva simulations with a round size multiple of $1.5$.}
-\label{fig:minerva1p5_sprob}
-\end{figure}
-
-Figure~\ref{fig:minerva1p5_sprob} shows that the round size estimate for the first round achieved the desired
-stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
-probability, but it was within  roughly $10\%$. 
-% we could suggest that predicting accurate multiples is possible (same curve rather than line)

### 2021-11-01 MultipleRoundSims/sections/03-software.tex
Jupyter notebooks links

@@
 in to any other Python library. See the following Jupyter Notebooks for information on
-the usage of R2B2:
+the usage of R2B2: Basic Usage~\cite{jupyterBasic}, Generating Graphs~\cite{jupyterGraphs}.
 
-\begin{itemize}
-    \item \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/R2B2%20Basics.ipynb}{Basic Usage}
-    \item \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/Generating%20Graphs.ipynb}{Generating Graphs}
-\end{itemize}
+% \begin{itemize} 
+%     \item {Basic Usage} \cite{} % \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/R2B2%20Basics.ipynb}{Basic Usage}
+%     \item \href{https://github.com/gwexploratoryaudits/r2b2/blob/notebooks/notebook/Generating%20Graphs.ipynb}{Generating Graphs}
+% \end{itemize}
 

### 2021-11-02 MultipleRoundSims/draft.tex
definitions

@@
 \begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO.
+    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO and End-of-Round (EoR) \BRAVO.
 We also
@@
 
-    \BRAVO is the most commonly used ballot-polling RLA.
-\Minerva was recently proposed, and requires fewer first-round ballots, on average, than both Selection-Ordered \BRAVO and End-of-Round \BRAVO when the first-round stopping probability is large.
+    \BRAVO is the most commonly used ballot-polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
+In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
+\Minerva was recently proposed for use when the first-round stopping probability is large.
+It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
 
-    An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.
-In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $90\%$, a common choice among election officials.
+    An open question, however, is how \Minerva compares to SO \BRAVO and EoR \BRAVO over multiple rounds, and for lower stopping probabilities.
+In this paper, we present results from simulations of multiple round audits.
+We examine both: first-round stopping probabilities of $90\%$, a common choice among election officials, and $25\%$, which would be more favorable to \BRAVO.
 Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).
@@
 We show that the simulation results are consistent with predictions of the R2B2 open-source library for ballot polling audits.
-We also observe that both \BRAVO audits are more conservative  than needed,
-    while \Minerva audits stop with fewer ballots.
+We also observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots.
 
-\keywords{Risk-limiting audit  \and Ballot polling audit}
+\keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
 \end{abstract}

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
definitions

@@
 For this reason,
-{\em evidence-based elections} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability.
-One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA).
+{\em evidence-based elections} \cite{evidence-based} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability.
+One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA) \cite{RLA}.
+An RLA is an audit which guarantees that the probability of concluding that an election outcome is correct, given that it is not, is below a pre-determined value known as the risk limit of the audit, independent of the true, unknown vote distribution of the underlying election.
+Over a dozen states have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.
+
 This paper provides insight into the main approach to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
 While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
-This paper examines the number of ballots drawn over multiple rounds of both audits, and also examines the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.
+This paper examines the number of ballots drawn over multiple rounds of both audits, the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.
 
 \subsection{Background}
-This paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology.
-In the general ballot-polling risk limiting audit (RLA), a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
+Of the multiple types of RLAs, this paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology.
+In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
 A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit.
@@
 Thus it does not require the tracking of ballots that SO \BRAVO does.
-It has been proven that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds.
-Simulations have also been presented to understand better its first round properties; these show that \Minerva also draws fewer ballots than SO \BRAVO for large first rounds.
-There are no results, however, either theoretical or based on simulations, regarding the expected number of ballots drawn over multiple rounds in a \Minerva audit.
-Further, there is no literature comparing the number of ballots drawn by \Minerva and EoR or SO \BRAVO over multiple rounds.
+Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds and the round schedule is pre-determined (before any ballots are drawn).
+They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first rounds when the (true) underlying election is as announced.
+
+There are no results, either theoretical or based on simulations, regarding the number of ballots drawn over multiple rounds in a \Minerva audit with a pre-determined schedule.
+Because \BRAVO does not need to work on a pre-determined round schedule, it can optimize the size of the next round based on the sample drawn so far.
+Thus an open question is whether the constraint of a predetermined round schedule limits the efficacy of \Minerva in future rounds, and there is no literature comparing the number of ballots drawn by \Minerva and SO \BRAVO over multiple rounds.
+Note that the Average Sample Number (ASN) computations for \BRAVO \cite{bravo} apply only for B2 audits and are especially misleading as estimates of the number of ballots drawn over multiple rounds when first round sizes are large.
 
@@
 
-\subsection{Organization}
+\subsection{Organization} Section \ref{sec:related} describes related work and \ref{sec:software} our open source software.
+The experiments we performed are described in section \ref{sec:expts} and sections \ref{sec:results1} and \ref{sec:results2} present our results.
+Section \ref{sec:conc} has our conclusions.

### 2021-11-02 MultipleRoundSims/sections/02-related.tex
definitions

@@
 It is possible to apply \BRAVO to the sequence of ballots if the sequential order is retained.
-This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to a set of ballots drawn earlier.
+This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to ballots that were drawn earlier in the sample.
 

### 2021-11-02 MultipleRoundSims/sections/03-software.tex
definitions

@@
 \code{Audit} can be divided into two basic groups: \textit{interactive}
-and \textit{bulk}.\\
-\\
+and \textit{bulk}.
+
 The interactive implementation allows users to execute an audit step-by-step as it
@@
 R2B2 also provides a significant amount of functionality `out-of-the-box' for educational
-or exploratory use. For those who wish learn about RLAs without having to write any
+or exploratory use. For those who wish to learn about RLAs without having to write any
 code themselves, R2B2 provides a command line tool for both interactive auditing and
@@
 of the next round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, [need to check].
+For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator.
 given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
definitions

@@
 All of the audits discussed in this paper are modeled as binary hypothesis tests.
-
 Under the alternate hypothesis, $H_a$, the announced outcome is correct.
 That is, the true underlying ballot distribution is given by the announced ballot tallies.
-
-Under the null hypothesis, $H_0$, the true outcome is a tie 
-(or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
+Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
 
@@
 
-It is useful to discuss the probability of an audit stopping in
-the $j^{th}$ round, given that the underlying election is as announced.
+We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
 \begin{definition}[Stopping Probability]
-The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
+The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
 $$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
 \end{definition}
-The notion of stopping probability can be useful for selecting round sizes.
+Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
 
-We performed simulations to study $S_j$ and obtain estimates of the expected number of ballots drawn for the various audits.
-We know that \BRAVO uses the smallest expected number of ballots when drawn ballot-by-ballot.
-However, this may not be the case when \BRAVO is used as an R2 audit for a large first round size.
-We also studied the risk values.
-While all the audits we studied have been proven to be RLAs, it is useful to observe how close to the risk limit each audit gets over a small, finite number of rounds as would be the case in a real election.
+Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
+\begin{definition}[Cumulative Stopping Probability]
+The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is 
+$$C_j(\mathcal{A})= \sum_{i=1}^j S_j$$
+\end{definition}
+Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$.
+
+Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
+\begin{definition}[Conditional Stopping Probability]
+The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
+$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
+\end{definition}
+Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e. those that did not stop before round $j$.
+
+We simulated audits using margins from the 2020 Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
+Round sizes increase roughly proportional to the inverse
+square of the margin, so 
+smaller margins are computationally much more expensive to simulate.
+For each of these states, we simulated 
+$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
+and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$).
+
+We ran simulations for: (a) $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $25\%$ probability of stopping in the first round which is more favorable to \BRAVO.
+We ran our simulations for up to five rounds.
+For \Minerva, we chose two round schedules: one where the additional number of ballots chosen in a round is the same as in the previous round (multiplying factor of $1.0$) and the second where the multiplying factor is $1.5$ (this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $90\%$ conditional stopping probability in the second round for a first round stopping probability of $90\%$).
+For both versions of \BRAVO, we chose a single round schedule: each round size has the same conditional stopping probability as the first one.

### 2021-11-02 MultipleRoundSims/sections/05-results.tex
definitions

@@
 %Results
-For this paper, we simulated audits for 
-the 2020 Presidential election
-in all US states where the pairwise margin for the two main candidates was at least $5\%$. 
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally much more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits assuming the underlying election was as announced,  
-and an additional $10,000=10^4$ audits assuming the underlying election was a tie.
-
-We ran simulations for a $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced.
-We ran our simulations for up to five rounds. We assume that the audit results in a hand count if it does not stop in any of $5$ rounds.
-
 \subsection{End-of-Round \BRAVO}

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
Lots of stuff

@@
 
-We simulated audits using margins from the 2020 Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
+We simulated audits for a risk limit of $0.1$ using margins from the 2020 Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
 Round sizes increase roughly proportional to the inverse
@@
 We ran our simulations for up to five rounds.
-For \Minerva, we chose two round schedules: one where the additional number of ballots chosen in a round is the same as in the previous round (multiplying factor of $1.0$) and the second where the multiplying factor is $1.5$ (this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $90\%$ conditional stopping probability in the second round for a first round stopping probability of $90\%$).
+
+For rounds after the first one, we chose the round schedule as follows.
 For both versions of \BRAVO, we chose a single round schedule: each round size has the same conditional stopping probability as the first one.
+As the proof of the risk-limiting property of \Minerva assumes that its round schedule is determined before any ballots are drawn, we could not use this approach for \Minerva round sizes.
+Instead, we chose to compare two fixed round schedules for \Minerva: one where the additional number of ballots drawn in a round is the same as in the previous round (multiplying factor of $1.0$) and the second where the multiplying factor is $1.5$.
+We consider the case of drawing samples of the same size because it may reflect a practical way to continue an
+audit; if election officials have selected some first round size within
+reasonable logistical bounds, drawing the same number of 
+ballots in subsequent rounds may be practical.
+We also consider round sizes with samples increasing by a multiple
+of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $90\%$ conditional stopping probability in the second round for a first round stopping probability of $90\%$.

### 2021-11-02 MultipleRoundSims/sections/05-results1.tex
Lots of stuff

--- 2021-11-02 MultipleRoundSims/sections/05-results.tex
@@
 %Results
-\subsection{End-of-Round \BRAVO}
-For the EoR \BRAVO simulations, our software estimated and used for each round
-the round size that would achieve a $90\%$ probability of stopping if the true election outcome were as announced.
+\subsection{Stopping Probability as a Function of Round and Margin}
+For both SO and EoR \BRAVO simulations, our software estimated round sizes that would give $\chi_j(\mathcal{A}) = 0.9$ and used those for the simulations.
 In Figure \ref{fig:eor_bravo_sprob}, we display the proportion of EoR \BRAVO audits that stopped in the $j^{th}$ round
-to all audits which had not yet stopped before the $j^{th}$ round, for only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
-make it to the $j^{th}$ round. The proportions give an estimate
-of the true probability of an EoR \BRAVO audit stopping in the $j$th round,
-given that it has not already stopped in a previous round, when the underlying election is as announced. 
+to all audits which had not stopped before the $j^{th}$ round, for $j=1,2,3$.
+Though we carried out the simulations for $5$ rounds we show only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average,
+make it to the $j^{th}$ round for $j \geq 4$. In Figure \ref{fig:so_bravo_sprob}, we display the same proportions for SO \BRAVO audits. 
+In both cases, these proportions are estimates of the true value of $\chi_j(\mathcal{A})$ for $j=1,2,3$ as a function of margin. 
 We see that, especially in earlier rounds for which 
-the values are more representative of true audit behavior, 
-our round size predictions are accurate.
-In particular, the average across all margins is just above $.9=90\%$ for
-all three rounds.
-
-We also study the proportion of audits that stopped when the underlying election was a tie.
-This proportion should approach a value less than the risk limit, $0.1$, as more audits are performed.
+the values are more representative of true audit behavior because fewer simulated audits have stopped, 
+our round size predictions are accurate (the proportions are close to $0.9$).
 
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
-\caption{This plot shows, for each state margin,
-the fraction of EoR \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
-\label{fig:eor_bravo_risk}
-\end{figure}
-
-We observe (see Figure~\ref{fig:eor_bravo_risk}) that the risk of EoR \BRAVO is roughly
-an order of magnitude less than the risk limit. 
-These results are as expected, because EoR \BRAVO is known to be too conservative \cite{usenix_minerva}.
-
-\subsection{Selection-Ordered \BRAVO}
-% next_sample_size code is slow... tie simulations running still!
-For the SO \BRAVO simulations, our software estimated and used for each round
-the round size that would achieve a $90\%$ probability of stopping.
-
-In Figure \ref{fig:so_bravo_sprob}, we again display proportions of SO \BRAVO audits that stopped in the $j^{th}$ round
-to all audits which had not yet stopped before the $j^{th}$ round for only the first three rounds.
-The proportions shown in Figure \ref{fig:so_bravo_sprob}, like those in \ref{fig:eor_bravo_sprob}, give an estimate
-of the true probability of an SO \BRAVO audit stopping in the $j^{th}$ round,
-given that it has not already stopped in a previous round. 
-In \ref{fig:so_bravo_sprob}, we see that our round size predictions are relatively accurate,
-all three rounds being near $90\%$.
-
-\begin{figure}
 \includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
@@
 
-As before, we will next consider the proportion of audits that stopped with an underlying tie over all five rounds.
-This proportion, for a risk-limiting audit, should approach a value less than the risk limit, 
-$0.1$, as more audits are performed.
-
-In Figure~\ref{fig:so_bravo_risk} we show only the results for the $13$
-states whose simulations with an underlying tie have finished running.
-To estimate the next round size that achieves a desired stopping probability,
-the SO \BRAVO software generates the probability distribution on the number of ballots in the sample ballot by ballot (see \cite{usenix_minerva}) since
-the stopping condition needs to be evaluated for each individual ballot drawn.
-because the underlying tied election causes audits to move on to larger rounds, 
-the simulations are computationally expensive. SO \BRAVO is proven to be a Risk-Limiting Audit,
-and we observe in Figure~\ref{fig:so_bravo_risk},
-that the risk of SO \BRAVO is much
-nearer the risk limit than that of EoR \BRAVO, as expected.
-
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/total_risk.png}
-\caption{This plot shows, for each state margin,
-the fraction of SO \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
-\label{fig:so_bravo_risk}
-\end{figure}
-
-%TODO SO and EoR macros
-
-\subsection{\Minerva Simulations}
-The proof that \Minerva is an RLA \cite{usenix_minerva} assumes that the round schedule is pre-determined. For this reason, we have to choose the round sizes of a \Minerva 
-audit {\em a priori}.
-For this paper, we consider two choices of round sizes.
-For both, we estimate and use a first round size 
-for
-a $90\%$ probability of stopping.
-Then, for subsequent rounds, we either (i) 
-draw the same number of ballots in each round or (ii)
-multiply the previous round size by a factor of $1.5$ and 
-sample this many new ballots.
-We consider the case of drawing samples of the same size
-because it may reflect a practical way to continue an
-audit; if election officials have selected some first round size within
-reasonable logistical bounds, drawing the same number of 
-ballots in subsequent rounds may be practical.
-We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this multiple gives a very rough approximation of 
-round sizes with a $90\%$ probability of stopping for \Minerva.
-
-As with the preceding simulations, we ran $10,000=10^4$ trials
-per state for both an underlying election as reported and an underlying tied election.
-
-Figure~\ref{fig:minerva1_sprob} and Figure~\ref{fig:minerva1p5_sprob} show that the first round size estimates were fairly accurate for multipliers of $10.$ and $1.5$ respectively, with first round stopping probabilities being very close to $90\%$.
-For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively did not consistently achieve $90\%$ stopping
-probability, as they were not accurate estimates of $90\%$ stopping probabilities.
+Figure~\ref{fig:minerva1_sprob} and Figure~\ref{fig:minerva1p5_sprob} show the same proportions for \Minerva round multipliers of $1.0$ and $1.5$ respectively.
+We see that the first round size estimates were fairly accurate, with first round stopping probabilities being very close to $90\%$.
+For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively achieved smaller stopping probabilities, as the $1.0$ multiplier was not chosen so as to obtain $\chi_j({\mathcal A}) = 0.9$, and the $1.5$ multiplier is a very rough estimate.
 Note that we chose a simple multiplier for future rounds, but one could make more accurate round size estimates before the audit begins.
@@
 \end{figure}
+
+\subsection{Maximum Risk as a Function of Round and Margin}
+We also study the proportion of audits that stopped when the underlying election was a tie.
+This proportion should approach a value less than the risk limit, $0.1$, as more audits are performed.
+
+%\begin{figure}
+%\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
+%\caption{This plot shows, for each state margin,
+%the fraction of EoR \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
+%\label{fig:eor_bravo_risk}
+%\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{bravo_risks_same_plot.png}
+\caption{This plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:bravo_risk}
+\end{figure}
+
+We observe that the risk of EoR \BRAVO is roughly
+an order of magnitude less than the risk limit. 
+These results are as expected, because EoR \BRAVO is known to be too conservative \cite{usenix_minerva}.
+
+In Figure~\ref{fig:bravo_risk} we show only the results for the $13$
+states for which our simulations with an underlying tied election have completed.
+To estimate the next round size that achieves a desired stopping probability,
+the SO \BRAVO software generates the probability distribution on the number of ballots in the sample ballot by ballot (see \cite{usenix_minerva}) since
+the stopping condition needs to be evaluated for each individual ballot drawn.
+Because the underlying tied election causes audits to move on to larger rounds, 
+the simulations are computationally expensive. SO \BRAVO is proven to be a Risk-Limiting Audit,
+and we observe in Figure~\ref{fig:bravo_risk},
+that the risk of SO \BRAVO is much
+nearer the risk limit than that of EoR \BRAVO, as expected.
+
+%\begin{figure}
+%\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/total_risk.png}
+%\caption{This plot shows, for each state margin,
+%the fraction of SO \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
+%\label{fig:so_bravo_risk}
+%\end{figure}
+
+%TODO SO and EoR macros
 

### 2021-11-02 MultipleRoundSims/draft.tex
Lots of stuff
abstract .9, .25

@@
 In this paper, we present results from simulations of multiple round audits.
-We examine both: first-round stopping probabilities of $90\%$, a common choice among election officials, and $25\%$, which would be more favorable to \BRAVO.
+We examine both: first-round stopping probabilities of $.9$, a common choice among election officials, and $.25$, which would be more favorable to \BRAVO.
 Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).
@@
 
-\section{Simulation Results}
-\label{sec:results}
-\input{sections/05-results}
+\section{Stopping Probability and Risk}
+\label{sec:results1}
+\input{sections/05-results1}
+
+\section{Number of Ballots}
+\label{sec:results2}
+\input{sections/06-results2}
 
@@
 \label{sec:conc}
-\input{sections/06-conc}
+\input{sections/07-conc}
 

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
section 04-expts .9, .25

@@
 
-We ran simulations for: (a) $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $25\%$ probability of stopping in the first round which is more favorable to \BRAVO.
+We ran simulations for: (a) $.9$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $.25$ probability of stopping in the first round which is more favorable to \BRAVO.
 We ran our simulations for up to five rounds.
@@
 We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $90\%$ conditional stopping probability in the second round for a first round stopping probability of $90\%$.
+of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $.9$ conditional stopping probability in the second round for a first round stopping probability of $.9$.

### 2021-11-02 MultipleRoundSims/sections/06-results2.tex
first draft: results 2

In this section we present our data on the expected number of ballots drawn as the number of rounds increases, and on the fraction of audits that stop (an estimate of cumulative stopping probability, $C_j$) for the states of Texas, Missouri and Massachusetts, with margins of $0.057$, $0.157$ and $0.342$ respectively. Interestingly, we observe that the advantage $\Minerva$ has for a first round size with stopping probability $S_1=0.9$ is retained for $S_1=0.25$.

\begin{figure}
\includegraphics[width=0.6\textwidth]{texas2.png}
\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
\label{fig:texas_25}
\end{figure}

We observe that the behavior of both \Minerva audits is similar, and that the plot for SO \BRAVO is to the right (more ballots) and below (lower probability of stopping) those for \Minerva, even for a stopping probability as low as $0.25$. We observe that the plot for EoR \BRAVO shows the worst performance, which is not surprising. We observe similar behavior across margins (see Figures \ref{fig:missouri_25} and \ref{fig:mass_25}), though the improvement due to \Minerva reduces margins get larger. We see also that the behavior is similar to that seen for $S_1=0.9$ (see Figure \ref{fig:texas_90}).

\begin{figure}
\includegraphics[width=0.6\textwidth]{missouri2.png}
\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
\label{fig:missouri_25}
\end{figure}

\begin{figure}
\includegraphics[width=0.6\textwidth]{massachusetts2.png}
\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $S_1=0.25$.}
\label{fig:mass_25}
\end{figure}

\begin{figure}
\includegraphics[width=0.6\textwidth]{texas1.png}
\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.9$.}
\label{fig:texas_90}
\end{figure}

### 2021-11-02 MultipleRoundSims/sections/02-related.tex
section 04-expts .9, .25
replace ; with ,

@@
 Huang {\em et al.} present a framework generalizing a number of ballot polling audits and compare their performance (round sizes and stopping probabilities) using simulations \cite{DBLP:conf/evoteid/HuangRSTV20}.
-This work was prior to the development of \Minerva, and focuses on the comparison between Bayesian audits \cite{bayesian-audits} and \BRAVO; essentially studying the impact of the prior of the Bayesian RLA.
+This work was prior to the development of \Minerva, and focuses on the comparison between Bayesian audits \cite{bayesian-audits} and \BRAVO, essentially studying the impact of the prior of the Bayesian RLA.

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
rephrase

@@
 Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds and the round schedule is pre-determined (before any ballots are drawn).
-They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first rounds when the (true) underlying election is as announced.
+They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
 

### 2021-11-02 MultipleRoundSims/sections/03-software.tex
typo

@@
 of the next round size to achieve a desired stopping probability. 
-For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator.
-given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).
+For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be as announced).
 

### 2021-11-02 MultipleRoundSims/sections/05-results1.tex
.9 and .25 for 05-results1.tex
changed figure size

@@
 \begin{figure}
+\begin{centering}
 \includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
@@
 \label{fig:eor_bravo_sprob}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
+\begin{centering}
 \includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
@@
 \label{fig:so_bravo_sprob}
+\end{centering}
 \end{figure}
@@
 Figure~\ref{fig:minerva1_sprob} and Figure~\ref{fig:minerva1p5_sprob} show the same proportions for \Minerva round multipliers of $1.0$ and $1.5$ respectively.
-We see that the first round size estimates were fairly accurate, with first round stopping probabilities being very close to $90\%$.
+We see that the first round size estimates were fairly accurate, with first round stopping probabilities being very close to $.9$.
 For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively achieved smaller stopping probabilities, as the $1.0$ multiplier was not chosen so as to obtain $\chi_j({\mathcal A}) = 0.9$, and the $1.5$ multiplier is a very rough estimate.
@@
 \begin{figure}
+\begin{centering}
 \includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
@@
 \label{fig:minerva1_sprob}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
+\begin{centering}
 \includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
@@
 \label{fig:minerva1p5_sprob}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{bravo_risks_same_plot.png}
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{bravo_risks_same_plot.png}
 \caption{This plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.}
 \label{fig:bravo_risk}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
 \caption{This plot shows, for each state margin,
@@
 \label{fig:minerva1_risk}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
 \caption{This plot shows, for each state margin,
@@
 \label{fig:minerva1p5_risk}
+\end{centering}
 \end{figure}

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
typo

@@
 That is, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
+Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser).
 

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
added some contributions

@@
 Thus it does not require the tracking of ballots that SO \BRAVO does.
-Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds and the round schedule is pre-determined (before any ballots are drawn).
+Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
 They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
@@
 \subsection{Our Results}
+We show the following: 
+\begin{enumerate}
+\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO is by a fraction of .
+Compare this to:
+\item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three.
+When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively.
+\item Our simulations may be used to study the stopping probabilities and maximum risks of the different audits for different stopping probabilities for the first round, we provide examples. 
+\end{enumerate}
 

### 2021-11-02 MultipleRoundSims/sections/05-results1.tex
added one more graph

@@
 This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
-as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
+as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
 \label{fig:eor_bravo_sprob}
@@
 This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
-as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
+as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
 \label{fig:so_bravo_sprob}
@@
 We see that the first round size estimates were fairly accurate, with first round stopping probabilities being very close to $.9$.
-For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively achieved smaller stopping probabilities, as the $1.0$ multiplier was not chosen so as to obtain $\chi_j({\mathcal A}) = 0.9$, and the $1.5$ multiplier is a very rough estimate.
+For subsequent rounds, the multipliers of $1.0$ achieved smaller stopping probabilities, as it was not chosen so as to obtain $\chi_j({\mathcal A}) = 0.9$.
+The $1.5$ multiplier is a good estimate for $j=2$, but the stopping probability for $j=3$ is slightly smaller than $0.9$.
 Note that we chose a simple multiplier for future rounds, but one could make more accurate round size estimates before the audit begins.
@@
 \caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
-as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.0$.}
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.0$ and $S_1=0.9$.}
 \label{fig:minerva1_sprob}
@@
 \caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
-as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.5$.}
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $S_1=0.9$.}
 \label{fig:minerva1p5_sprob}
@@
 
+Finally, we can perform a similar study for $S_1=0.25$. See \ref{fig:minerva_25} for an example, \Minerva with round mutiplier $1.5$.
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.8\textwidth]{minerva25percthen1p5_sprob.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $S_1 = 0.25$.}
+\label{fig:minerva1p5_sprob}
+\end{centering}
+\end{figure}
 \subsection{Maximum Risk as a Function of Round and Margin}

### 2021-11-02 MultipleRoundSims/draft.tex
shortened abstract

@@
 \begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO and End-of-Round (EoR) \BRAVO.
-We also
-    present details on the R2B2 open source software library and the related simulation software.
+    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot polling risk limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO and End-of-Round (EoR) \BRAVO.
 
-    \BRAVO is the most commonly used ballot-polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
+    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
 In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
@@
 We examine both: first-round stopping probabilities of $.9$, a common choice among election officials, and $.25$, which would be more favorable to \BRAVO.
-Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).
 
 We show that the simulation results are consistent with predictions of the R2B2 open-source library for ballot polling audits.
-We also observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots.
+Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule.
+Most importantly, we observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots, for both first round stopping probabilities, though using \Minerva is not as advantageous for the smaller first round stopping probability.
 
@@
 \end{abstract}
-%
 %

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
ratios in results2

@@
 of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $.9$ conditional stopping probability in the second round for a first round stopping probability of $.9$.
+
+We used the R2B2 library \cite{r2b2_anon}, which provides a framework for the exploration of round-by-round
+and ballot-by-ballot RLAs. It has implementations of several ballot polling risk-limiting audits as well as a simulator, 
+all written in Python. For each of these audits, the software can compute the stopping condition for a given sample and estimates
+of the next round size to achieve a desired stopping probability. 
+For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be as announced).
+
+When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes.
+
+After drawing a simulated sample of ballots, the simulator evaluates the given audit's stopping condition for this sample.
+If the audit stops, the simulation stops, and if the audit continues, the simulation draws another round. 
+The abstract simulator class does not prescribe any one method for choosing round sizes. 
+We implement several classes to support various round size choices: 
+round sizes from an estimate to achieve a desired probability of stopping, 
+predetermined round sizes, and pseudorandomly-generated round sizes.

### 2021-11-02 MultipleRoundSims/sections/06-results2.tex
ratios in results2

@@
 In this section we present our data on the expected number of ballots drawn as the number of rounds increases, and on the fraction of audits that stop (an estimate of cumulative stopping probability, $C_j$) for the states of Texas, Missouri and Massachusetts, with margins of $0.057$, $0.157$ and $0.342$ respectively.
-Interestingly, we observe that the advantage $\Minerva$ has for a first round size with stopping probability $S_1=0.9$ is retained for $S_1=0.25$.
+Interestingly, we observe that $\Minerva$ has an advantage for a first round size with stopping probability $S_1=0.25$, but it is not as large as that for $S_1=0.9$.
 
 \begin{figure}
-\includegraphics[width=0.6\textwidth]{texas2.png}
+\begin{centering}
+\includegraphics[width=0.75\textwidth]{texas2.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
 \label{fig:texas_25}
+\end{centering}
 \end{figure}
@@
 We observe that the plot for EoR \BRAVO shows the worst performance, which is not surprising.
-We observe similar behavior across margins (see Figures \ref{fig:missouri_25} and \ref{fig:mass_25}), though the improvement due to \Minerva reduces margins get larger.
-We see also that the behavior is similar to that seen for $S_1=0.9$ (see Figure \ref{fig:texas_90}).
+We observe similar behavior across margins (see Figures \ref{fig:missouri_25} and \ref{fig:mass_25}), though the improvement due to \Minerva reduces as margins get larger.
+We see also that the improvement due to using \Minerva is not as large as that seen for $S_1=0.9$ (see Figure \ref{fig:texas_90}).
+
+For $S_1=0.25$, the ratio of first round size of EoR \BRAVO to \Minerva is 1.45, 1.37, 1.23 for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+This may be compared to $2.03$, $1.99$ and $1.8$ respectively for $S_1=0.9$.
+
+Similarly, for $S_1=0.25$, the ratio of first round size of SO \BRAVO to \Minerva is 1.13, 1.08, 1.12 for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+This may be compared to $1.38$, $1.38$ and $1.30$ respectively for $S_1=0.9$.
 
 \begin{figure}
-\includegraphics[width=0.6\textwidth]{missouri2.png}
+\begin{centering}
+\includegraphics[width=0.75\textwidth]{missouri2.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
 \label{fig:missouri_25}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.6\textwidth]{massachusetts2.png}
+\begin{centering}
+\includegraphics[width=0.75\textwidth]{massachusetts2.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $S_1=0.25$.}
 \label{fig:mass_25}
+\end{centering}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.6\textwidth]{texas1.png}
+\begin{centering}
+\includegraphics[width=0.75\textwidth]{texas1.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.9$.}
 \label{fig:texas_90}
+\end{centering}
 \end{figure}

### 2021-11-02 MultipleRoundSims/draft.tex
ratios in results2
Update draft.tex

@@
 Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule.
-Most importantly, we observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots, for both first round stopping probabilities, though using \Minerva is not as advantageous for the smaller first round stopping probability.
+Most importantly, we observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots, for both first round stopping probabilities.
+However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.
 
@@
 
-\section{Software}
-\label{sec:software}
-\input{sections/03-software}
+%\section{Software}
+%\label{sec:software}
+%\input{sections/03-software}
 

### 2021-11-02 MultipleRoundSims/sections/06-results2.tex
typo

@@
 
-We observe that the behavior of both \Minerva audits is similar, and that the plot for SO \BRAVO is to the right (more ballots) and below (lower probability of stopping) those for \Minerva, even for a stopping probability as low as $0.25$.
+We observe that the behavior of both \Minerva audits is similar, and that the plot for SO \BRAVO is to the right of (more ballots) and below (lower probability of stopping) those for \Minerva, even for a stopping probability as low as $0.25$.
 We observe that the plot for EoR \BRAVO shows the worst performance, which is not surprising.

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
ratios in results2
intro

@@
 \begin{enumerate}
-\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO is by a fraction of .
-Compare this to:
+\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO.
+However, the improvement is considerably smaller than that when the stopping probability is $0.9$.
+\begin{itemize}
+\item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva.
+On the other hand, for a first round stoping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
+\item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva.
+On the other hand, for a first round stoping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.
+\end{itemize}
 \item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three.
@@
 
-\subsection{Organization} Section \ref{sec:related} describes related work and \ref{sec:software} our open source software.
+\subsection{Organization} Section \ref{sec:related} describes related work.
 The experiments we performed are described in section \ref{sec:expts} and sections \ref{sec:results1} and \ref{sec:results2} present our results.

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
intro

@@
 
-\subsection{Our Results}
+\subsection{Our Contributions}
 We show the following: 
@@
 \item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva.
-On the other hand, for a first round stoping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
+On the other hand, for a first round stopping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.
 \item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva.
-On the other hand, for a first round stoping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.
+On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
 \end{itemize}

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
typo

--- 2021-11-02 MultipleRoundSims/sections/01-intro.tex
@@
 \item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva.
-On the other hand, for a first round stoping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
+On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
 \item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva.

### 2021-11-02 MultipleRoundSims/sections/05-results1.tex
deleted risk graph

@@
 as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $S_1 = 0.25$.}
-\label{fig:minerva1p5_sprob}
+\label{fig:minerva_25}
 \end{centering}
@@
 
-Figures \ref{fig:minerva1_risk} and Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples of $1.0$ and $1.5$ respectively, as would be expected for an RLA with risk limit $0.1$.
+Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples $1.5$, as would be expected for an RLA with risk limit $0.1$.
 Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
-showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.
+showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper. The plot for round multiple $1.0$ is very similar.
 
-\begin{figure}
-\begin{centering}
-\includegraphics[width=0.6\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
-\caption{This plot shows, for each state margin,
-the fraction of \Minerva audits with a round size multiple of $1.0$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
-\label{fig:minerva1_risk}
-\end{centering}
-\end{figure}
+%\begin{figure}
+%\begin{centering}
+%\includegraphics[width=0.6\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
+%\caption{This plot shows, for each state margin,
+%the fraction of \Minerva audits with a round size multiple of $1.0$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
+%\label{fig:minerva1_risk}
+%\end{centering}
+%\end{figure}
 

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
fixed conflicts

--- 2021-11-02 MultipleRoundSims/sections/01-intro.tex
@@
 When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively.
-\item Our simulations may be used to study the stopping probabilities and maximum risks of the different audits for different stopping probabilities for the first round, we provide examples. 
+Both our simulator and the code estimating probabilities and round sizes are flexible enough to enable the study of various predetermined round schedules.
 \end{enumerate}

### 2021-11-02 MultipleRoundSims/sections/07-conc.tex
final

%Conclusions
We describe the use of the R2B2 library and simulator to characterize the maximum risk, stopping probability and average number of ballots required across round schedules which may be specified through conditional stopping probabilities (as with the \BRAVO audits) or pre-determined round sizes (as with \Minerva). We use simulations to study the number of ballots drawn when the first round size is small (stopping probability of 0.25) and when it is large (stopping probability of 0.9). We observe that the advantage of using \Minerva is smaller for the smaller stopping probability of the first round, as would be expected. We also observe that \Minerva does still require fewer ballots all the way through five rounds.

A promising direction for future work would be a more detailed study of the impact of first round stopping probability and different round schedules on overall stopping probability and number of ballots for both \Minerva and \BRAVO.

### 2021-11-02 MultipleRoundSims/draft.tex
Final?

@@
 In this paper, we present results from simulations of multiple round audits.
-We examine both: first-round stopping probabilities of $.9$, a common choice among election officials, and $.25$, which would be more favorable to \BRAVO.
+We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO.
 

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
Final?

@@
 
-We ran simulations for: (a) $.9$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $.25$ probability of stopping in the first round which is more favorable to \BRAVO.
+We ran simulations for: (a) $0.9$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $.25$ probability of stopping in the first round which is more favorable to \BRAVO.
 We ran our simulations for up to five rounds.
@@
 We also consider round sizes with samples increasing by a multiple
-of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $.9$ conditional stopping probability in the second round for a first round stopping probability of $.9$.
+of $1.5$ because this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $0.9$ conditional stopping probability in the second round for a first round stopping probability of $0.9$.
 

### 2021-11-02 MultipleRoundSims/sections/07-conc.tex
Final?

@@
 We describe the use of the R2B2 library and simulator to characterize the maximum risk, stopping probability and average number of ballots required across round schedules which may be specified through conditional stopping probabilities (as with the \BRAVO audits) or pre-determined round sizes (as with \Minerva).
-We use simulations to study the number of ballots drawn when the first round size is small (stopping probability of 0.25) and when it is large (stopping probability of 0.9).
+We use simulations to study the number of ballots drawn when the first round size is small (stopping probability of 0.25) and when it is large (stopping probability of 0.9) for a risk limit of $0.1$.
 We observe that the advantage of using \Minerva is smaller for the smaller stopping probability of the first round, as would be expected.

### 2022-01-07 MultipleRoundSims/sections/01-intro.tex
Final?
typo in intro

@@
 
-Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, available for use in real election audits.
+Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, are available for use in real election audits.
 Both have been used in real election audits.
@@
 \subsection{Our Contributions}
-We show the following: 
+We show the following for a risk limit of $0.1$: 
 \begin{enumerate}

### 2022-01-09 MultipleRoundSims/sections/01-intro.tex
updates according to reviews

@@
 An RLA is an audit which guarantees that the probability of concluding that an election outcome is correct, given that it is not, is below a pre-determined value known as the risk limit of the audit, independent of the true, unknown vote distribution of the underlying election.
-Over a dozen states have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.
+Over a dozen states in the US have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.
 
-This paper provides insight into the main approach to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
+This paper provides insight into the main approaches to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
 While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
@@
 Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, are available for use in real election audits.
-Both have been used in real election audits.
-For this reason, it is very important to understand their properties over multiple rounds.
+Both have been used in real election audits\cite{usenix_minerva, virginia_rla}.
+For this reason, it is important to understand their properties over multiple rounds.
 
 \subsection{Our Contributions}
-We show the following for a risk limit of $0.1$: 
+For a two candidate plurality contest with a risk limit of $10\%$, we show the following
 \begin{enumerate}

### 2022-01-09 MultipleRoundSims/sections/04-expts.tex
updates according to reviews

@@
 %Definitions
-In this section, we motivate and describe the experiments. We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
+In this section, we motivate and describe the experiments.
+We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
+Note that \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise audits between the winners and the losers\cite{RLA, arxiv_athena}.
+Therefore, a two candidate plurality contest can be thought of as the most general case, and our simulations provide insight for multiple-candidate and multiple-winner contests too.
 
@@
 \begin{definition}[Risk]
-The maximum risk $R$ of an audit $\mathcal{A}$ is
+The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
+the true underlying distribution of ballots is
 $$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
@@
 
-We simulated audits for a risk limit of $0.1$ using margins from the 2020 Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
+We simulated audits for a risk limit of $0.1$ using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
 Round sizes increase roughly proportional to the inverse

### 2022-01-09 MultipleRoundSims/sections/05-results1.tex
updates according to reviews

@@
 This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
-as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
+as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
 \label{fig:so_bravo_sprob}
@@
 
-Finally, we can perform a similar study for $S_1=0.25$. See \ref{fig:minerva_25} for an example, \Minerva with round mutiplier $1.5$.
+Finally, we can perform a similar study for $S_1=0.25$. See Figure~\ref{fig:minerva_25} for an example, \Minerva with round mutiplier $1.5$.
 

### 2022-01-09 MultipleRoundSims/sections/02-related.tex
updates according to reviews
review changes

@@
 The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken.
-Real election audits invest in drawing a large number of ballots before making the decision.
-It is possible to apply \BRAVO to the sequence of ballots if the sequential order is retained.
+Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing and searching for specific ballots).
+It is possible to apply \BRAVO to the sequence of ballots in a round if the sequential order is retained.
 This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to ballots that were drawn earlier in the sample.
@@
 The risk limiting property of \BRAVO follows from the similar property of the SPRT.
-Stopping probabilities for \BRAVO may be computed as described by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena}.
-The results of the computations match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
+Stopping probabilities for \BRAVO may be estimated as implemented in \cite{arlo}.
+A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.} uses a similar technique for narrow margins and a separate algorithm for wider margins\cite{usenix_minerva,arxiv_athena}.
+The results of these computations match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
 
@@
 This work was prior to the development of \Minerva, and focuses on the comparison between Bayesian audits \cite{bayesian-audits} and \BRAVO, essentially studying the impact of the prior of the Bayesian RLA.
+Some workload measurements have been made\cite{RI-report}.
+While workload can be estimated with simple metrics like number of ballots examined\cite{bernoulli-ballot-polling}, Bernhard presents a more complex workload estimation model\cite{bernhard-diss}.

### 2022-01-09 MultipleRoundSims/sections/02-related.tex
change risk limit 0.1 to 10%

@@
 Stopping probabilities for \BRAVO may be estimated as implemented in \cite{arlo}.
-A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.} uses a similar technique for narrow margins and a separate algorithm for wider margins\cite{usenix_minerva,arxiv_athena}.
+A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena} uses a similar technique for narrow margins and a separate algorithm for wider margins.
 The results of these computations match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.

### 2022-01-09 MultipleRoundSims/sections/04-expts.tex
change risk limit 0.1 to 10%

@@
 
-We simulated audits for a risk limit of $0.1$ using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
+We simulated audits for a risk limit of $10\%$ using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
 Round sizes increase roughly proportional to the inverse

### 2022-01-09 MultipleRoundSims/sections/05-results1.tex
change risk limit 0.1 to 10%

@@
 We also study the proportion of audits that stopped when the underlying election was a tie.
-This proportion should approach a value less than the risk limit, $0.1$, as more audits are performed.
+This proportion should approach a value less than the risk limit, $10\%$, as more audits are performed.
 
@@
 
-Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples $1.5$, as would be expected for an RLA with risk limit $0.1$.
+Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples $1.5$, as would be expected for an RLA with risk limit $10\%$.
 Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,

### 2022-01-09 MultipleRoundSims/sections/01-intro.tex
polling vs comparison context added

@@
 \subsection{Background}
-Of the multiple types of RLAs, this paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology.
+This paper focuses on ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
+Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Goergia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
 In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.

### 2022-01-09 MultipleRoundSims/sections/04-expts.tex
reviews cont.

@@
 We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
-Note that \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise audits between the winners and the losers\cite{RLA, arxiv_athena}.
+Note that sampling without replacement is more efficient as sampling fractions grow, but \Minerva has not been extended for sampling without replacement.
+Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise audits between the winners and the losers\cite{RLA, arxiv_athena}.
 Therefore, a two candidate plurality contest can be thought of as the most general case, and our simulations provide insight for multiple-candidate and multiple-winner contests too.

### 2022-01-10 MultipleRoundSims/draft.tex
updates according to reviews
abstract shortening

--- 2021-11-02 MultipleRoundSims/draft.tex
@@
 \usepackage{listings,hyperref}
+\usepackage{cite}
 
@@
 \begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot polling risk limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO and End-of-Round (EoR) \BRAVO.
-
+    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
     \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
+    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
+    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in the first round but requires a predetermined round schedule.
+    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
 In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
@@
 It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
-
-    An open question, however, is how \Minerva compares to SO \BRAVO and EoR \BRAVO over multiple rounds, and for lower stopping probabilities.
-In this paper, we present results from simulations of multiple round audits.
-We examine both: first-round stopping probabilities of $.9$, a common choice among election officials, and $.25$, which would be more favorable to \BRAVO.
-
-We show that the simulation results are consistent with predictions of the R2B2 open-source library for ballot polling audits.
-Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule.
-Most importantly, we observe that both \BRAVO audits are more conservative than needed, while \Minerva audits stop with fewer ballots, for both first round stopping probabilities.
+    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
+    Our simulations use stopping probabilities of $0.9$ and $0.25$.
+    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
+    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
+    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
+    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
+    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
     However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.

### 2022-01-10 MultipleRoundSims/sections/06-results2.tex
switch to new plots w asn
added about ASN
abstract shortening

@@
 Interestingly, we observe that $\Minerva$ has an advantage for a first round size with stopping probability $S_1=0.25$, but it is not as large as that for $S_1=0.9$.
+On all our plots we mark ASN, the Average Sample Number for B2 \BRAVO for context.
+Notice that, in all the plots, both instances of \Minerva show a higher probability of completion than does either \BRAVO audit when the average number of ballots drawn is ASN.
 
@@
 \begin{centering}
-\includegraphics[width=0.75\textwidth]{texas2.png}
+\includegraphics[width=0.75\textwidth]{texas25.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
@@
 
-For $S_1=0.25$, the ratio of first round size of EoR \BRAVO to \Minerva is 1.45, 1.37, 1.23 for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+For $S_1=0.25$, the ratio of first round size of EoR \BRAVO to \Minerva is $1.45$, $1.37$, $1.23$ for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
 This may be compared to $2.03$, $1.99$ and $1.8$ respectively for $S_1=0.9$.
-
-Similarly, for $S_1=0.25$, the ratio of first round size of SO \BRAVO to \Minerva is 1.13, 1.08, 1.12 for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+Similarly, for $S_1=0.25$, the ratio of first round size of SO \BRAVO to \Minerva is $1.13$, $1.08$, $1.12$ for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
 This may be compared to $1.38$, $1.38$ and $1.30$ respectively for $S_1=0.9$.
+Note that the effect of such improvements on workload depends greatly on the number of ballots being sampled.
+For example, a $20\%$ reduction in sample size in Massachusetts might save election officials $10$ ballots, whereas the same reduction in Texas could save thousands.
 
@@
 \begin{centering}
-\includegraphics[width=0.75\textwidth]{missouri2.png}
+\includegraphics[width=0.75\textwidth]{missouri25.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
@@
 \begin{centering}
-\includegraphics[width=0.75\textwidth]{massachusetts2.png}
+\includegraphics[width=0.75\textwidth]{massachusetts25.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $S_1=0.25$.}
@@
 \begin{centering}
-\includegraphics[width=0.75\textwidth]{texas1.png}
+\includegraphics[width=0.75\textwidth]{texas90.png}
 \caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.9$.}

### 2022-01-10 MultipleRoundSims/sections/04-expts.tex
abstract shortening
minor edits

--- 2022-01-09 MultipleRoundSims/sections/04-expts.tex
@@
 We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
-Note that \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise audits between the winners and the losers\cite{RLA, arxiv_athena}.
-Therefore, a two candidate plurality contest can be thought of as the most general case, and our simulations provide insight for multiple-candidate and multiple-winner contests too.
-
+Note that sampling without replacement is more efficient for large sampling fractions, but \Minerva has not been extended for sampling without replacement.
 We first present relevant definitions.
@@
 All of the audits discussed in this paper are modeled as binary hypothesis tests.
-Under the alternate hypothesis, $H_a$, the announced outcome is correct.
+Under the alternative hypothesis, $H_a$, the announced outcome is correct.
 That is, the true underlying ballot distribution is given by the announced ballot tallies.
@@
 Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
-
-Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
+Note that $\sum _j S_j(\mathcal{A})=1$.
+We can also consider the cumulative stopping probability:
 \begin{definition}[Cumulative Stopping Probability]
@@
 We simulated audits for a risk limit of $10\%$ using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger. 
+Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}.
+Therefore, the two candidate plurality contest is a general case, and these simulations provide insight for multiple-candidate and multiple-winner contests too.
 Round sizes increase roughly proportional to the inverse
@@
 
-For rounds after the first one, we chose the round schedule as follows.
+For rounds after the first, we chose the round schedule as follows.
 For both versions of \BRAVO, we chose a single round schedule: each round size has the same conditional stopping probability as the first one.
@@
 For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be as announced).
-
-When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes.
+When there are more than two candidates, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes.
 

### 2022-01-11 MultipleRoundSims/draft.tex
minor edits

@@
     In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
-    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
+    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announc/for ed.
     In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
-    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in the first round but requires a predetermined round schedule.
+    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
     %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.

### 2022-01-11 MultipleRoundSims/sections/02-related.tex
minor edits

@@
 The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken.
-Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing and searching for specific ballots).
+Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing and searching for individual ballots).
 It is possible to apply \BRAVO to the sequence of ballots in a round if the sequential order is retained.
@@
 It uses information from the entire sample, and has been proven to be risk limiting when the round schedule for the audit is determined before the audit begins.
-That is, information about the actual ballots drawn in the first round cannot be used to determine future round sizes.
-First-round sizes for a $0.9$ stopping probability when the election is as announced have been computed for a wide range of margins and shown to be smaller than those for both EoR and SO \BRAVO.
+That is, information about the actual ballots drawn in the first round cannot inform future round sizes.
+First-round sizes for a $0.9$ stopping probability when the election is as announced have been computed for a wide range of margins and are smaller than those for EoR and SO \BRAVO.
 First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability $0.9$.
@@
 Some workload measurements have been made\cite{RI-report}.
-While workload can be estimated with simple metrics like number of ballots examined\cite{bernoulli-ballot-polling}, Bernhard presents a more complex workload estimation model\cite{bernhard-diss}.
+While total ballots sampled can give naive workload estimates\cite{bernoulli-ballot-polling}, Bernhard presents a more complex workload estimation model\cite{bernhard-diss}.

### 2022-01-13 MultipleRoundSims/draft.tex
few changes in Intro

--- 2022-01-10 MultipleRoundSims/draft.tex
@@
     In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
-    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in the first round but requires a predetermined round schedule.
+    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
     %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.

### 2022-01-13 MultipleRoundSims/sections/01-intro.tex
few changes in Intro

@@
 While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
-This paper examines the number of ballots drawn over multiple rounds of both audits, the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.
+This paper examines the number of ballots drawn over multiple rounds of both audits, for two chosen probabilities of stopping (one high: 90\%; the other low: 25\%) if the election is as announced.
 
@@
 This paper focuses on ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
-Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Goergia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
+Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
 In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
@@
 
-The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced, and stopping decisions are made after each ballot draw.
+The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced and stopping decisions are made after each ballot draw.
 In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2.
@@
 Thus it does not require the tracking of ballots that SO \BRAVO does.
-Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit\footnote{Their proof assumes that the number of ballots drawn in each round is know beforehand} and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
 They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.

### 2022-01-15 MultipleRoundSims/draft.tex
updates after meeting

@@
 %
-\begin{comment}
 \author{Oliver Broadrick\inst{1} \and
@@
 Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
-Filip Zag{\'o}rski\inst{3} }
+Filip Zag{\'o}rski\inst{3}\inst{4} }
 %
@@
 Department of Computer Science, Stanford University 
-\email{grantmcc@stanford.edu}\\
-\and Wroclaw University of Science and Technology\\
-\email{filip.zagorski@gmail.com}}
+\email{grantmcc@stanford.edu}
+\\
+\and Wroclaw University of Science and Technology and \thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927}
+\and Votifica
+\\
+\email{filip.zagorski@gmail.com}
+}
 %
-\end{comment}
 \maketitle              % typeset the header of the contribution
@@
 
+\section{Acknowledgements}
+\label{sec:ack}
+This research was supported in part by NSF Awards 2015253
+and 1421373.
+
 %

### 2022-01-15 MultipleRoundSims/sections/04-expts.tex
updates after meeting

@@
 \begin{definition}
-An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives one of the 
-following decisions
-\begin{enumerate}
-\item
-$Correct$: the audit is complete
-\item
-$Uncertain$: continue the audit
-\end{enumerate}
+An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
+(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
 \end{definition}
@@
 Under the alternative hypothesis, $H_a$, the announced outcome is correct.
-That is, the true underlying ballot distribution is given by the announced ballot tallies.
+In particular, the true underlying ballot distribution is given by the announced ballot tallies.
 Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser).
-
-The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora show that this is the maximum risk \cite{Bayesian-RLA}.) 
+The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora shows that this is the maximum risk \cite{Bayesian-RLA}.) 
 \begin{definition}[Risk]
@@
 the true underlying distribution of ballots is
-$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
+$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
 \end{definition}
@@
 risk limit $\alpha$ iff 
-$$R(\mathcal{A}) \le \alpha.$$
+$R(\mathcal{A}) \le \alpha.$
 \end{definition}
+Note that an audit $\mathcal{A}$ includes all audit parameters (risk limit, round sizes, etc.).
 
@@
 
-We simulated audits for a risk limit of $10\%$ using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger. 
+We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
 Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}.

### 2022-01-15 MultipleRoundSims/sections/05-results1.tex
updates after meeting

@@
 
+\begin{comment}
+%\begin{minipage}{.5\textwidth}
 \begin{figure}
-\begin{centering}
-\includegraphics[width=0.6\textwidth]{bravo_risks_same_plot.png}
+%\begin{centering}
+\begin{minipage}{.5\textwidth}
+\includegraphics[width=\textwidth]{bravo_risks_same_plot.png}
 \caption{This plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.}
 \label{fig:bravo_risk}
-\end{centering}
+\end{minipage}
+%\end{centering}
+%\end{figure}
+%\end{minipage}
+%\begin{minipage}{.5\textwidth}
+%\begin{figure}
+%\begin{centering}
+\begin{minipage}{.5\textwidth}
+\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
+\caption{This plot shows, for each state margin,
+the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:risk}
+\end{minipage}
+%\end{centering}
 \end{figure}
+%\end{minipage}
+\end{comment}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{both_risk_plots.png}
+\caption{The left hand plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.
+The right hand plot, for each state margin, shows the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:risk}
+%\end{centering}
+\end{figure}
+%
 
@@
 
-In Figure~\ref{fig:bravo_risk} we show only the results for the $13$
+In Figure~\ref{fig:risk} we show only the results for the $13$
 states for which our simulations with an underlying tied election have completed.
@@
 the simulations are computationally expensive. SO \BRAVO is proven to be a Risk-Limiting Audit,
-and we observe in Figure~\ref{fig:bravo_risk},
+and we observe in Figure~\ref{fig:risk},
 that the risk of SO \BRAVO is much
@@
 
-Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples $1.5$, as would be expected for an RLA with risk limit $10\%$.
+Figure~\ref{fig:risk} shows that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples $1.5$, as would be expected for an RLA with risk limit $10\%$. 
 Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
@@
 % we could suggest that predicting accurate multiples is possible (same curve rather than line)
-
-\begin{figure}
-\begin{centering}
-\includegraphics[width=0.6\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
-\caption{This plot shows, for each state margin,
-the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
-\label{fig:minerva1p5_risk}
-\end{centering}
-\end{figure}

### 2022-01-16 MultipleRoundSims/draft.tex
minor changes; size of latter figures

--- 2022-01-13 MultipleRoundSims/draft.tex
@@
 Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
-Filip Zag{\'o}rski\inst{3} }
+Filip Zag{\'o}rski\inst{3}\inst{4}}
 %
@@
 %
-\institute{Department of Computer Science, The George Washington University 
+\institute{Department of Computer Science, The George Washington University\thanks{Supported in part by NSF Award 2015253} 
 \email{obroadrick@gwmail.gwu.edu}\\
@@
 Department of Computer Science, Stanford University 
-\email{grantmcc@stanford.edu}\\
-\and Wroclaw University of Science and Technology\\
-\email{filip.zagorski@gmail.com}}
+\email{grantmcc@stanford.edu}
+\\
+\and Wroclaw University of Science and Technology\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
+\and Votifica
+\\
+\email{filip.zagorski@gmail.com}
+}
 %
-\end{comment}
+
 \maketitle              % typeset the header of the contribution
@@
 
+%\section{Acknowledgements}
+%\label{sec:ack}
+%This research was supported in part by NSF Awards 2015253
+%and 1421373.
+
 %

### 2022-01-16 MultipleRoundSims/sections/01-intro.tex
minor changes; size of latter figures

@@
 Thus it does not require the tracking of ballots that SO \BRAVO does.
-Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit\footnote{Their proof assumes that the number of ballots drawn in each round is know beforehand} and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand.
+In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn.
+Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general.
+(We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.} .
 They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
@@
 \subsection{Our Contributions}
-For a two candidate plurality contest with a risk limit of $10\%$, we show the following
+For a two candidate plurality contest with a risk limit of $10\%$, we observe the following about the total number of ballots drawn over five rounds:
 \begin{enumerate}

### 2022-01-16 MultipleRoundSims/sections/04-expts.tex
minor changes; size of latter figures

@@
 In particular, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser).
-The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora shows that this is the maximum risk \cite{Bayesian-RLA}.) 
+Under the null hypothesis, $H_0$, a tie is the correct outcome \footnote{or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser}.
+The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}.
+Note that an audit $\mathcal{A}$ includes all audit parameters (maximum risk, round sizes, etc.).
+
 \begin{definition}[Risk]
@@
 \end{definition}
-Note that an audit $\mathcal{A}$ includes all audit parameters (risk limit, round sizes, etc.).
 
@@
 \begin{definition}[Cumulative Stopping Probability]
-The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is 
-$$C_j(\mathcal{A})= \sum_{i=1}^j S_j$$
+The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
 \end{definition}

### 2022-04-10 MultipleRoundSims/poster.tex
poster

% Gemini theme
% See: https://rev.cs.uchicago.edu/k4rtik/gemini-uccs
% A fork of https://github.com/anishathalye/gemini

\documentclass[final]{beamer}

% ====================
% Packages
% ====================
\usepackage{amsfonts}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[size=custom,width=120,height=72,scale=1.0]{beamerposter}
\geometry{paperwidth=42in,paperheight=32.5in}
\usetheme{gemini}
\usecolortheme{ucf}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}

% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.025\paperwidth}
\setlength{\colwidth}{0.3\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

% ====================
% Title
% ====================

\title{Title}

%\author{Nitish A. Gupta \inst{1} \and Nitish A. Gupta \inst{2}}
\author{Oliver Broadrick\inst{1} \and
Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{2}\inst{3}}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute[shortinst]{\inst{1} Department of Computer Science, The George Washington University[odbroadrick@gmail.com]%\thanks{Supported in part by NSF Award 2015253} 
\samelineand \inst{2} Wroclaw University of Science and Technology%\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
\samelineand \inst{3} Votifica
}
%

%\institute[shortinst]{\inst{1} \textit{University of Central Florida, Orlando, FL} \samelineand \inst{2} Another Institute}

% ====================
% Footer (optional)
% ====================

\footercontent{
  \href{https://www.example.com}{https://www.example.com} \hfill
  ABC Conference 2025, New York --- XYZ-1234 \hfill
  \href{mailto:alyssa.p.hacker@example.com}{nitish.gupta@example.com}}
% (can be left out to remove footer)

% ====================
% Logo (optional)
% ====================

% use this to include logos on the left and/or right side of the header:
% \logoright{\includegraphics[height=7cm]{logo1.pdf}}
% \logoleft{\includegraphics[height=7cm]{logo2.pdf}}

% ====================
% Body
% ====================

\begin{document}
\addtobeamertemplate{headline}{}
{
    \begin{tikzpicture}[remember picture,overlay]
      \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
      {\includegraphics[height=8.5cm]{logos/ucf_logo2.png}}; % also try shield-white.eps
      \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
      {\includegraphics[height=8.5cm]{logos/lab_logo.jpg}};
    \end{tikzpicture}
}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{A block title}

    Some block contents, followed by a diagram, followed by a dummy paragraph.

    \begin{figure}
      \centering
      \begin{tikzpicture}[scale=6]
        \draw[step=0.25cm,color=gray] (-1,-1) grid (1,1);
        \draw (1,0) -- (0.2,0.2) -- (0,1) -- (-0.2,0.2) -- (-1,0)
          -- (-0.2,-0.2) -- (0,-1) -- (0.2,-0.2) -- cycle;
      \end{tikzpicture}
      \caption{A figure caption.}
    \end{figure}

    Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi ultricies
    eget libero ac ullamcorper. Integer et euismod ante. Aenean vestibulum
    lobortis augue, ut lobortis turpis rhoncus sed. Proin feugiat nibh a
    lacinia dignissim. Proin scelerisque, risus eget tempor fermentum, ex
    turpis condimentum urna, quis malesuada sapien arcu eu purus.

  \end{block}

% here
definitions.

\begin{definition}
An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
\end{definition}

All of the audits discussed in this paper are modeled as binary hypothesis tests. Under the alternative hypothesis, $H_a$, the announced outcome is correct. In particular, the true underlying ballot distribution is given by the announced ballot tallies. Under the null hypothesis, $H_0$, a tie is the correct outcome \footnote{or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser}.
The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}. Note that an audit $\mathcal{A}$ includes all audit parameters (maximum risk, round sizes, etc.).

\begin{definition}[Risk]
The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
the true underlying distribution of ballots is
$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
\end{definition}

This leads us to the following definition of an $\alpha$-RLA.
\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$R(\mathcal{A}) \le \alpha.$
\end{definition}

We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
\begin{definition}[Stopping Probability]
The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}
Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$. Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
\begin{definition}[Cumulative Stopping Probability]
The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
\end{definition}
Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$.

Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
\begin{definition}[Conditional Stopping Probability]
The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
\end{definition}
Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e. those that did not stop before round $j$.

We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger. 
Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}. Therefore, the two candidate plurality contest is a general case, and these simulations provide insight for multiple-candidate and multiple-winner contests too.
Round sizes increase roughly proportional to the inverse
square of the margin, so 
smaller margins are computationally much more expensive to simulate.
For each of these states, we simulated 
$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$). 
  \end{block}
  \begin{block}{A block containing a list}

    Nam vulputate nunc felis, non condimentum lacus porta ultrices. Nullam sed
    sagittis metus. Etiam consectetur gravida urna quis suscipit.

    \begin{itemize}
      \item \textbf{Mauris tempor} risus nulla, sed ornare
      \item \textbf{Libero tincidunt} a duis congue vitae
      \item \textbf{Dui ac pretium} morbi justo neque, ullamcorper
    \end{itemize}

    Eget augue porta, bibendum venenatis tortor.

  \end{block}

  \begin{alertblock}{A highlighted block}

    This block catches your eye, so \textbf{important stuff} should probably go
    here.

    Curabitur eu libero vehicula, cursus est fringilla, luctus est. Morbi
    consectetur mauris quam, at finibus elit auctor ac. Aliquam erat volutpat.
    Aenean at nisl ut ex ullamcorper eleifend et eu augue. Aenean quis velit
    tristique odio convallis ultrices a ac odio.

    \begin{itemize}
      \item \textbf{Fusce dapibus tellus} vel tellus semper finibus. In
        consequat, nibh sed mattis luctus, augue diam fermentum lectus.
      \item \textbf{In euismod erat metus} non ex. Vestibulum luctus augue in
        mi condimentum, at sollicitudin lorem viverra.
      \item \textbf{Suspendisse vulputate} mauris vel placerat consectetur.
        Mauris semper, purus ac hendrerit molestie, elit mi dignissim odio, in
        suscipit felis sapien vel ex.
    \end{itemize}

    Aenean tincidunt risus eros, at gravida lorem sagittis vel. Vestibulum ante
    ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae.

  \end{alertblock}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{A block containing an enumerated list}

    Vivamus congue volutpat elit non semper. Praesent molestie nec erat ac
    interdum. In quis suscipit erat. \textbf{Phasellus mauris felis, molestie
    ac pharetra quis}, tempus nec ante. Donec finibus ante vel purus mollis
    fermentum. Sed felis mi, pharetra eget nibh a, feugiat eleifend dolor. Nam
    mollis condimentum purus quis sodales. Nullam eu felis eu nulla eleifend
    bibendum nec eu lorem. Vivamus felis velit, volutpat ut facilisis ac,
    commodo in metus.

    \begin{enumerate}
      \item \textbf{Morbi mauris purus}, egestas at vehicula et, convallis
        accumsan orci. Orci varius natoque penatibus et magnis dis parturient
        montes, nascetur ridiculus mus.
      \item \textbf{Cras vehicula blandit urna ut maximus}. Aliquam blandit nec
        massa ac sollicitudin. Curabitur cursus, metus nec imperdiet bibendum,
        velit lectus faucibus dolor, quis gravida metus mauris gravida turpis.
      \item \textbf{Vestibulum et massa diam}. Phasellus fermentum augue non
        nulla accumsan, non rhoncus lectus condimentum.
    \end{enumerate}

  \end{block}

  \begin{block}{Fusce aliquam magna velit}

    Et rutrum ex euismod vel. Pellentesque ultricies, velit in fermentum
    vestibulum, lectus nisi pretium nibh, sit amet aliquam lectus augue vel
    velit. Suspendisse rhoncus massa porttitor augue feugiat molestie. Sed
    molestie ut orci nec malesuada. Sed ultricies feugiat est fringilla
    posuere.

    \begin{figure}
      \centering
      \begin{tikzpicture}
        \begin{axis}[
            scale only axis,
            no markers,
            domain=0:2*pi,
            samples=100,
            axis lines=center,
            axis line style={-},
            ticks=none]
          \addplot[red] {sin(deg(x))};
          \addplot[blue] {cos(deg(x))};
        \end{axis}
      \end{tikzpicture}
      \caption{Another figure caption.}
    \end{figure}

  \end{block}

  \begin{block}{Nam cursus consequat egestas}

    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
    sit amet sollicitudin risus tincidunt non.

    Etiam sit amet tempus lorem, aliquet condimentum velit. Donec et nibh
    consequat, sagittis ex eget, dictum orci. Etiam quis semper ante. Ut eu
    mauris purus. Proin nec consectetur ligula. Mauris pretium molestie
    ullamcorper. Integer nisi neque, aliquet et odio non, sagittis porta justo.

    \begin{itemize}
      \item \textbf{Sed consequat} id ante vel efficitur. Praesent congue massa
        sed est scelerisque, elementum mollis augue iaculis.
        \begin{itemize}
          \item In sed est finibus, vulputate
            nunc gravida, pulvinar lorem. In maximus nunc dolor, sed auctor eros
            porttitor quis.
          \item Fusce ornare dignissim nisi. Nam sit amet risus vel lacus
            tempor tincidunt eu a arcu.
          \item Donec rhoncus vestibulum erat, quis aliquam leo
            gravida egestas.
        \end{itemize}
      \item \textbf{Sed luctus, elit sit amet} dictum maximus, diam dolor
        faucibus purus, sed lobortis justo erat id turpis.
      \item \textbf{Pellentesque facilisis dolor in leo} bibendum congue.
        Maecenas congue finibus justo, vitae eleifend urna facilisis at.
    \end{itemize}

    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
    sit amet sollicitudin risus tincidunt non.
  \end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{A block containing some math}

    Nullam non est elit. In eu ornare justo. Maecenas porttitor sodales lacus,
    ut cursus augue sodales ac.

    $$
    \int_{-\infty}^{\infty} e^{-x^2}\,dx = \sqrt{\pi}
    $$

    Interdum et malesuada fames $\{1, 4, 9, \ldots\}$ ac ante ipsum primis in
    faucibus. Cras eleifend dolor eu nulla suscipit suscipit. Sed lobortis non
    felis id vulputate.

    \heading{A heading inside a block}

    Praesent consectetur mi $x^2 + y^2$ metus, nec vestibulum justo viverra
    nec. Proin eget nulla pretium, egestas magna aliquam, mollis neque. Vivamus
    dictum $\mathbf{u}^\intercal\mathbf{v}$ sagittis odio, vel porta erat
    congue sed. Maecenas ut dolor quis arcu auctor porttitor.

    \heading{Another heading inside a block}

    Sed augue erat, scelerisque a purus ultricies, placerat porttitor neque.
    Donec $P(y \mid x)$ fermentum consectetur $\nabla_x P(y \mid x)$ sapien
    sagittis egestas. Duis eget leo euismod nunc viverra imperdiet nec id
    justo.

  \end{block}

  \begin{block}{Nullam vel erat at velit convallis laoreet}

    Class aptent taciti sociosqu ad litora torquent per conubia nostra, per
    inceptos himenaeos. Phasellus libero enim, gravida sed erat sit amet,
    scelerisque congue diam. Fusce dapibus dui ut augue pulvinar iaculis.

    \begin{table}
      \centering
      \begin{tabular}{l r r c}
        \toprule
        \textbf{First column} & \textbf{Second column} & \textbf{Third column} & \textbf{Fourth} \\
        \midrule
        Foo & 13.37 & 384,394 & $\alpha$ \\
        Bar & 2.17 & 1,392 & $\beta$ \\
        Baz & 3.14 & 83,742 & $\delta$ \\
        Qux & 7.59 & 974 & $\gamma$ \\
        \bottomrule
      \end{tabular}
      \caption{A table caption.}
    \end{table}

    Donec quis posuere ligula. Nunc feugiat elit a mi malesuada consequat. Sed
    imperdiet augue ac nibh aliquet tristique. Aenean eu tortor vulputate,
    eleifend lorem in, dictum urna. Proin auctor ante in augue tincidunt
    tempor. Proin pellentesque vulputate odio, ac gravida nulla posuere
    efficitur. Aenean at velit vel dolor blandit molestie. Mauris laoreet
    commodo quam, non luctus nibh ullamcorper in. Class aptent taciti sociosqu
    ad litora torquent per conubia nostra, per inceptos himenaeos.

    Nulla varius finibus volutpat. Mauris molestie lorem tincidunt, iaculis
    libero at, gravida ante. Phasellus at felis eu neque suscipit suscipit.
    Integer ullamcorper, dui nec pretium ornare, urna dolor consequat libero,
    in feugiat elit lorem euismod lacus. Pellentesque sit amet dolor mollis,
    auctor urna non, tempus sem.

  \end{block}

  \begin{block}{References}

    \nocite{*}
    \footnotesize{\bibliographystyle{plain}\bibliography{poster}}

  \end{block}

\end{column}

\separatorcolumn
\end{columns}
\end{frame}

\end{document}

### 2022-04-10 MultipleRoundSims/poster.tex
a

@@
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.17}
+\pgfplotsset{compat=1.14}
+
+\usepackage{float}
+\usepackage{xspace}
+\usepackage{graphicx}
+\graphicspath{{./imgs/}}
+\usepackage{comment}
+\usepackage{listings,hyperref}
+\usepackage{cite}
+
+% nice looking audit titles
+\newcommand{\Minerva}{\textsc{Minerva}\xspace}
+\newcommand{\B}{{{B2}}\xspace}
+\newcommand{\R}{{{R2}}\xspace}
+\newcommand{\BRAVO}{\textsc{Bravo}\xspace}
 
@@
       \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
-      {\includegraphics[height=8.5cm]{logos/ucf_logo2.png}}; % also try shield-white.eps
+      {\includegraphics[height=8.5cm]{logos/gw.png}}; % also try shield-white.eps
       \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
-      {\includegraphics[height=8.5cm]{logos/lab_logo.jpg}};
+      {\includegraphics[height=8.5cm]{logos/gw.png}};
     \end{tikzpicture}
@@
   \begin{block}{A block title}
+    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
+    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
+    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
+    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
+    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
+In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
+\Minerva was recently proposed for use when the first-round stopping probability is large.
+It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
+    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
+    Our simulations use stopping probabilities of $0.9$ and $0.25$.
+    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
+    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
+    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
+    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
+    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
+    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.
 
-    Some block contents, followed by a diagram, followed by a dummy paragraph.
-
-    \begin{figure}
-      \centering
-      \begin{tikzpicture}[scale=6]
-        \draw[step=0.25cm,color=gray] (-1,-1) grid (1,1);
-        \draw (1,0) -- (0.2,0.2) -- (0,1) -- (-0.2,0.2) -- (-1,0)
-          -- (-0.2,-0.2) -- (0,-1) -- (0.2,-0.2) -- cycle;
-      \end{tikzpicture}
-      \caption{A figure caption.}
-    \end{figure}
-
-    Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi ultricies
-    eget libero ac ullamcorper. Integer et euismod ante. Aenean vestibulum
-    lobortis augue, ut lobortis turpis rhoncus sed. Proin feugiat nibh a
-    lacinia dignissim. Proin scelerisque, risus eget tempor fermentum, ex
-    turpis condimentum urna, quis malesuada sapien arcu eu purus.
+%
+%    Some block contents, followed by a diagram, followed by a dummy paragraph.
+%
+%    \begin{figure}
+%      \centering
+%      \begin{tikzpicture}[scale=6]
+%%        \draw[step=0.25cm,color=gray] (-1,-1) grid (1,1);
+%        \draw (1,0) -- (0.2,0.2) -- (0,1) -- (-0.2,0.2) -- (-1,0)
+%          -- (-0.2,-0.2) -- (0,-1) -- (0.2,-0.2) -- cycle;
+%      \end{tikzpicture}
+%      \caption{A figure caption.}
+%%%    \end{figure}
 
@@
 
-% here
-definitions.
+\begin{block}{Introduction}
+%Definitions
+In this section, we motivate and describe the experiments.
+We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
+Note that sampling without replacement is more efficient for large sampling fractions, but \Minerva has not been extended for sampling without replacement.
+We first present relevant definitions.
 

### 2022-04-11 MultipleRoundSims/poster.tex
4 cols

@@
 \newlength{\colwidth}
-\setlength{\sepwidth}{0.025\paperwidth}
-\setlength{\colwidth}{0.3\paperwidth}
+\setlength{\sepwidth}{0.01\paperwidth}
+\setlength{\colwidth}{0.22\paperwidth}
 
@@
 
-\title{Title}
+\title{Election Security with Risk-Limiting Audits}
 
@@
 %
-\institute[shortinst]{\inst{1} Department of Computer Science, The George Washington University[odbroadrick@gmail.com]%\thanks{Supported in part by NSF Award 2015253} 
+\institute[shortinst]{\inst{1} Department of Computer Science, The George Washington University [odbroadrick@gmail.com]%\thanks{Supported in part by NSF Award 2015253} 
 \samelineand \inst{2} Wroclaw University of Science and Technology%\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
@@
       \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
-      {\includegraphics[height=8.5cm]{logos/gw.png}}; % also try shield-white.eps
+      {\includegraphics[height=8cm]{logos/gw.png}}; % also try shield-white.eps
       \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
-      {\includegraphics[height=8.5cm]{logos/gw.png}};
+      {\includegraphics[height=8cm]{logos/gw.png}};
     \end{tikzpicture}
@@
 
-  \begin{block}{A block title}
+  \begin{block}{Abstract}
     In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
@@
     \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
-    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
-In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
-\Minerva was recently proposed for use when the first-round stopping probability is large.
-It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
     It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
     Our simulations use stopping probabilities of $0.9$ and $0.25$.
-    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
-    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
     The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
-    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
     We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
@@
 %Definitions
-In this section, we motivate and describe the experiments.
+%In this section, we motivate and describe the experiments.
 We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
@@
 
-\begin{definition}
+\heading{Definition: Audit}
+%\begin{definition}
 An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
 (1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
-\end{definition}
+%\end{definition}
 
-All of the audits discussed in this paper are modeled as binary hypothesis tests.
+%All of the audits discussed in this paper are modeled as binary hypothesis tests.
 Under the alternative hypothesis, $H_a$, the announced outcome is correct.
@@
 Under the null hypothesis, $H_0$, a tie is the correct outcome \footnote{or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser}.
-The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}.
+%The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}.
 Note that an audit $\mathcal{A}$ includes all audit parameters (maximum risk, round sizes, etc.).
 
-\begin{definition}[Risk]
+%\begin{definition}[Risk]
+\heading{Definition: Risk}
 The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
@@
 $R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
-\end{definition}
+%\end{definition}
 
-This leads us to the following definition of an $\alpha$-RLA.
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+%This leads us to the following definition of an $\alpha$-RLA.
+\heading{Definition: Risk-Limiting Audit ($\alpha$-RLA)}
+%\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk-Limiting Audit with 
 risk limit $\alpha$ iff 
 $R(\mathcal{A}) \le \alpha.$
-\end{definition}
+%\end{definition}
 
 We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
-\begin{definition}[Stopping Probability]
+%\begin{definition}[Stopping Probability]
+\heading{Definition: Stopping Probability}
 The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
 $$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
-\end{definition}
+%\end{definition}
 Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
@@
 We can also consider the cumulative stopping probability:
-\begin{definition}[Cumulative Stopping Probability]
+%\begin{definition}[Cumulative Stopping Probability]
 The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
-\end{definition}
+%\end{definition}
 Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$. 
-
+%
 Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
-\begin{definition}[Conditional Stopping Probability]
+%\begin{definition}[Conditional Stopping Probability]
 The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
 $$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
-\end{definition}
-Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e. those that did not stop before round $j$.
-
+%\end{definition}
+Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
+those that did not stop before round $j$.
+%
 We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
@@
   \end{block}
-  \begin{block}{A block containing a list}
 
-    Nam vulputate nunc felis, non condimentum lacus porta ultrices. Nullam sed
-    sagittis metus. Etiam consectetur gravida urna quis suscipit.
-
-    \begin{itemize}
-      \item \textbf{Mauris tempor} risus nulla, sed ornare
-      \item \textbf{Libero tincidunt} a duis congue vitae
-      \item \textbf{Dui ac pretium} morbi justo neque, ullamcorper
-    \end{itemize}
-
-    Eget augue porta, bibendum venenatis tortor.
-
-  \end{block}
-
-  \begin{alertblock}{A highlighted block}
-
-    This block catches your eye, so \textbf{important stuff} should probably go
-    here.
-
-    Curabitur eu libero vehicula, cursus est fringilla, luctus est. Morbi
-    consectetur mauris quam, at finibus elit auctor ac. Aliquam erat volutpat.
-    Aenean at nisl ut ex ullamcorper eleifend et eu augue. Aenean quis velit
-    tristique odio convallis ultrices a ac odio.
-
-    \begin{itemize}
-      \item \textbf{Fusce dapibus tellus} vel tellus semper finibus. In
-        consequat, nibh sed mattis luctus, augue diam fermentum lectus.
-      \item \textbf{In euismod erat metus} non ex. Vestibulum luctus augue in
-        mi condimentum, at sollicitudin lorem viverra.
-      \item \textbf{Suspendisse vulputate} mauris vel placerat consectetur.
-        Mauris semper, purus ac hendrerit molestie, elit mi dignissim odio, in
-        suscipit felis sapien vel ex.
-    \end{itemize}
-
-    Aenean tincidunt risus eros, at gravida lorem sagittis vel. Vestibulum ante
-    ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae.
-
-  \end{alertblock}
+%  \begin{block}{A block containing a list}
+%
+%    Nam vulputate nunc felis, non condimentum lacus porta ultrices. Nullam sed
+%    sagittis metus. Etiam consectetur gravida urna quis suscipit.
+%
+%    \begin{itemize}
+%      \item \textbf{Mauris tempor} risus nulla, sed ornare
+%      \item \textbf{Libero tincidunt} a duis congue vitae
+%      \item \textbf{Dui ac pretium} morbi justo neque, ullamcorper
+%    \end{itemize}
+%
+%    Eget augue porta, bibendum venenatis tortor.
+%
+%  \end{block}
+%
+%  \begin{alertblock}{A highlighted block}
+%
+%    This block catches your eye, so \textbf{important stuff} should probably go
+%%    here.
+%%
+%%%    Curabitur eu libero vehicula, cursus est fringilla, luctus est. Morbi
+%%    consectetur mauris quam, at finibus elit auctor ac. Aliquam erat volutpat.
+%%%    Aenean at nisl ut ex ullamcorper eleifend et eu augue. Aenean quis velit
+%    tristique odio convallis ultrices a ac odio.
+%
+%    \begin{itemize}
+%      \item \textbf{Fusce dapibus tellus} vel tellus semper finibus. In
+%        consequat, nibh sed mattis luctus, augue diam fermentum lectus.
+%      \item \textbf{In euismod erat metus} non ex. Vestibulum luctus augue in
+%        mi condimentum, at sollicitudin lorem viverra.
+%      \item \textbf{Suspendisse vulputate} mauris vel placerat consectetur.
+%        Mauris semper, purus ac hendrerit molestie, elit mi dignissim odio, in
+%        suscipit felis sapien vel ex.
+%    \end{itemize}
+%
+%    Aenean tincidunt risus eros, at gravida lorem sagittis vel. Vestibulum ante
+%    ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae.
+%
+%  \end{alertblock}
 
@@
 
-  \begin{block}{A block containing an enumerated list}
+  \begin{block}{Stopping Probability}
 
-    Vivamus congue volutpat elit non semper. Praesent molestie nec erat ac
-    interdum. In quis suscipit erat. \textbf{Phasellus mauris felis, molestie
-    ac pharetra quis}, tempus nec ante. Donec finibus ante vel purus mollis
-    fermentum. Sed felis mi, pharetra eget nibh a, feugiat eleifend dolor. Nam
-    mollis condimentum purus quis sodales. Nullam eu felis eu nulla eleifend
-    bibendum nec eu lorem. Vivamus felis velit, volutpat ut facilisis ac,
-    commodo in metus.
-
-    \begin{enumerate}
-      \item \textbf{Morbi mauris purus}, egestas at vehicula et, convallis
-        accumsan orci. Orci varius natoque penatibus et magnis dis parturient
-        montes, nascetur ridiculus mus.
-      \item \textbf{Cras vehicula blandit urna ut maximus}. Aliquam blandit nec
-        massa ac sollicitudin. Curabitur cursus, metus nec imperdiet bibendum,
-        velit lectus faucibus dolor, quis gravida metus mauris gravida turpis.
-      \item \textbf{Vestibulum et massa diam}. Phasellus fermentum augue non
-        nulla accumsan, non rhoncus lectus condimentum.
-    \end{enumerate}
+Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
+%\begin{definition}[Conditional Stopping Probability]
+The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
+$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
+%\end{definition}
+Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
+those that did not stop before round $j$.
+%
 
@@
 
+\end{column}
+
+\separatorcolumn
+
+\begin{column}{\colwidth}
+
+\begin{block}{Providence}
+
+Providence is great! Efficient like Minerva but flexible like BRAVO.
+
+\end{block}
+
+\begin{block}{Simulations}
+
+Look, Providence does pretty ok!
+
+\end{block}
+
   \begin{block}{References}
@@
 \separatorcolumn
+
 \end{columns}
+
 \end{frame}

### 2022-04-11 MultipleRoundSims/poster.tex
filling in slowly

@@
 \usepackage[size=custom,width=120,height=72,scale=1.0]{beamerposter}
-\geometry{paperwidth=42in,paperheight=32.5in}
+\geometry{paperwidth=48in,paperheight=36in}
 \usetheme{gemini}
@@
 \newcommand{\R}{{{R2}}\xspace}
-\newcommand{\BRAVO}{\textsc{Bravo}\xspace}
+\newcommand{\BRAVO}{\textsc{BRAVO}\xspace}
 
@@
 \newlength{\colwidth}
-\setlength{\sepwidth}{0.01\paperwidth}
-\setlength{\colwidth}{0.22\paperwidth}
+\setlength{\sepwidth}{0.001\paperwidth}
+\setlength{\colwidth}{0.23\paperwidth}
 
@@
 %
-\institute[shortinst]{\inst{1} Department of Computer Science, The George Washington University [odbroadrick@gmail.com]%\thanks{Supported in part by NSF Award 2015253} 
+\institute[shortinst]{\inst{1} Department of Computer Science, The George Washington University%\thanks{Supported in part by NSF Award 2015253} 
 \samelineand \inst{2} Wroclaw University of Science and Technology%\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
@@
 \footercontent{
-  \href{https://www.example.com}{https://www.example.com} \hfill
-  ABC Conference 2025, New York --- XYZ-1234 \hfill
-  \href{mailto:alyssa.p.hacker@example.com}{nitish.gupta@example.com}}
+  \href{odbroadrick@gmail.com}{odbroadrick@gmail.com}
+}
 % (can be left out to remove footer)
@@
 
-  \begin{block}{Abstract}
-    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
-    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
-    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
-    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
-    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
-    Our simulations use stopping probabilities of $0.9$ and $0.25$.
-    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
-    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
-    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.
+\begin{block}{Risk-Limiting Audits}
+\begin{itemize}
+\item
+In election security, a risk-limiting audit (RLA) draws trusted paper ballots in rounds, stopping if a rigorous statistical criterion is satisfied, or proceeding to a full hand count.
+If the announced outcome of the election is erroneous, an RLA will detect the error with high, predetermined minimum probability.
 
+\item
+An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
+(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
+
+%\item
 %
-%    Some block contents, followed by a diagram, followed by a dummy paragraph.
-%
-%    \begin{figure}
-%      \centering
-%      \begin{tikzpicture}[scale=6]
-%%        \draw[step=0.25cm,color=gray] (-1,-1) grid (1,1);
-%        \draw (1,0) -- (0.2,0.2) -- (0,1) -- (-0.2,0.2) -- (-1,0)
-%          -- (-0.2,-0.2) -- (0,-1) -- (0.2,-0.2) -- cycle;
-%      \end{tikzpicture}
-%      \caption{A figure caption.}
-%%%    \end{figure}
+%In this work, we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of RLAs Minerva, Selection-Ordered (SO) Bravo, and End-of-Round (EoR) Bravo.
+We show that Minerva requires fewer ballots on average than either version of Bravo for round sizes that are chosen to give probabilities of stopping 0.9 and 0.25.
+In particular, the advantage of Minerva is less for the smaller stopping probabil- ity.
+We also present Providence, a novel ballot polling RLA.
+Based on initial simulation results, Providence has efficiency similar to Minerva but is also resistant to an adversary who can choose future round sizes with knowledge of previous samples.
+Thus, unlike Minerva, Providence allows election officials to choose round sizes during the audit to optimize for workload and safety.
+%\end{block}
 
-  \end{block}
-
-\begin{block}{Introduction}
+%\begin{block}{Risk}
 %Definitions
@@
 
-\heading{Definition: Audit}
 %\begin{definition}
-An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
-(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
+%An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
+%(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
 %\end{definition}
@@
 %\begin{definition}[Risk]
-\heading{Definition: Risk}
+%\heading{Definition: Risk}
+\item
 The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
 the true underlying distribution of ballots is
-$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
+$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0],$$ where $H_0$ is the null
+hypothesis: the true underlying election is a tie.
 %\end{definition}
@@
 %This leads us to the following definition of an $\alpha$-RLA.
-\heading{Definition: Risk-Limiting Audit ($\alpha$-RLA)}
+%\heading{Definition: Risk-Limiting Audit ($\alpha$-RLA)}
 %\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+\item
 An audit $\mathcal{A}$ is a Risk-Limiting Audit with 
 risk limit $\alpha$ iff 
-$R(\mathcal{A}) \le \alpha.$
+$$R(\mathcal{A}) \le \alpha.$$
 %\end{definition}
+\end{itemize}
+\end{block}
+\begin{block}{\BRAVO and \Minerva}
+\begin{itemize}
+\item Existing RLAs include \BRAVO (which can be implemented as End-of-Round (EoR) \BRAVO or Selection-Ordered (SO) \BRAVO)
+and \Minerva. In a first round with a large sample size (probabilty of stopping at least 90\%), \Minerva is known to require roughly
+half as many ballots as EoR \BRAVO. 
+\item We provide the first simulations that compare these audits beyond a first round and for various
+round schedules.
+\end{itemize}
+\end{block}
 
-We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
-%\begin{definition}[Stopping Probability]
-\heading{Definition: Stopping Probability}
-The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
-$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
-%\end{definition}
-Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
+\begin{block}{Simulations}
+\begin{itemize}
+%\item
+%Simulations can be used to provide evidence for theoretical claims about risk-limiting audits.
+\item
+We ran simulations to gain further insight into audit behavior 
+and provide additional evidence for theoretical claims.
+\item 
+We simulated audits for a risk limit of 10\% using margins
+from the 2020 US Presidential election, limiting ourselves to pairwise margins
+for the two main candidates of $0.05$ or larger.
+\item 
+We used the R2B2 library, which provides a framework for the exploration
+of round-by-round and ballot-by-ballot RLAs. It has implementations of several
+ballot polling risk-limiting audits as well as a simulator, all written in Python.
+\item
+For each of these states, we simulated 
+$10,000=10^4$ audits assuming the underlying election was as announced,  
+and an additional $10,000=10^4$ audits assuming the underlying election was a tie. 
+\end{itemize}
+\end{block}
+
+\heading{Round Schedules}
+\begin{itemize}
+\item
+A round schedule is a list of positive integers corresponding to the number of ballots 
+sampled in each round.
+To begin, first round sizes are selected (by search) to achieve a $0.90$
+probability of stopping assuming the underlying election is as announced.
+\item
+Subsequent EoR and SO \BRAVO round sizes can be found given the preceding evidence to 
+again achieve a probability of stopping $0.90$. It is not known whether \Minerva is 
+risk-limiting for round sizes chosen based on preceding evidence like this, and 
+so subsequent round sizes in \Minerva must be predetermined. We run simulations with
+\Minerva round schedules where each round has size given by multiplying the previous
+round size by $1.5$ or by $1.0$.
+
+\end{itemize}
+
+\begin{block}{Risk}
+\begin{figure}
+\includegraphics[width=\textwidth]{both_risk_plots.png}
+\caption{The left hand plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.
+The right hand plot, for each state margin, shows the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
+\label{fig:risk}
+%\end{centering}
+\end{figure}
+%
+\end{block}
+
+%\begin{block}{more}
+%We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
+%%%\begin{definition}[Stopping Probability]
+%\heading{Definition: Stopping Probability}
+%The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
+%$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
+%%\end{definition}
+%Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
 Note that $\sum _j S_j(\mathcal{A})=1$.
 We can also consider the cumulative stopping probability:
-%\begin{definition}[Cumulative Stopping Probability]
-The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
-%\end{definition}
-Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$. 
-%
-Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
-%\begin{definition}[Conditional Stopping Probability]
-The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
-$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
-%\end{definition}
-Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
+%%\begin{definition}[Cumulative Stopping Probability]
+%The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
+%%\end{definition}
+%Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$. 
+%%
+%We are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
+%%\begin{definition}[Conditional Stopping Probability]
+%The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
+%$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
+%%\end{definition}
+%Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
 those that did not stop before round $j$.
-%
-We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
-Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}.
+%%
+%We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
+%Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}.
 Therefore, the two candidate plurality contest is a general case, and these simulations provide insight for multiple-candidate and multiple-winner contests too.
-Round sizes increase roughly proportional to the inverse
-square of the margin, so 
-smaller margins are computationally much more expensive to simulate.
-For each of these states, we simulated 
-$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
-and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$). 
-  \end{block}
+%%Round sizes increase roughly proportional to the inverse
+%square of the margin, so 
+%smaller margins are computationally much more expensive to simulate.
+%For each of these states, we simulated 
+%$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
+%and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$). 
+%  \end{block}
 
@@
 
-Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
+We are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
 %\begin{definition}[Conditional Stopping Probability]
@@
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+\centering
 
-  \end{block}
+\begin{minipage}{.49\textwidth}
+\begin{centering}
+\includegraphics[width=1.0\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
+This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
+as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\label{fig:eor_bravo_sprob}
+\end{centering}\end{minipage}
+\begin{minipage}{.49\textwidth}
+\begin{centering}
+\includegraphics[width=1.0\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
+This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
+as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\label{fig:so_bravo_sprob}
+\end{centering}
+\end{minipage}
 
-  \begin{block}{Fusce aliquam magna velit}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{figure}
+%\begin{centering}
+%\includegraphics[width=0.6\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
+%This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
+%as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
+%%\label{fig:eor_bravo_sprob}
+%\end{centering}
+%\end{figure}
+%
+%\begin{figure}
+%\begin{centering}
+%\includegraphics[width=0.6\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
+%This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
+%as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
+%%\label{fig:so_bravo_sprob}
+%\end{centering}
+%\end{figure}
+%
+\begin{figure}[h]
+\centering
 
-    Et rutrum ex euismod vel. Pellentesque ultricies, velit in fermentum
-    vestibulum, lectus nisi pretium nibh, sit amet aliquam lectus augue vel
-    velit. Suspendisse rhoncus massa porttitor augue feugiat molestie. Sed
-    molestie ut orci nec malesuada. Sed ultricies feugiat est fringilla
-    posuere.
+\begin{minipage}{.49\textwidth}
+%\begin{figure}
+\begin{centering}
+\includegraphics[width=1\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.0$ and $\chi_1=0.9$.}
+\label{fig:minerva1_sprob}
+\end{centering}
+%\end{figure}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+%\begin{figure}
+\begin{centering}
+\includegraphics[width=1\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1=0.9$.}
+\label{fig:minerva1p5_sprob}
+\end{centering}
+%\end{figure}
+\end{minipage}
+\end{figure}
+
+We can perform a similar study for a first round size with $\chi_1=0.25$. 
+See Figure~\ref{fig:minerva_25} for an example, \Minerva with round mutiplier $1.5$.
 
 \begin{figure}
-      \centering
-      \begin{tikzpicture}
-        \begin{axis}[
-            scale only axis,
-            no markers,
-            domain=0:2*pi,
-            samples=100,
-            axis lines=center,
-            axis line style={-},
-            ticks=none]
-          \addplot[red] {sin(deg(x))};
-          \addplot[blue] {cos(deg(x))};
-        \end{axis}
-      \end{tikzpicture}
-      \caption{Another figure caption.}
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{minerva25percthen1p5_sprob.png}
+\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1 = 0.25$.}
+\label{fig:minerva_25}
+\end{centering}
 \end{figure}
@@
 
-  \begin{block}{Nam cursus consequat egestas}
-
-    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
-    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
-    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
-    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
-    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
-    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
-    sit amet sollicitudin risus tincidunt non.
-
-    Etiam sit amet tempus lorem, aliquet condimentum velit. Donec et nibh
-    consequat, sagittis ex eget, dictum orci. Etiam quis semper ante. Ut eu
-    mauris purus. Proin nec consectetur ligula. Mauris pretium molestie
-    ullamcorper. Integer nisi neque, aliquet et odio non, sagittis porta justo.
-
-    \begin{itemize}
-      \item \textbf{Sed consequat} id ante vel efficitur. Praesent congue massa
-        sed est scelerisque, elementum mollis augue iaculis.
-        \begin{itemize}
-          \item In sed est finibus, vulputate
-            nunc gravida, pulvinar lorem. In maximus nunc dolor, sed auctor eros
-            porttitor quis.
-          \item Fusce ornare dignissim nisi. Nam sit amet risus vel lacus
-            tempor tincidunt eu a arcu.
-          \item Donec rhoncus vestibulum erat, quis aliquam leo
-            gravida egestas.
-        \end{itemize}
-      \item \textbf{Sed luctus, elit sit amet} dictum maximus, diam dolor
-        faucibus purus, sed lobortis justo erat id turpis.
-      \item \textbf{Pellentesque facilisis dolor in leo} bibendum congue.
-        Maecenas congue finibus justo, vitae eleifend urna facilisis at.
-    \end{itemize}
-
-    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
-    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
-    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
-    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
-    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
-    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
-    sit amet sollicitudin risus tincidunt non.
-  \end{block}
-
+%  \begin{block}{Fusce aliquam magna velit}
+%
+%    Et rutrum ex euismod vel. Pellentesque ultricies, velit in fermentum
+%    vestibulum, lectus nisi pretium nibh, sit amet aliquam lectus augue vel
+%    velit. Suspendisse rhoncus massa porttitor augue feugiat molestie. Sed
+%    molestie ut orci nec malesuada. Sed ultricies feugiat est fringilla
+%    posuere.
+%
+%    \begin{figure}
+%      \centering
+%      \begin{tikzpicture}
+%        \begin{axis}[
+%            scale only axis,
+%            no markers,
+%            domain=0:2*pi,
+%            samples=100,
+%            axis lines=center,
+%            axis line style={-},
+%            ticks=none]
+%          \addplot[red] {sin(deg(x))};
+%          \addplot[blue] {cos(deg(x))};
+%        \end{axis}
+%      \end{tikzpicture}
+%      \caption{Another figure caption.}
+%    \end{figure}
+%
+%  \end{block}
+%
+%  \begin{block}{Nam cursus consequat egestas}
+%
+%    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
+%    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
+%    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
+%    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
+%    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
+%    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
+%    sit amet sollicitudin risus tincidunt non.
+%
+%    Etiam sit amet tempus lorem, aliquet condimentum velit. Donec et nibh
+%    consequat, sagittis ex eget, dictum orci. Etiam quis semper ante. Ut eu
+%    mauris purus. Proin nec consectetur ligula. Mauris pretium molestie
+%    ullamcorper. Integer nisi neque, aliquet et odio non, sagittis porta justo.
+%
+%    \begin{itemize}
+%      \item \textbf{Sed consequat} id ante vel efficitur. Praesent congue massa
+%        sed est scelerisque, elementum mollis augue iaculis.
+%        \begin{itemize}
+%          \item In sed est finibus, vulputate
+%            nunc gravida, pulvinar lorem. In maximus nunc dolor, sed auctor eros
+%            porttitor quis.
+%          \item Fusce ornare dignissim nisi. Nam sit amet risus vel lacus
+%            tempor tincidunt eu a arcu.
+%          \item Donec rhoncus vestibulum erat, quis aliquam leo
+%            gravida egestas.
+%        \end{itemize}
+%      \item \textbf{Sed luctus, elit sit amet} dictum maximus, diam dolor
+%        faucibus purus, sed lobortis justo erat id turpis.
+%      \item \textbf{Pellentesque facilisis dolor in leo} bibendum congue.
+%        Maecenas congue finibus justo, vitae eleifend urna facilisis at.
+%    \end{itemize}
+%    
+%    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
+%    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
+%    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
+%    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
+%    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
+%    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
+%    sit amet sollicitudin risus tincidunt non.
+%  \end{block}
+%
 \end{column}
@@
 
-  \begin{block}{A block containing some math}
+\begin{block}{Number of Ballots}
 
-    Nullam non est elit. In eu ornare justo. Maecenas porttitor sodales lacus,
-    ut cursus augue sodales ac.
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.6\textwidth]{texas25.png}
+\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
+\label{fig:texas_25}
+\end{centering}
+\end{figure}
 
-    $$
-    \int_{-\infty}^{\infty} e^{-x^2}\,dx = \sqrt{\pi}
-    $$
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.65\textwidth]{missouri25.png}
+\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
+\label{fig:missouri_25}
+\end{centering}
+\end{figure}
 
-    Interdum et malesuada fames $\{1, 4, 9, \ldots\}$ ac ante ipsum primis in
-    faucibus. Cras eleifend dolor eu nulla suscipit suscipit. Sed lobortis non
-    felis id vulputate.
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.65\textwidth]{massachusetts25.png}
+\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $S_1=0.25$.}
+\label{fig:mass_25}
+\end{centering}
+\end{figure}
 
-    \heading{A heading inside a block}
-
-    Praesent consectetur mi $x^2 + y^2$ metus, nec vestibulum justo viverra
-    nec. Proin eget nulla pretium, egestas magna aliquam, mollis neque. Vivamus
-    dictum $\mathbf{u}^\intercal\mathbf{v}$ sagittis odio, vel porta erat
-    congue sed. Maecenas ut dolor quis arcu auctor porttitor.
-
-    \heading{Another heading inside a block}
-
-    Sed augue erat, scelerisque a purus ultricies, placerat porttitor neque.
-    Donec $P(y \mid x)$ fermentum consectetur $\nabla_x P(y \mid x)$ sapien
-    sagittis egestas. Duis eget leo euismod nunc viverra imperdiet nec id
-    justo.
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.605\textwidth]{texas90.png}
+\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.9$.}
+\label{fig:texas_90}
+\end{centering}
+\end{figure}
 
@@
 
-  \begin{block}{Nullam vel erat at velit convallis laoreet}
-
-    Class aptent taciti sociosqu ad litora torquent per conubia nostra, per
-    inceptos himenaeos. Phasellus libero enim, gravida sed erat sit amet,
-    scelerisque congue diam. Fusce dapibus dui ut augue pulvinar iaculis.
-
-    \begin{table}
-      \centering
-      \begin{tabular}{l r r c}
-        \toprule
-        \textbf{First column} & \textbf{Second column} & \textbf{Third column} & \textbf{Fourth} \\
-        \midrule
-        Foo & 13.37 & 384,394 & $\alpha$ \\
-        Bar & 2.17 & 1,392 & $\beta$ \\
-        Baz & 3.14 & 83,742 & $\delta$ \\
-        Qux & 7.59 & 974 & $\gamma$ \\
-        \bottomrule
-      \end{tabular}
-      \caption{A table caption.}
-    \end{table}
-
-    Donec quis posuere ligula. Nunc feugiat elit a mi malesuada consequat. Sed
-    imperdiet augue ac nibh aliquet tristique. Aenean eu tortor vulputate,
-    eleifend lorem in, dictum urna. Proin auctor ante in augue tincidunt
-    tempor. Proin pellentesque vulputate odio, ac gravida nulla posuere
-    efficitur. Aenean at velit vel dolor blandit molestie. Mauris laoreet
-    commodo quam, non luctus nibh ullamcorper in. Class aptent taciti sociosqu
-    ad litora torquent per conubia nostra, per inceptos himenaeos.
-
-    Nulla varius finibus volutpat. Mauris molestie lorem tincidunt, iaculis
-    libero at, gravida ante. Phasellus at felis eu neque suscipit suscipit.
-    Integer ullamcorper, dui nec pretium ornare, urna dolor consequat libero,
-    in feugiat elit lorem euismod lacus. Pellentesque sit amet dolor mollis,
-    auctor urna non, tempus sem.
-
-  \end{block}
+%  \begin{block}{A block containing some math}
+%
+%    Nullam non est elit. In eu ornare justo. Maecenas porttitor sodales lacus,
+%    ut cursus augue sodales ac.
+%
+%    $$
+%    \int_{-\infty}^{\infty} e^{-x^2}\,dx = \sqrt{\pi}
+%    $$
+%
+%    Interdum et malesuada fames $\{1, 4, 9, \ldots\}$ ac ante ipsum primis in
+%    faucibus. Cras eleifend dolor eu nulla suscipit suscipit. Sed lobortis non
+%    felis id vulputate.
+%
+%    \heading{A heading inside a block}
+%
+%    Praesent consectetur mi $x^2 + y^2$ metus, nec vestibulum justo viverra
+%    nec. Proin eget nulla pretium, egestas magna aliquam, mollis neque. Vivamus
+%    dictum $\mathbf{u}^\intercal\mathbf{v}$ sagittis odio, vel porta erat
+%    congue sed. Maecenas ut dolor quis arcu auctor porttitor.
+%
+%    \heading{Another heading inside a block}
+%
+%    Sed augue erat, scelerisque a purus ultricies, placerat porttitor neque.
+%    Donec $P(y \mid x)$ fermentum consectetur $\nabla_x P(y \mid x)$ sapien
+%    sagittis egestas. Duis eget leo euismod nunc viverra imperdiet nec id
+%    justo.
+%
+%  \end{block}
+%
+%  \begin{block}{Nullam vel erat at velit convallis laoreet}
+%
+%%    Class aptent taciti sociosqu ad litora torquent per conubia nostra, per
+%    inceptos himenaeos. Phasellus libero enim, gravida sed erat sit amet,
+%    scelerisque congue diam. Fusce dapibus dui ut augue pulvinar iaculis.
+%%
+%    \begin{table}
+%      \centering
+%%      \begin{tabular}{l r r c}
+%        \toprule
+%%        \textbf{First column} & \textbf{Second column} & \textbf{Third column} & \textbf{Fourth} \\
+%        \midrule
+%        Foo & 13.37 & 384,394 & $\alpha$ \\
+%%        Bar & 2.17 & 1,392 & $\beta$ \\
+%        Baz & 3.14 & 83,742 & $\delta$ \\
+%        Qux & 7.59 & 974 & $\gamma$ \\
+%%        \bottomrule
+%      \end{tabular}
+%      \caption{A table caption.}
+%%    \end{table}
+%
+%    Donec quis posuere ligula. Nunc feugiat elit a mi malesuada consequat. Sed
+%%    imperdiet augue ac nibh aliquet tristique. Aenean eu tortor vulputate,
+%    eleifend lorem in, dictum urna. Proin auctor ante in augue tincidunt
+%%    tempor. Proin pellentesque vulputate odio, ac gravida nulla posuere
+%    efficitur. Aenean at velit vel dolor blandit molestie. Mauris laoreet
+%%    commodo quam, non luctus nibh ullamcorper in. Class aptent taciti sociosqu
+%    ad litora torquent per conubia nostra, per inceptos himenaeos.
+%
+%    Nulla varius finibus volutpat. Mauris molestie lorem tincidunt, iaculis
+%%    libero at, gravida ante. Phasellus at felis eu neque suscipit suscipit.
+%    Integer ullamcorper, dui nec pretium ornare, urna dolor consequat libero,
+%    in feugiat elit lorem euismod lacus. Pellentesque sit amet dolor mollis,
+%%    auctor urna non, tempus sem.
+%
+%%  \end{block}
 

### 2022-04-12 MultipleRoundSims/poster.tex
progress

@@
 \newcommand{\Minerva}{\textsc{Minerva}\xspace}
+\newcommand{\Prov}{\textsc{Providence}\xspace}
 \newcommand{\B}{{{B2}}\xspace}
@@
 \item 
-We used the R2B2 library, which provides a framework for the exploration
+We used the R2B2 library\cite{r2b2}, which provides a framework for the exploration
 of round-by-round and ballot-by-ballot RLAs. It has implementations of several
@@
 \begin{block}{Risk}
+
 \begin{figure}
@@
 \begin{centering}
-\includegraphics[width=1.0\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
-This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
-as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\includegraphics[width=1.0\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round, as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
 \label{fig:eor_bravo_sprob}
@@
 \begin{centering}
-\includegraphics[width=1.0\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
-This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
-as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\includegraphics[width=1.0\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
 \label{fig:so_bravo_sprob}
@@
 \end{minipage}
-
+\caption{These plots show, for each state margin, when the underlying election is as announced, the number of audits that stopped in the $j^{th}$ round, as a fraction of all audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.
+The left shows EoR \BRAVO and the right shows SO \BRAVO.}
 \end{figure}
@@
 \includegraphics[width=1\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
-as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.0$ and $\chi_1=0.9$.}
+%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.0$ and $\chi_1=0.9$.}
 \label{fig:minerva1_sprob}
@@
 \includegraphics[width=1\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
-\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
-as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1=0.9$.}
+%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1=0.9$.}
 \label{fig:minerva1p5_sprob}
@@
 \end{minipage}
+\caption{These plots show, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.
+The left shows \Minerva with the round schedule obtained by $\chi_1=0.9$ and round size multiple $1.5$.}
 \end{figure}
@@
 \begin{block}{Number of Ballots}
+\begin{itemize}
+\item
+The number of ballots sampled is one crude measure of the workload of an audit.
+To keep the costs of RLAs low, audits should be designed to stop with as few ballots as possible.
+\item
+The following plots show the probability of stopping as a function of the average number of ballots
+sampled by round in our simulations.
+\end{itemize}
 
-\begin{figure}
+\begin{figure}[h]
+\centering
+\begin{minipage}{.49\textwidth}
 \begin{centering}
-\includegraphics[width=0.6\textwidth]{texas25.png}
-\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
+\includegraphics[width=1\textwidth]{texas25.png}
+%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
 \label{fig:texas_25}
 \end{centering}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\begin{centering}
+\includegraphics[width=1\textwidth]{missouri25.png}
+%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
+\label{fig:missouri_25}
+\end{centering}
+\end{minipage}
+\caption{These plots show the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states of Texas with margin $.057$ (left) and Missouri with margin $0.157$ (right), both with first round stopping probability $\chi_1=0.25$.}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[h]
+\centering
+\begin{minipage}{.49\textwidth}
 \begin{centering}
-\includegraphics[width=0.65\textwidth]{missouri25.png}
-\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
-\label{fig:missouri_25}
+\includegraphics[width=1\textwidth]{massachusetts25.png}
+%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $\chi_1=0.25$.}
+\label{fig:mass_25}
 \end{centering}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\begin{centering}
+\includegraphics[width=1\textwidth]{texas90.png}
+%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $\chi_1=0.9$.}
+\label{fig:texas_90}
+\end{centering}
+\end{minipage}
+\caption{These plots show the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states Massachusetts (left) with margin $0.342$ and $\chi_1=0.25$ and Texas (right) with margin $0.057$ and $\chi_1=0.9$.}
 \end{figure}
 
-\begin{figure}
-\begin{centering}
-\includegraphics[width=0.65\textwidth]{massachusetts25.png}
-\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $S_1=0.25$.}
-\label{fig:mass_25}
-\end{centering}
-\end{figure}
+For $\chi_1=0.25$, the ratio of first round size of EoR \BRAVO to \Minerva is $1.45$, $1.37$, $1.23$ for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+This may be compared to $2.03$, $1.99$ and $1.8$ respectively for $\chi_1=0.9$.
+Similarly, for $\chi_1=0.25$, the ratio of first round size of SO \BRAVO to \Minerva is $1.13$, $1.08$, $1.12$ for states Texas, Missouri and Massachusetts, and margins $0.057$, $0.157$ and $0.342$ respectively.
+This may be compared to $1.38$, $1.38$ and $1.30$ respectively for $\chi_1=0.9$.
 
-\begin{figure}
-\begin{centering}
-\includegraphics[width=0.605\textwidth]{texas90.png}
-\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.9$.}
-\label{fig:texas_90}
-\end{centering}
-\end{figure}
+\end{block}
+
+\begin{block}{Modeling Workload}
+
+\BRAVO requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
+In real audits, election officials draw ballots in rounds because their is some overhead with each round (e.g. opening ballot boxes).
+Therefore a more sophisticated workload model has a per ballot cost $c_b$ and a per round cost $c_r$. 
+Let $n$ be the number of ballots sampled and $j$ be the number of rounds. Then the cost $C$ is given by 
+$$C(n,j) = c_b\cdot n + c_r\cdot j,$$
+and depending on the ratio between $c_b$ and $c_r$, we should expect that certain round schedules
+achieve lower expected cost.
 
@@
 \begin{column}{\colwidth}
-
 \begin{block}{Providence}
 
-Providence is great! Efficient like Minerva but flexible like BRAVO.
+\begin{itemize}
+\item
+The efficiency of \Minerva is great, but it lacks the flexibility of \BRAVO in choosing round sizes based on previous samples.
+\item
+\Prov is our novel RLA which has the efficiency of \Minerva and the flexibility of \BRAVO.
+\item For alternative hypothesis $H_a$ that the election is truly as announced and null
+hypothesis $H_0$ that the true election is a tie, \BRAVO has the stopping condition that
+for $k$ cumulative ballots for the winner and $n$ cumulative sampled ballots,
+$$\sigma(k,n,p_a,p_0) \triangleq \frac{\Pr[K=k\mid H_a]}{\Pr[K=k\mid H_0]}\ge \frac{1}{\alpha}.$$
+\item \Minerva has the stopping condition that in round $j$ with cumulative winner ballots $k_j$ and round sizes $n_1,n_2,\ldots,n_j$
+$$\tau_j(k_j,n_1,n_2,\ldots,n_j,p_a,p_0) \triangleq \frac{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_a]}{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_0]}\ge \frac{1}{\alpha}.$$ Testing this stopping condition requires computationaly expensive convolutions.
+\item The \Prov stopping condition uses ideas from both \BRAVO and \Minerva and requires no convolution to test:
+$$\omega_j(k_{j-1},k_j,n_{j-1},n_j,p_a,p_0)\triangleq \sigma(k_{j-1},n_{j-1},p_a,p_0)\cdot \tau_1(k_j,n_j,p_a,p_0) \ge \frac{1}{\alpha}.$$
+
+\end{itemize}
+
+\heading {Resistance against an adversary} 
+If adversary $\mathcal{A}$ has knowledge of previous samples and chooses round size $n_j$ in round $j$, 
+\Prov is still risk-limiting.
+
+\heading {\Prov pilot}
+In February 2022, The Rhode Island Board of Elections hosted a public pilot \Prov RLA which passed 
+in the first round.
+
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{dice.jpg}
+\caption{Professor Vora tosses her random-seed-generating 10-sided die, in a roll she dedicated to election officials everywhere.}
+\end{figure}
 
@@
 
-Look, Providence does pretty ok!
+Need to fill in simulation results here. I am thinking (a) risk simulations as evidence that \Prov is an RLA, 
+and (b) number of ballot results if I get them done quickly enough.
 
@@
 
-    \nocite{*}
-    \footnotesize{\bibliographystyle{plain}\bibliography{poster}}
+    %\nocite{*}
+    \footnotesize{\bibliographystyle{plain}\bibliography{audits}}
+    %\bibliographystyle{splncs04}
+    %\bibliography{audits}
 

### 2022-04-12 MultipleRoundSims/poster.tex
notes

@@
 
+\emph{I initially wrote this section since I'd really like to have some evidence that for certain workload parameters, we see an optimal
+number of rounds greater than 1 with Providence. I am working on that code now.}
+
 \BRAVO requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
@@
 achieve lower expected cost.
-
 \end{block}
@@
 for $k$ cumulative ballots for the winner and $n$ cumulative sampled ballots,
-$$\sigma(k,n,p_a,p_0) \triangleq \frac{\Pr[K=k\mid H_a]}{\Pr[K=k\mid H_0]}\ge \frac{1}{\alpha}.$$
-\item \Minerva has the stopping condition that in round $j$ with cumulative winner ballots $k_j$ and round sizes $n_1,n_2,\ldots,n_j$
-$$\tau_j(k_j,n_1,n_2,\ldots,n_j,p_a,p_0) \triangleq \frac{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_a]}{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_0]}\ge \frac{1}{\alpha}.$$ Testing this stopping condition requires computationaly expensive convolutions.
+$$\sigma(k,n,p_a,p_0) \triangleq \frac{\Pr[K=k\mid H_a, n]}{\Pr[K=k\mid H_0, n]}\ge \frac{1}{\alpha}.$$
+\item \Minerva has the stopping condition that in round $j$ with cumulative winner ballots $k_j$ and round sizes $\bar n_j = n_1,n_2,\ldots,n_j$
+$$\tau_j(k_j,\bar n_j,p_a,p_0) \triangleq \frac{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_a, \bar n_j]}{\Pr[K_j\ge k_j \wedge \mathcal{A}_{i<j}(X)\neq Correct \mid H_0, \bar n_j]}\ge \frac{1}{\alpha}.$$ Testing this stopping condition requires computationaly expensive convolutions.
 \item The \Prov stopping condition uses ideas from both \BRAVO and \Minerva and requires no convolution to test:
@@
 If adversary $\mathcal{A}$ has knowledge of previous samples and chooses round size $n_j$ in round $j$, 
-\Prov is still risk-limiting.
+\Prov is still risk-limiting. Proof of this and the RLA property of \Prov are currently being revised for submission.
 
@@
 
-Need to fill in simulation results here. I am thinking (a) risk simulations as evidence that \Prov is an RLA, 
-and (b) number of ballot results if I get them done quickly enough.
+\emph{Need to fill in simulation results here. I am thinking (a) risk simulations as evidence that \Prov is an RLA, and (b) number of ballot results if I get them done quickly enough.}
+
+Simulations are currently running to explore the risk, stopping probability, and number of ballots used by \Prov.
 
@@
     %\nocite{*}
-    \footnotesize{\bibliographystyle{plain}\bibliography{audits}}
-    %\bibliographystyle{splncs04}
-    %\bibliography{audits}
+    %\footnotesize{\bibliographystyle{plain}\bibliography{audits}}
+    \bibliographystyle{splncs04}
+    %\bibliographystyle{plain}
+    \bibliography{audits}
+    \emph{Having some latex trouble here, but should include r2b2, the recent simulations paper, bravo paper, athena paper, ...}
 

### 2022-04-13 MultipleRoundSims/poster.tex
poster refactoring

@@
 
-\begin{block}{Risk-Limiting Audits}
+\begin{block}{Original Contributions}
+
+Risk-Limiting Audits (RLAs) are rigorous election audits performed in rounds. 
+Most commonly used RLAs EoR \BRAVO and SO \BRAVO (Stark) rely on the Sequential Probability Ratio Test (SPRT), using
+a ratio of sample probabilities conditioned on alternative and null hypotheses.
+Newer RLA \Minerva (Vora) uses a ratio of the tails of those distributions, requiring half as many
+ballots as EoR \BRAVO in the first round when probability of stopping is $0.9$.
+\Minerva was used in a pilot RLA in Montgomery County, Ohio in 2020,
+and is recommended by the largest US voting organizations: Verified Voting, the Brennan Center, and Common Cause, 
+It is unknown how the audits compare in later rounds or with lower stopping probabilities, a problem with no easy analytical approach.
+Unlike EoR and SO \BRAVO, \Minerva is proven to be risk-limiting only when round sizes are predetermined, 
+meaning round sizes cannot be chosen based on previous samples.
 \begin{itemize}
 \item
-In election security, a risk-limiting audit (RLA) draws trusted paper ballots in rounds, stopping if a rigorous statistical criterion is satisfied, or proceeding to a full hand count.
+We provide simulations of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
+first round stopping probability; this information can be used to advise election officials.
+\item
+We introduce \Prov, an RLA which has the efficient tail ratio approach of \Minerva but is resistant to an adversary who can pick
+future round sizes with knowledge of previous samples. \Prov was named after the Rhode Island city where it was used for a public
+pilot audit in February 2022.
+\end{itemize}
+\end{block}
+
+\begin{block}{Results}
+
+\begin{itemize}
+\item Both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
+However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability.
+\item Proofs that \Prov is risk-limiting and resistant to an adversary who can pick future round sizes with knowledge of previous samples.
+\end{itemize}
+\end{block}
+
+\begin{block}{Risk-Limiting Audits}
+
+\begin{itemize}
+\item
+A risk-limiting audit (RLA) draws paper ballots in rounds, stopping if a rigorous statistical criterion is satisfied, or proceeding to a full hand count.
 If the announced outcome of the election is erroneous, an RLA will detect the error with high, predetermined minimum probability.
-
 \item
@@
 (1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
-
-%\item
-%
-%In this work, we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of RLAs Minerva, Selection-Ordered (SO) Bravo, and End-of-Round (EoR) Bravo.
-We show that Minerva requires fewer ballots on average than either version of Bravo for round sizes that are chosen to give probabilities of stopping 0.9 and 0.25.
-In particular, the advantage of Minerva is less for the smaller stopping probabil- ity.
-We also present Providence, a novel ballot polling RLA.
-Based on initial simulation results, Providence has efficiency similar to Minerva but is also resistant to an adversary who can choose future round sizes with knowledge of previous samples.
-Thus, unlike Minerva, Providence allows election officials to choose round sizes during the audit to optimize for workload and safety.
-%\end{block}
-
-%\begin{block}{Risk}
-%Definitions
-%In this section, we motivate and describe the experiments.
-We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.
-Note that sampling without replacement is more efficient for large sampling fractions, but \Minerva has not been extended for sampling without replacement.
-We first present relevant definitions.
-
-%\begin{definition}
-%An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
-%(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
-%\end{definition}
-
-%All of the audits discussed in this paper are modeled as binary hypothesis tests.
-Under the alternative hypothesis, $H_a$, the announced outcome is correct.
-In particular, the true underlying ballot distribution is given by the announced ballot tallies.
-Under the null hypothesis, $H_0$, a tie is the correct outcome \footnote{or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser}.
-%The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}.
-Note that an audit $\mathcal{A}$ includes all audit parameters (maximum risk, round sizes, etc.).
-
-%\begin{definition}[Risk]
-%\heading{Definition: Risk}
 \item
@@
 hypothesis: the true underlying election is a tie.
-%\end{definition}
-
-%This leads us to the following definition of an $\alpha$-RLA.
-%\heading{Definition: Risk-Limiting Audit ($\alpha$-RLA)}
-%\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
 \item
@@
 $$R(\mathcal{A}) \le \alpha.$$
-%\end{definition}
-\end{itemize}
-\end{block}
-\begin{block}{\BRAVO and \Minerva}
-\begin{itemize}
-\item Existing RLAs include \BRAVO (which can be implemented as End-of-Round (EoR) \BRAVO or Selection-Ordered (SO) \BRAVO)
-and \Minerva. In a first round with a large sample size (probabilty of stopping at least 90\%), \Minerva is known to require roughly
-half as many ballots as EoR \BRAVO. 
-\item We provide the first simulations that compare these audits beyond a first round and for various
-round schedules.
 \end{itemize}
@@
 \begin{block}{Simulations}
+
 \begin{itemize}
-%\item
-%Simulations can be used to provide evidence for theoretical claims about risk-limiting audits.
 \item 
-We ran simulations to gain further insight into audit behavior 
-and provide additional evidence for theoretical claims.
+Software: R2B2 library\cite{r2b2} which has Python implementations of several RLAs as well as a simulator
 \item 
-We simulated audits for a risk limit of 10\% using margins
-from the 2020 US Presidential election, limiting ourselves to pairwise margins
-for the two main candidates of $0.05$ or larger.
+Risk limit: 10\%
 \item
-We used the R2B2 library\cite{r2b2}, which provides a framework for the exploration
-of round-by-round and ballot-by-ballot RLAs. It has implementations of several
-ballot polling risk-limiting audits as well as a simulator, all written in Python.
+Margins: 2020 US Presidential election statewide results, limiting ourselves to pairwise margins
+for the two main candidates of $0.05$ or larger
 \item 
-For each of these states, we simulated 
-$10,000=10^4$ audits assuming the underlying election was as announced,  
-and an additional $10,000=10^4$ audits assuming the underlying election was a tie. 
+Trials per state: $10,000=10^4$ audits assuming the underlying election is as announced,  
+and $10,000=10^4$ audits assuming the underlying election is a tie
+\item 
+Round schedules: EoR \BRAVO and SO \BRAVO round sizes are chosen to achieve the same
+probability of stopping in each round given preceding samples. \Minerva first round sizes
+are chosen to achieve some probability of stopping and subsequent round sizes are given by multiplying the previous
+round size by $1.5$ or by $1.0$.
 \end{itemize}
 \end{block}
-
-\heading{Round Schedules}
-\begin{itemize}
-\item
-A round schedule is a list of positive integers corresponding to the number of ballots 
-sampled in each round.
-To begin, first round sizes are selected (by search) to achieve a $0.90$
-probability of stopping assuming the underlying election is as announced.
-\item
-Subsequent EoR and SO \BRAVO round sizes can be found given the preceding evidence to 
-again achieve a probability of stopping $0.90$. It is not known whether \Minerva is 
-risk-limiting for round sizes chosen based on preceding evidence like this, and 
-so subsequent round sizes in \Minerva must be predetermined. We run simulations with
-\Minerva round schedules where each round has size given by multiplying the previous
-round size by $1.5$ or by $1.0$.
-
-\end{itemize}
-
-\begin{block}{Risk}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{both_risk_plots.png}
-\caption{The left hand plot shows the fraction of EoR \BRAVO audits (all states with margins at least $0.05$) and SO \BRAVO audits (the 13 states for which our simulations are complete so far) that stopped in any of the $5$ rounds when the underlying election was a tie.
-The right hand plot, for each state margin, shows the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
-\label{fig:risk}
-%\end{centering}
-\end{figure}
-%
-\end{block}
-
-%\begin{block}{more}
-%We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
-%%%\begin{definition}[Stopping Probability]
-%\heading{Definition: Stopping Probability}
-%The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
-%$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
-%%\end{definition}
-%Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.
-Note that $\sum _j S_j(\mathcal{A})=1$.
-We can also consider the cumulative stopping probability:
-%%\begin{definition}[Cumulative Stopping Probability]
-%The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
-%%\end{definition}
-%Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$. 
-%%
-%We are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
-%%\begin{definition}[Conditional Stopping Probability]
-%The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
-%$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
-%%\end{definition}
-%Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
-those that did not stop before round $j$.
-%%
-%We simulated audits for a risk limit of $10\%$ (as in \cite{bravo} and \cite{usenix_minerva}) using margins from the 2020 US Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger.
-%Note that both \BRAVO and \Minerva can be extended for multiple-candidate, multiple-winner plurality contests by performing pairwise tests between the winners and the losers\cite{RLA, arxiv_athena}.
-Therefore, the two candidate plurality contest is a general case, and these simulations provide insight for multiple-candidate and multiple-winner contests too.
-%%Round sizes increase roughly proportional to the inverse
-%square of the margin, so 
-%smaller margins are computationally much more expensive to simulate.
-%For each of these states, we simulated 
-%$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
-%and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$). 
-%  \end{block}
-
-%  \begin{block}{A block containing a list}
-%
-%    Nam vulputate nunc felis, non condimentum lacus porta ultrices. Nullam sed
-%    sagittis metus. Etiam consectetur gravida urna quis suscipit.
-%
-%    \begin{itemize}
-%      \item \textbf{Mauris tempor} risus nulla, sed ornare
-%      \item \textbf{Libero tincidunt} a duis congue vitae
-%      \item \textbf{Dui ac pretium} morbi justo neque, ullamcorper
-%    \end{itemize}
-%
-%    Eget augue porta, bibendum venenatis tortor.
-%
-%  \end{block}
-%
-%  \begin{alertblock}{A highlighted block}
-%
-%    This block catches your eye, so \textbf{important stuff} should probably go
-%%    here.
-%%
-%%%    Curabitur eu libero vehicula, cursus est fringilla, luctus est. Morbi
-%%    consectetur mauris quam, at finibus elit auctor ac. Aliquam erat volutpat.
-%%%    Aenean at nisl ut ex ullamcorper eleifend et eu augue. Aenean quis velit
-%    tristique odio convallis ultrices a ac odio.
-%
-%    \begin{itemize}
-%      \item \textbf{Fusce dapibus tellus} vel tellus semper finibus. In
-%        consequat, nibh sed mattis luctus, augue diam fermentum lectus.
-%      \item \textbf{In euismod erat metus} non ex. Vestibulum luctus augue in
-%        mi condimentum, at sollicitudin lorem viverra.
-%      \item \textbf{Suspendisse vulputate} mauris vel placerat consectetur.
-%        Mauris semper, purus ac hendrerit molestie, elit mi dignissim odio, in
-%        suscipit felis sapien vel ex.
-%    \end{itemize}
-%
-%    Aenean tincidunt risus eros, at gravida lorem sagittis vel. Vestibulum ante
-%    ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae.
-%
-%  \end{alertblock}
 
@@
 
+\begin{block}{Risk}
+
+\begin{itemize}
+\item 
+All audits have estimated risks below the risk limit
+\item
+EoR \BRAVO falls an order of magnitude less than the others, unnecessarily conservative
+\end{itemize}
+
+\begin{figure}[h]
+\centering
+\begin{minipage}{.49\textwidth}
+\includegraphics[width=\textwidth]{bravo_risks.png}
+\label{fig:bravo_risk}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\includegraphics[width=1.0\textwidth]{riskminerva1p5_10t4.png}
+\label{fig:minerva_risk}
+\end{minipage}
+\caption{These plots show for $\chi_1=0.9$ the fraction of audits (EoR \BRAVO (left), SO \BRAVO (left), and \Minerva with multiplier $1.5$ (right)) that stopped in any of the $5$ rounds when the underlying election was a tie.
+(For SO \BRAVO audits we show the 13 states for which simulations are complete.)}
+%\label{fig:risk}
+\end{figure}
+
+\end{block}
+
 \begin{block}{Stopping Probability}
 
-We are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
-%\begin{definition}[Conditional Stopping Probability]
-The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
-$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
-%\end{definition}
-Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e.
-those that did not stop before round $j$.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}
+\item
+Conditional stopping probability, $\chi_j$, is the probability that an audit will stop in round $j$ given that it did not stop earlier, and in 
+simulations $\chi_j$ is estimated by the proportion of audits that stop in a round to those that ``entered'' that round
+i.e. those that did not stop before round $j$. 
+%The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
+%$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
+\end{itemize}
+
 \begin{figure}[h]
 \centering
-
 \begin{minipage}{.49\textwidth}
-\begin{centering}
-\includegraphics[width=1.0\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round, as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\includegraphics[width=1.0\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}
 \label{fig:eor_bravo_sprob}
-\end{centering}\end{minipage}
+\end{minipage}
 \begin{minipage}{.49\textwidth}
-\begin{centering}
-\includegraphics[width=1.0\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.}
+\includegraphics[width=1.0\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}
 \label{fig:so_bravo_sprob}
-\end{centering}
 \end{minipage}
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{figure}
-%\begin{centering}
-%\includegraphics[width=0.6\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
-%This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
-%as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
-%%\label{fig:eor_bravo_sprob}
-%\end{centering}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{centering}
-%\includegraphics[width=0.6\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
-%This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
-%as a fraction of all SO \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $S_1=0.9$.}
-%%\label{fig:so_bravo_sprob}
-%\end{centering}
-%\end{figure}
-%
+
 \begin{figure}[h]
 \centering
-
 \begin{minipage}{.49\textwidth}
-%\begin{figure}
-\begin{centering}
 \includegraphics[width=1\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.0$ and $\chi_1=0.9$.}
 \label{fig:minerva1_sprob}
-\end{centering}
-%\end{figure}
 \end{minipage}
 \begin{minipage}{.49\textwidth}
-%\begin{figure}
-\begin{centering}
 \includegraphics[width=1\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
-%\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1=0.9$.}
 \label{fig:minerva1p5_sprob}
-\end{centering}
-%\end{figure}
 \end{minipage}
@@
 
-We can perform a similar study for a first round size with $\chi_1=0.25$. 
-See Figure~\ref{fig:minerva_25} for an example, \Minerva with round mutiplier $1.5$.
+We also consider a first round size with $\chi_1=0.25$.
 
@@
 
-\begin{block}{Modeling Workload}
-
-\emph{I initially wrote this section since I'd really like to have some evidence that for certain workload parameters, we see an optimal
-number of rounds greater than 1 with Providence. I am working on that code now.}
-
-\BRAVO requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
-In real audits, election officials draw ballots in rounds because their is some overhead with each round (e.g. opening ballot boxes).
-Therefore a more sophisticated workload model has a per ballot cost $c_b$ and a per round cost $c_r$. 
-Let $n$ be the number of ballots sampled and $j$ be the number of rounds. Then the cost $C$ is given by 
-$$C(n,j) = c_b\cdot n + c_r\cdot j,$$
-and depending on the ratio between $c_b$ and $c_r$, we should expect that certain round schedules
-achieve lower expected cost.
-\end{block}
+%work load workload WORKLOAD
+%\begin{block}{Modeling Workload}
+%
+%\emph{I initially wrote this section since I'd really like to have some evidence that for certain workload parameters, we see an optimal
+%number of rounds greater than 1 with Providence. I am working on that code now.}
+%
+%\BRAVO requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
+%In real audits, election officials draw ballots in rounds because their is some overhead with each round (e.g. opening ballot boxes).
+%Therefore a more sophisticated workload model has a per ballot cost $c_b$ and a per round cost $c_r$. 
+%Let $n$ be the number of ballots sampled and $j$ be the number of rounds. Then the cost $C$ is given by 
+%$$C(n,j) = c_b\cdot n + c_r\cdot j,$$
+%and depending on the ratio between $c_b$ and $c_r$, we should expect that certain round schedules
+%achieve lower expected cost.
+%\end{block}
 
@@
 \includegraphics[width=1.0\textwidth]{dice.jpg}
-\caption{Professor Vora tosses her random-seed-generating 10-sided die, in a roll she dedicated to election officials everywhere.}
+\caption{
+Professor Vora tosses her random-seed-generating 10-sided die, in a roll she dedicated to election officials everywhere.
+The standing people from left to right are (1) Miguel Nunez, Rhode Island Deputy Director of Elections, 
+(2) Professor Vora, and (3) Mark Lindeman, Verified Voting Director.
+}
 \end{figure}

### 2022-04-13 MultipleRoundSims/slides.tex
poster refactoring

--- 2022-01-16 MultipleRoundSims/draft.tex
@@
 %
-\documentclass[runningheads]{llncs}
+\documentclass{beamer}
 %
@@
 Sarah Morin\inst{1} \and
-Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
+Grant McClearn\inst{2}\\ \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
 Filip Zag{\'o}rski\inst{3}\inst{4}}
 %
-\authorrunning{Broadrick, Morin et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Department of Computer Science, The George Washington University\thanks{Supported in part by NSF Award 2015253} 
-\email{obroadrick@gwmail.gwu.edu}\\
-\and
-Department of Computer Science, Stanford University 
-\email{grantmcc@stanford.edu}
-\\
-\and Wroclaw University of Science and Technology\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
-\and Votifica
-\\
-\email{filip.zagorski@gmail.com}
+\institute{\inst{1} Department of Computer Science, The George Washington University (odbroadrick@gmail.com)
+%\thanks{Supported in part by NSF Award 2015253} 
+\and \inst{2} Department of Computer Science, Stanford University (grantmcc@stanford.edu)
+\and \inst{3} Wroclaw University of Science and Technology (filip.zagorski@gmail.com)
+%\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
+\and \inst{4} Votifica
 }
@@
 
-\maketitle              % typeset the header of the contribution
+%\maketitle              % typeset the header of the contribution
+\frame{\titlepage}
 %
-\begin{abstract}
+
+\begin{frame}
+\frametitle{Outline}
+
+\begin{itemize}
+\item Risk-Limiting Audits
+\begin{itemize}
+\item \BRAVO and \Minerva
+\end{itemize}
+\item 
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+%\begin{abstract}
     In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
@@
 \keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
-\end{abstract}
+\end{frame}
+%\end{abstract}
 %
@@
 \end{document}
-
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}

### 2022-04-14 MultipleRoundSims/poster.tex
poster update, missing ASN for providence only and reference bug fix

@@
 Risk-Limiting Audits (RLAs) are rigorous election audits performed in rounds. 
-Most commonly used RLAs EoR \BRAVO and SO \BRAVO (Stark) rely on the Sequential Probability Ratio Test (SPRT), using
+RLAs EoR \BRAVO and SO \BRAVO \cite{bravo} rely on the Sequential Probability Ratio Test (SPRT), using
 a ratio of sample probabilities conditioned on alternative and null hypotheses.
-Newer RLA \Minerva (Vora) uses a ratio of the tails of those distributions, requiring half as many
-ballots as EoR \BRAVO in the first round when probability of stopping is $0.9$.
+More recent RLA \Minerva \cite{usenix_minerva} uses a ratio of the tails of the same distributions, and requires half as many
+ballots as EoR \BRAVO in a first round with probability of stopping $0.9$.
+Understanding audit behavior for later rounds has no easy analytical approach.
 \Minerva was used in a pilot RLA in Montgomery County, Ohio in 2020,
-and is recommended by the largest US voting organizations: Verified Voting, the Brennan Center, and Common Cause, 
-It is unknown how the audits compare in later rounds or with lower stopping probabilities, a problem with no easy analytical approach.
-Unlike EoR and SO \BRAVO, \Minerva is proven to be risk-limiting only when round sizes are predetermined, 
-meaning round sizes cannot be chosen based on previous samples.
+and is recommended by the largest US voting organizations: Verified Voting, the Brennan Center, and Common Cause.
+%It is unknown how the audits compare in later rounds or with lower stopping probabilities, a problem with no easy analytical approach.
+
 \begin{itemize}
 \item
-We provide simulations of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
+We provide simulations \cite{simulations} of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
 first round stopping probability; this information can be used to advise election officials.
-\item
-We introduce \Prov, an RLA which has the efficient tail ratio approach of \Minerva but is resistant to an adversary who can pick
-future round sizes with knowledge of previous samples. \Prov was named after the Rhode Island city where it was used for a public
+\end{itemize}
+
+Unlike EoR and SO \BRAVO, \Minerva is proven to be risk-limiting only when round sizes are predetermined, 
+meaning round sizes cannot be chosen based on previous samples, flexibility that may have meaningful impact on workload.
+
+\begin{itemize}
+\item We introduce \Prov, an RLA which has the efficient tail ratio approach of \Minerva but is resistant to an adversary who can pick
+future round sizes with knowledge of previous samples. 
+\item \Prov was named after the Rhode Island city where it was used for a public
 pilot audit in February 2022.
@@
 \begin{itemize}
-\item Both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
-However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability.
-\item Proofs that \Prov is risk-limiting and resistant to an adversary who can pick future round sizes with knowledge of previous samples.
+\item \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities
+\item Advantage of using \Minerva decreases considerably for smaller first round stopping probability
+\item Proofs that \Prov is risk-limiting and resistant to an adversary who can pick future round sizes with knowledge of previous samples
 \end{itemize}
@@
 i.e. those that did not stop before round $j$. 
+%TODO move this more rigorous definition somewhere else
 %The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
@@
 
-We also consider a first round size with $\chi_1=0.25$.
-
 \begin{figure}
 \begin{centering}
-\includegraphics[width=0.6\textwidth]{minerva25percthen1p5_sprob.png}
+\includegraphics[width=0.5\textwidth]{minerva25percthen1p5_sprob.png}
 \caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
@@
 
-%  \begin{block}{Fusce aliquam magna velit}
-%
-%    Et rutrum ex euismod vel. Pellentesque ultricies, velit in fermentum
-%    vestibulum, lectus nisi pretium nibh, sit amet aliquam lectus augue vel
-%    velit. Suspendisse rhoncus massa porttitor augue feugiat molestie. Sed
-%    molestie ut orci nec malesuada. Sed ultricies feugiat est fringilla
-%    posuere.
-%
-%    \begin{figure}
-%      \centering
-%      \begin{tikzpicture}
-%        \begin{axis}[
-%            scale only axis,
-%            no markers,
-%            domain=0:2*pi,
-%            samples=100,
-%            axis lines=center,
-%            axis line style={-},
-%            ticks=none]
-%          \addplot[red] {sin(deg(x))};
-%          \addplot[blue] {cos(deg(x))};
-%        \end{axis}
-%      \end{tikzpicture}
-%      \caption{Another figure caption.}
-%    \end{figure}
-%
-%  \end{block}
-%
-%  \begin{block}{Nam cursus consequat egestas}
-%
-%    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
-%    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
-%    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
-%    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
-%    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
-%    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
-%    sit amet sollicitudin risus tincidunt non.
-%
-%    Etiam sit amet tempus lorem, aliquet condimentum velit. Donec et nibh
-%    consequat, sagittis ex eget, dictum orci. Etiam quis semper ante. Ut eu
-%    mauris purus. Proin nec consectetur ligula. Mauris pretium molestie
-%    ullamcorper. Integer nisi neque, aliquet et odio non, sagittis porta justo.
-%
-%    \begin{itemize}
-%      \item \textbf{Sed consequat} id ante vel efficitur. Praesent congue massa
-%        sed est scelerisque, elementum mollis augue iaculis.
-%        \begin{itemize}
-%          \item In sed est finibus, vulputate
-%            nunc gravida, pulvinar lorem. In maximus nunc dolor, sed auctor eros
-%            porttitor quis.
-%          \item Fusce ornare dignissim nisi. Nam sit amet risus vel lacus
-%            tempor tincidunt eu a arcu.
-%          \item Donec rhoncus vestibulum erat, quis aliquam leo
-%            gravida egestas.
-%        \end{itemize}
-%      \item \textbf{Sed luctus, elit sit amet} dictum maximus, diam dolor
-%        faucibus purus, sed lobortis justo erat id turpis.
-%      \item \textbf{Pellentesque facilisis dolor in leo} bibendum congue.
-%        Maecenas congue finibus justo, vitae eleifend urna facilisis at.
-%    \end{itemize}
-%    
-%    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
-%    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
-%    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
-%    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
-%    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
-%    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
-%    sit amet sollicitudin risus tincidunt non.
-%  \end{block}
-%
 \end{column}
@@
 \begin{block}{Number of Ballots}
+
 \begin{itemize}
@@
 \begin{minipage}{.49\textwidth}
-\begin{centering}
 \includegraphics[width=1\textwidth]{texas25.png}
-%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $S_1=0.25$.}
 \label{fig:texas_25}
-\end{centering}
 \end{minipage}
 \begin{minipage}{.49\textwidth}
-\begin{centering}
 \includegraphics[width=1\textwidth]{missouri25.png}
-%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Missouri, margin $0.157$, and first round stopping probability $S_1=0.25$.}
 \label{fig:missouri_25}
-\end{centering}
 \end{minipage}
@@
 \begin{minipage}{.49\textwidth}
-\begin{centering}
 \includegraphics[width=1\textwidth]{massachusetts25.png}
-%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Massachusetts, margin $0.342$, and first round stopping probability $\chi_1=0.25$.}
 \label{fig:mass_25}
-\end{centering}
 \end{minipage}
 \begin{minipage}{.49\textwidth}
-\begin{centering}
 \includegraphics[width=1\textwidth]{texas90.png}
-%\caption{This plot shows the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the state of Texas, margin $0.057$, and first round stopping probability $\chi_1=0.9$.}
 \label{fig:texas_90}
-\end{centering}
 \end{minipage}
@@
 
-%work load workload WORKLOAD
-%\begin{block}{Modeling Workload}
-%
-%\emph{I initially wrote this section since I'd really like to have some evidence that for certain workload parameters, we see an optimal
-%number of rounds greater than 1 with Providence. I am working on that code now.}
-%
-%\BRAVO requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
-%In real audits, election officials draw ballots in rounds because their is some overhead with each round (e.g. opening ballot boxes).
-%Therefore a more sophisticated workload model has a per ballot cost $c_b$ and a per round cost $c_r$. 
-%Let $n$ be the number of ballots sampled and $j$ be the number of rounds. Then the cost $C$ is given by 
-%$$C(n,j) = c_b\cdot n + c_r\cdot j,$$
-%and depending on the ratio between $c_b$ and $c_r$, we should expect that certain round schedules
-%achieve lower expected cost.
-%\end{block}
-
-%  \begin{block}{A block containing some math}
-%
-%    Nullam non est elit. In eu ornare justo. Maecenas porttitor sodales lacus,
-%    ut cursus augue sodales ac.
-%
-%    $$
-%    \int_{-\infty}^{\infty} e^{-x^2}\,dx = \sqrt{\pi}
-%    $$
-%
-%    Interdum et malesuada fames $\{1, 4, 9, \ldots\}$ ac ante ipsum primis in
-%    faucibus. Cras eleifend dolor eu nulla suscipit suscipit. Sed lobortis non
-%    felis id vulputate.
-%
-%    \heading{A heading inside a block}
-%
-%    Praesent consectetur mi $x^2 + y^2$ metus, nec vestibulum justo viverra
-%    nec. Proin eget nulla pretium, egestas magna aliquam, mollis neque. Vivamus
-%    dictum $\mathbf{u}^\intercal\mathbf{v}$ sagittis odio, vel porta erat
-%    congue sed. Maecenas ut dolor quis arcu auctor porttitor.
-%
-%    \heading{Another heading inside a block}
-%
-%    Sed augue erat, scelerisque a purus ultricies, placerat porttitor neque.
-%    Donec $P(y \mid x)$ fermentum consectetur $\nabla_x P(y \mid x)$ sapien
-%    sagittis egestas. Duis eget leo euismod nunc viverra imperdiet nec id
-%    justo.
-%
-%  \end{block}
-%
-%  \begin{block}{Nullam vel erat at velit convallis laoreet}
-%
-%%    Class aptent taciti sociosqu ad litora torquent per conubia nostra, per
-%    inceptos himenaeos. Phasellus libero enim, gravida sed erat sit amet,
-%    scelerisque congue diam. Fusce dapibus dui ut augue pulvinar iaculis.
-%%
-%    \begin{table}
-%      \centering
-%%      \begin{tabular}{l r r c}
-%        \toprule
-%%        \textbf{First column} & \textbf{Second column} & \textbf{Third column} & \textbf{Fourth} \\
-%        \midrule
-%        Foo & 13.37 & 384,394 & $\alpha$ \\
-%%        Bar & 2.17 & 1,392 & $\beta$ \\
-%        Baz & 3.14 & 83,742 & $\delta$ \\
-%        Qux & 7.59 & 974 & $\gamma$ \\
-%%        \bottomrule
-%      \end{tabular}
-%      \caption{A table caption.}
-%%    \end{table}
-%
-%    Donec quis posuere ligula. Nunc feugiat elit a mi malesuada consequat. Sed
-%%    imperdiet augue ac nibh aliquet tristique. Aenean eu tortor vulputate,
-%    eleifend lorem in, dictum urna. Proin auctor ante in augue tincidunt
-%%    tempor. Proin pellentesque vulputate odio, ac gravida nulla posuere
-%    efficitur. Aenean at velit vel dolor blandit molestie. Mauris laoreet
-%%    commodo quam, non luctus nibh ullamcorper in. Class aptent taciti sociosqu
-%    ad litora torquent per conubia nostra, per inceptos himenaeos.
-%
-%    Nulla varius finibus volutpat. Mauris molestie lorem tincidunt, iaculis
-%%    libero at, gravida ante. Phasellus at felis eu neque suscipit suscipit.
-%    Integer ullamcorper, dui nec pretium ornare, urna dolor consequat libero,
-%    in feugiat elit lorem euismod lacus. Pellentesque sit amet dolor mollis,
-%%    auctor urna non, tempus sem.
-%
-%%  \end{block}
-
-\end{column}
-
-\separatorcolumn
-
-\begin{column}{\colwidth}
 \begin{block}{Providence}
@@
 $$\omega_j(k_{j-1},k_j,n_{j-1},n_j,p_a,p_0)\triangleq \sigma(k_{j-1},n_{j-1},p_a,p_0)\cdot \tau_1(k_j,n_j,p_a,p_0) \ge \frac{1}{\alpha}.$$ 
+\end{itemize}
 
-\end{itemize}
+\end{block}
+
+\end{column}
+
+\separatorcolumn
+
+\begin{column}{\colwidth}
+
+\begin{block}{Providence}
 
@@
 
-\emph{Need to fill in simulation results here. I am thinking (a) risk simulations as evidence that \Prov is an RLA, and (b) number of ballot results if I get them done quickly enough.}
-
-Simulations are currently running to explore the risk, stopping probability, and number of ballots used by \Prov.
+\begin{figure}[h]
+\centering
+\begin{minipage}{.49\textwidth}
+\includegraphics[width=\textwidth]{prov_risk.png}
+\label{fig:bravo_risk}
+\end{minipage}
+\begin{minipage}{.49\textwidth}
+\includegraphics[width=1.0\textwidth]{prov_sprob.png}
+\label{fig:minerva_risk}
+\end{minipage}
+\caption{
+The left plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie
+where each round size is selected to achieve $\chi_j=0.9$.
+The right plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie for rounds
+$j=1,2,3$ and $\chi_j=0.9$.
+}
+\label{fig:prov-risk-sprob}
+\end{figure}
 
@@
 
-    %\nocite{*}
-    %\footnotesize{\bibliographystyle{plain}\bibliography{audits}}
-    \bibliographystyle{splncs04}
-    %\bibliographystyle{plain}
-    \bibliography{audits}
-    \emph{Having some latex trouble here, but should include r2b2, the recent simulations paper, bravo paper, athena paper, ...}
+\footnotesize{\bibliographystyle{plain}\bibliography{audits}}
 

### 2022-04-14 MultipleRoundSims/poster.tex
minor changes
full draft done except annoying bibtex error...

@@
 \usepackage{cite}
+\usepackage{hyperref}
 
@@
 \end{minipage}
-\caption{These plots show for $\chi_1=0.9$ the fraction of audits (EoR \BRAVO (left), SO \BRAVO (left), and \Minerva with multiplier $1.5$ (right)) that stopped in any of the $5$ rounds when the underlying election was a tie.
+\caption{For $\chi_1=0.9$ the fraction of audits (EoR \BRAVO (left), SO \BRAVO (left), and \Minerva with multiplier $1.5$ (right)) that stopped in any of the $5$ rounds when the underlying election was a tie.
 (For SO \BRAVO audits we show the 13 states for which simulations are complete.)}
@@
 \end{minipage}
-\caption{These plots show, for each state margin, when the underlying election is as announced, the number of audits that stopped in the $j^{th}$ round, as a fraction of all audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.
+\caption{For each state margin, when the underlying election is as announced, the number of audits that stopped in the $j^{th}$ round, as a fraction of all audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and $\chi_j=0.9$.
 The left shows EoR \BRAVO and the right shows SO \BRAVO.}
@@
 \end{minipage}
-\caption{These plots show, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.
+\caption{When the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round, as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.
 The left shows \Minerva with the round schedule obtained by $\chi_1=0.9$ and round size multiple $1.5$.}
@@
 \includegraphics[width=0.5\textwidth]{minerva25percthen1p5_sprob.png}
-\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
+\caption{For each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
 as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$, round size multiple of $1.5$ and $\chi_1 = 0.25$.}
@@
 \end{minipage}
-\caption{These plots show the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states of Texas with margin $.057$ (left) and Missouri with margin $0.157$ (right), both with first round stopping probability $\chi_1=0.25$.}
+\caption{Cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states of Texas with margin $.057$ (left) and Missouri with margin $0.157$ (right), both with first round stopping probability $\chi_1=0.25$.}
 \end{figure}
@@
 \end{minipage}
-\caption{These plots show the cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states Massachusetts (left) with margin $0.342$ and $\chi_1=0.25$ and Texas (right) with margin $0.057$ and $\chi_1=0.9$.}
+\caption{Cumulative fraction of audits that stopped as a function of average number of sampled ballots for all four audits we studied, for the states Massachusetts (left) with margin $0.342$ and $\chi_1=0.25$ and Texas (right) with margin $0.057$ and $\chi_1=0.9$.}
 \end{figure}
@@
 
+\heading {Resistance against an adversary} 
+If for each round $j=1,2,\ldots$, 
+adversary $\mathcal{A}$ has knowledge of all previous samples 
+and is allowed to choose round size $n_{j+1}$ for round $j+1$, 
+\Prov is still risk-limiting.
+
 \end{block}
@@
 
-\begin{block}{Providence}
+\begin{block}{Providence Pilot}
 
-\heading {Resistance against an adversary} 
-If adversary $\mathcal{A}$ has knowledge of previous samples and chooses round size $n_j$ in round $j$, 
-\Prov is still risk-limiting. Proof of this and the RLA property of \Prov are currently being revised for submission.
-
-\heading {\Prov pilot}
-In February 2022, The Rhode Island Board of Elections hosted a public pilot \Prov RLA which passed 
-in the first round.
+In February 2022, The Rhode Island Board of Elections hosted a public pilot \Prov RLA which 
+passed (met the risk limit) in the first round.
 
@@
 
+\begin{figure}
+\includegraphics[width=.65\textwidth]{prov_asn.png}
+\caption{Cumulative fraction of audits that stopped as a function of average number of sampled ballots for the state of Texas with margin $.057$ with first round stopping probability $\chi_1=0.9$.
+\Prov, \Minerva with multiplier $1.5$, and both implementations of \BRAVO are included.}
+\end{figure}
+
 \end{block}
@@
 
-\footnotesize{\bibliographystyle{plain}\bibliography{audits}}
+%\footnotesize{
+\bibliographystyle{plain}
+\bibliography{audits}
+%}
 
@@
 \separatorcolumn
-
 \end{columns}
-
 \end{frame}
-
 \end{document}

### 2022-04-21 MultipleRoundSims/poster.tex
a

@@
 \footercontent{
-  \href{odbroadrick@gmail.com}{odbroadrick@gmail.com}
+  odbroadrick@gmail.com
 }
@@
       \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
-      {\includegraphics[height=8cm]{logos/gw.png}}; % also try shield-white.eps
+      {\includegraphics[height=9cm]{logos/gw.png}}; % also try shield-white.eps
       \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
-      {\includegraphics[height=8cm]{logos/gw.png}};
+      {\includegraphics[height=9cm]{logos/gw.png}};
     \end{tikzpicture}
@@
 \item
-We provide simulations \cite{simulations} of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
+We provide simulations~\cite{simulations} of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
 first round stopping probability; this information can be used to advise election officials.
@@
 \item 
-Software: R2B2 library\cite{r2b2} which has Python implementations of several RLAs as well as a simulator
+Software: R2B2 library \cite{r2b2} which has Python implementations of several RLAs as well as a simulator
 \item 
@@
 \includegraphics[width=\textwidth]{prov_risk.png}
-\label{fig:bravo_risk}
+\label{fig:prov_risk}
 \end{minipage}
@@
 \includegraphics[width=1.0\textwidth]{prov_sprob.png}
-\label{fig:minerva_risk}
+\label{fig:prov_sprob}
 \end{minipage}
@@
 \Prov, \Minerva with multiplier $1.5$, and both implementations of \BRAVO are included.}
+\label{fig:prov-asn}
 \end{figure}

### 2022-04-21 MultipleRoundSims/slides.tex
a

@@
 \end{itemize}
-\item 
+\item Experiments
+\item Results
+\begin{itemize}
+\item Stopping Probability
+\item Risk
+\item Number of Ballots
+\end{itemize}
+\item Discussion and Future Work
 \end{itemize}
@@
 \begin{frame}
-%\begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
-    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
-    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
-    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
-    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
-In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
-\Minerva was recently proposed for use when the first-round stopping probability is large.
-It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
-    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
-    Our simulations use stopping probabilities of $0.9$ and $0.25$.
-    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
-    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
-    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
-    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
-    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
-    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.
+\frametitle{Post-tabulation audits}
 
-\keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
+\begin{itemize}
+\item Scanning machines are used to tabulate ballots
+\begin{itemize}
+\item Cannot trust the machines: bugs, configuration errors, hacking
+\end{itemize}
+\item Post-tabulation audits
+\item Risk-Limiting Audits
+\end{itemize}
 \end{frame}
-%\end{abstract}
-%
-%
-\section{Introduction}
-\label{sec:intro}
-\input{sections/01-intro}
 
-\section{Related work}
-\label{sec:related}
-\input{sections/02-related}
+\begin{frame}
+\frametitle{Risk-Limiting Audits}
 
-%\section{Software}
-%\label{sec:software}
-%\input{sections/03-software}
+\begin{itemize}
+\item Risk-Limiting Audit (RLA) is a post-tabulation audit that manualy
+checks a random sample of voters' ballots
+\item Relies on a voter-verified paper trail
+\item Sketch:
+\begin{enumerate}
+\item election results announced
+\item sample ballots randomly
+\item check if the sample is 'statistically similar' to the announced tally
+\begin{itemize}
+\item "Yes" - correct - stop the audit
+\item "No" - incorrect - proceed to a full hand count
+\item "Don't know yet" - undetermined - draw more samples (goto: 2)
+\end{itemize}
+\end{enumerate}
+\end{itemize}
+\end{frame}
 
-\section{Experiments}
-\label{sec:expts}
-\input{sections/04-expts}
+\begin{frame}
+\frametitle{\BRAVO}
+\begin{itemize}
+\item Most commonly used ballot polling RLA
+\item In the two candidate case is an instance of Wald's Sequential Probability Ratio Test (SPRT)
+\item Is thus the most efficient RLA when ballots are drawn sequentially (i.e. ballot-by-ballot)
+\item Real audits are performed in rounds for which \BRAVO can be implemented as:
+\begin{itemize}
+\item Selection-Ordered (SO) \BRAVO
+\item End-of-Round (EoR) \BRAVO
+\end{itemize}
+\end{itemize}
+\end{frame}
 
-\section{Stopping Probability and Risk}
-\label{sec:results1}
-\input{sections/05-results1}
+\begin{frame}
+\frametitle{\Minerva}
+\begin{itemize}
+\item Recent RLA designed for round-by-round use
+\item Uses a ratio of the \emph{tails} of the probability distributions used in \BRAVO
+\item Known to require half the number of ballots as EoR \BRAVO in a first round to achieve a large ($0.90$) probability of stopping
+\item Unknown how the audits compare for smaller stopping probability or for rounds after the first
+\end{itemize}
+\end{frame}
 
-\section{Number of Ballots}
-\label{sec:results2}
-\input{sections/06-results2}
+\begin{frame}
+\frametitle{Experiments}
+\begin{itemize}
+\item Use simulations to provide evidence for theoretical claims
+\item R2B2 software library for round-by-round and ballot-by-ballot RLAs
+\item Simulate RLAs for election results from the 2020 Presidential election (all margins above $0.05$)
+\begin{itemize}
+\item $10000=10^4$ trials assuming the underlying election is as announced
+\item $10000=10^4$ trials assuming the underlying election is a tie
+\end{itemize}
+\item Risk limit: $10\%$
+\item Round schedules:
+\begin{itemize}
+\item \BRAVO round sizes to achieve a chosen probability of stopping in each round given that the audit has already reached that round
+\item \Minerva first round sizes to achieve a chosen probability of stopping, and subsequent round sizes found by multiplying the previous round size by a constant ($1.5$ and $1$)
+\end{itemize}
+\item Stopping probabilities: $0.90$ and $0.25$
+\end{itemize}
+\end{frame}
 
-\section{Conclusions and Future Work}
-\label{sec:conc}
-\input{sections/07-conc}
+\begin{frame}
+\frametitle{Experiments}
+\begin{definition}
+An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
+(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
+\end{definition}
 
-%\section{Acknowledgements}
-%\label{sec:ack}
-%This research was supported in part by NSF Awards 2015253
-%and 1421373.
+\begin{itemize}
+\item Binary hypothesis test: $H_0$ (a tie) and $H_a$ (announced results)
+\end{itemize}
 
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-\bibliographystyle{splncs04}
-\bibliography{audits}
-%
+\begin{definition}[Risk]
+The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
+the true underlying distribution of ballots is
+$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
+\end{definition}
+
+\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+risk limit $\alpha$ iff 
+$R(\mathcal{A}) \le \alpha.$
+\end{definition}
+\end{frame}
+
+\begin{frame}
+\frametitle{Experiments}
+\begin{definition}[Stopping Probability]
+The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
+$S_j(\mathcal{A})=$
+$$\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
+\end{definition}
+
+\begin{definition}[Cumulative Stopping Probability]
+The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
+\end{definition}
+
+\begin{definition}[Conditional Stopping Probability]
+The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
+$\chi_j (\mathcal{A})=$
+$$\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
+\end{definition}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Stopping Probability ($\chi_1=0.9$)}
+
+\includegraphics[width=0.5\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}
+\includegraphics[width=0.5\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}
+
+\includegraphics[width=0.5\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
+\includegraphics[width=0.5\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Stopping Probability ($\chi_1=0.25$)}
+
+\includegraphics[width=1\textwidth]{minerva25percthen1p5_sprob.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Risk}
+
+\centering
+\noindent \includegraphics[width=1.07\textwidth]{both_risk_plots.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Number of Ballots}
+
+\includegraphics[width=0.5\textwidth]{texas25.png}
+\includegraphics[width=0.5\textwidth]{missouri25.png}
+
+\includegraphics[width=0.5\textwidth]{massachusetts25.png}
+\includegraphics[width=0.5\textwidth]{texas90.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Round Size Proportions}
+
+For $\chi_1 = 0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO
+\begin{itemize}
+\item Improvement considerably smaller than that when the stopping probability is $0.9$
+\item Number of ballots for SO \BRAVO for $\chi_1 =0.9$ is about a third more than that required by \Minerva, but for $\chi_1 = 0.25$, it requires only about a tenth more ballots than does \Minerva
+\item Number of ballots for EoR \BRAVO for $\chi_1 = 0.9$ is about twice those required by \Minerva, but for $\chi_1 =0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva
+
+\end{itemize}
+For $\chi_1 = 0.9$ with multiplying factor $1$, \Minerva consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three. 
+\begin{itemize}
+\item When the multiplying factor is $1.5$, we see $\chi_2\approx 0.91$
+and $\chi_3=0.83$
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusion}
+
+\begin{itemize}
+
+\item We describe use of the R2B2 library and simulator to characterize: 
+\begin{itemize}
+\item risk, 
+\item stopping probability, and 
+\item number of ballots 
+\item for various round schedules.
+\end{itemize}
+\item \Minerva requires fewer ballots than either implementation of \BRAVO in all cases we study, but the advantage decreases for a smaller stopping probability
+
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Future Work}
+\begin{itemize}
+\item More detailed study of the impact of different round schedules
+\item Simulations with other underlying distributions
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\centering 
+{Thank you}
+\end{frame}
+
+% References?
+
 \end{document}

### 2022-04-24 MultipleRoundSims/slides.tex
minor edits
updated slides

@@
 \begin{itemize}
-\item Risk-Limiting Audits
+\item Risk-Limiting Audits (RLAs)
 \begin{itemize}
 \item \BRAVO and \Minerva
+\pause
+\item Both are options in Arlo, statistical election audit software used by election officials across the US
 \end{itemize}
-\item Experiments
-\item Results
+\pause 
+\item Experiments: simulated $10000=10^4$ audits for various margins with both
 \begin{itemize}
-\item Stopping Probability
-\item Risk
-\item Number of Ballots
+\item a correctly announced outcome 
+\item an underlying tie
 \end{itemize}
+\pause 
+\item Observed: stopping probability, maximum risk, number of ballots
+\pause 
+\item Results:
+\begin{itemize}
+\item \Minerva requires fewer ballots over multiple rounds for both 
+\begin{itemize}
+\item high stopping probability (0.90)
+\item low stopping probability (0.25)
+\end{itemize}
+\item less advantage for the lower stopping probability
+\end{itemize}
+\pause 
 \item Discussion and Future Work
@@
 \begin{frame}
-\frametitle{Post-tabulation audits}
+\frametitle{Risk-Limiting Audits}
 
 \begin{itemize}
-\item Scanning machines are used to tabulate ballots
+\item Scanners are used to tabulate ballots
+\pause 
 \begin{itemize}
 \item Cannot trust the machines: bugs, configuration errors, hacking
+\pause 
 \end{itemize}
-\item Post-tabulation audits
+\item Compliance and tabulation audits
+\pause 
 \item Risk-Limiting Audits
+\begin{itemize}
+\item
+Given that the election outcome is incorrect,
+the probability with which the audit stops, declaring the outcome correct, is at most the risk limit, $\alpha$.
+\end{itemize}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Risk-Limiting Audits}
+\frametitle{Ballot Polling Risk-Limiting Audits: A Procedure}
 
 \begin{itemize}
-\item Risk-Limiting Audit (RLA) is a post-tabulation audit that manualy
-checks a random sample of voters' ballots
-\item Relies on a voter-verified paper trail
+\item Is a manual audit, which relies on a voter-verified paper trail and successfully completed compliance audits
+\pause 
 \item Sketch:
 \begin{enumerate}
-\item election results announced
-\item sample ballots randomly
-\item check if the sample is 'statistically similar' to the announced tally
+\item Election results announced
+\pause 
+\item In a public procedure, sample ballots at random and manually interpret them
+\pause 
+\item Compute a pre-specified error measure, the maximum risk, and compare to the risk limit
 \begin{itemize}
-\item "Yes" - correct - stop the audit
-\item "No" - incorrect - proceed to a full hand count
-\item "Don't know yet" - undetermined - draw more samples (goto: 2)
+\item If smaller, stop the audit
+\item Else, sample more (goto 2)
 \end{itemize}
@@
 \item Most commonly used ballot polling RLA
+\pause 
 \item In the two candidate case is an instance of Wald's Sequential Probability Ratio Test (SPRT)
-\item Is thus the most efficient RLA when ballots are drawn sequentially (i.e. ballot-by-ballot)
-\item Real audits are performed in rounds for which \BRAVO can be implemented as:
+\pause 
+\item Is thus the most efficient RLA when 
+the decision of whether to stop the audit is made after each ballot is drawn (ballot-by-ballot)
+\pause 
+\item In real audits, decisions are taken after many ballots are drawn (round-by-round)
+\pause 
+\item \BRAVO can be implemented as:
 \begin{itemize}
-\item Selection-Ordered (SO) \BRAVO
-\item End-of-Round (EoR) \BRAVO
+\item Selection-Ordered (SO) \BRAVO, 
+where ballot selection order is retained, and the decisions are taken as though the audit were ballot-by-ballot
+\item End-of-Round (EoR) \BRAVO, where
+the decision using the \BRAVO stopping rule is taken once, after the entire round of ballots is drawn
 \end{itemize}
@@
 \item Recent RLA designed for round-by-round use
-\item Uses a ratio of the \emph{tails} of the probability distributions used in \BRAVO
-\item Known to require half the number of ballots as EoR \BRAVO in a first round to achieve a large ($0.90$) probability of stopping
+\pause 
+\item \BRAVO uses the likelihood ratio; that is the ratio of points on two probability distribution functions
+\pause 
+\item \Minerva uses a ratio of the \emph{tails} of the pdfs used in \BRAVO
+\pause 
+\begin{itemize}
+\item Shown to be risk-limiting if all round sizes are pre-committed, before the audit begins
+\pause 
+\item In a first round chosen to give a $0.90$ probability of stopping, \Minerva requires
+\begin{itemize}
+\item $~50\%$ as many ballots as EoR \BRAVO 
+\item $~70\text{-}80\%$ as many ballots as SO \BRAVO 
+\end{itemize}
+\pause 
 \item Unknown how the audits compare for smaller stopping probability or for rounds after the first
+\end{itemize}
 \end{itemize}
@@
 \item Use simulations to provide evidence for theoretical claims
+\pause 
 \item R2B2 software library for round-by-round and ballot-by-ballot RLAs
+\pause 
 \item Simulate RLAs for election results from the 2020 Presidential election (all margins above $0.05$)
 \begin{itemize}
-\item $10000=10^4$ trials assuming the underlying election is as announced
-\item $10000=10^4$ trials assuming the underlying election is a tie
+\item $10000=10^4$ trials assuming the underlying election was as announced
+\item $10000=10^4$ trials assuming the underlying election was a tie
 \end{itemize}
+\pause 
 \item Risk limit: $10\%$
+\pause 
 \item Round schedules:
@@
 \item \BRAVO round sizes to achieve a chosen probability of stopping in each round given that the audit has already reached that round
+\pause 
 \item \Minerva first round sizes to achieve a chosen probability of stopping, and subsequent round sizes found by multiplying the previous round size by a constant ($1.5$ and $1$)
 \end{itemize}
+\pause 
 \item Stopping probabilities: $0.90$ and $0.25$
@@
 
+\pause 
 \begin{itemize}
 \item Binary hypothesis test: $H_0$ (a tie) and $H_a$ (announced results)
+\pause 
+\item The tie is the hardest incorrect outcome to detect 
+\pause 
+\item Probability of stopping given a tie should be low
+\pause 
+\item Probability of stopping given a correctly announced outcome should be high for as few ballots as possible 
 \end{itemize}
+\end{frame}
 
-\begin{definition}[Risk]
+\begin{frame}
+\frametitle{Experiments}
+\begin{definition}[Maximum Risk]
 The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
-the true underlying distribution of ballots is
+the ballots is
 $R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
@@
 
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+\pause 
+\begin{definition}[Risk-Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk-Limiting Audit with 
 risk limit $\alpha$ iff 
@@
 
+\pause 
 \begin{definition}[Cumulative Stopping Probability]
@@
 
+\pause 
 \begin{definition}[Conditional Stopping Probability]
@@
 \frametitle{Results: Stopping Probability ($\chi_1=0.9$)}
+\pause
 
-\includegraphics[width=0.5\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}
-\includegraphics[width=0.5\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}
+\includegraphics[width=0.5\textwidth]{scale/eor.png}
+\pause 
+\includegraphics[width=0.5\textwidth]{scale/so.png}
 
-\includegraphics[width=0.5\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
-\includegraphics[width=0.5\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
+\pause 
+\includegraphics[width=0.5\textwidth]{scale/minerva1.png}
+\pause 
+\includegraphics[width=0.5\textwidth]{scale/minerva1p5.png}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Results: Risk}
+\frametitle{Results: Risk ($\chi_1=0.9$)}
 
 \centering
-\noindent \includegraphics[width=1.07\textwidth]{both_risk_plots.png}
+\includegraphics[width=.5\textwidth]{bravo_risks.png}
+\pause 
+\includegraphics[width=.5\textwidth]{riskminerva1p5_10t4.png}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Results: Number of Ballots}
+\frametitle{Results: Number of Ballots ($\chi_1=0.9$)}
+\centering
 
-\includegraphics[width=0.5\textwidth]{texas25.png}
-\includegraphics[width=0.5\textwidth]{missouri25.png}
+\pause 
+\includegraphics[width=1\textwidth]{texas90.png}
+\end{frame}
 
-\includegraphics[width=0.5\textwidth]{massachusetts25.png}
-\includegraphics[width=0.5\textwidth]{texas90.png}
+\begin{frame}
+\frametitle{Results: Number of Ballots ($\chi_1=0.25$)}
+
+\centering
+\includegraphics[width=1\textwidth]{texas25.png}
 \end{frame}
@@
 For $\chi_1 = 0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO
+\pause 
 \begin{itemize}
-\item Improvement considerably smaller than that when the stopping probability is $0.9$
-\item Number of ballots for SO \BRAVO for $\chi_1 =0.9$ is about a third more than that required by \Minerva, but for $\chi_1 = 0.25$, it requires only about a tenth more ballots than does \Minerva
-\item Number of ballots for EoR \BRAVO for $\chi_1 = 0.9$ is about twice those required by \Minerva, but for $\chi_1 =0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva
+\item Improvement considerably smaller than that when $\chi_1=0.9$
+\pause 
+\item SO \BRAVO: 
+\begin{itemize}
+\item for $\chi_1 =0.9$ requires a third more than does \Minerva
+\item for $\chi_1 = 0.25$ requires a tenth more than does \Minerva
+\end{itemize}
+\pause 
+\item EoR \BRAVO:
+\begin{itemize}
+\item for $\chi_1 = 0.9$ requires twice as many as \Minerva
+\item for $\chi_1 =0.25$ requires a fourth to a half more (depending on margin) than does \Minerva
+\end{itemize}
 
 \end{itemize}
-For $\chi_1 = 0.9$ with multiplying factor $1$, \Minerva consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three. 
+\end{frame}
+\begin{frame}
+\frametitle{Results: \Minerva Stopping Probabilities}
+
+For $\chi_1 = 0.9$, \Minerva consequent conditional stopping probabilities for rounds two and three are respectively:
 \begin{itemize} 
-\item When the multiplying factor is $1.5$, we see $\chi_2\approx 0.91$
-and $\chi_3=0.83$
+\pause
+\item with multiplying factor $1$, $\chi_2\approx 0.75$ and $\chi_3\approx 0.74$
+\pause
+\item with multiplying factor $1.5$, $\chi_2\approx 0.91$ and $\chi_3\approx 0.83$
 \end{itemize}
@@
 \begin{itemize}
-\item risk, 
+\pause
+\item maximum risk, 
+\pause
 \item stopping probability, and 
+\pause
 \item number of ballots 
+\pause
 \item for various round schedules.
+\pause
 \end{itemize}
-\item \Minerva requires fewer ballots than either implementation of \BRAVO in all cases we study, but the advantage decreases for a smaller stopping probability
+\item \Minerva requires fewer ballots than either implementation of \BRAVO in all cases we study, but the advantage decreases for a smaller stopping probability for each round
 
@@
 \item More detailed study of the impact of different round schedules
+\pause
 \item Simulations with other underlying distributions
@@
 \centering 
-{Thank you}
+\bigskip
+\bigskip
+\bigskip
+Thank you
+
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+
+odbroadrick@gmail.com
 \end{frame}

### 2022-05-05 multiple-round-sims/sections/02-related.tex
new paper

--- 2022-01-11 MultipleRoundSims/sections/02-related.tex
@@
 The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken.
-Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing and searching for individual ballots).
+Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing storage boxes and searching for individual ballots).
 It is possible to apply \BRAVO to the sequence of ballots in a round if the sequential order is retained.
@@
 The risk limiting property of \BRAVO follows from the similar property of the SPRT.
-Stopping probabilities for \BRAVO may be estimated as implemented in \cite{arlo}.
-A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena} uses a similar technique for narrow margins and a separate algorithm for wider margins.
-The results of these computations match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
+Stopping probabilities for \BRAVO may be estimated as implemented in \cite{arlo}; this method is due to Mark Lindeman and uses quadratic approximations.
+A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena} uses a similar technique for narrow margins and a separate algorithm for wider margins, the results of which match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
 

### 2022-05-05 providence/draft.tex
new paper

--- 2022-01-16 MultipleRoundSims/draft.tex
@@
-% This is a draft of the coming paper "Multiple Round Ballot Polling Risk-Limiting Audit Simulations".
+% This is a draft of the coming paper "Providence: A Flexible Round-by-Round Risk-Limiting Audit"
 % 2021.
@@
 \newcommand{\Minerva}{\textsc{Minerva}\xspace}
+\newcommand{\Providence}{\textsc{Providence}\xspace}
 \newcommand{\B}{{{B2}}\xspace}
@@
 %
-\author{Oliver Broadrick\inst{1} \and
-Sarah Morin\inst{1} \and
-Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
-Filip Zag{\'o}rski\inst{3}\inst{4}}
+%\author{Oliver Broadrick\inst{1} \and
+%Sarah Morin\inst{1} \and
+%Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
+%Filip Zag{\'o}rski\inst{3}\inst{4}}
 %
-\authorrunning{Broadrick, Morin et al.}
+%\authorrunning{Broadrick, Morin et al.}
 % First names are abbreviated in the running head.
@@
 \institute{Department of Computer Science, The George Washington University\thanks{Supported in part by NSF Award 2015253} 
-\email{obroadrick@gwmail.gwu.edu}\\
+\email{odbroadrick@gmail.com}\\
 \and
@@
 \begin{abstract}
-    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.
-    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
-    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
-    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
-    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced.
-In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes.
-\Minerva was recently proposed for use when the first-round stopping probability is large.
-It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.
-    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
-    Our simulations use stopping probabilities of $0.9$ and $0.25$.
-    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
-    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
-    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
-    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
-    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
-    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.
-
+A Risk-Limiting Audit (RLA) draws paper election ballots in subsequent samples called rounds. 
+After each round, a rigorous error criterion decides whether to stop (confirming the announced outcome) or continue (draw another round).
+The most commonly used RLA, \BRAVO (Stark), requires the fewest ballots when round sizes are 1. 
+Recent RLA \Minerva requires fewer ballots than \BRAVO but requires round sizes to be fixed before the audit begins.
+We present \Providence, and audit that has similar efficiency to \Minerva and supports choosing round sizes during the audit.
+We prove \Providence is Risk-Limiting and resistant to an adversary who can choose subsequent round sizes given previous samples.
+We present simulations in which \Providence has similar efficiency to \Minerva.
+Finally, we describe the pilot use of \Providence by the Rhode Island Board of Elections.
+Our implementation of \Providence in the R2B2 software library provides standard functionality 
+plus a function which gives the probability with which an announced loser will appear to be winning
+after a round, a potentially useful metric to election officials when choosing round sizes.
+% and integrated into Arlo, election audit software used across the US.
 \keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
@@
 \label{sec:intro}
-\input{sections/01-intro}
+\input{sections/intro}
 
@@
 \label{sec:related}
-\input{sections/02-related}
+\input{sections/related}
 
-%\section{Software}
-%\label{sec:software}
-%\input{sections/03-software}
+\section{\Providence}
+\label{sec:prov}
+\input{sections/prov}
 
-\section{Experiments}
-\label{sec:expts}
-\input{sections/04-expts}
+\section{Simulations}
+\label{sec:sims}
+\input{sections/sims}
 
-\section{Stopping Probability and Risk}
-\label{sec:results1}
-\input{sections/05-results1}
+\section{Pilot}
+\label{sec:pilot}
+\input{sections/pilot}
 
-\section{Number of Ballots}
-\label{sec:results2}
-\input{sections/06-results2}
-
-\section{Conclusions and Future Work}
+\section{Conclusion}
 \label{sec:conc}
-\input{sections/07-conc}
+\input{sections/conc}
 

### 2022-05-05 providence/sections/intro.tex
new paper

--- 2022-01-13 MultipleRoundSims/sections/01-intro.tex
@@
 
-This paper provides insight into the main approaches to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
-While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
-This paper examines the number of ballots drawn over multiple rounds of both audits, for two chosen probabilities of stopping (one high: 90\%; the other low: 25\%) if the election is as announced.
-
-\subsection{Background}
 This paper focuses on ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
@@
 Thus it does not require the tracking of ballots that SO \BRAVO does.
-Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit\footnote{Their proof assumes that the number of ballots drawn in each round is know beforehand} and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+%\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand.
+In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn.
+Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general.
+(We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
 They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
-
-There are no results, either theoretical or based on simulations, regarding the number of ballots drawn over multiple rounds in a \Minerva audit with a pre-determined schedule.
-Because \BRAVO does not need to work on a pre-determined round schedule, it can optimize the size of the next round based on the sample drawn so far.
-Thus an open question is whether the constraint of a predetermined round schedule limits the efficacy of \Minerva in future rounds, and there is no literature comparing the number of ballots drawn by \Minerva and SO \BRAVO over multiple rounds.
-Note that the Average Sample Number (ASN) computations for \BRAVO \cite{bravo} apply only for B2 audits and are especially misleading as estimates of the number of ballots drawn over multiple rounds when first round sizes are large.
+Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability.
 
@@
 \subsection{Our Contributions}
-For a two candidate plurality contest with a risk limit of $10\%$, we show the following
+We present \Providence, and provide the following:
 \begin{enumerate}
-\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO.
-However, the improvement is considerably smaller than that when the stopping probability is $0.9$.
-\begin{itemize}
-\item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva.
-On the other hand, for a first round stopping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.
-\item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva.
-On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
-\end{itemize}
-\item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three.
-When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively.
-Both our simulator and the code estimating probabilities and round sizes are flexible enough to enable the study of various predetermined round schedules.
+\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
+\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
+\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
+\item Open source implementation of \Providence including the novel metric Probability of Misleading(name?)
 \end{enumerate}
 
-\subsection{Organization} Section \ref{sec:related} describes related work.
-The experiments we performed are described in section \ref{sec:expts} and sections \ref{sec:results1} and \ref{sec:results2} present our results.
+\subsection{Organization} 
+Section \ref{sec:related} describes related work.
+The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results.
 Section \ref{sec:conc} has our conclusions.

### 2022-05-05 providence/sections/pilot.tex
new paper

% pilot
A pilot audit was performed in Providence, Rhode Island in February 2022 from November 2021 special elections.
The audited contest was a question on School Construction and Renovation Projects which had an announced $25.67\%$ margin.
The risk-limit was $10\%$.
A large first round size of $140$ with probability of stopping $95\%$ was selected in order to give many ballots for more interesting analysis afterwards.

### 2022-05-05 providence/sections/prov.tex
new paper

\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
    \label{def:minervatwo}
    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
$$
\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
        % & \\
        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
        % & \\
        Undetermined ~~else \\
    \end{array}
    \right .
$$
where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
\begin{equation}
    \begin{aligned}
    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})\triangleq
    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
    \end{aligned}
\end{equation}
\end{definition}
% old definition of omega_j for j\ge 2

Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution.

\section{Proofs: Minerva 2.0 is Risk Limiting}
\begin{lemma}
    \label{lemma:sigma_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\sigma(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
From $0<p_0< p_a< 1$, we get
$$
\frac{p_a}{p_0}>1
$$
and
$$
1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
$$
and thus
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
$$
Now simply observe that
$$
\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
=\sigma(k+1, p_a, p_0, n).
$$
\end{proof}

\begin{lemma}
    \label{lemma:frac_sums_increasing}
    Given a monotone increasing sequence: $\frac{a_1}{b_1}, \frac{a_2}{b_2}, \ldots, \frac{a_n}{b_n}$, for $a_i, b_i > 0$, the sequence:
    $$z_i = \frac{\sum_{j=i}^n a_j}{\sum_{j=i}^n b_j}$$
    is also monotone increasing.
\end{lemma}

\begin{proof}
    Note that $z_i$ is a weighted average of the values of $\frac{a_j}{b_j}$ for $j \geq i$:
    $$z_i = \sum_{j=i}^n y_j \frac{a_j}{b_j}$$  ~for~ $$y_j = \frac{b_j}{\sum _{j=i}^n b_j} > 0.$$
    Further, $\sum _{j=i}^n y_j = 1$ and hence  $y_j \leq 1$ ~and~ $y_j = 1 \Leftrightarrow i=j=n$.
    Observe that, because $\frac{a_i}{b_i}$ is monotone increasing,
    $z_i \geq \frac{a_i}{b_i}$
    with equality if and only if $i=n$.
    Suppose $i < n$. Then
    $$z_{i+1} \geq \frac{a_{i+1}}{b_{i+1}} > \frac{a_i}{b_i}, $$
    and
    $$z_i = y_i \frac{a_i}{b_i} + (1-y_i) z_{i+1} < z_{i+1}.$$
    Thus $z_i$ is also monotone increasing.
\end{proof}

\begin{lemma}
    \label{lemma:tau1_increasing}
For $0<p_0< p_a< 1$ and $n>0$, the ratio $\tau_1(k,p_a,p_0,n)$ is strictly increasing as a function $k$ for $0\le k\le n$.
\end{lemma}
\begin{proof}
    Apply Lemmas \ref{lemma:sigma_increasing}-\ref{lemma:frac_sums_increasing}.
\end{proof}
\begin{lemma}
    \label{lemma:imin-exists}
    Given a strictly monotone increasing sequence: $x_1, x_2, \ldots x_n $ and some constant $A$,
    %there exists an $i_{min}$ such that
    $$A \le x_i \Leftrightarrow \exists i_{min} \le i ~\text{s.t.}~   x_{i_{min} -1} < A \le x_{i_{min}} \le x_{i},$$
    unless $A\le x_1$, in which case $i_{min} =1 $.
\end{lemma}
\begin{proof}
    Evident.
\end{proof}

\begin{lemma}
    \label{lemma:minerva2_kmin_exists}
    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a $k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
\end{lemma}
\begin{proof}
    From Definition~\ref{def:minervatwo}, $$\mathcal{A}(X_j)=\text{Correct}\iff \omega_j(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \ge \frac{1}{\alpha}.$$
    Now to apply Lemma~\ref{lemma:imin-exists}, it suffices to show that
    $\omega_j$ is monotone increasing with respect to $k_j$.
    For $j=1$, we have $\omega_1=\tau_1$, so $\omega_1$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. For $j\ge 2$,
    $$\omega_j(k_j,k_{j-1},p_a,p_0,n_j,n_{j-1},\alpha)=\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1}).$$
    As a function of $k_j$, $\sigma$ is constant, and thus $\omega$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. Therefore by Lemma \ref{lemma:imin-exists}, we have the desired property.
\end{proof}

%Here we introduce the notion of the history of an audit.
%In round $j+1$, all previous round sizes and cumulative
%tallies of winner ballots are together relevant information to the new round.
%We call this information the history of the audit through round $j$,
%and we denote it as follows:
%$$\bm{k}_j\triangleq(k_1,k_2,\ldots,k_j,n_1,n_2,\ldots,n_j)$$
\begin{lemma}
\label{lemma:any_ratio_is_sigma_simple}
For $j\ge 1$,
$$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
%$$\frac{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_a]}{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
\end{lemma}
\begin{proof}
We induct on the number of rounds.
For $j=1$, we have
$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
%
% Suppose the lemma is true for $j=m$.
% That is,
% $$\frac{Pr[\bm{K_m}=\bm{k_m} \mid H_a]}{Pr[\bm{K_m}=\bm{k_m} \mid  H_0]} = \sigma(k_m, p_a, p_0, n_m).$$
%
Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
%We will show it is true for $j=m+1$.
Observe that
 $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
 $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$
 by the induction hypothesis.
Then this is simply equal to
 $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
 $$$$
 =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
 \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
 $$
 $$
 =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
 $$
\end{proof}

\begin{theorem}
\label{thm:minerva2_is_rla_new}
An $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence audit is an
$\alpha$-RLA.
\end{theorem}
\begin{proof}
Let $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence.
Let $\bm{n_j}$ be the cumulative roundsizes used in this
audit, with corresponding cumulative tallies of
ballots for the reported winner $\bm{k_j}$.
For round $j=1$, by Definitions \ref{def:minervatwo}
and \ref{def:minerva_ratio}, we see that
the $\mathcal{A}=\text{Correct}$ (the audit stops) only when
$$
\tau_1(k_{1},p_a,p_0,n_1)
=\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
\ge \frac{1}{\alpha}.
$$
By Lemma \ref{lemma:minerva2_kmin_exists}, we see that this
is equivalent to the following:
$$
\frac{Pr[K_{1} \geq k_{min,1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{min, 1} \mid H_0, n_1]}
\ge \frac{1}{\alpha}.
$$

For any round $j\ge 2$, by Definition \ref{def:minervatwo}
and Lemma \ref{lemma:minerva2_kmin_exists},
$\mathcal{A}=\text{Correct}$ (the audit stops) only when
$$
\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq
\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
\ge \frac{1}{\alpha}.
$$
%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
%could do this for simpler notation^
By Lemma \ref{lemma:any_ratio_is_sigma_simple}
and Definition \ref{def:minerva_ratio}, this is equivalent to
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
we see that there exists a $k_{min, j}\le k_j$ for which
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$

%Equivalently,
%$$
%\frac{Pr[K_{j} \ge k_{min, j}\wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_a]}{Pr[K_{j}\ge k_{min, j} \wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_0]}\ge \frac{1}{\alpha}.
%$$
%Taking the sum over all possible rounds and corresponding preceding values
%of $k$, we get
Taking the sum over all possible audit histories, we get
$$
\frac{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
Finally, because the total probability of stopping the audit under
the alternative hypothesis is less than 1, we get
$$
\frac{Pr[\mathcal{A}=\text{Correct} \mid H_a]}{Pr[\mathcal{A}=\text{Correct} \mid H_0]}\ge \frac{1}{\alpha}
$$
$$
Pr[\mathcal{A}=\text{Correct} \mid H_0]
\le
Pr[\mathcal{A}=\text{Correct} \mid H_a] \cdot \alpha
\le
\alpha.
$$
\end{proof}

\subsection{Resistance against an adversary choosing round sizes}

### 2022-05-05 providence/sections/related.tex
new paper

--- 2022-05-05 multiple-round-sims/sections/02-related.tex
@@
 While total ballots sampled can give naive workload estimates\cite{bernoulli-ballot-polling}, Bernhard presents a more complex workload estimation model\cite{bernhard-diss}.
+
+\subsection{Model}
+
+\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+risk limit $\alpha$ iff for sample $X$
+$$
+Pr[\mathcal{A}(X) 
+= \text{Correct} \,|\, H_0]\le \alpha
+$$
+\end{definition}
+
+\begin{definition}[\Bravo Ratio] \label{def:bravo-ratio} The \Bravo audit uses the ratio $\sigma$.
+Consider a sample size of $n$ ballots with $k$ for the reported winner.
+The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
+\begin{equation}
+    \sigma(k, p_a, p_0, n) \triangleq \frac{p_a^{k} (1-p_a)^{n-k}}{p_0^{k} (1-p_0)^{n-k}} 
+    \label{eqn:bravoratio}
+\end{equation}
+\end{definition}
+
+In \Bravo, $p_0=\frac{1}{2}$. If testing the \Bravo stopping condition after each individual ballot
+is drawn (a \B \Bravo audit), clearly $\sigma$ is equivalent to the 
+likelihood ratio:
+$$
+\frac{Pr[K=k|H_a,n]}{Pr[K=k|H_0,n]}= \frac{\binom{n}{k}p_a^{k} (1-p_a)^{n-k}}{\binom{n}{k}(\frac{1}{2})^n} =\sigma(k, p_a, \frac{1}{2}, n)
+$$
+
+\begin{definition}[$(\alpha,p)$-\Bravo]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\Bravo audit iff the following stopping condition is tested at each ballot draw.
+If the sample $X$ is of size $n$ and has $k$ ballots for the winner,
+\begin{equation}
+    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \text{Correct} & ~\sigma(k, p, \frac{1}{2}, n) 
+         %\triangleq \frac{p^{k} (1-p)^{n-k}}{(\frac{1}{2})^n} 
+        \geq \frac{1}{\alpha}\\
+        Undetermined & ~else 
+    \end{array}
+    \right .
+    \label{eqn:bravo}
+\end{equation}
+\end{definition}
+
+It becomes useful to have shorthand for a sequence of round sizes and a sequence
+of winner ballot tallies.
+We use:
+$$\bm{k_j}\triangleq(k_1,k_2,\ldots,k_j)$$
+$$\bm{n_j}\triangleq(n_1,n_2,\ldots,n_j)$$
+
+\begin{definition}[\Minerva Ratio] \label{def:minerva_ratio} The \R \Minerva audit uses the ratio $\tau_j$.
+We use cumulative round sizes $\bm{n_j}$, with corresponding $\bm{k_j}$ ballots for the reported winner in reach round.
+The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
+         \begin{equation}
+             \label{eqn:tau}
+                 \tau_{j}(k_{j}, p_a,p_0, \bm{n_j}, \alpha )  \triangleq
+                 \frac{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_a, \bm{n_j}]}{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_0, \bm{n_j}]}
+         \end{equation}
+\end{definition}
+
+\begin{definition}[$ (\alpha, p, \bm{n_j} ) $-\Minerva]
+     \label{def:minerva}
+     Given \B $(\alpha, p)$-\Bravo and cumulative round sizes\\ $\bm{n_j}$, the corresponding \R \Minerva stopping rule for the $j^{th}$ round is:
+ \begin{equation}
+     \mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \tau_{j}(k_{j}, p_a, \frac{1}{2}, \bm{n_j}, \alpha ) \geq \frac{1}{\alpha}\\
+             % & \\
+             % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
+             % & \\
+             Undetermined ~~else \\
+         \end{array}
+         \right .
+         \label{eqn:minerva-test}
+ \end{equation}
+\end{definition}

### 2022-05-05 providence/sections/sims.tex
new paper

% simulations
We run simulations assuming announced outcomes are correct to give insight into stopping probability and number of ballots drawn.

### 2022-05-11 providence/sections/related.tex
prog

@@
 
-\begin{definition}[\Bravo Ratio] \label{def:bravo-ratio} The \Bravo audit uses the ratio $\sigma$.
+\begin{definition}[\BRAVO Ratio] \label{def:bravo-ratio} The \BRAVO audit uses the ratio $\sigma$.
 Consider a sample size of $n$ ballots with $k$ for the reported winner.
@@
 
-In \Bravo, $p_0=\frac{1}{2}$. If testing the \Bravo stopping condition after each individual ballot
-is drawn (a \B \Bravo audit), clearly $\sigma$ is equivalent to the 
+In \BRAVO, $p_0=\frac{1}{2}$. If testing the \BRAVO stopping condition after each individual ballot
+is drawn (a \B \BRAVO audit), clearly $\sigma$ is equivalent to the 
 likelihood ratio:
@@
 
-\begin{definition}[$(\alpha,p)$-\Bravo]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\Bravo audit iff the following stopping condition is tested at each ballot draw.
+\begin{definition}[$(\alpha,p)$-\BRAVO ]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\BRAVO audit iff the following stopping condition is tested at each ballot draw.
 If the sample $X$ is of size $n$ and has $k$ ballots for the winner,
@@
      \label{def:minerva}
-     Given \B $(\alpha, p)$-\Bravo and cumulative round sizes\\ $\bm{n_j}$, the corresponding \R \Minerva stopping rule for the $j^{th}$ round is:
+     Given \B $(\alpha, p)$-\BRAVO and cumulative round sizes\\ $\bm{n_j}$, the corresponding \R \Minerva stopping rule for the $j^{th}$ round is:
  \begin{equation}

### 2022-05-11 providence/sections/sims.tex
...

@@
 % simulations
-We run simulations assuming announced outcomes are correct to give insight into stopping probability and number of ballots drawn.
+We use simulations to provide additional evidence for theoretical claims and provide insight into audit behavior.
+As in \cite{sims}, we use margins from the 2020 US Presidential election, state-wide pairwise margins between the leading two candidates of 5\% or more.
+Narrower margins are computationally expensive, especially for the simulations with an underyling tie which quickly increase in sample size.
+We use the simulator in the R2B2 software library\cite{r2b2}.
+We perform $10000=10^4$ trials per margin for both an underyling announced outcomes and an underyling tie.
+
+The simulations with an underlying tie give an estimate of maximum risk as shown in Figure~\ref{fig:prov-risk}. For all margins, the estimated risks were less than the risk limit.
+
+Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn.
+Figure~\ref{prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
+Figure~\ref{prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
+Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
+As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
+Note that a difference in twice as many ballots could be negligible in a contest with a wide margin.
+In a contest with a narrow margin (in the 2020 US Presidential election, several states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work.
+Section~\ref{sec:workload} discusses workload in more depth.
+%TODO add a more specific number of states for small margin
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{prov_risk.png}
+\caption{The proportion of simulated audits with an underlying tie that stopped in one of the five simulated rounds.}
+\label{fig:prov-risk}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{prov_sprob.png}
+\caption{The proportion of simulated audits with an margin as announced that stopped in one of the five simulated rounds.}
+\label{fig:prov-sprob}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{prov_asn.png}
+\caption{For all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.}
+\label{fig:prov-asn}
+\end{figure}

### 2022-05-11 providence/sections/workload.tex
...

%TODO
This section could be great, but I'll have to write it first.

### 2022-05-12 providence/sections/pilot.tex
pilot table

@@
 % pilot
-A pilot audit was performed in Providence, Rhode Island in February 2022 from November 2021 special elections.
-The audited contest was a question on School Construction and Renovation Projects which had an announced $25.67\%$ margin.
+A pilot audit was performed in Providence, Rhode Island in February 2022 of November 2021 special elections.
+The audited contest was a yes-or-no question on School Construction and Renovation Projects and had an announced $25.67\%$ margin.
 The risk-limit was $10\%$.
-A large first round size of $140$ with probability of stopping $95\%$ was selected in order to give many ballots for more interesting analysis afterwards.
+A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected, and selection order was tracked, in order to give the potential for more interesting analysis afterwards. 
+As expected, the audit concluded in the first round with a \Providence risk of $4.18\%$.
+Table~\ref{fig:pilot-risks} shows risk measures for the drawn sample using \Minerva and \BRAVO (both EoR and SO).
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| } 
+\hline
+%\diagbox[dir=NW]{First \\Round \\Size}{RLA}
+ballots& \rotatebox{45}{\Providence} & \rotatebox{45}{\Minerva} & \rotatebox{45}{EoR \BRAVO} & \rotatebox{45}{SO \BRAVO} \\
+\hline
+140 & \bf{4.18\%} & \bf{4.18\%} & \bf{5.41\%} & 36.6\% \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Risk measures for the drawn first round of $140$ ballots in the Providence, RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
+\label{fig:pilot-risks}
+\end{figure}
+
+TODO: Add examples of how the audits perform for various hypothetical round schedules.
+I wait to do this until I'm done with the workload estimates since the examples here should be chosen to motivate that section.

### 2022-05-12 providence/sections/related.tex
pilot table

@@
 In \BRAVO, $p_0=\frac{1}{2}$. If testing the \BRAVO stopping condition after each individual ballot
-is drawn (a \B \BRAVO audit), clearly $\sigma$ is equivalent to the 
+is drawn (a \B \BRAVO audit), $\sigma$ is equivalent to the 
 likelihood ratio:

### 2022-05-16 providence/sections/appendix.tex
prog

--- 2022-05-05 providence/sections/prov.tex
@@
-\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
-    \label{def:minervatwo}
-    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
-$$
-\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
-        % & \\
-        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
-        % & \\
-        Undetermined ~~else \\
-    \end{array}
-    \right .
-$$
-where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
-\begin{equation}
-    \begin{aligned}
-    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})\triangleq
-    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
-    \end{aligned}
-\end{equation}
-\end{definition}
-% old definition of omega_j for j\ge 2
-
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution.
-
-\section{Proofs: Minerva 2.0 is Risk Limiting}
 \begin{lemma}
@@
 
-%Here we introduce the notion of the history of an audit.
-%In round $j+1$, all previous round sizes and cumulative
-%tallies of winner ballots are together relevant information to the new round.
-%We call this information the history of the audit through round $j$,
-%and we denote it as follows:
-%$$\bm{k}_j\triangleq(k_1,k_2,\ldots,k_j,n_1,n_2,\ldots,n_j)$$
 \begin{lemma}
@@
 $$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
-%$$\frac{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_a]}{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
 \end{lemma}
@@
 $$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
-%
-% Suppose the lemma is true for $j=m$.
-% That is,
-% $$\frac{Pr[\bm{K_m}=\bm{k_m} \mid H_a]}{Pr[\bm{K_m}=\bm{k_m} \mid  H_0]} = \sigma(k_m, p_a, p_0, n_m).$$
-%
 Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
-%We will show it is true for $j=m+1$.
 Observe that
@@
 \end{proof}
-
-\begin{theorem}
-\label{thm:minerva2_is_rla_new}
-An $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence audit is an
-$\alpha$-RLA.
-\end{theorem}
-\begin{proof}
-Let $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence.
-Let $\bm{n_j}$ be the cumulative roundsizes used in this
-audit, with corresponding cumulative tallies of
-ballots for the reported winner $\bm{k_j}$.
-For round $j=1$, by Definitions \ref{def:minervatwo}
-and \ref{def:minerva_ratio}, we see that
-the $\mathcal{A}=\text{Correct}$ (the audit stops) only when
-$$
-\tau_1(k_{1},p_a,p_0,n_1)
-=\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
-\ge \frac{1}{\alpha}.
-$$
-By Lemma \ref{lemma:minerva2_kmin_exists}, we see that this
-is equivalent to the following:
-$$
-\frac{Pr[K_{1} \geq k_{min,1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{min, 1} \mid H_0, n_1]}
-\ge \frac{1}{\alpha}.
-$$
-
-For any round $j\ge 2$, by Definition \ref{def:minervatwo}
-and Lemma \ref{lemma:minerva2_kmin_exists},
-$\mathcal{A}=\text{Correct}$ (the audit stops) only when
-$$
-\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq
-\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
-\ge \frac{1}{\alpha}.
-$$
-%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
-%could do this for simpler notation^
-By Lemma \ref{lemma:any_ratio_is_sigma_simple}
-and Definition \ref{def:minerva_ratio}, this is equivalent to
-$$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
-$$
-By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
-we see that there exists a $k_{min, j}\le k_j$ for which
-$$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
-$$
-
-%Equivalently,
-%$$
-%\frac{Pr[K_{j} \ge k_{min, j}\wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_a]}{Pr[K_{j}\ge k_{min, j} \wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_0]}\ge \frac{1}{\alpha}.
-%$$
-%Taking the sum over all possible rounds and corresponding preceding values
-%of $k$, we get
-Taking the sum over all possible audit histories, we get
-$$
-\frac{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
-$$
-Finally, because the total probability of stopping the audit under
-the alternative hypothesis is less than 1, we get
-$$
-\frac{Pr[\mathcal{A}=\text{Correct} \mid H_a]}{Pr[\mathcal{A}=\text{Correct} \mid H_0]}\ge \frac{1}{\alpha}
-$$
-$$
-Pr[\mathcal{A}=\text{Correct} \mid H_0]
-\le
-Pr[\mathcal{A}=\text{Correct} \mid H_a] \cdot \alpha
-\le
-\alpha.
-$$
-\end{proof}
-
-\subsection{Resistance against an adversary choosing round sizes}

### 2022-05-16 providence/sections/intro.tex
prog

@@
 
-Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, are available for use in real election audits.
-Both have been used in real election audits\cite{usenix_minerva, virginia_rla}.
-For this reason, it is important to understand their properties over multiple rounds.
+While more efficient than \BRAVO, \Minerva requires that the round schedule is fixed in advance of the audit.
+This may lead to significant unnecessary work.
+For example, suppose the fixed round schedule for \Minerva is $10,000$ ballots per round (a reasonable number in a state-wide contest with a narrow margin), and the first round sample finds a number of ballots for the winner just short of that necessary for the audit to stop.
+It then may be sufficient to draw a very small second round and still stop with high probability in the second round, but the additional $10,000$ ballots must be drawn.
+In contrast, subsequent \BRAVO round sizes can be chosen based on preceding samples.
+
+An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.
 
@@
 \item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
-\item Open source implementation of \Providence including the novel metric Probability of Misleading(name?)
+\item Open source implementation of \Providence 
+%including the novel metric Probability of Misleading(name?)
 \end{enumerate}
 
-\subsection{Organization} 
-Section \ref{sec:related} describes related work.
+%\subsection{Organization} 
+%Section \ref{sec:related} describes related work.
 The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results.

### 2022-05-16 providence/sections/prov.tex
prog

@@
 
-\section{Proofs: Minerva 2.0 is Risk Limiting}
-\begin{lemma}
-    \label{lemma:sigma_increasing}
-For $0<p_0< p_a< 1$ and $n>0$, the ratio $\sigma(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
-\end{lemma}
-\begin{proof}
-From $0<p_0< p_a< 1$, we get
-$$
-\frac{p_a}{p_0}>1
-$$
-and
-$$
-1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
-$$
-and thus
-$$
-\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
-$$
-Now simply observe that
-$$
-\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
-=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
-=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
-=\sigma(k+1, p_a, p_0, n).
-$$
-\end{proof}
-
-\begin{lemma}
-    \label{lemma:frac_sums_increasing}
-    Given a monotone increasing sequence: $\frac{a_1}{b_1}, \frac{a_2}{b_2}, \ldots, \frac{a_n}{b_n}$, for $a_i, b_i > 0$, the sequence:
-    $$z_i = \frac{\sum_{j=i}^n a_j}{\sum_{j=i}^n b_j}$$
-    is also monotone increasing.
-\end{lemma}
-
-\begin{proof}
-    Note that $z_i$ is a weighted average of the values of $\frac{a_j}{b_j}$ for $j \geq i$:
-    $$z_i = \sum_{j=i}^n y_j \frac{a_j}{b_j}$$  ~for~ $$y_j = \frac{b_j}{\sum _{j=i}^n b_j} > 0.$$
-    Further, $\sum _{j=i}^n y_j = 1$ and hence  $y_j \leq 1$ ~and~ $y_j = 1 \Leftrightarrow i=j=n$.
-    Observe that, because $\frac{a_i}{b_i}$ is monotone increasing,
-    $z_i \geq \frac{a_i}{b_i}$
-    with equality if and only if $i=n$.
-    Suppose $i < n$. Then
-    $$z_{i+1} \geq \frac{a_{i+1}}{b_{i+1}} > \frac{a_i}{b_i}, $$
-    and
-    $$z_i = y_i \frac{a_i}{b_i} + (1-y_i) z_{i+1} < z_{i+1}.$$
-    Thus $z_i$ is also monotone increasing.
-\end{proof}
-
-\begin{lemma}
-    \label{lemma:tau1_increasing}
-For $0<p_0< p_a< 1$ and $n>0$, the ratio $\tau_1(k,p_a,p_0,n)$ is strictly increasing as a function $k$ for $0\le k\le n$.
-\end{lemma}
-\begin{proof}
-    Apply Lemmas \ref{lemma:sigma_increasing}-\ref{lemma:frac_sums_increasing}.
-\end{proof}
-\begin{lemma}
-    \label{lemma:imin-exists}
-    Given a strictly monotone increasing sequence: $x_1, x_2, \ldots x_n $ and some constant $A$,
-    %there exists an $i_{min}$ such that
-    $$A \le x_i \Leftrightarrow \exists i_{min} \le i ~\text{s.t.}~   x_{i_{min} -1} < A \le x_{i_{min}} \le x_{i},$$
-    unless $A\le x_1$, in which case $i_{min} =1 $.
-\end{lemma}
-\begin{proof}
-    Evident.
-\end{proof}
-
-\begin{lemma}
-    \label{lemma:minerva2_kmin_exists}
-    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a $k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
-\end{lemma}
-\begin{proof}
-    From Definition~\ref{def:minervatwo}, $$\mathcal{A}(X_j)=\text{Correct}\iff \omega_j(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \ge \frac{1}{\alpha}.$$
-    Now to apply Lemma~\ref{lemma:imin-exists}, it suffices to show that
-    $\omega_j$ is monotone increasing with respect to $k_j$.
-    For $j=1$, we have $\omega_1=\tau_1$, so $\omega_1$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. For $j\ge 2$,
-    $$\omega_j(k_j,k_{j-1},p_a,p_0,n_j,n_{j-1},\alpha)=\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1}).$$
-    As a function of $k_j$, $\sigma$ is constant, and thus $\omega$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}.
-Therefore by Lemma \ref{lemma:imin-exists}, we have the desired property.
-\end{proof}
-
-%Here we introduce the notion of the history of an audit.
-%In round $j+1$, all previous round sizes and cumulative
-%tallies of winner ballots are together relevant information to the new round.
-%We call this information the history of the audit through round $j$,
-%and we denote it as follows:
-%$$\bm{k}_j\triangleq(k_1,k_2,\ldots,k_j,n_1,n_2,\ldots,n_j)$$
-\begin{lemma}
-\label{lemma:any_ratio_is_sigma_simple}
-For $j\ge 1$,
-$$\frac{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_a]}{Pr[\bm{K_j}=\bm{k_j} \mid \bm{n_j}, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
-%$$\frac{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_a]}{Pr[K_j = k_{j} \mid n_{j-1}^*, k_{j-1}^*, H_0]} = \sigma(k_j, p_a, p_0, n_j).$$
-\end{lemma}
-\begin{proof}
-We induct on the number of rounds.
-For $j=1$, we have
-$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
-%
-% Suppose the lemma is true for $j=m$.
-% That is,
-% $$\frac{Pr[\bm{K_m}=\bm{k_m} \mid H_a]}{Pr[\bm{K_m}=\bm{k_m} \mid  H_0]} = \sigma(k_m, p_a, p_0, n_m).$$
-%
-Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
-%We will show it is true for $j=m+1$.
-Observe that
- $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
- $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$
- by the induction hypothesis.
-Then this is simply equal to
- $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
- $$$$
- =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
- \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
- $$
- $$
- =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
- $$
-\end{proof}
+Now we prove that \Providence is risk-limiting.
 

### 2022-05-16 providence/sections/sims.tex
prog

@@
 % simulations
-We use simulations to provide additional evidence for theoretical claims and provide insight into audit behavior.
+We use simulations to provide additional evidence for theoretical claims and gain insight into audit behavior.
 As in \cite{sims}, we use margins from the 2020 US Presidential election, state-wide pairwise margins between the leading two candidates of 5\% or more.
@@
 Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn.
-Figure~\ref{prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
-Figure~\ref{prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
+Figure~\ref{fig:prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
+Figure~\ref{fig:prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
 Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.

### 2022-05-16 providence/sections/appendix.tex
prog

@@
 =\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
+$$$$
 =\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
@@
     For $j=1$, we have $\omega_1=\tau_1$, so $\omega_1$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}. For $j\ge 2$,
-    $$\omega_j(k_j,k_{j-1},p_a,p_0,n_j,n_{j-1},\alpha)=\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1}).$$
+    $$\omega_j(k_j,k_{j-1},p_a,p_0,n_j,n_{j-1},\alpha)=$$$$\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1}).$$
     As a function of $k_j$, $\sigma$ is constant, and thus $\omega$ is strictly increasing by Lemma \ref{lemma:tau1_increasing}.
@@
 For $j=1$, we have
-$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} = \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
+$$\frac{Pr[\bm{K_1}=\bm{k_1} \mid \bm{n_1},H_a]}{Pr[\bm{K_1}=\bm{k_1} \mid  \bm{n_1},H_0]} =\frac{Pr[K_1 = k_{1} \mid n_1,H_a]}{Pr[K_1 = k_1 \mid n_1,H_0]} $$$$= \frac{\text{Bin}(k_1,n_1,p_a)}{\text{Bin}(k_1,n_1,p_0)}=\sigma(k_1, p_a, p_0, n_1).$$
 Suppose the lemma is true for round $j=m$ with history $\bm{k_m}$.
 Observe that
- $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} = \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
+ $$\frac{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}},H_a]}{Pr[\bm{K_{m+1}}=\bm{k_{m+1}} \mid \bm{n_{m+1}}, H_0]} $$$$= \frac{ Pr[\bm{K_{m}}=\bm{k_{m}}\mid \bm{n_{m+1}},H_a] \cdot Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_a]}{ Pr[\bm{K_{m}}= \bm{k_{m}} \mid  \bm{n_{m+1}},H_0]  \cdot  Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]  }$$
  $$=\sigma(k_m, p_a, p_0, n_m) \cdot \frac{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m}, \bm{n_{m+1}}, H_a]}{Pr[K_{m+1}'=k_{m+1}'|\bm{k_m},\bm{n_{m+1}},H_0]}$$

### 2022-05-16 providence/sections/pilot.tex
a table is a table, not a figure
prog

@@
 % pilot
-A pilot audit was performed in Providence, Rhode Island in February 2022 of November 2021 special elections.
-The audited contest was a yes-or-no question on School Construction and Renovation Projects and had an announced $25.67\%$ margin.
-The risk-limit was $10\%$.
+The Rhode Island Board of Elections performed a pilot audit in Providence 
+in February 2022. The contest audited was a single yes-or-no question 
+%on School Construction and Renovation Projects
+from the November 2021 special election. 
+The question had announced margin $25.67\%$.
+The audit used risk-limit was $10\%$.
 A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected, and selection order was tracked, in order to give the potential for more interesting analysis afterwards. 
 As expected, the audit concluded in the first round with a \Providence risk of $4.18\%$.
-Table~\ref{fig:pilot-risks} shows risk measures for the drawn sample using \Minerva and \BRAVO (both EoR and SO).
+Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Minerva and \BRAVO (both EoR and SO).
 
-\begin{figure}
+\begin{table}
 \begin{center}
@@
 \caption{Risk measures for the drawn first round of $140$ ballots in the Providence, RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
-\label{fig:pilot-risks}
-\end{figure}
+\label{tab:pilot-risks}
+\end{table}
 

### 2022-05-16 providence/sections/prov.tex
prog

@@
     \begin{aligned}
-    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})\triangleq
+    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})
+    \triangleq\\
     \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
@@
 \end{definition}
-% old definition of omega_j for j\ge 2
 
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution.
-
-Now we prove that \Providence is risk-limiting.
+Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution. We now we prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.
 
@@
 $$
-\tau_1(k_{1},p_a,p_0,n_1)
+\tau_1(k_{1},p_a,p_0,n_1)\\
 =\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
@@
 $$
+\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge
+$$
+$$
 \frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge 
+$$
+$$
 \frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
@@
 \subsection{Resistance against an adversary choosing round sizes}
+I don't feel very comfortable with how to word/present this argument... I have not yet made an attempt at doing so.

### 2022-05-16 providence/sections/related.tex
prog

@@
 First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability $0.9$.
+Additional simulations \cite{simulations} have shown that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
+As expected, the advantage of \Minerva decreases for smaller stopping probability (smaller round sizes) as such round schedules approach the B2 round schedule (1,1,1,\ldots) for which \BRAVO is known to be most efficient.
 
-Ballot polling audit simulations have been used to familiarize election officials and the public with the approach \cite{dice}.
-McLaughlin and Stark \cite{mclaughlin_thesis,simulations_house} compare the workload for the Canvass Audits by Sampling and Testing (CAST) and Kaplan-Markov (KM) audits using simulations.
-Blom {\em et al.} demonstrate the efficiency of their ballot polling approach to audit instant runoff voting (IRV) using simulations \cite{blom_IRV}.
-Huang {\em et al.} present a framework generalizing a number of ballot polling audits and compare their performance (round sizes and stopping probabilities) using simulations \cite{DBLP:conf/evoteid/HuangRSTV20}.
-This work was prior to the development of \Minerva, and focuses on the comparison between Bayesian audits \cite{bayesian-audits} and \BRAVO, essentially studying the impact of the prior of the Bayesian RLA.
+Ballot polling audit simulations have been used to familiarize election officials and the public with the approach \cite{dice} and to investigate the efficiency of audits \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
 Some workload measurements have been made\cite{RI-report}. 
@@
 \subsection{Model}
+An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
+\BRAVO and \Minerva are modeled as binary hypothesis tests where the null hypothesis $H_0$ corresponds to a true underlying tie and the alternative hypothesis $H_a$ corresponds to the underyling election being as announced.
+(With an even total number of ballots, $H_0$ corresponds to the announced loser winning by one ballot.)
+Thus the null hypothesis is the incorrectly announced outcome which is most difficult to detect.
+A risk-limiting audit requires that in the event of an underlying tie, the probability of failing to detect the error is below the risk limit.
 
@@
 
+The stopping conditions of \BRAVO and \Minerva rely on the following ratios.
+
 \begin{definition}[\BRAVO Ratio] \label{def:bravo-ratio} The \BRAVO audit uses the ratio $\sigma$.
@@
 
-In \BRAVO, $p_0=\frac{1}{2}$. If testing the \BRAVO stopping condition after each individual ballot
+In \BRAVO, $p_0=\frac{1}{2}$. A \BRAVO audit outputs correct if and only if
+$$\sigma(k,p_a,\frac{1}{2},n)\ge \frac{1}{\alpha}.$$
+
+If testing the \BRAVO stopping condition after each individual ballot
 is drawn (a \B \BRAVO audit), $\sigma$ is equivalent to the 
@@
 
+\begin{comment}
 \begin{definition}[$(\alpha,p)$-\BRAVO ]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\BRAVO audit iff the following stopping condition is tested at each ballot draw.
@@
 \end{definition}
+\end{comment}
 
-It becomes useful to have shorthand for a sequence of round sizes and a sequence
+It now becomes useful to have shorthand for a sequence of round sizes and a sequence
 of winner ballot tallies.
@@
 
+Where \BRAVO uses a ratio of points on conditional probability distributions, \Minerva uses a ratio of \emph{tails} of conditional probability distributions.
+
 \begin{definition}[\Minerva Ratio] \label{def:minerva_ratio} The \R \Minerva audit uses the ratio $\tau_j$.
@@
          \begin{equation}
+         \begin{aligned}
              \label{eqn:tau}
-                 \tau_{j}(k_{j}, p_a,p_0, \bm{n_j}, \alpha )  \triangleq
+                 \tau_{j}(k_{j}, p_a,p_0, \bm{n_j}, \alpha )  \triangleq\\
                  \frac{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_a, \bm{n_j}]}{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_0, \bm{n_j}]}
+         \end{aligned}
          \end{equation}
@@
 
+\begin{comment}
 \begin{definition}[$ (\alpha, p, \bm{n_j} ) $-\Minerva]
@@
 \end{definition}
+\end{comment}

### 2022-07-16 providence/sections/pilot.tex
outlining for my own understanding

@@
 The question had announced margin $25.67\%$.
-The audit used risk-limit was $10\%$.
-A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected, and selection order was tracked, in order to give the potential for more interesting analysis afterwards. 
+The audit used risk-limit $10\%$.
+A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected
+in order to give the potential for more interesting analysis afterwards; the large margin 
+made a first round with large stopping probability practical.
+Selection order was tracked for the sake of analysis.
 As expected, the audit concluded in the first round with a \Providence risk of $4.18\%$.

### 2022-07-16 providence/sections/workload.tex
outlining for my own understanding

@@
-%TODO
-This section could be great, but I'll have to write it first.
+Some election audits have benefited from a one-and-done approach: draw a large sample with high probability of stopping in the first round and usually avoid a second round altogether.
+This is appealing for two reasons.
+Firstly, rounds have some overhead in time and effort.
+Thus the workload (cost and time) of an audit grows not just with the number of ballots sampled but also with the number of rounds.
+On the other hand, a one-and-done audit often draws more ballots than are necessary.
+We propose a simple workload model under which our simulation results suggest that a few smaller rounds gives lower average workload.
+
+We begin with the simple assumption that part of the workload grows linearly with the number of ballots, giving us the per ballot cost parameter $c_{b}$.
+It is less clear how cost is related to the number of rounds, but we propose a simple model in which the round cost is also linear, with constant per round cost $c_{r}$.
+This is a reasonable assumption for contests with small margin in which all or nearly all ballot boxes need to be opened in each round.
+[@Poorvi, I need help justifying the workload model.]
+
+So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we get that the cost $C$ of the audit is
+$$C(\mathcal{A}) = E_b * c_b + E_r * c_r.$$
+
+Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) is challenging since the number of possible sequences of samples grows exponentially with the number of rounds. 
+%A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
+%We will use the more standard approach of simulations to give an example here.
+Therefore we again use the standard approach of simulations.
+
+Consider the US state of Pennsylvania. In the 2020 US Presidential contest, Pennsylvania had a margin of $99999\%$.
+
+We let the per ballot cost be one, $c_b=1$.
+Then the per round cost $c_r$ tells us the cost of a round as a number of ballots.
+We begin with an example in which $c_r=1000$.
+That is, we suppose that the overhead of a round is equal to the workload of sampling $1000$ ballots.
+
+We now consider a simple round schedule for our example.
+Each round is selected to give the same probability of stopping, $p$.
+If the audit does not stop in the first round, we find a round size which, given the sample drawn in the first round, will give a probability of stopping $p$ in the second round.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
+
+The simulation results for our example are given in Figure~\ref{fig:workload}. 
+For various round schedules, given by the varying parameter $p$ on the horizontal axis, we see that the average audit cost varies with a minimum at roughly $.xx$(TODO).
+In particular, the one-and-done $p=.9$ audits have an average of $zz\%$ higher cost than the $p=.xx$ audits.
+
+Without extensive workload measurements, it is hard to know the most accurate value for $c_r$, and so we also evaluate the simulation results with varying $c_r$.

### 2022-07-19 providence/sections/workload.tex
workload progress

@@
 So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we get that the cost $C$ of the audit is
-$$C(\mathcal{A}) = E_b * c_b + E_r * c_r.$$
+\begin{equation}
+C(\mathcal{A}) = E_b c_b + E_r c_r
+\label{eq:cost}
+\end{equation}
 
@@
 Then the per round cost $c_r$ tells us the cost of a round as a number of ballots.
-We begin with an example in which $c_r=1000$.
-That is, we suppose that the overhead of a round is equal to the workload of sampling $1000$ ballots.
+We begin with an example in which $c_r=10$.
+That is, we suppose that the overhead of a round is equal to the workload of sampling $10$ ballots.
 
@@
 For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
+For this particular round schedule scheme, computing the expected number of rounds is straightforward; .
+Our simulations are then used to estimate the expected number of ballots.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{workload.png}
+\caption{For cost parameters $c_b=1$ and $c_r=10$, this plot shows the expected cost for various round schedule parameters $p$.
+Expected cost is found using Equation~\ref{eq:cost} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
+\label{fig:workload}
+\end{figure}
 
 The simulation results for our example are given in Figure~\ref{fig:workload}. 
-For various round schedules, given by the varying parameter $p$ on the horizontal axis, we see that the average audit cost varies with a minimum at roughly $.xx$(TODO).
-In particular, the one-and-done $p=.9$ audits have an average of $zz\%$ higher cost than the $p=.xx$ audits.
+For various round schedules, given by the varying parameters $p$ on the horizontal axis, we see that the average audit cost varies with a minimum at roughly $.8$.
+In particular, the one-and-done $p=.9$ audits have an average of $zz\%$TODO higher cost than the $p=.8$ audits.
+Especially for low margin contests, using the cost-minimizing round schedule could translate to significant reduction in cost or savings in time, a particularly important factor given tight certification deadlines.
 
 Without extensive workload measurements, it is hard to know the most accurate value for $c_r$, and so we also evaluate the simulation results with varying $c_r$. 
+For each $c_r$, we find the value of $p$ which minimizes the expected workload as we did in Figure~\ref{fig:workload}.
+Figure~\ref{fig:optimal_ps} presents, for varying $c_r$ the round schedule parameter $p$ which minimizes expected workload.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{optimal_ps.png}
+\caption{The optimal (cost-minimizing) stopping probability $p$ for varying cost model parameters $c_r$.
+With $c_b=1$, the varying value of $c_r$ can equivalently be thought of as the ratio of the cost of a round to the cost of a ballot.}
+\label{fig:optimal_ps}
+\end{figure}

### 2022-07-20 providence/sections/conc.tex
more

%Conclusions
A rigorous tabulation audit is an important part of a secure election. Ballot polling RLAs are commonly used and simple, not relying on special election equipment like comparison RLAs. We present \Providence which is the most efficient and secure ballot polling RLA, as efficient as \Minerva  and flexible as \BRAVO. We present proofs and simulation results to verify the claimed properties of \Providence, and we provide an open source implementation of the stopping condition and useful related functionality.

### 2022-07-20 providence/sections/intro.tex
more

@@
 
-This paper focuses on ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
+This paper concerns ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
 Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).

### 2022-07-20 providence/sections/workload.tex
more

@@
 Thus the workload (cost and time) of an audit grows not just with the number of ballots sampled but also with the number of rounds.
-On the other hand, a one-and-done audit often draws more ballots than are necessary.
-We propose a simple workload model under which our simulation results suggest that a few smaller rounds gives lower average workload.
+Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading in the sense that the true winner is losing by the tally of the sampled ballots.
+On the other hand, a one-and-done audit may draw more ballots than are necessary.
+To evaluate the efficiency of various round schedules, we propose a simple workload model.
+Under this model, our simulation results suggest that slightly smaller round sizes give lower average workload.
 

### 2022-07-20 providence/sections/sims.tex
sim edits

@@
 We use the simulator in the R2B2 software library\cite{r2b2}.
-We perform $10000=10^4$ trials per margin for both an underyling announced outcomes and an underyling tie.
-
-The simulations with an underlying tie give an estimate of maximum risk as shown in Figure~\ref{fig:prov-risk}. For all margins, the estimated risks were less than the risk limit.
-
-Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn.
-Figure~\ref{fig:prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
-Figure~\ref{fig:prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
-Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
-As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-Note that a difference in twice as many ballots could be negligible in a contest with a wide margin.
-In a contest with a narrow margin (in the 2020 US Presidential election, several states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work.
-Section~\ref{sec:workload} discusses workload in more depth.
-%TODO add a more specific number of states for small margin
+We perform $10000=10^4$ trials per margin for both an underyling outcomed as announced and an underyling tie.
 
@@
 
+In the simulations with an underlying tie, the proportion of audits that stop, as shown in Figure~\ref{fig:prov-risk}, give an estimate of maximum risk.
+For all margins, this estimated risk is less than the risk limit, supporting the claim that \Providence is risk-limiting.
+
 \begin{figure}
@@
 
+Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn.
+Figure~\ref{fig:prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
+Figure~\ref{fig:prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
+Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
+As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
+Note that a difference in twice as many ballots could be negligible in a contest with a wide margin.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work.
+Section~\ref{sec:workload} discusses workload in more depth.
+
 \begin{figure}

### 2022-07-20 providence/sections/workload.tex
explaining

@@
 
-Consider the US state of Pennsylvania. In the 2020 US Presidential contest, Pennsylvania had a margin of $99999\%$.
+Consider the US state of Michigan. In the 2020 US Presidential contest, Michigan had a margin of $2.8\%$.
 
@@
 For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
-For this particular round schedule scheme, computing the expected number of rounds is straightforward; .
+For this particular round schedule scheme, computing the expected number of rounds is possible analytically(TODO: I worked it out, just need to add here).
 Our simulations are then used to estimate the expected number of ballots.
@@
 
-Without extensive workload measurements, it is hard to know the most accurate value for $c_r$, and so we also evaluate the simulation results with varying $c_r$. 
+Without workload measurements from multiple round audits, it is difficult to know an accurate value for $c_r$, and so we also evaluate the simulation results with varying $c_r$. 
 For each $c_r$, we find the value of $p$ which minimizes the expected workload as we did in Figure~\ref{fig:workload}.
+Rather than use the minimum cost estimate among the sparsely sampled points, we fit a quadratic polynomial and use its minimum to get a smoother relationship between $c_r$ and optimal $p$.
 Figure~\ref{fig:optimal_ps} presents, for varying $c_r$ the round schedule parameter $p$ which minimizes expected workload.
+We observe that for a workload model with no overhead for each round (i.e.
+$c_r=0$), the optimal round schedule is that with $p=0.72$.
+As $c_r$ increases, the optimal value of $p$ also increases, up to as much as $p=.9$ for $c_r=10^4$.
+
+Analysis like this can be used to select round schedules for audits.
+
+% check that report to see if we can at least suggest what might be reasonable
 

### 2022-07-20 providence/sections/prov.tex
nicer plots
notes

@@
+We now introduce the stopping condition of \Providence.
+(TODO: add some more motivation and description of Providence...
+where minerva insists on a bound on the average risk providence only bounds the current path of the audit...
+it is nice because it sort of combines the main ratios of bravo and minerva)
+
 \begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]

### 2022-07-20 providence/sections/sims.tex
better plots

@@
 \includegraphics[width=.5\textwidth]{prov_risk.png}
-\caption{The proportion of simulated audits with an underlying tie that stopped in one of the five simulated rounds.}
+\caption{The percent of simulated audits \Providence with an underlying tie that stopped in any of the five simulated rounds. This percent gives an estimate of the risk of the \Providence.}
 \label{fig:prov-risk}
@@
 
-In the simulations with an underlying tie, the proportion of audits that stop, as shown in Figure~\ref{fig:prov-risk}, give an estimate of maximum risk.
+In the simulations with an underlying tie, the percent of audits that stop, as shown in Figure~\ref{fig:prov-risk}, give an estimate of maximum risk.
 For all margins, this estimated risk is less than the risk limit, supporting the claim that \Providence is risk-limiting.
@@
 \includegraphics[width=.5\textwidth]{prov_sprob.png}
-\caption{The proportion of simulated audits with an margin as announced that stopped in one of the five simulated rounds.}
+\caption{For the simulated \Providence audits with underlying margin as announced, the percent of audits that stopped among those that entered a round.
+This percent gives an estimate of the stopping probability conditioned on the sample of the previous round.
+The average percent for rounds 1, 2, and 3 is $89.96\%$, $90.52\%$, and $90.98\%$ respectively.
+We show only the first three rounds since so few audits make it to rounds 4 and 5.}
 \label{fig:prov-sprob}

### 2022-07-20 providence/sections/workload.tex
notes
better plots

@@
 
-Consider the US state of Michigan. In the 2020 US Presidential contest, Michigan had a margin of $2.8\%$.
+Consider the US state of Michigan. In the 2020 US Presidential contest, Michigan had a margin of $2.8\%$. We again use risk-limit $\alpha = .1$.
 
@@
 Our simulations are then used to estimate the expected number of ballots.
+We run $10^4$ trials for each value of $p$.
 
@@
 As $c_r$ increases, the optimal value of $p$ also increases, up to as much as $p=.9$ for $c_r=10^4$.
-
-Analysis like this can be used to select round schedules for audits.
-
-% check that report to see if we can at least suggest what might be reasonable
+(TODO: check report(s) for any bit of an estimate we can give of what these values might be in practice, noting the issues/error expected)
 
@@
 \end{figure}
+
+Analysis like this may be useful to election officials for selecting round schedules for audits. 
+It remains an open question whether a fast, accurate analytical approach for estimating $E_b$ and $E_r$ exists.
+
+Since the simulations are relatively computationally expensive, an analytical alternative is desirable. (TODO: need to compare my best analytical approach to the simulation results now)

### 2022-10-07 providence/sections/workload.tex
more workload section

--- 2022-07-20 providence/sections/workload.tex
@@
 This is appealing for two reasons.
-Firstly, rounds have some overhead in time and effort.
-Thus the workload (cost and time) of an audit grows not just with the number of ballots sampled but also with the number of rounds.
-Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading in the sense that the true winner is losing by the tally of the sampled ballots.
-On the other hand, a one-and-done audit may draw more ballots than are necessary.
-To evaluate the efficiency of various round schedules, we propose a simple workload model.
-Under this model, our simulation results suggest that slightly smaller round sizes give lower average workload.
+Firstly, rounds have some overhead in both time and effort.
+Thus the time and personhours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
+Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading in the sense that the true winner recieves has fewer votest than some other candidate according to the tally of the sampled ballots.
+On the other hand, a one-and-done audit may draw more ballots than are necessary; a more efficient round schedule could require less effort and time pre-certification.
+To evaluate the quality of various round schedules, we construct a simple workload model.
+Under this model we show how optimal round schedules can be chosen.
+We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like desired probability of a nonmisleading result.
 
-We begin with the simple assumption that part of the workload grows linearly with the number of ballots, giving us the per ballot cost parameter $c_{b}$.
-It is less clear how cost is related to the number of rounds, but we propose a simple model in which the round cost is also linear, with constant per round cost $c_{r}$.
-This is a reasonable assumption for contests with small margin in which all or nearly all ballot boxes need to be opened in each round.
-[@Poorvi, I need help justifying the workload model.]
+As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $5.3\%$ between the two candidates with the most votes.
+Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds. 
+%A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
+%We will use the more standard approach of simulations to give an example here.
+Therefore we use the typical approach of simulations, again with risk-limit $10\%$.
+We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$.
+That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again give a probability of stopping $p$ in the second round.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is possible analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trials for each value of $p$, assuming the announced results are correct.
 
-So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we get that the cost $C$ of the audit is
+\subsection{Person-hours}
+The simplest workload models are a function of just the total number of a ballots sampled.
+Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
+Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the Providence values.
+It is straightforward to show that Providence and both forms of BRAVO collapse to the same test in the case where each round is a single ballot.
+Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e.
+larger rounds), Providence requires fewer ballots on average.
+In particular, the savings of Providence become larger as $p$ increases; for $p=.95$, EoR BRAVO and SO BRAVO require more than $2$ and $1.4$ times as many ballots as Providence respectively.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{avg_bals.png}
+\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
+\label{fig:avg_bals}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
+\caption{The total number of ballots sampled on average in our simulations given as a fraction of those sampled by Providence, for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
+\label{fig:avg_bals_ratio}
+\end{figure}
+
+It is clear that average number of ballots alone is an inadequate workload measure. 
+(Consider a state conducting its audit by selecting a single ballot at random, 
+notifying just the county where the ballot is located, and then waiting to hear back for the manual interpretation of the ballot before moving on to the next one. 
+This of course is preposterous and is why audits are actually performed in rounds.
+
+In a US state-wide RLA, the state organizes the audit by determining the random sample and communicating with the counties, but election officials at the county level physically sample and inspect the ballots from their precincts.
+Therefore each audit round requires some number of person-hours for set up and communication between state and county.
+This overhead for a round includes choosing the round size, generating the random sample, and communicating that random sample to the counties, as well as the communication of the results back to the state afterwards.
+%To make it clear that this overhead exists,
+
+Consequently, we now consider a model with a constant per-ballot cost $c_b$ and a constant per-round cost $c_r$.
+So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the cost $C$ of the audit is
 \begin{equation}
 C(\mathcal{A}) = E_b c_b + E_r c_r
-\label{eq:cost}
+\label{eq:round_cost}
 \end{equation}
 
-Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) is challenging since the number of possible sequences of samples grows exponentially with the number of rounds. 
-%A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
-%We will use the more standard approach of simulations to give an example here.
-Therefore we again use the standard approach of simulations.
-
-Consider the US state of Pennsylvania. In the 2020 US Presidential contest, Pennsylvania had a margin of $99999\%$.
-
-We let the per ballot cost be one, $c_b=1$.
+For simplicity, (and without loss of generality), we assume the per ballot cost is one, $c_b=1$.
 Then the per round cost $c_r$ tells us the cost of a round as a number of ballots.
-We begin with an example in which $c_r=10$.
-That is, we suppose that the overhead of a round is equal to the workload of sampling $10$ ballots.
-
-We now consider a simple round schedule for our example.
-Each round is selected to give the same probability of stopping, $p$.
-If the audit does not stop in the first round, we find a round size which, given the sample drawn in the first round, will give a probability of stopping $p$ in the second round.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
-For this particular round schedule scheme, computing the expected number of rounds is straightforward; .
-Our simulations are then used to estimate the expected number of ballots.
+We begin with $c_r=1000$ as a conservative example.
+That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
+Based on available data[rhode island nums], the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload.
+This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
+As shown in Figure~\ref{fig:with_round_cost}, lower average costs are achieved by selecting higher stopping probability; Providence achieves the lowest minimum average cost is achieved at roughly $0.7$.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{workload.png}
-\caption{For cost parameters $c_b=1$ and $c_r=10$, this plot shows the expected cost for various round schedule parameters $p$.
+\includegraphics[width=.5\textwidth]{with_round_cost.png}
+\caption{For cost parameters $c_b=1$ and $c_r=1000$, this plot shows the expected cost for various round schedule parameters $p$.
 Expected cost is found using Equation~\ref{eq:cost} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
-\label{fig:workload}
+\label{fig:with_round_cost}
 \end{figure}
 
-The simulation results for our example are given in Figure~\ref{fig:workload}. 
-For various round schedules, given by the varying parameters $p$ on the horizontal axis, we see that the average audit cost varies with a minimum at roughly $.8$.
-In particular, the one-and-done $p=.9$ audits have an average of $zz\%$TODO higher cost than the $p=.8$ audits.
-Especially for low margin contests, using the cost-minimizing round schedule could translate to significant reduction in cost or savings in time, a particularly important factor given tight certification deadlines.
+Importantly, this gives us a way to estimate the expected cost, as well as which round schedule value $p$ achieves it, for arbitrary round cost.
+For each round cost $c_r$, we estimate the minimum of the data like that shown in Figure~\ref{fig:with_round_cost} by fitting a lower order polynomial.
+Figure~\ref{fig:optimal_costs} shows the optimal achievable workload for a wide range of per round costs.
+Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve the minimal workloads.
+For any reasonable round costs, Providence minimizes cost with at least roughly $10\%$ and $30\%$ improvement over SO BRAVO and EoR BRAVO respectively.
 
-Without extensive workload measurements, it is hard to know the most accurate value for $c_r$, and so we also evaluate the simulation results with varying $c_r$. 
-For each $c_r$, we find the value of $p$ which minimizes the expected workload as we did in Figure~\ref{fig:workload}.
-Figure~\ref{fig:optimal_ps} presents, for varying $c_r$ the round schedule parameter $p$ which minimizes expected workload.
+\begin{figure}
+\includegraphics[width=.5\textwidth]{optimal_costs.png}
+\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the Providence values.}
+\label{fig:optimal_costs}
+\end{figure}
 
@@
 \end{figure}
+
+Another consideration for election officals is the sensitivit
+Election officials may also want to choose larger sample sizes in order to avoid small samples with misleading tallies.
+In particular, it may be desirable to avoid with very high probability a case in which a sample shows a true loser winning.
+Assuming a correctly announced outcome, it is straightforward to compute the probability with which a reported loser has more votes than the true winner in a sample, and so we incorporate this into our tool as well.
+As an illustrating example, our simulations ...
+TODO
+
+For a more complete model, we can also introduce container-level workload.
+The time to sample a ballot from an entirely new box is typically much greater than to sample a ballot from an already-open box.
+[rhode islandd numbers shows it is abou tthis] Available election results give per-precinct granularity of vote tallies, as opposed to individual box informatoin.
+In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct constant cost, $c_p$.
+In this model, the workload estimate incurs an additional cost of $c_p$ every time a precinct is sampled from for the first time in a round.
+That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$.
+Then the new model is
+\begin{equation}
+C(E_b, E_r, E_p) = E_b c_b + E_r c_r + E_p c_p
+\label{eq:round_and_precinct_cost}
+\end{equation}
+
+We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_costs_with_precinct}.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{optimal_costs_with_precinct.png}
+\caption{The average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost}.}
+\label{fig:optimal_ps}
+\end{figure}
+
+\subsection{Real time}
+Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time spent is also an important factor to consider when planning audits.
+
+TODO

### 2022-10-07 providence/sections/workload.tex
unstyled sub headers
more workload

@@
 \subsection{Person-hours}
+
+Average ballots
+
 The simplest workload models are a function of just the total number of a ballots sampled.
@@
 \end{figure}
+
+Round overhead
 
@@
 That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
-Based on available data[rhode island nums], the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload.
+Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload.
 This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
@@
 
+Misleading probability
+
 Another consideration for election officals is the sensitivit
@@
 
+Precinct overhead
+
 For a more complete model, we can also introduce container-level workload.
-The time to sample a ballot from an entirely new box is typically much greater than to sample a ballot from an already-open box.
-[rhode islandd numbers shows it is abou tthis] Available election results give per-precinct granularity of vote tallies, as opposed to individual box informatoin.
-In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct constant cost, $c_p$.
+The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box.
+Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container.
+Typically available election results give per-precinct granularity of vote tallies, rather than individual container information.
+In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant cost, $c_p$.
 In this model, the workload estimate incurs an additional cost of $c_p$ every time a precinct is sampled from for the first time in a round.

### 2022-10-07 providence/sections/workload.tex
misleading

@@
 
-Another consideration for election officals is the sensitivit
 Election officials may also want to choose larger sample sizes in order to avoid small samples with misleading tallies.
-In particular, it may be desirable to avoid with very high probability a case in which a sample shows a true loser winning.
+It is desirable to avoid, with very high probability, a case where a sample shows a true loser winning.
 Assuming a correctly announced outcome, it is straightforward to compute the probability with which a reported loser has more votes than the true winner in a sample, and so we incorporate this into our tool as well.
-As an illustrating example, our simulations ...
-TODO
+As an illustrating example, we consider the expected occurence of misleading samples using our simulations.
+Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a misleading cumulative sample at any point.
+Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
+Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.
+In Figure~\reg{fig:misleading}, a horizontal line is included to illustrate the implication of a small \emph{misleading limit} of $0.1$.
+Observe that to achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
+It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US.
+Consequently, the desired \emph{misleading limit} may be a decided constraint in the choice of round schedule.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{misleading.png}
+\caption{The proportion of simulated Providence audits that had a \emph{misleading} cumulative sample in any round.}
+\label{fig:misleading}
+\end{figure}
 

### 2022-10-07 providence/sections/workload.tex
more workload

@@
 \includegraphics[width=.5\textwidth]{optimal_costs.png}
-\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the Providence values.}
+\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the Providence values.
+(We show the value for varying $c_r$ since the value of $c_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $c_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
 \label{fig:optimal_costs}
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{optimal_costs_with_precinct.png}
+\includegraphics[width=.5\textwidth]{optima_with_precinct_cost.png}
 \caption{The average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost}.}

### 2022-10-08 providence/sections/workload.tex
more

@@
 
-Average ballots
-
-The simplest workload models are a function of just the total number of a ballots sampled.
+\emph{Average total ballots.} The simplest workload models are a function of just the total number of a ballots sampled.
 Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
@@
 
-Round overhead
-
-It is clear that average number of ballots alone is an inadequate workload measure. 
+\emph{Round overhead.} It is clear that average number of ballots alone is an inadequate workload measure. 
 (Consider a state conducting its audit by selecting a single ballot at random, 
@@
 This overhead for a round includes choosing the round size, generating the random sample, and communicating that random sample to the counties, as well as the communication of the results back to the state afterwards.
-%To make it clear that this overhead exists,
 
@@
 
-Misleading probability
-
-Election officials may also want to choose larger sample sizes in order to avoid small samples with misleading tallies.
-It is desirable to avoid, with very high probability, a case where a sample shows a true loser winning.
-Assuming a correctly announced outcome, it is straightforward to compute the probability with which a reported loser has more votes than the true winner in a sample, and so we incorporate this into our tool as well.
-As an illustrating example, we consider the expected occurence of misleading samples using our simulations.
-Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a misleading cumulative sample at any point.
-Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
-Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.
-In Figure~\reg{fig:misleading}, a horizontal line is included to illustrate the implication of a small \emph{misleading limit} of $0.1$.
-Observe that to achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
-It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US.
-Consequently, the desired \emph{misleading limit} may be a decided constraint in the choice of round schedule.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{misleading.png}
-\caption{The proportion of simulated Providence audits that had a \emph{misleading} cumulative sample in any round.}
-\label{fig:misleading}
-\end{figure}
-
-Precinct overhead
-
-For a more complete model, we can also introduce container-level workload.
+\emph{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
 The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box.
@@
 
-We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_costs_with_precinct}.
+We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_cost_precinct_cost_ratio}.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{optima_with_precinct_cost.png}
+\includegraphics[width=.5\textwidth]{optimal_cost_precinct_cost_ratio.png}
 \caption{The average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost}.}
-\label{fig:optimal_ps}
+\label{fig:optimal_cost_precinct_cost_ratio}
 \end{figure}
@@
 TODO
+
+\subsection{Misleading samples}
+
+Election officials may also want to choose larger sample sizes in order to avoid small samples with misleading tallies.
+It is desirable to avoid, with very high probability, a case where a sample shows a true loser winning.
+Assuming a correctly announced outcome, it is straightforward to compute the probability with which a reported loser has more votes than the true winner in a sample, and so we incorporate this into our tool as well.
+As an illustrating example, we consider the expected occurence of misleading samples using our simulations.
+Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a misleading cumulative sample at any point.
+Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
+Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.
+In Figure~\ref{fig:misleading}, a horizontal line is included to illustrate the implication of a small \emph{misleading limit} of $0.1$.
+Observe that to achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
+It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US.
+Consequently, the desired \emph{misleading limit} may be a decided constraint in the choice of round schedule.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{misleading.png}
+\caption{The proportion of simulated Providence audits that had a \emph{misleading} cumulative sample in any round.}
+\label{fig:misleading}
+\end{figure}

### 2022-10-08 providence/sections/workload.tex
real time

@@
 \includegraphics[width=.5\textwidth]{optimal_cost_precinct_cost_ratio.png}
-\caption{The average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost}.}
+\caption{The optimal average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost} for varying $c_p$, given as a fraction of the value for Providence}
 \label{fig:optimal_cost_precinct_cost_ratio}
@@
 \subsection{Real time}
-Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time spent is also an important factor to consider when planning audits.
+Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
+Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
+In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
+Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
+For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
+Similarly, we use real time per-ballot, per-round, and per-precinct cost variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
+\begin{equation}
+T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p
+\label{eq:real_time}
+\end{equation}
 
-TODO
+As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
+Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.
+In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
+Figure~\ref{fig:real_time} shows how the esimated real time for these values differs as a function of $p$.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{real_time.png}
+\caption{The real time as estimated by Equation~\ref{eq:real_time} for varying $p$ with expected values as estimated by our simulations.}
+\label{fig:real_time}
+\end{figure}
 

### 2022-10-08 providence/sections/workload.tex
workload

@@
 Importantly, this gives us a way to estimate the expected cost, as well as which round schedule value $p$ achieves it, for arbitrary round cost.
-For each round cost $c_r$, we estimate the minimum of the data like that shown in Figure~\ref{fig:with_round_cost} by fitting a lower order polynomial.
+For each round cost $c_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_cost} and then find the minimum average cost achieved for each of the audits and its corresponding stopping probability $p$.
 Figure~\ref{fig:optimal_costs} shows the optimal achievable workload for a wide range of per round costs.
-Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve the minimal workloads.
-For any reasonable round costs, Providence minimizes cost with at least roughly $10\%$ and $30\%$ improvement over SO BRAVO and EoR BRAVO respectively.
+For very low round costs, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less.
+On the other hand, for extremely larger values of round cost, the average number of ballots has little impact on the workload function, and so the three audits again have similar values.
+For more reasonable values of the round cost $c_r$, SO BRAVO and EoR BRAVO achieve minimum cost roughly $1.1$ and $1.3$ times greater than that of Providence.
+Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
+As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $c_r$.
+Notice that Providence is generally above and to the left of SO BRAVO, and SO BRAVO is generally above and to the left of EOR BRAVO.
+This relationship reflects the fact that for the same round cost, Providence can get away with a larger stopping probability because it requires fewer ballots.
 

### 2022-10-09 providence/sections/intro.tex
small changes to abstract and intro

@@
 This paper concerns ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
-Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
+Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere).
 In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
 A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit.
-Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
+If so, the audit stops; else another round is drawn.
+Election officials would typically decide to do a full manual hand count if the audit does not stop in spite of drawing a large number of ballots.
+Because the decision to stop or draw more ballots is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
 The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.

### 2022-10-10 providence/sections/workload.tex
table

@@
 As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
-Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.
-In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
+Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
 Figure~\ref{fig:real_time} shows how the esimated real time for these values differs as a function of $p$.
+It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retreiving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors.
+We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function.
+Use of this optimal scheduling tool would depend on parameter estimates tailored to each case.
 
@@
 
-Election officials may also want to choose larger sample sizes in order to avoid small samples with misleading tallies.
-It is desirable to avoid, with very high probability, a case where a sample shows a true loser winning.
-Assuming a correctly announced outcome, it is straightforward to compute the probability with which a reported loser has more votes than the true winner in a sample, and so we incorporate this into our tool as well.
-As an illustrating example, we consider the expected occurence of misleading samples using our simulations.
-Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a misleading cumulative sample at any point.
+Unfortunately, efficiency alone is not sufficient for planning audits. In the US today, election officials need to make legitimate safety considerations.
+When drawing a random sample of the ballots, it is always possible that the tally of the sample provides misleading information.
+In a random sample, a true loser may recieve more votes than the true winner.
+This happens more often when the sample sizes are small.
+In these RLAs, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
+
+Imagine that Alice beats Bob in an election contest both truly and by the announced results, but Bob's supporters are insistent he really won.
+When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner.
+After the first round, by chance, there are more votes for Bob than for Alice in the sample.
+Bob's supporters celebrate their victory that the audit has in fact revealed that Bob really won, but the election officials have to explain that they are moving on to a second round.
+After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
+This is an undesirable situation.
+
+We introduce the notion of a \emph{misleading sample}, any cumulative sample which, assuming the announced outcome is correct, contains more ballots for a loser than for the winner.
+We can again use our simulations to gain insight into the frequency of \emph{misleading samples}.
+For each stopping probability $p$, Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a \emph{misleading sample} at any point. 
 Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
 Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.
-In Figure~\ref{fig:misleading}, a horizontal line is included to illustrate the implication of a small \emph{misleading limit} of $0.1$.
-Observe that to achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
-It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US.
-Consequently, the desired \emph{misleading limit} may be a decided constraint in the choice of round schedule.
+
+In Figure~\ref{fig:misleading}, horizontal lines are included to show \emph{misleading limits} of $0.1$, $0.01$, and $0.001$.
+To achieve a probability of a misleading sample of at most $0.1$, a round schedule with at least roughly $p=.3$ is needed.
+To achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
+It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US and the associated threats of violence.
+Consequently, the desired \emph{misleading limit} may be a deciding constraint in the choice of round schedule.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{misleading.png}
+\includegraphics[width=.5\textwidth]{misleading_limits.png}
 \caption{The proportion of simulated Providence audits that had a \emph{misleading} cumulative sample in any round.}
@@
 \end{figure}
+
+If election officials wish to achieve a maximum \emph{misleading limit} for all the rounds, our simulation analysis could help.
+On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample.
+Table~\ref{tab:misleading} shows
+
+\begin{figure}
+\begin{tabular}{ |l|l|r|c|c|c| }
+\hline
+$M$ & margin & $n$ & Prov & SO & EOR \\
+\hline
+0.1&0.25&25&0.221&0.152&0.115\\
+&0.2&41&0.178&0.169&0.105\\
+&0.15&73&0.202&0.186&0.141\\
+&0.1&163&0.222&0.182&0.107\\
+&0.05&657&0.227&0.192&0.127\\
+&0.04&1027&0.237&0.193&0.124\\
+&0.03&1825&0.246&0.194&0.124\\
+&0.02&4105&0.246&0.195&0.124\\
+&0.01&16423&0.246&0.196&0.124\\
+\hline
+0.01&0.25&85&0.792&0.707&0.559\\
+&0.2&133&0.826&0.71&0.593\\
+&0.15&239&0.817&0.712&0.549\\
+&0.1&539&0.805&0.717&0.567\\
+&0.05&2163&0.817&0.721&0.569\\
+&0.04&3381&0.82&0.722&0.563\\
+&0.03&6011&0.824&0.723&0.573\\
+&0.02&13527&0.824&0.723&0.57\\
+&0.01&54117&0.824&0.724&0.57\\
+\hline
+0.001&0.25&149&0.962&0.889&0.783\\
+&0.2&235&0.963&0.89&0.768\\
+&0.15&421&0.958&0.894&0.801\\
+&0.1&951&0.958&0.894&0.793\\
+&0.05&3815&0.96&0.896&0.785\\
+&0.04&5965&0.961&0.896&0.791\\
+&0.03&10607&0.961&0.897&0.787\\
+&0.02&23869&0.962&0.897&0.787\\
+&0.01&95491&0.962&0.897&0.787\\
+\hline
+\end{tabular}
+\caption{For various margins, this table gives the minimum first round size $n$ to achieve at most a probability $M$ of a \emph{misleading sample} in the first round.
+The corresponding stopping probabilities of Providence, SO BRAVO, and EoR BRAVO are given for each value of $n$.}
+\label{fig:misleading}
+\end{figure}
+
+\textbf{Other misleading RLA results.}

### 2022-10-10 providence/sections/workload.tex
more

@@
 
-\emph{Average total ballots.} The simplest workload models are a function of just the total number of a ballots sampled.
+\noindent\textbf{Average total ballots.} 
+The simplest workload models are a function of just the total number of a ballots sampled.
 Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
-Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the Providence values.
-It is straightforward to show that Providence and both forms of BRAVO collapse to the same test in the case where each round is a single ballot.
+Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the \Providence values.
+It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test in the case where each round is a single ballot.
 Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e.
-larger rounds), Providence requires fewer ballots on average.
-In particular, the savings of Providence become larger as $p$ increases; for $p=.95$, EoR BRAVO and SO BRAVO require more than $2$ and $1.4$ times as many ballots as Providence respectively.
+larger rounds), \Providence requires fewer ballots on average.
+In particular, the savings of \Providence become larger as $p$ increases; for $p=.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
 
@@
 \includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
-\caption{The total number of ballots sampled on average in our simulations given as a fraction of those sampled by Providence, for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
+\caption{The total number of ballots sampled on average in our simulations given as a fraction of those sampled by \Providence, for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
 \label{fig:avg_bals_ratio}
@@
 
-\emph{Round overhead.} It is clear that average number of ballots alone is an inadequate workload measure. 
+\noindent\textbf{Round overhead.} 
+It is clear that average number of ballots alone is an inadequate workload measure. 
 (Consider a state conducting its audit by selecting a single ballot at random, 
 notifying just the county where the ballot is located, and then waiting to hear back for the manual interpretation of the ballot before moving on to the next one. 
-This of course is preposterous and is why audits are actually performed in rounds.
+This of course is inefficient and is why audits are actually performed in rounds.)
 
@@
 This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
-As shown in Figure~\ref{fig:with_round_cost}, lower average costs are achieved by selecting higher stopping probability; Providence achieves the lowest minimum average cost is achieved at roughly $0.7$.
+As shown in Figure~\ref{fig:with_round_cost}, lower average costs are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average cost is achieved at roughly $0.7$.
 
@@
 On the other hand, for extremely larger values of round cost, the average number of ballots has little impact on the workload function, and so the three audits again have similar values.
-For more reasonable values of the round cost $c_r$, SO BRAVO and EoR BRAVO achieve minimum cost roughly $1.1$ and $1.3$ times greater than that of Providence.
+For more reasonable values of the round cost $c_r$, SO \BRAVO and EoR \BRAVO achieve minimum cost roughly $1.1$ and $1.3$ times greater than that of \Providence.
 Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
 As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $c_r$.
-Notice that Providence is generally above and to the left of SO BRAVO, and SO BRAVO is generally above and to the left of EOR BRAVO.
-This relationship reflects the fact that for the same round cost, Providence can get away with a larger stopping probability because it requires fewer ballots.
+Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EOR \BRAVO.
+This relationship reflects the fact that for the same round cost, \Providence can get away with a larger stopping probability because it requires fewer ballots.
 
@@
 \includegraphics[width=.5\textwidth]{optimal_costs.png}
-\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the Providence values.
+\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the \Providence values.
 (We show the value for varying $c_r$ since the value of $c_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $c_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
@@
 
-\emph{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
+\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
 The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box.
@@
 \includegraphics[width=.5\textwidth]{optimal_cost_precinct_cost_ratio.png}
-\caption{The optimal average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost} for varying $c_p$, given as a fraction of the value for Providence}
+\caption{The optimal average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost} for varying $c_p$, given as a fraction of the value for \Providence}
 \label{fig:optimal_cost_precinct_cost_ratio}
@@
 \includegraphics[width=.5\textwidth]{misleading_limits.png}
-\caption{The proportion of simulated Providence audits that had a \emph{misleading} cumulative sample in any round.}
+\caption{The proportion of simulated \Providence audits that had a \emph{misleading sample} in any round.}
 \label{fig:misleading}
@@
 On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample.
-Table~\ref{tab:misleading} shows
+Table~\ref{tab:misleading} shows for various margins the minimum first round size $n$ that guarantees a probability of a \emph{misleading sample} at most $M\in\{0.1,0.01,0.001\}$.
+For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO.
+    As seen in the Table~\ref{tab:misleading}, for $M=0.01$, the the minimum round sizes require at least roughly a $0.8$ probability of stopping in the first round.
+Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$, the election officials may use this constraint on the probability of a \emph{misleading sample} as the deciding factor in selecting a first round size.
 
-\begin{figure}
+\begin{table}
+\center
 \begin{tabular}{ |l|l|r|c|c|c| }
@@
 \caption{For various margins, this table gives the minimum first round size $n$ to achieve at most a probability $M$ of a \emph{misleading sample} in the first round.
-The corresponding stopping probabilities of Providence, SO BRAVO, and EoR BRAVO are given for each value of $n$.}
-\label{fig:misleading}
-\end{figure}
+The corresponding stopping probabilities of \Providence, SO \BRAVO, and EoR \BRAVO are given for each value of $n$.}
+\label{tab:misleading}
+\end{table}
 

### 2022-10-10 providence/sections/workload.tex
more
quantum

@@
 As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $c_r$.
-Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EOR \BRAVO.
+Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO.
 This relationship reflects the fact that for the same round cost, \Providence can get away with a larger stopping probability because it requires fewer ballots.
@@
 \caption{The optimal (cost-minimizing) stopping probability $p$ for varying cost model parameters $c_r$.
-With $c_b=1$, the varying value of $c_r$ can equivalently be thought of as the ratio of the cost of a round to the cost of a ballot.}
+With $c_b=1$, the varying value of $c_r$ can equivalently be thought of as the ratio of the cost of a round to the cost of a ballot.
+(Note that the steps in this function are a consequence of our subsampling the workload function.
+That is, the workload-minimizing value of $p$ for each $c_r$ is only allowed to take on values at increments of $0.5$.)}
 \label{fig:optimal_ps}
@@
 
-If election officials wish to achieve a maximum \emph{misleading limit} for all the rounds, our simulation analysis could help.
+If election officials wish to enforce a \emph{misleading limit} for all the rounds, our simulation analysis could help.
 On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample.
@@
 \hline
-$M$ & margin & $n$ & Prov & SO & EOR \\
+$M$ & margin & $n$ & Prov & SO & EoR \\
 \hline

### 2022-10-10 providence/sections/intro.tex
rewrote intro
kept a copy of the original intro

--- 2022-07-20 providence/sections/intro.tex
@@
 %Intro
-The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances.
-For this reason,
-{\em evidence-based elections} \cite{evidence-based} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability.
-One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA) \cite{RLA}.
-An RLA is an audit which guarantees that the probability of concluding that an election outcome is correct, given that it is not, is below a pre-determined value known as the risk limit of the audit, independent of the true, unknown vote distribution of the underlying election.
-Over a dozen states in the US have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.
+It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected.
+Errors and/or manipulation may not always change an election outcome, but we would want to know when they do.
+{\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetected change in the election outcome.
+{\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; incorrect outcomes are detected with high probability when the evidence is examined.
+One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
+Many US states have had pilot RLA programs.
+Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
 
-This paper concerns ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
-Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
-In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
-A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit.
-Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
-The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.
+We propose the \Providence audit, a new approach to the ballot polling RLA, and propose a new model for the work load of an election.
+We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections.
+Our implementation of \Providence is likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
 
-The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced and stopping decisions are made after each ballot draw.
-In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2.
-There are two obvious approaches.
-The B2 stopping condition can be applied once at the end of each round: End-of Round (EoR) \BRAVO.
-Alternatively, the order of ballots in the sample can be tracked by election officials and the B2 \BRAVO stopping condition can be applied retroactively after each ballot drawn: Selection-Ordered (SO) \BRAVO.
-SO \BRAVO requires fewer ballots on average than EoR \BRAVO but requires the work of tracking the order of ballots rather than just their tally.
+\subsection{Background}
+Ballot comparison RLAs require the fewest ballots of all known RLA approaches.
+On the other hand, they require the use of special election technology and are not always feasible.
+Ballot-polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system.
+What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).
 
-\Minerva was designed for R2 audits and applies its stopping rule once for each round.
-Thus it does not require the tracking of ballots that SO \BRAVO does.
-Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+A ballot polling audit begins when a {\em round} \cite{usenix_minerva} of multiple randomly-chosen ballots is drawn.
+A risk measure is then computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn.
+Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
+Ballot polling audits have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+
+A {\em round-by-round (R2)} audit is one where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections.
+A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
+The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
+Election officials typically draw ballots in large round sizes, and \BRAVO needs to be modified for use in this manner.
+For use as an R2 audit, the \BRAVO stopping condition can be applied once at the end of each round (End-of Round (EoR)), or retroactively after each ballot drawn if ballot order is retained (Selection-Ordered (SO)).
+SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
+But it requires the additional effort of tracking the order of ballots, and not simply the sample tally.
+
+Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which uses only the sample and round tallies, and does not need ballot order.
+They prove that it requires fewer ballots than EoR \BRAVO when both audits have the same pre-determined round schedule and the true tally is as announced.
 %\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand.
@@
 (We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
-They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
-Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability.
+They also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced. 
+Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability \cite{simulations}, though the improvement from using \Minerva over either version of \BRAVO decreases with round size.
 
-While more efficient than \BRAVO, \Minerva requires that the round schedule is fixed in advance of the audit.
-This may lead to significant unnecessary work.
-For example, suppose the fixed round schedule for \Minerva is $10,000$ ballots per round (a reasonable number in a state-wide contest with a narrow margin), and the first round sample finds a number of ballots for the winner just short of that necessary for the audit to stop.
-It then may be sufficient to draw a very small second round and still stop with high probability in the second round, but the additional $10,000$ ballots must be drawn.
-In contrast, subsequent \BRAVO round sizes can be chosen based on preceding samples.
+A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins, because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
+\BRAVO, on the other hand, is not limited in this manner.
+This allows \BRAVO audits the flexibility of choosing smaller subsequent round sizes if the sample drawn so far is a ``good'' sample.
+An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.
 
-An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.
+A major limitation of our understanding of the ballot polling problem as a community is that we use the number of ballots drawn or values proportional to this number \cite{mclaughlin_thesis,bernhard-diss,RI-report} as measures of the workload of an audit.
+If this were a correct measure of the workload of an audit, we would want to use B2 audits (round size is one) and make decisions about stopping the audit after drawing each ballot, because this leads to the smallest expected number of ballots.
+Election officials, on the other hand, greatly prefer drawing many ballots at once.
+This preference is likely because each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
+Further, there is an overhead to finding a storage box and opening it.
+Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large.
+Thus the workload of an audit is not simply a linear (or affine) function of the number of ballots drawn.
+Relatedly, an optimal round schedule is not completely determined by the expected number of ballots drawn.
+It depends on other variables as well.
 
@@
 \begin{enumerate}
-\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
-\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
-\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
-\item Open source implementation of \Providence 
+\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples.
+\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence uses number of ballots similar to those of \Minerva, both fewer than either version of \BRAVO.
+\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island.
+\item Open source implementation of \Providence. 
+\item A model of workload that includes the overhead effort of each round and the overhead effort of retrieving a storage unit of ballots; simulations that illustrate the use of this model to compare the different types of ballot polling audits and to plan an audit with minimal workload.
+\item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
 %including the novel metric Probability of Misleading(name?)
 \end{enumerate}
+
+Our results demonstrate the superiority of \Providence over the other audits.
+We hope that our work will help election officials better plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
 

### 2022-10-10 providence/sections/intro.tex
small changes to intro

@@
 SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
-But it requires the additional effort of tracking the order of ballots, and not simply the sample tally.
+But it requires the additional effort of tracking the order of ballots.
 
-Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which uses only the sample and round tallies, and does not need ballot order.
+Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which does not need ballot order and relies only on sample and round tallies.
 They prove that it requires fewer ballots than EoR \BRAVO when both audits have the same pre-determined round schedule and the true tally is as announced.
@@
 Further, there is an overhead to finding a storage box and opening it.
+For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each).
+For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
 Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large.
@@
 It depends on other variables as well.
+The consideration of all these variables is necessary while planning an audit.
 

### 2022-10-11 providence/sections/workload.tex
acknowledgements, etc

@@
 
-Consequently, we now consider a model with a constant per-ballot cost $c_b$ and a constant per-round cost $c_r$.
-So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the cost $C$ of the audit is
+Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
+So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $C$ of the audit is
 \begin{equation}
-C(\mathcal{A}) = E_b c_b + E_r c_r
-\label{eq:round_cost}
+W(E_b,E_r) = E_b w_b + E_r w_r
+\label{eq:round_workload}
 \end{equation}
 
-For simplicity, (and without loss of generality), we assume the per ballot cost is one, $c_b=1$.
-Then the per round cost $c_r$ tells us the cost of a round as a number of ballots.
-We begin with $c_r=1000$ as a conservative example.
+For simplicity, (and without loss of generality), we assume the per ballot workload is one, $w_b=1$.
+Then the per round workload $w_r$ tells us the workload of a round as a number of ballots.
+We begin with $w_r=1000$ as a conservative example.
 That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
-Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload.
+Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload.
 This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
-As shown in Figure~\ref{fig:with_round_cost}, lower average costs are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average cost is achieved at roughly $0.7$.
+As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{with_round_cost.png}
-\caption{For cost parameters $c_b=1$ and $c_r=1000$, this plot shows the expected cost for various round schedule parameters $p$.
-Expected cost is found using Equation~\ref{eq:cost} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
-\label{fig:with_round_cost}
+\includegraphics[width=.5\textwidth]{with_round_workload.png}
+\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$.
+Expected workload is found using Equation~\ref{eq:workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
+\label{fig:with_round_workload}
 \end{figure}
 
-Importantly, this gives us a way to estimate the expected cost, as well as which round schedule value $p$ achieves it, for arbitrary round cost.
-For each round cost $c_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_cost} and then find the minimum average cost achieved for each of the audits and its corresponding stopping probability $p$.
-Figure~\ref{fig:optimal_costs} shows the optimal achievable workload for a wide range of per round costs.
-For very low round costs, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less.
-On the other hand, for extremely larger values of round cost, the average number of ballots has little impact on the workload function, and so the three audits again have similar values.
-For more reasonable values of the round cost $c_r$, SO \BRAVO and EoR \BRAVO achieve minimum cost roughly $1.1$ and $1.3$ times greater than that of \Providence.
+Importantly, this gives us a way to estimate the expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload.
+For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.
+Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads.
+For very low round workloads, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less.
+On the other hand, for extremely larger values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values.
+For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
 Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
-As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $c_r$.
+As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $w_r$.
 Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO.
-This relationship reflects the fact that for the same round cost, \Providence can get away with a larger stopping probability because it requires fewer ballots.
+This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{optimal_costs.png}
-\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the \Providence values.
-(We show the value for varying $c_r$ since the value of $c_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $c_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
-\label{fig:optimal_costs}
+\includegraphics[width=.5\textwidth]{optimal_workloads.png}
+\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.
+(We show the value for varying $w_r$ since the value of $w_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $w_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
+\label{fig:optimal_workloads}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{optimal_ps.png}
-\caption{The optimal (cost-minimizing) stopping probability $p$ for varying cost model parameters $c_r$.
-With $c_b=1$, the varying value of $c_r$ can equivalently be thought of as the ratio of the cost of a round to the cost of a ballot.
+\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$.
+With $w_b=1$, the varying value of $w_r$ can equivalently be thought of as the ratio of the workload of a round to the workload of a ballot.
 (Note that the steps in this function are a consequence of our subsampling the workload function.
-That is, the workload-minimizing value of $p$ for each $c_r$ is only allowed to take on values at increments of $0.5$.)}
+That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.5$.)}
 \label{fig:optimal_ps}
@@
 Typically available election results give per-precinct granularity of vote tallies, rather than individual container information.
-In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant cost, $c_p$.
-In this model, the workload estimate incurs an additional cost of $c_p$ every time a precinct is sampled from for the first time in a round.
+In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant workload, $w_p$.
+In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round.
 That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$.
@@
 \begin{equation}
-C(E_b, E_r, E_p) = E_b c_b + E_r c_r + E_p c_p
-\label{eq:round_and_precinct_cost}
+W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p
+\label{eq:round_and_precinct_workload}
 \end{equation}
 
-We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_cost_precinct_cost_ratio}.
+We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_workload_precinct_workload_ratio}.
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{optimal_cost_precinct_cost_ratio.png}
-\caption{The optimal average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost} for varying $c_p$, given as a fraction of the value for \Providence}
-\label{fig:optimal_cost_precinct_cost_ratio}
+\includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
+\caption{The optimal average workload of the audits we simulated using workload function \ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence}
+\label{fig:optimal_workload_precinct_workload_ratio}
 \end{figure}
@@
 For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
-Similarly, we use real time per-ballot, per-round, and per-precinct cost variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
+Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
 \begin{equation}
@@
 For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO.
-    As seen in the Table~\ref{tab:misleading}, for $M=0.01$, the the minimum round sizes require at least roughly a $0.8$ probability of stopping in the first round.
-Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$, the election officials may use this constraint on the probability of a \emph{misleading sample} as the deciding factor in selecting a first round size.
+    As seen in the Table~\ref{tab:misleading}, to enforce $M=0.01$ requires minimum round sizes with at least roughly a $0.8$ probability of stopping in the first round.
+Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$ to choose the first round size, the election officials may opt to use this constraint on the probability of a \emph{misleading sample} as the deciding factor in planning their audits.
 

### 2022-10-11 providence/sections/related.tex

@@
 %Related Work
-The \BRAVO audit \cite{bravo} is a well-known ballot polling audit which has been used in numerous pilot and real audits.
-When used to audit a two-candidate election, it is an instance of Wald's sequential probability ratio test (SPRT) \cite{wald}, and inherits the SPRT property of being the most efficient test (requiring the smallest expected number of ballots) if the election is as announced.
-The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken.
-Real election audits invest in drawing large numbers of ballot, called rounds, before making stopping decisions because sequentially sampling individual ballots has significant overhead (unsealing storage boxes and searching for individual ballots).
-It is possible to apply \BRAVO to the sequence of ballots in a round if the sequential order is retained.
-This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to ballots that were drawn earlier in the sample.
+The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn.
 
-We do know a great deal about the properties of \BRAVO.
-The risk limiting property of \BRAVO follows from the similar property of the SPRT.
-Stopping probabilities for \BRAVO may be estimated as implemented in \cite{arlo}; this method is due to Mark Lindeman and uses quadratic approximations.
-A later method for stopping probability estimates presented by Zag{\'o}rski {\em et al.}\cite{usenix_minerva,arxiv_athena} uses a similar technique for narrow margins and a separate algorithm for wider margins, the results of which match simulation results reported by Lindeman {\em et al.} \cite[Table 1]{bravo}.
-
-The \Minerva audit \cite{usenix_minerva,arxiv_athena} was developed for large first round sizes which enable election officials to be done in one round with large probability.
-It uses information from the entire sample, and has been proven to be risk limiting when the round schedule for the audit is determined before the audit begins.
-That is, information about the actual ballots drawn in the first round cannot inform future round sizes.
-First-round sizes for a $0.9$ stopping probability when the election is as announced have been computed for a wide range of margins and are smaller than those for EoR and SO \BRAVO.
-First round simulations of \Minerva \cite{arxiv_athena} demonstrate that its first-round properties---regarding the probabilities of stopping when the underlying election is tied and when it is as announced---are as predicted for first round sizes with stopping probability $0.9$.
+The \Minerva audit \cite{usenix_minerva,arxiv_athena} was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
+First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins; simulations \cite{arxiv_athena} support these observations.
 Additional simulations \cite{simulations} have shown that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
@@
 
-Ballot polling audit simulations have been used to familiarize election officials and the public with the approach \cite{dice} and to investigate the efficiency of audits \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
-Some workload measurements have been made\cite{RI-report}. 
-While total ballots sampled can give naive workload estimates\cite{bernoulli-ballot-polling}, Bernhard presents a more complex workload estimation model\cite{bernhard-diss}.
+Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and to understand audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
+There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
+Simple workload estimates may be obtained by using the number of ballots drawn \cite{bernoulli-ballot-polling}, a more thorough workload estimation model includes the time taken to access individual ballots\cite{bernhard-diss}.
 

### 2022-10-11 providence/sections/intro.tex
replace "open" with "unseal" in intro

@@
 This preference is likely because each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
-Further, there is an overhead to finding a storage box and opening it.
+Further, there is an overhead to finding a storage box and unsealing it.
 For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each).
@@
 
-Our results demonstrate the superiority of \Providence over the other audits.
-We hope that our work will help election officials better plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
+Our results demonstrate the superiority of \Providence over the other audits. Our work may be used by election officials to plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
 

### 2022-10-11 providence/sections/related.tex
small changes

@@
 
-\subsection{Model}
+We now summarize the model drawing largely from the notation and terminology of \cite{usenix_minerva,arxiv_athena,simulations,bravo}.
+The model is related work and not claimed to be original to this work.
+
 An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
-\BRAVO and \Minerva are modeled as binary hypothesis tests where the null hypothesis $H_0$ corresponds to a true underlying tie and the alternative hypothesis $H_a$ corresponds to the underyling election being as announced.
-(With an even total number of ballots, $H_0$ corresponds to the announced loser winning by one ballot.)
-Thus the null hypothesis is the incorrectly announced outcome which is most difficult to detect.
-A risk-limiting audit requires that in the event of an underlying tie, the probability of failing to detect the error is below the risk limit.
+\BRAVO and \Minerva are modeled as binary hypothesis tests where the null hypothesis $H_0$ corresponds to a tied election and the alternative hypothesis $H_a$ to an election tally as announced. 
+(When the number of ballots is odd, $H_0$ corresponds to the announced loser winning by one ballot.)
+Thus the null hypothesis is the outcome distinct from the announced one which is most difficult to detect; the probability of failing to detect it, given that the null hypothesis is true, is the worst case such probability and should be below the risk limit \cite{Bayesian-RLA}.
 
@@
 
-If testing the \BRAVO stopping condition after each individual ballot
-is drawn (a \B \BRAVO audit), $\sigma$ is equivalent to the 
+%If testing the \BRAVO stopping condition after each individual ballot
+% is drawn (a \B \BRAVO audit), 
+It is easy to see that the ratio $\sigma$ is the 
 likelihood ratio:
@@
 
-It now becomes useful to have shorthand for a sequence of round sizes and a sequence
-of winner ballot tallies.
+It now becomes useful to have shorthand for a sequence of cumulative round sizes and the corresponding sequence
+of cumulative winner ballot tallies.
 We use:
@@
 
-Where \BRAVO uses a ratio of points on conditional probability distributions, \Minerva uses a ratio of \emph{tails} of conditional probability distributions.
+Where \BRAVO uses the ratio of the values of the conditional probability distributions, \Minerva uses the ratio of their \emph{tails}.
 
 \begin{definition}[\Minerva Ratio] \label{def:minerva_ratio} The \R \Minerva audit uses the ratio $\tau_j$.
-We use cumulative round sizes $\bm{n_j}$, with corresponding $\bm{k_j}$ ballots for the reported winner in reach round.
+We use cumulative round sizes $\bm{n_j}$, with corresponding $\bm{k_j}$ ballots for the reported winner in each round.
 The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.

### 2022-10-11 providence/sections/intro.tex

@@
 
-%\subsection{Organization} 
-%Section \ref{sec:related} describes related work.
-The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results.
-Section \ref{sec:conc} has our conclusions.
+\subsection{Organization} 
+Section \ref{sec:related} describes related work.
+Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
+Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2020 US Presidential election in the state of Virginia.
+Our conclusions, the availability of an audit implementation and acknowledgements may be found in sections \ref{sec:conc}, \ref{sec:avail} and \ref{sec:ack} respectively.

### 2022-10-11 providence/sections/workload.tex
add constant to workload functions

@@
 Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
-So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $C$ of the audit is
+So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $W$ of the audit is
 \begin{equation}
-W(E_b,E_r) = E_b w_b + E_r w_r
+W(E_b,E_r) = E_b w_b + E_r w_r + C
 \label{eq:round_workload}
@@
 
+Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials for their estimates.
 For simplicity, (and without loss of generality), we assume the per ballot workload is one, $w_b=1$.
@@
 \begin{equation}
-W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p
+W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
 \label{eq:round_and_precinct_workload}

### 2022-10-11 providence/sections/workload.tex
beginning to work out the so misleading case

@@
 
-\textbf{Other misleading RLA results.}
+\textbf{Misleading SO Risk Measures.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
+
+After drawing $n>1$ ballots in a round, some number $k$ of them are votes for the announced winner.
+There are $\binom{n}{k}$ possible sequences of ballots which can lead to such a sample.
+Given the value of $k$, however, the particular sequence of the sample contains no additional information about whether the sample is more likely under one the alternative or null hypotheses.
+That said, an SO \BRAVO RLA will not stop as a function of $n$ and $k$ but also as a function of the sequence.
+In particular, if the sequence of ballots is such that the standard \BRAVO stopping condition was met for some $n'<n$ and corresponding $k'<n$, the audit will stop, even if by the end of the sequence the values $k$ and $n$ no longer meet the \BRAVO condition.
+To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result.
+The misleading nature of such stoppages is the note we are making.
+
+It is easy to use our simulations to see how often this occurs.

### 2022-10-11 providence/sections/prov.tex

--- 2022-05-16 providence/sections/prov.tex
@@
+In this section we motivate and introduce the stopping condition of \Providence.
+
+The \Minerva stopping condition tests the tail ratios of the probability distributions (given the null and alternate hypotheses) of winner votes $K_j$ in round $j$.
+Note that these distributions are not conditioned on $k_{j-1}$.
+Thus they may be viewed as a weighted average of various distributions, each conditioned on a possible value of $K_{j-1}$.
+The proof of \Minerva's risk-limiting property relies on the fact that, at every stopping point, the tail of the probability distribution function given $H_0$ is at most $\alpha$ times that given $H_a$.
+However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if the next round size is the same for all values of $K_{j-1}$.
+If not, this average pdf will not correctly represent the pdf of $K_j$.
+For example, if some values of $K_{j-1}$ would result in larger next round sizes, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller round sizes.
+
+For the \Providence audit we test the tail ratios of the distributions conditioned on $k_{j-1}$ and multiplied by the weighting factor.
+That is, we test this tail ratio separately for each weighted distribution corresponding to each possible value of $K_{j-1}$.
+Because of this, we no longer require that all values of $k_{j-1}$ should result in the same next round size.
+This results in a great deal of flexibility in multiple round audits.
+
+\subsection{Definition}
+\label{sec:prov_def}
 \begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
@@
 
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution. We now we prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.
+Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence considerably more computationally efficient.
+
+\subsection{Proof of Risk-Limiting Property}
+\label{sec:proof}
+We now prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.
 
@@
 $\mathcal{A}=\text{Correct}$ (the audit stops) only when
-$$
-\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq
+\begin{equation*}
+\begin{aligned}
+\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq\\
 \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
 \ge \frac{1}{\alpha}.
-$$
+\end{aligned}
+\end{equation*}
 %Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
@@
 
-\subsection{Resistance against an adversary choosing round sizes}
-I don't feel very comfortable with how to word/present this argument... I have not yet made an attempt at doing so.
+\subsection{Resistance to an adversary choosing round sizes}
+\label{sec:adversary}
+Filip TBD

### 2022-10-11 providence/sections/sims.tex
some corrections

--- 2022-05-16 providence/sections/sims.tex
@@
 We use simulations to provide additional evidence for theoretical claims and gain insight into audit behavior.
-As in \cite{sims}, we use margins from the 2020 US Presidential election, state-wide pairwise margins between the leading two candidates of 5\% or more.
-Narrower margins are computationally expensive, especially for the simulations with an underyling tie which quickly increase in sample size.
+As in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins between the leading two candidates of 5\% or more.
+Narrower margins are computationally expensive, especially for the simulations with an underyling tie, which, by design, have a low probability of stopping and hence require all rounds and quickly increase in sample size.
 We use the simulator in the R2B2 software library\cite{r2b2}.
-We perform $10000=10^4$ trials per margin for both an underyling announced outcomes and an underyling tie.
-
-The simulations with an underlying tie give an estimate of maximum risk as shown in Figure~\ref{fig:prov-risk}. For all margins, the estimated risks were less than the risk limit.
-
-Simulations with the underlying ballot distribution as announced provide insight into stopping probability and number of ballots drawn.
-Figure~\ref{fig:prov-sprob} shows that the stoping probabilities over the first rounds are near and slightly above $90\%$ as expected, since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
-Figure~\ref{fig:prov-asn} shows that the probability of stopping as a function of number of ballots sampled.
-Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
-As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-Note that a difference in twice as many ballots could be negligible in a contest with a wide margin.
-In a contest with a narrow margin (in the 2020 US Presidential election, several states had margins less than $3\%$) the difference in number of ballots sampled could be weeks of work.
-Section~\ref{sec:workload} discusses workload in more depth.
-%TODO add a more specific number of states for small margin
+For each margin, for each hypothesis, we perform $10^4$ trials.
+RISK LIMIT.
+NUMBER OF ROUNDS.
+STOPPING PROB.
 
@@
 \includegraphics[width=.5\textwidth]{prov_risk.png}
-\caption{The proportion of simulated audits with an underlying tie that stopped in one of the five simulated rounds.}
+\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds AT WHAT STOPPING PROB?
+RISK LIMIT?).
+This value is an estimate of the maximum risk of the \Providence audit.}
 \label{fig:prov-risk}
@@
 
+In the simulations of \Providence audits of a tied election, the fraction of audits that stop, as shown in Figure~\ref{fig:prov-risk}, is an estimate of maximum risk.
+For all margins, this estimated maximum risk is less than the risk limit, supporting the claim that \Providence is risk-limiting.
+
 \begin{figure}
 \includegraphics[width=.5\textwidth]{prov_sprob.png}
-\caption{The proportion of simulated audits with an margin as announced that stopped in one of the five simulated rounds.}
+\caption{The fraction of simulated \Providence audits of the election as announced that stopped for each round.
+This value is an estimate of the stopping probability conditioned on the sample of the previous round.
+The average fraction for rounds 1, 2, and 3 is $89.96\%$, $90.52\%$, and $90.98\%$ respectively.
+We show only the first three rounds since so few audits make it to rounds 4 and 5.}
 \label{fig:prov-sprob}
@@
 
+Simulations of audits of the election as announced provide insight into stopping probability and number of ballots drawn when the election is as announced.
+We wish the stopping probability to be as predicted, and the number of ballots drawn to be small.
+Figure~\ref{fig:prov-sprob} shows that the stopping probabilities over the first rounds are near and slightly above $90\%$ as expected since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
+
+Figure~\ref{fig:prov-asn} plots the probability of stopping as a function of the number of ballots sampled.
+Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
+As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work.
+% Section~\ref{sec:workload} discusses workload in more depth.
+
 \begin{figure}
 \includegraphics[width=.5\textwidth]{prov_asn.png}
-\caption{For all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.}
+\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.}
 \label{fig:prov-asn}

### 2022-10-11 providence/sections/workload.tex

@@
 Thus the time and personhours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
-Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading in the sense that the true winner recieves has fewer votest than some other candidate according to the tally of the sampled ballots.
+Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading: the true winner has fewer votes in the audit sample than some other candidate.
 On the other hand, a one-and-done audit may draw more ballots than are necessary; a more efficient round schedule could require less effort and time pre-certification.
@@
 Under this model we show how optimal round schedules can be chosen.
-We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like desired probability of a nonmisleading result.
+We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like maximum allowed probability of a misleading audit sample.
 
 As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $5.3\%$ between the two candidates with the most votes.
-Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds. 
+Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) WHAT ARE Eb AND Er?
+is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
 %A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
 %We will use the more standard approach of simulations to give an example here.
-Therefore we use the typical approach of simulations, again with risk-limit $10\%$.
+Therefore we use the typical approach of simulations, again with risk limit $10\%$.
 We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$.
-That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again give a probability of stopping $p$ in the second round.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is possible analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trials for each value of $p$, assuming the announced results are correct.
+That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again have a probability of stopping $p$ in the second round.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is possible analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the announced results are correct.
 

### 2022-10-11 providence/sections/pilot.tex
correct names in table, thanks filip
misleading sequences for SO BRAVO

@@
 The Rhode Island Board of Elections performed a pilot audit in Providence 
-in February 2022. The contest audited was a single yes-or-no question 
-%on School Construction and Renovation Projects
-from the November 2021 special election. 
-The question had announced margin $25.67\%$.
-The audit used risk-limit $10\%$.
+in February 2022. The contest audited was a single yes-or-no question in the November 202 1 election: Portsmouth's
+Issue 1, "School Construction and Renovation Projects". The question had announced margin $25.67\%$ and the audit used risk-limit $10\%$.
+
 A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected
@@
 %\diagbox[dir=NW]{First \\Round \\Size}{RLA}
-ballots& \rotatebox{45}{\Providence} & \rotatebox{45}{\Minerva} & \rotatebox{45}{EoR \BRAVO} & \rotatebox{45}{SO \BRAVO} \\
+ballots& \rotatebox{45}{\Providence} & \rotatebox{45}{\Minerva} & \rotatebox{45}{SO \BRAVO} & \rotatebox{45}{EoR \BRAVO} \\
 \hline
@@
 
-TODO: Add examples of how the audits perform for various hypothetical round schedules.
-I wait to do this until I'm done with the workload estimates since the examples here should be chosen to motivate that section.
+Note that the risk measures shown in Table~\ref{tab:pilot-risk} imply that, for the sample obtained in the pilot audit, an EoR \BRAVO audit would not have been able to stop in the first round, despite the large round size.
+Further, if the risk limit had been $0.05$ instead of $0.10$, SO \BRAVO also would have required moving on to a second round.
+
+We can use simulations to better understand typical audit behavior for the margin of this pilot audit and contextualize the results we obtained in the pilot.
+We run $10^4$ trial audits for several stopping probabilities $p$.
+Each round size is chosen to give a probability of stopping $p$ assuming the announced tally and given the results of previous rounds.
+We use the same $10\%$ risk limit and margin of $25.57\%$.
+Figure~\ref{fig:pilot_sims} shows the average number of ballots sampled for each value of $p$.
+
+Section~\ref{sec:workload} addresses more appropriate measures of workload

### 2022-10-11 providence/sections/workload.tex
misleading sequences for SO BRAVO

@@
 There are $\binom{n}{k}$ possible sequences of ballots which can lead to such a sample.
-Given the value of $k$, however, the particular sequence of the sample contains no additional information about whether the sample is more likely under one the alternative or null hypotheses.
+Given a value of $k$, however, the particular sequence of the sample that led to that value of $k$ contains no additional information about whether the sample is more likely under one the alternative or null hypotheses.
+That is to say, $\Pr[K=k|H_a]$ and $\Pr[K=k|H_0]$ have the same value regardless of the sequence.
 That said, an SO \BRAVO RLA will not stop as a function of $n$ and $k$ but also as a function of the sequence.
 In particular, if the sequence of ballots is such that the standard \BRAVO stopping condition was met for some $n'<n$ and corresponding $k'<n$, the audit will stop, even if by the end of the sequence the values $k$ and $n$ no longer meet the \BRAVO condition.
+We refer to such sequences which stop under SO \BRAVO, but not under EoR \BRAVO, \emph{misleading sequences}.
 To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result.
 The misleading nature of such stoppages is the note we are making.
+This is another case where election officials might have difficulty explaining the misleading situation to the public.
 
-It is easy to use our simulations to see how often this occurs.
+It is easy to use our simulations to see how often this occurs by checking whether the final cumulative sample of each SO \BRAVO trial meets the EoR \BRAVO stopping condition and counting those which do not.
+Figure~\ref{so_misleading} shows the proportion of simulated SO \BRAVO audits that stopped with a \emph{misleading sequence}.
+Unlike the more general \emph{misleading sample} discussed so far, these \emph{misleading sequences} are unique to SO \BRAVO audits, and Figure~\ref{so_misleading} only shows the proportion of audits that stopped with a \emph{misleading sequence}; additional SO \BRAVO audits also contained \emph{misleading samples}.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{so_misleading.png}
+\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
+\label{fig:avg_bals}
+\end{figure}

### 2022-10-11 providence/sections/pilot.tex
more pilot discussion

@@
 We use the same $10\%$ risk limit and margin of $25.57\%$.
-Figure~\ref{fig:pilot_sims} shows the average number of ballots sampled for each value of $p$.
 
-Section~\ref{sec:workload} addresses more appropriate measures of workload
+Figure~\ref{fig:pilot_sims} shows the average number of ballots sampled for each value of $p$ in the simulations.
+The vertical line denotes the stopping probability of the first round size chosen in the pilot of $140$ ballots.
+The average number of ballots for such a high stopping probability round schedule is high.
+blah blah based on plot and some numbers illustrating that point...
+Of course, average number of ballots is not the only metric that matters when planning an audit.
+For this pilot audit, extensive planning of the round schedule was not necessary because the margin was high enough the relatively few ballots were needed even to achieve the high probability of stopping.
+In Section~\ref{sec:workload} we consider a larger, state-wide contest where selecting the round schedule has more significant implications.
+Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.

### 2022-10-11 providence/sections/pilot.tex
providence numbers

@@
 Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{pilot_sims.png}
+\caption{The total number of ballots sampled on average in our simulations of audits using the same contest parameters and risk limit as the pilot audit.
+The average total ballots are shown for various round schedules for which we ran simulations, parameterized by $p$ the conditional stopping probability used to select each round size.}
+\label{fig:pilot_sims}
+\end{figure}

### 2022-10-11 providence/sections/pilot.tex

@@
 Of course, average number of ballots is not the only metric that matters when planning an audit.
-For this pilot audit, extensive planning of the round schedule was not necessary because the margin was high enough the relatively few ballots were needed even to achieve the high probability of stopping.
-In Section~\ref{sec:workload} we consider a larger, state-wide contest where selecting the round schedule has more significant implications.
-Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.
 
@@
 \end{figure}
+
+For this pilot audit, extensive planning of the round schedule was not necessary because the margin was high enough the relatively few ballots were needed even to achieve the high probability of stopping.
+In Section~\ref{sec:workload} we consider a larger, state-wide contest in Virginia where selecting the round schedule has more significant implications.
+Virginia also currently uses ballot polling RLAs whereas Rhode Island primarily uses batch comparison RLAs.
+Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.

### 2022-10-11 providence/sections/sims.tex

@@
 For each margin, for each hypothesis, we perform $10^4$ trials.
-RISK LIMIT.
-NUMBER OF ROUNDS.
-STOPPING PROB.
+All trials use a $10\%$ risk limit, a maximum of $5$ rounds, and a conditional stopping probability of $0.90$ in each round.
+That is, each subsequent round size is selected to be large enough, assuming the announced tally is correct and given the tally of previous rounds, to give a $0.90$ probability of stopping in the current round.
 
@@
 \includegraphics[width=.5\textwidth]{prov_risk.png}
-\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds AT WHAT STOPPING PROB?
-RISK LIMIT?).
+\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds at a $10\%$ risk limit).
 This value is an estimate of the maximum risk of the \Providence audit.}

### 2022-10-11 providence/sections/adversary.tex
rla proof

% \subsection{Motivation for considering adaptive adversary}
% \fpo{(old notes)}

\Minerva was proven to be risk-limiting audit for pre-commited round schedule.
It is not clear that \Minerva is still risk-limiting if an adversary can select
adaptively round schedule as the audit goes.
Problem: does stopping condition holds for adaptive choice of round sizes?

\begin{definition}
Let $d \in \{0, a\}$ stand for the null or the alternative hypothesis respectively.
Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner at the end of the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,} 
\[
s_{i, d, [n_1, \ldots, n_r]} = P\left[K = i | (n_1, \ldots, n_r), H_d\right]. 
%  = s_{i, d, [n_r]},
\]
For $r = 1$ 
\[
 s_{i, d, [n_r]} = \mathsf{Bin}(n_r, p_d, i).
\]
\end{definition}

\begin{example}[Adaptive round size selection] 
Let us assume a malicious auditor with the following strategy for \Minerva audit:
\begin{itemize}
 \item select round size $n_2$ if $k = k^* = k^{p_a, p_0, \alpha}_{min, 1} - 1$,
 \item select round size $n'_2$ for $k \leq k^{p_a, p_0, \alpha}_{min, 1} - 2$.
\end{itemize}
This leads to the situation that computed values of $r_{i, d}$ are not correct anymore \textit{i.e.,} value $s_{k^*}$ is missing for probabilities that go into $n'_2$.

\[
 s^{k^*, up}_{i, d, [n_1, n_{2}]} = \sum_{j = x}^{\min\{i, k^{p_a, p_0, \alpha}_{min, 1}-1\}} Bin(n_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
\]

\[
 s^{k^*, down}_{i, d, [n_1, n'_{2}]} = \sum_{j = \max\{0, i - (n'_2 - n_1)\}}^{k^*-1} Bin(n'_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
\]

One would need to to prove that for $n_2, n'_2$ and for any choice of $k^*$ the following holds:
\[
\sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*, up}_{i, 0, [n_1, n_2]} \le \alpha \sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*}_{i, A, [n_1, n_2]}.
\]
\end{example}

\begin{definition}[Strong RLA]
 An audit $\mathcal{A}$ is a Strong Risk Limiting Audit with risk limit $\alpha$ iff for
 all strategies of selecting round schedule and for sample $X$ 
 \[
  \Pr\left[\mathcal{A}(X) = \text{Correct} | H_0\right] \leq \alpha.
 \]

\end{definition}

\begin{lemma}
\Providence is Strong RLA. 
\end{lemma}
\begin{proof}
 The property comes as a conclusion of Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
\end{proof}

% \newpage

% \subsection{Round extensions}

% The end of a round of size $n_1$, with $k^{p_a, p_0, \alpha}_{min, 1}$ and observed ballots for
% the winner $k$.

% The size of the next round is selected to be $n_{r+1}$.
% The \Providence k-mins are computed as follows. 
% 
% (1) The probability distribution at the end of $(r+1)^{st}$ round is computed:
% % \[
% %  s_{i, d, [n_1, \ldots, n_{r+1}]} = \sum_{j = \max\{0, i - (n_{r+1} - n_r)\}}^{\min\{i, k^{p_a, p_0, \alpha}_{min, r}-1\}} Bin(n_{r+1} - n_r, p_d, i - j) \cdot s_{j, d, [n_1, \ldots, n_r]},
% % \]
% \[
%  s_{i, d, [n_1, \ldots, n_{r+1}]} = \sum_{j = \max\{0, i - (n_{r+1} - n_r)\}}^{\min\{i, k^{p_a, p_0, \alpha}_{min, r}-1\}} Bin(n_{r+1} - n_r, p_d, i - j) \cdot s_{j, d, [n_1, \ldots, n_r]},
% \]
% where $p_d = p_0 = \frac{1}{2}$ or $p_d = p_A$.
% 
% (2) The $k^{p_a, p_0, \alpha}_{min, r}$ is found as:
% \[
%  k^{p_a, p_0, \alpha}_{min, r} =  k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} = \min\left\{ k: \sum_{i = k}^{n_r} s_{i, 0, [n_1, \ldots, n_r]} \le \alpha \sum_{i = k}^{n_r} s_{i, A, [n_1, \ldots, n_r]} \right\}.
% \]

% 
% \begin{definition} 
% Let $[n_1, \ldots, n_r]$ be the round schedule of an audit that has not stopped by the round $r-1$. Let us define 
% \begin{small}
% \begin{equation}\label{eq:kMin}
% k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} =
%   \min\left\{k : \omega_r(k, k_{r-1},p_a,p_0,n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\}.
% %  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\}$
% \end{equation}
% \end{small}
% Then if $k_r \geq k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]}$ then the result of the audit is Correct (\textit{i.e.,} stopping condition in Definition~\ref{def:minervatwo} holds).
% \end{definition}
% 
% 
% 
% \begin{lemma}
% Let $[n_1, \ldots, n_{r-1}, n_r]$ be a round schedule for an execution of  \Providence audit that has not stopped
% in any of its first $r-1$ rounds (\textit{i.e.,} for every $j = 1, \ldots, r-1$:
% $k_j < k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_j]}$), then: 
% 
% \[ 
% k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_{r-1}, n_r]} = k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_{r-1}, n_r]}.
% \]
% \end{lemma}
% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
% \begin{proof}
% Let $k_{r-1}$ denote the number of ballots drawn for the declared winner up to the round $r-1$ (out of $n_{r-1}$ sampled ballots). The stopping decision for the round $r$ is made as follows:
% 
% \[
%  k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
% \]
% \[
%   =  k^{p_a, p_0, \alpha, k_{r-1}}_{min, 2, [n_{r-1}, n_r]}
% \]
% 
% \end{proof}
% Then, from (\ref{eq:prov}) one can write it as:
% 
% % \[
% %     \begin{aligned}
% %     \omega_{j}(k, k_{r-1}, p_a, p_0, n_{r}, n_{r-1})
% %     \triangleq\\
% %     \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1})
% %     \end{aligned}
% % \]
% \begin{small}
% \[
% \begin{aligned}
%  k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
%  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\} =\\
%  \min\left\{k :
% \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}} \frac{
% \sum_{i = k}^{n_r} s_{i - k_{r-1}, a, [n_r - n_{r-1}]}
% }{
% \sum_{i = k}^{n_r} s_{i - k_{r-1}, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\}
%  \end{aligned}
%  \]
% \end{small}
% 
% \fpo{should be added: for $k > k_{r-1}$ but it seems to be obvious (from monotonicity of k-min)}
% % \[
% % \sigma(k_{r-1},p_a,p_0,n_{r-1}) = \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}}
% % \]
% 
% % \begin{small}
% % \begin{equation}\label{eq:kMinLong}
% % k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
% %  \min\left\{k :
% % \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}} \frac{
% % \sum_{i = k}^{n_r} s_{i - k_{r-1}, a, [n_r - n_{r-1}]}
% % }{
% % \sum_{i = k}^{n_r} s_{i - k_{r-1}, 0, [n_r - n_{r-1}]}
% % }\geq \frac{1}{\alpha}  \right\}
% % \end{equation}
% % \end{small}
% %  = \min\left\{ k: \sum_{i = k}^{n_r} s_{i, 0, [n_1, \ldots, n_r]} \le \alpha \sum_{i = k}^{n_r} s_{i, A, [n_1, \ldots, n_r]} \right\}.
% % \]
% 
% Now, one can rewrite the enumerator and the denominator as  follows ($d \in \{a, 0\}$):
% 
% \begin{small}
% \[
% \begin{aligned}
%  s_{k_{r-1}, d, n_{r-1}} \sum_{i = k}^{n_r} s_{i - k_{r-1}, d, [n_r - n_{r-1}]} =\\
%  \mathsf{Bin}(n_{r-1}, p_d, k_{r-1}) \sum_{i = k}^{n_r} 
%  \mathsf{Bin}({n_r - n_{r-1}, d, i - k_{r-1}}) =\\
%  {n_{r-1} \choose k_{r-1}} p_d^{k_{r-1}} p_d^{n_{r-1} - k_{r-1}}
%  \sum_{i = k}^{n_r} 
%  {n_r - n_{r-1} \choose i - k_{r-1}} p_d^{i - k_{r-1}} p_d^{n_r - n_{r-1} - (i - k_{r-1})} =\\
%   \sum_{i = k}^{n_r} {n_{r-1} \choose k_{r-1}}
%  {n_r - n_{r-1} \choose i - k_{r-1}} p_d^{i} p_d^{n_r - i} =
%  \sum_{i = k}^{n_r} {n_{r} \choose i}
%  p_d^{i} p_d^{n_r - i} = \sum_{i = k}^{n_r} s_{i, d, [n_r]}.
%  \end{aligned}
% \]
% \end{small}
% 
% \begin{small}
% \[
% k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
%  \min\left\{k :
% \frac{
% \sum_{i = k}^{n_r} s_{i, a, [n_r]}
% }{
% \sum_{i = k}^{n_r} s_{i, 0, [n_r]}
% }\geq \frac{1}{\alpha}  \right\}
% = k^{p_a, p_0, \alpha}_{min, 1, [n_r]}
% \]
% \end{small}
% \fpo{There is no need to put $r$ in $k_{min}$ definition $[n_1, \ldots, n_r]$ is enough}
% 
% \end{proof}

% \fpo{Corollary 1: For RO-Bravo you need to pay attention.}
% 
% \fpo{Corollary 2: For \Providence you can (1) be late for an audit (round schedule does not
% matter much); (2) you can be distracted: only the end-of-round matters}

% \begin{lemma}
%  I
% \end{lemma}

% \begin{theorem}
%  \Providence is risk limiting for adversarial choices of round schedules.
% \end{theorem}
% \begin{proof}
% 
% \end{proof}

\subsection{Efficiency}
\begin{lemma}
% For any round schedule %$[n_1, \ldots, n_r]$
For any risk-limit $\alpha \in (0, 1)$, for any margin
and for any round schedule $[n_1, \ldots, n_r]$, 
 \Providence RLA is more efficient than \R \BRAVO.
\end{lemma}
\begin{proof}
 Let $[n_1, \ldots, n_r]$ be a round schedule, and  lets assume that \R \BRAVO audit stopped exactly at
 round $r$, 
 after observing $k_1, \ldots, k_r$ ballots for the announced winner in each round respectively. This means that
 the following conditions hold:
 \begin{enumerate}
  \item for all $j = 1, \ldots, r-1$: 
%   $k_j < k^{p_a, p_0, \alpha}_{min, \BRAVO, n_j}$,
  $\frac{P[{K_j = k_{j}} | H_a, n_j]}{P[K_j = {k_{j}} | H_0, n_j]} < \frac{1}{\alpha}$,
  \item for $j = r$ it holds: % $k_r \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}.$
  $\frac{P[{K_r = k_{r}} | H_a, n_r]}{P[K_r = {k_{r}} | H_0, n_r]} \geq \frac{1}{\alpha}$,
 \end{enumerate}

 From Lemma~\ref{lemma:tau1_increasing} 
 \[
 \tau_1(k, p_a, p_0, n_r) = 
  \frac{P[K_r = {k} | H_a, n_r]}{P[K_r = {k} | H_0, n_r]}
 \]
is a stricly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
% \textit{i.e.,}
% \begin{equation}\label{eq_incr}
% % \frac{1}{\alpha} \leq
%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}. 
% \end{equation}

If \BRAVO test is satisfied for $k_r$ then:

% From the properties of binomial distribution we obtain:
% \[
%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
% \]
\[
 \frac{1}{\alpha} \leq \sigma(k_{r}, p_a, p_0, n_{r}) =
\]
\[
 = \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \sigma(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1})  \leq
\]
\[
 \leq^{(*)} \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \tau_1(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1}) =
\]
\[
 = \omega_r(k_r, k_{r-1}, p_a, p_0, n_r, n_{r-1}).
\]

Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.

% 
% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
% 
% 
%  \[
%  k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_r]}  
%   = \min\left\{k :
% \frac{
% s_{k_{r-1}, a, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% }{
% s_{k_{r-1}, 0, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]
% \[
%   = \min\left\{k :
% \frac{
% s_{k_{r-1}, a, [n_{r-1}]}
% \left(
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% + \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% \right)
% }{
% s_{k_{r-1}, 0, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]

%  \[
%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
%         = \min\left\{k :
% \frac{
% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]

% 
%  
%  Now, let us consider $j$-{th} round for $j = 1, \ldots, r$. We are going to show that
%  $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.
%  
%  From Lemma~\ref{eq:kMin} we have:
%  \[
%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
%   = \min\left\{k :
% \frac{
% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]
% 
% From the definition of $k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ we have that for every $k \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ it holds: $s_{k, a, [n_j]}/s_{k, 0, [n_j]} \geq 1/\alpha$ which means that:
% \[
% \frac{
% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}
% \]
% which means that $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.

\end{proof}

### 2022-10-11 providence/sections/prov.tex
fix extra line in prov proof
rla proof

@@
 Let $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence.
-Let $\bm{n_j}$ be the cumulative roundsizes used in this
+Let $\bm{n_j}$ be the cumulative round sizes used in this
 audit, with corresponding cumulative tallies of
@@
 $$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
 $$
 By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
-we see that there exists a $k_{min, j}\le k_j$ for which
+we see that there exists a $k_{min, j} = k_{min, j, n_{j-1}, n_j}^{p_a, p_0, k_{j-1}} \leq k_j$ 
+%$k_{min, j}\le k_j$ 
+for which
 $$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge
 $$
 $$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge 
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]} \ge
+\frac{1}{\alpha}
 $$
-$$
-\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
-$$
-
 %Equivalently,
@@
 %of $k$, we get
-Taking the sum over all possible audit histories, we get
+% Taking the sum over all possible audit histories, we get
+It can be rewritten as
 $$
-\frac{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\sum_{\bm{k_j}}\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
 $$
@@
 \label{sec:adversary}
-Filip TBD
+% Filip TBD
+\input{sections/adversary}

### 2022-10-11 providence/sections/related.tex
added bernhard sok

@@
 %Related Work
+Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
+
 The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn.
@@
 
-Where \BRAVO uses the ratio of the values of the conditional probability distributions, \Minerva uses the ratio of their \emph{tails}.
+Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
 

### 2022-10-11 providence/sections/prov.tex
modified summary of providence rationale

@@
 
-The \Minerva stopping condition tests the tail ratios of the probability distributions (given the null and alternate hypotheses) of winner votes $K_j$ in round $j$.
+The \Minerva stopping condition tests the tail ratios of the probability distributions\footnote{given the alternate and null hypotheses respectively} of winner votes $K_j$ in round $j$.
 Note that these distributions are not conditioned on $k_{j-1}$.
-Thus they may be viewed as a weighted average of various distributions, each conditioned on a possible value of $K_{j-1}$.
+Thus each may be viewed as a weighted average of various distributions given the corresponding hypothesis, each conditioned on a possible value of $K_{j-1}$, with the weights corresponding to the probabilities of drawing that value of $K_{j-1}$.
+
 The proof of \Minerva's risk-limiting property relies on the fact that, at every stopping point, the tail of the probability distribution function given $H_0$ is at most $\alpha$ times that given $H_a$.
-However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if the next round size is the same for all values of $K_{j-1}$.
+However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if $n_j$ is the same for all values of $K_{j-1}$.
 If not, this average pdf will not correctly represent the pdf of $K_j$.
-For example, if some values of $K_{j-1}$ would result in larger next round sizes, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller round sizes.
+For example, if some values of $K_{j-1}$ would lead to a choice of larger values of $n_j$, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller values of $n_j$.
 
@@
 That is, we test this tail ratio separately for each weighted distribution corresponding to each possible value of $K_{j-1}$.
-Because of this, we no longer require that all values of $k_{j-1}$ should result in the same next round size.
+Because of this, we no longer require that all values of $k_{j-1}$ should result in the same values of $n_j$.
 This results in a great deal of flexibility in multiple round audits.

### 2022-10-11 providence/sections/workload.tex
typo

@@
 
-\textbf{Misleading SO Risk Measures.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
+\textbf{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
 

### 2022-10-11 providence/sections/prov.tex
notes on proofs in the pdf

@@
 
+\subsection{Motivation}
+\label{sec:prov_motivation}
 The \Minerva stopping condition tests the tail ratios of the probability distributions\footnote{given the alternate and null hypotheses respectively} of winner votes $K_j$ in round $j$.
 Note that these distributions are not conditioned on $k_{j-1}$.
-Thus each may be viewed as a weighted average of various distributions given the corresponding hypothesis, each conditioned on a possible value of $K_{j-1}$, with the weights corresponding to the probabilities of drawing that value of $K_{j-1}$.
+Thus each may be viewed as a weighted average of the corresponding conditional distributions---each conditioned on a possible value of $K_{j-1}$, with the weights corresponding to the probabilities of drawing that value of $K_{j-1}$.
+(See \cite{usenix_minerva} for more details).
 
@@
 However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if $n_j$ is the same for all values of $K_{j-1}$.
-If not, this average pdf will not correctly represent the pdf of $K_j$.
-For example, if some values of $K_{j-1}$ would lead to a choice of larger values of $n_j$, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller values of $n_j$.
+If not, the average distribution will not correctly represent the distribution of $K_j$.
+
+For example, if the current sample is used to determine a next round size of a certain fixed stopping probability, small values of $K_{j-1}$ would be followed by larger values of $n_j$.
+In this case, the pdf for the small values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller values of $n_j$.
+This is the reason why the round schedule is predetermined for \Minerva, in order to ensure in an enforceable manner that all values of $k_{j-1}$ lead to the same value of $n_j$ if the audit does not stop in round $j-1$.
 
@@
 That is, we test this tail ratio separately for each weighted distribution corresponding to each possible value of $K_{j-1}$.
-Because of this, we no longer require that all values of $k_{j-1}$ should result in the same values of $n_j$.
+Because of this, we no longer require that all values of $k_{j-1}$ should be followed by the same values of $n_j$ if the audit does not stop in round $j-1$.
 This results in a great deal of flexibility in multiple round audits.
@@
 
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence considerably more computationally efficient.
+Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence computationally considerably more efficient.
 

### 2022-10-11 providence/sections/workload.tex

@@
 As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $5.3\%$ between the two candidates with the most votes.
-Analytical approximation of the expected audit behavior ($E_b$ and $E_r$) WHAT ARE Eb AND Er?
-is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
+Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
 %A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
@@
 Therefore we use the typical approach of simulations, again with risk limit $10\%$.
+We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website including irrelevant ballots.
 We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$.
 That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again have a probability of stopping $p$ in the second round.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is possible analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the announced results are correct.
+Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each sub-audit between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that sub-audit.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the announced results are correct.
 

### 2022-10-11 providence/sections/adversary.tex
small word changes in new proof

@@
 
-\Minerva was proven to be risk-limiting audit for pre-commited round schedule.
-It is not clear that \Minerva is still risk-limiting if an adversary can select
-adaptively round schedule as the audit goes.
-Problem: does stopping condition holds for adaptive choice of round sizes?
+\Minerva was proven to be risk-limiting audit for a predetermined round schedule.
+As explained in section \ref{sec:prov_motivation}, it is not clear that \Minerva is risk-limiting if an adversary can 
+adaptively select the round schedule as the audit proceeds.
+In this section we prove that \Providence does not have this problem, and is risk-limiting even when the adversary can choose next round sizes based on knowledge of the current sample.
 
@@
 \[
-s_{i, d, [n_1, \ldots, n_r]} = P\left[K = i | (n_1, \ldots, n_r), H_d\right]. 
+s_{i, d, [n_1, \ldots, n_r]} = P\left[K_r = i | (n_1, \ldots, n_r), H_d\right]. 
 %  = s_{i, d, [n_r]},

### 2022-10-11 providence/sections/sims.tex
asn comment

@@
 \includegraphics[width=.5\textwidth]{prov_asn.png}
-\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.}
+\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
+The average sample number (ASN) for \B \BRAVO is included for context.}
 \label{fig:prov-asn}

### 2022-10-11 providence/sections/workload.tex
asn comment

@@
 
-After drawing $n>1$ ballots in a round, some number $k$ of them are votes for the announced winner.
+After drawing a cumulative $n>1$ ballots in a round, some number $k$ of them are votes for the announced winner.
 There are $\binom{n}{k}$ possible sequences of ballots which can lead to such a sample.
-Given a value of $k$, however, the particular sequence of the sample that led to that value of $k$ contains no additional information about whether the sample is more likely under one the alternative or null hypotheses.
+Given a value of $k$, however, the particular sequence of the sample that led to that value of $k$ contains no additional information about whether the sample is more likely under the alternative or null hypotheses.
 That is to say, $\Pr[K=k|H_a]$ and $\Pr[K=k|H_0]$ have the same value regardless of the sequence.
-That said, an SO \BRAVO RLA will not stop as a function of $n$ and $k$ but also as a function of the sequence.
+Despite this, the SO \BRAVO RLA stopping condition is not just a function of $n$ and $k$ but also a function of the sequence, the selection order.
 In particular, if the sequence of ballots is such that the standard \BRAVO stopping condition was met for some $n'<n$ and corresponding $k'<n$, the audit will stop, even if by the end of the sequence the values $k$ and $n$ no longer meet the \BRAVO condition.
-We refer to such sequences which stop under SO \BRAVO, but not under EoR \BRAVO, \emph{misleading sequences}.
-To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result.
+We refer to such sequences which stop under SO \BRAVO, but not under EoR \BRAVO, as \emph{misleading sequences}.
+To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result\cite{wald}.
 The misleading nature of such stoppages is the note we are making.
 This is another case where election officials might have difficulty explaining the misleading situation to the public.
+
+As an example, consider
+
+Now recall the pilot \Providence RLA performed in Providence, Rhode Island described in Section~\ref{sec:pilot}. The sample drawn...
 

### 2022-10-11 providence/sections/workload.tex
more

@@
 \includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
-\caption{The optimal average workload of the audits we simulated using workload function \ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence}
+\caption{The optimal average workload of the audits we simulated using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence.
+Similar to Figure~\ref{optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case.
+If the time for a single ballot is $75$ seconds, then $c_p=50$ corresponds to over an hour of extra time to sample a ballot from a new container.}
 \label{fig:optimal_workload_precinct_workload_ratio}

### 2022-10-11 providence/sections/workload.tex
add distinct ballots footnote

@@
 \noindent\textbf{Average total ballots.} 
-The simplest workload models are a function of just the total number of a ballots sampled.
-Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
+The simplest workload models are a function of just the total number of a ballots sampled.\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is insignificant\cite{athena}.
+Modifying the model we discuss here to account for total distinct ballots is easy.} Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
 Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the \Providence values.

### 2022-10-11 providence/sections/pilot.tex
added pilot selection order

@@
 Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.
+
+For the sake of analysis, the selection order of the ballots sampled during the pilot was also recorded.
+Figure~\ref{fig:pilot_sequence} shows the cumulative tally of winner ballots after each new ballot in the selection order is added to the sample.
+We observe two interesting phenomena in this particular sample's selection order.
+First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
+It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this where the condition is met early in the sample but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
+Secondly, after the first $11$ ballots were drawn, only $5$ of them were for the announced winner, giving misleading evidence due to a too-small sample size.
+Both these ideas are addressed more thoroughly in Section~\ref{sec:workload}.
+%first draft, needed to be shorter: ...stopping condition for the cumulative tally at the end of the round would see that the \BRAVO condition is not met (and would be correct that an EoR \BRAVO audit would require a second round for this sample).
+On the other hand, an SO \BRAVO audit does stop already in the first round because the cumulative sample met the \BRAVO stopping condition at some point before all the ballots had been added to the sample.
+In Figure~\ref{fig:pilot_sequence}, this is seen when the observed sample (orange) line touches (or goes above) the minimum winner ballots (blue) line which gives the \BRAVO stopping condition (i.e.
+a \BRAVO condition is met whenever the number of winner ballots in the sample is at least the number indicated by that blue line).
+While relatively rare, there also possible cases of selection order where the SO \BRAVO condition is met very early but then bad luck is had thereafter and the ultimate cumulative sample at the end of the round gives very poor evidence that the announced tally is correct.
+There could even be fewer ballots for the winner than the loser.
+Such cases might be difficult for an election official to explain to the public, especially when the evidence at the end of the round provides poor support for the alternative hypothesis.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} In Section~\ref{sec:workload} we address such cases more thoroughly.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{pilot_sequence.png}
+\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner found in the sample is shown.}
+\label{fig:pilot_sequence}
+\end{figure}

### 2022-10-11 providence/sections/workload.tex
added pilot selection order

@@
 \includegraphics[width=.5\textwidth]{misleading_limits.png}
-\caption{The proportion of simulated \Providence audits that had a \emph{misleading sample} in any round.}
+\caption{The proportion of simulated \Providence audits for the Virginia contest parameters that had a \emph{misleading sample} in any round.}
 \label{fig:misleading}
+\end{figure}
+
+We observe a similar behavior in our simulations of audits on the contest from the pilot audit.
+Figure~\ref{fig:pilot_misleading} shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
+Despite the large difference in margin ($\sim 5\%$ in Virginia and $\sim 25\%$ in the pilot) we still observe that a \emph{misleading limit} of $0.01$ is first achieved at roughly $p=0.8$ and $0.001$ at $p=0.95$.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{pilot_misleading_prop.png}
+\caption{The proportion of simulated \Providence audits for the pilot audit parameters that had a \emph{misleading sample} in any round.}
+\label{fig:pilot_misleading}
 \end{figure}

### 2022-10-11 providence/sections/adversary.tex
fix: changes

@@
 \Minerva was proven to be risk-limiting audit for a predetermined round schedule.
-As explained in section \ref{sec:prov_motivation}, it is not clear that \Minerva is risk-limiting if an adversary can 
+As explained in Section~\ref{sec:prov_motivation}, it is not clear that \Minerva is risk-limiting if an adversary can 
 adaptively select the round schedule as the audit proceeds.
@@
 Let $d \in \{0, a\}$ stand for the null or the alternative hypothesis respectively.
-Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner at the end of the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,}
+Let $s_{i, d, (n_1, \ldots, n_j)}$ be the probability of observing $i$ ballots for the winner at the end of the $j^{th}$ round of an audit with round schedule $(n_1, \ldots, n_j)$ conditioned on hypothesis $H_d$ \textit{i.e.,}
 \[
-s_{i, d, [n_1, \ldots, n_r]} = P\left[K_r = i | (n_1, \ldots, n_r), H_d\right]. 
+s_{i, d, (n_1, \ldots, n_j)} = \Pr\left[K_j = i | H_d, n_1, \ldots, n_j\right]. 
 %  = s_{i, d, [n_r]},
 \]
-For $r = 1$ 
+For $j = 1$ 
 \[
- s_{i, d, [n_r]} = \mathsf{Bin}(n_r, p_d, i).
+ s_{i, d, (n_j)} = \mathsf{Bin}(n_j, p_d, i).
 \]
@@
 \[
- s^{k^*, up}_{i, d, [n_1, n_{2}]} = \sum_{j = x}^{\min\{i, k^{p_a, p_0, \alpha}_{min, 1}-1\}} Bin(n_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
+ s^{k^*, up}_{i, d, (n_1, n_{2})} = \sum_{j = x}^{\min\{i, k^{p_a, p_0, \alpha}_{min, 1}-1\}} Bin(n_{2} - n_1, p_d, i - j) \cdot s_{j, d, (n_1)},
 \]
@@
 \[
- s^{k^*, down}_{i, d, [n_1, n'_{2}]} = \sum_{j = \max\{0, i - (n'_2 - n_1)\}}^{k^*-1} Bin(n'_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
+ s^{k^*, down}_{i, d, (n_1, n'_{2})} = \sum_{j = \max\{0, i - (n'_2 - n_1)\}}^{k^*-1} Bin(n'_{2} - n_1, p_d, i - j) \cdot s_{j, d, (n_1)},
 \]
@@
 \[
-\sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*, up}_{i, 0, [n_1, n_2]} \le \alpha \sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*}_{i, A, [n_1, n_2]}.
+\sum_{i = k^{p_a, p_0, \alpha}_{min, 2, (n_1, n_2)}}^{n_2} s^{k^*, up}_{i, 0, [n_1, n_2]} \le \alpha \sum_{i = k^{p_a, p_0, \alpha}_{min, 2, (n_1, n_2)}}^{n_2} s^{k^*}_{i, a, (n_1, n_2)}.
 \]
@@
 
-\begin{definition}[Strong RLA]
+\begin{lemma}\label{lemma:markov}
+Let $(n_1, \ldots, n_{j-1}, n_j)$ be a round schedule for an execution of  \Providence audit that has not stopped
+in any of its first $j-1$ rounds (\textit{i.e.,} for every $i = 1, \ldots, j-1$:
+$k_i < k^{p_a, p_0, \alpha}_{min, j, (n_1, \ldots, n_j)}$), then:
+
+\[ 
+k^{p_a, p_0, \alpha, k_{j-1}}_{min, (n_1, \ldots, n_{j-1}, n_j)} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, (n_{j-1}, n_j)}.
+\]
+\end{lemma}
+Proof of Lemma~\ref{lemma:markov} is in the Appendix.
+
+\begin{definition}[Strategy-immune RLA]
  An audit $\mathcal{A}$ is a Strong Risk Limiting Audit with risk limit $\alpha$ iff for
- all strategies of selecting round schedule and for sample $X$ 
+ all strategies of selecting round schedule and for a sample $X$ 
  \[
@@
 \begin{lemma}
-\Providence is Strong RLA. 
+\Providence is Strategy-immune RLA. 
 \end{lemma}

### 2022-10-11 providence/sections/appendix.tex
fix: changes

@@
     \label{lemma:minerva2_kmin_exists}
-    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a $k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
+    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a 
+    $k_{min, j, n_{j-1}, n_j}^{p_a, p_0, k_{j-1}} = 
+    k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
 \end{lemma}
@@
 \end{proof}
+
+\begin{definition} 
+Let $[n_1, \ldots, n_r]$ be the round schedule of an audit that has not stopped by the round $r-1$. Let us define 
+\begin{small}
+\begin{equation}\label{eq:kMin}
+k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} =
+  \min\left\{k : \omega_r(k, k_{r-1},p_a,p_0,n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\}.
+%  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\}$
+\end{equation}
+\end{small}
+Then if $k_r \geq k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]}$ then the result of the audit is Correct (\textit{i.e.,} stopping condition in Definition~\ref{def:minervatwo} holds).
+\end{definition}
+
+The following lemma shows a Markov-like property of \Providence audit (\textit{i.e.,}
+for an audit that has not stopped in the first $r-1$ rounds, results of the round $r-1$ matter: cumulative sample size $n_{j-1}$ and the number of ballots for the winner $k_{r-1}$).
+
+\fpo{if we want to have lemmas in appendix formulated, we would need to set the lemma counter by hand}
+\setcounter{lemma}{0}
+\begin{lemma}
+Let $[n_1, \ldots, n_{r-1}, n_r]$ be a round schedule for an execution of  \Providence audit that has not stopped
+in any of its first $r-1$ rounds (\textit{i.e.,} for every $j = 1, \ldots, r-1$:
+$k_j < k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_j]}$), then:
+
+\[ 
+k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_{r-1}, n_r]} = k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_{r-1}, n_r]}.
+\]
+\end{lemma}
+% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
+\begin{proof}
+Let $k_{r-1}$ denote the number of ballots drawn for the declared winner up to the round $r-1$ (out of $n_{r-1}$ sampled ballots). The stopping decision for the round $r$ is made as follows:
+
+\[
+ k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
+\]
+\[
+  =  k^{p_a, p_0, \alpha, k_{r-1}}_{min, 2, [n_{r-1}, n_r]}
+\]
+
+\end{proof}

### 2022-10-12 providence/sections/adversary.tex
add constant to real time formula, begin working on minerva vs prov example
changed property to Strategy-Proof

--- 2022-10-11 providence/sections/adversary.tex
@@
 Let $d \in \{0, a\}$ stand for the null or the alternative hypothesis respectively.
-Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner at the end of the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,}
+Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner after drawing the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,}
 \[
@@
 \begin{itemize}
- \item select round size $n_2$ if $k = k^* = k^{p_a, p_0, \alpha}_{min, 1} - 1$,
- \item select round size $n'_2$ for $k \leq k^{p_a, p_0, \alpha}_{min, 1} - 2$.
+ \item select second round size $n_2$ if $k_{1} = k^* = k^{p_a, p_0, \alpha}_{min, 1} - 1$,
+ \item select round size $n'_2$ for $k_1 \leq k^{p_a, p_0, \alpha}_{min, 1} - 2$.
 \end{itemize}
-This leads to the situation that computed values of $r_{i, d}$ are not correct anymore \textit{i.e.,} value $s_{k^*}$ is missing for probabilities that go into $n'_2$.
+This leads to the situation where the true value of $s_{i, d, [n_1, n_2]}$ is missing any contributions from $K_1=k^*$.
 
@@
 
-\begin{definition}[Strong RLA]
- An audit $\mathcal{A}$ is a Strong Risk Limiting Audit with risk limit $\alpha$ iff for
+\begin{definition}[Strategy-Proof RLA]
+ An audit $\mathcal{A}$ is a Strategy-Proof Risk Limiting Audit with risk limit $\alpha$ iff for
  all strategies of selecting round schedule and for sample $X$ 
@@
 \begin{lemma}
-\Providence is Strong RLA. 
+\Providence is a Strategy-Proof RLA. 
 \end{lemma}
 \begin{proof}
- The property comes as a conclusion of Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
+This property follow from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
 \end{proof}
+
+% To see the practical implication of this result, we consider a toy example: an RLA of a two-candidate contest with margin $10\%$ and risk limit $10\%$. 
+% Suppose a first round of $100$ ballots is drawn and $60$ ballots in the sample are votes for the announced winner. Since $\tau_1(k,p_a,p_0,\ldots$...
 

### 2022-10-12 providence/sections/workload.tex
add constant to real time formula, begin working on minerva vs prov example
misleading sequences

@@
 \begin{equation}
-T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p
+T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p + C
 \label{eq:real_time}
@@
 In a random sample, a true loser may recieve more votes than the true winner.
-This happens more often when the sample sizes are small.
+This happens more often when the sample sizes are small, like for a hypothetical first round size of $11$ in the pilot audit, as seen in Figure~\ref{fig:pilot_sequence}.
 In these RLAs, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
@@
 
-As an example, consider
+Recall from Section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}.
+In particular, the audit passed with an SO \BRAVO risk measure of $5.41\%$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $36.6\%$.
 
-Now recall the pilot \Providence RLA performed in Providence, Rhode Island described in Section~\ref{sec:pilot}. The sample drawn...
-
-It is easy to use our simulations to see how often this occurs by checking whether the final cumulative sample of each SO \BRAVO trial meets the EoR \BRAVO stopping condition and counting those which do not.
+It is easy to use our simulations to see how often SO \BRAVO \emph{misleading sequences} occur by checking whether the final cumulative sample of each SO \BRAVO trial meets the EoR \BRAVO stopping condition and counting those which do not.
 Figure~\ref{so_misleading} shows the proportion of simulated SO \BRAVO audits that stopped with a \emph{misleading sequence}.

### 2022-10-12 providence/sections/pilot.tex
more

@@
 The average number of ballots for such a high stopping probability round schedule is high.
-blah blah based on plot and some numbers illustrating that point...
+For the first round size used in the pilot there was an approximately $95\%$ probability of stopping, and consequently in Figure~\ref{fig:pilot_sims}, this corresponds to a higher average total ballots sampled than lower values of $p$.
 Of course, average number of ballots is not the only metric that matters when planning an audit.

### 2022-10-12 providence/sections/Motivation_old.tex
RLA proof

--- 2022-10-11 providence/sections/prov.tex
@@
-In this section we motivate and introduce the stopping condition of \Providence.
-
 \subsection{Motivation}
@@
 This results in a great deal of flexibility in multiple round audits.
-
-\subsection{Definition}
-\label{sec:prov_def}
-\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
-    \label{def:minervatwo}
-    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
-$$
-\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
-        % & \\
-        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
-        % & \\
-        Undetermined ~~else \\
-    \end{array}
-    \right .
-$$
-where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
-\begin{equation}
-    \begin{aligned}
-    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})
-    \triangleq\\
-    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
-    \end{aligned}
-\end{equation}
-\end{definition}
-
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence computationally considerably more efficient.
-
-\subsection{Proof of Risk-Limiting Property}
-\label{sec:proof}
-We now prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.
-
-\begin{theorem}
-\label{thm:minerva2_is_rla_new}
-An $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence audit is an
-$\alpha$-RLA.
-\end{theorem}
-\begin{proof}
-Let $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence.
-Let $\bm{n_j}$ be the cumulative round sizes used in this
-audit, with corresponding cumulative tallies of
-ballots for the reported winner $\bm{k_j}$.
-For round $j=1$, by Definitions \ref{def:minervatwo}
-and \ref{def:minerva_ratio}, we see that
-the $\mathcal{A}=\text{Correct}$ (the audit stops) only when
-$$
-\tau_1(k_{1},p_a,p_0,n_1)\\
-=\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
-\ge \frac{1}{\alpha}.
-$$
-By Lemma \ref{lemma:minerva2_kmin_exists}, we see that this
-is equivalent to the following:
-$$
-\frac{Pr[K_{1} \geq k_{min,1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{min, 1} \mid H_0, n_1]}
-\ge \frac{1}{\alpha}.
-$$
-
-For any round $j\ge 2$, by Definition \ref{def:minervatwo}
-and Lemma \ref{lemma:minerva2_kmin_exists},
-$\mathcal{A}=\text{Correct}$ (the audit stops) only when
-\begin{equation*}
-\begin{aligned}
-\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq\\
-\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
-\ge \frac{1}{\alpha}.
-\end{aligned}
-\end{equation*}
-%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
-%could do this for simpler notation^
-By Lemma \ref{lemma:any_ratio_is_sigma_simple}
-and Definition \ref{def:minerva_ratio}, this is equivalent to
-$$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
-$$
-By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
-we see that there exists a $k_{min, j} = k_{min, j, n_{j-1}, n_j}^{p_a, p_0, k_{j-1}} \leq k_j$ 
-%$k_{min, j}\le k_j$ 
-for which
-$$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge
-$$
-$$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]} \ge
-\frac{1}{\alpha}
-$$
-%Equivalently,
-%$$
-%\frac{Pr[K_{j} \ge k_{min, j}\wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_a]}{Pr[K_{j}\ge k_{min, j} \wedge K_{j-1} = k_{j-1} \mid n_{j}^*, k_{j-2}^*, H_0]}\ge \frac{1}{\alpha}.
-%$$
-%Taking the sum over all possible rounds and corresponding preceding values
-%of $k$, we get
-% Taking the sum over all possible audit histories, we get
-It can be rewritten as
-$$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
-$$
-Finally, because the total probability of stopping the audit under
-the alternative hypothesis is less than 1, we get
-$$
-\frac{Pr[\mathcal{A}=\text{Correct} \mid H_a]}{Pr[\mathcal{A}=\text{Correct} \mid H_0]}\ge \frac{1}{\alpha}
-$$
-$$
-Pr[\mathcal{A}=\text{Correct} \mid H_0]
-\le
-Pr[\mathcal{A}=\text{Correct} \mid H_a] \cdot \alpha
-\le
-\alpha.
-$$
-\end{proof}
-
-\subsection{Resistance to an adversary choosing round sizes}
-\label{sec:adversary}
-% Filip TBD
-\input{sections/adversary}

### 2022-10-12 providence/sections/adversary.tex
moved adversary example

@@
 
-\Minerva was proven to be risk-limiting audit for a predetermined round schedule.
-As explained in section \ref{sec:prov_motivation}, it is not clear that \Minerva is risk-limiting if an adversary can 
+\Minerva was proven to be a risk-limiting audit for a predetermined round schedule.
+As explained earlier, it is not clear that \Minerva is risk-limiting if an adversary can 
 adaptively select the round schedule as the audit proceeds.
 In this section we prove that \Providence does not have this problem, and is risk-limiting even when the adversary can choose next round sizes based on knowledge of the current sample.
-
-\begin{definition}
-Let $d \in \{0, a\}$ stand for the null or the alternative hypothesis respectively.
-Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner after drawing the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,}
-\[
-s_{i, d, [n_1, \ldots, n_r]} = P\left[K_r = i | (n_1, \ldots, n_r), H_d\right]. 
-%  = s_{i, d, [n_r]},
-\]
-For $r = 1$ 
-\[
- s_{i, d, [n_r]} = \mathsf{Bin}(n_r, p_d, i).
-\]
-\end{definition}
-
-\begin{example}[Adaptive round size selection] 
-Let us assume a malicious auditor with the following strategy for \Minerva audit:
-\begin{itemize}
- \item select second round size $n_2$ if $k_{1} = k^* = k^{p_a, p_0, \alpha}_{min, 1} - 1$,
- \item select round size $n'_2$ for $k_1 \leq k^{p_a, p_0, \alpha}_{min, 1} - 2$.
-\end{itemize}
-This leads to the situation where the true value of $s_{i, d, [n_1, n_2]}$ is missing any contributions from $K_1=k^*$.
-
-\[
- s^{k^*, up}_{i, d, [n_1, n_{2}]} = \sum_{j = x}^{\min\{i, k^{p_a, p_0, \alpha}_{min, 1}-1\}} Bin(n_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
-\]
-
-\[
- s^{k^*, down}_{i, d, [n_1, n'_{2}]} = \sum_{j = \max\{0, i - (n'_2 - n_1)\}}^{k^*-1} Bin(n'_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
-\]
-
-One would need to to prove that for $n_2, n'_2$ and for any choice of $k^*$ the following holds:
-\[
-\sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*, up}_{i, 0, [n_1, n_2]} \le \alpha \sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*}_{i, A, [n_1, n_2]}.
-\]
-\end{example}
 
@@
 \begin{proof}
-This property follow from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
+This property follows from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
+Note that, as described in section \ref{sec:proof}, the proof of the risk-limiting nature of the audit does not rely on round sizes $n_j$ being identical for all values of $k_{j-1}$.
 \end{proof}

### 2022-10-12 providence/sections/adversary.tex
example

--- 2022-10-12 providence/sections/adversary.tex
@@
 
-% To see the practical implication of this result, we consider a toy example: an RLA of a two-candidate contest with margin $10\%$ and risk limit $10\%$. 
-% Suppose a first round of $100$ ballots is drawn and $60$ ballots in the sample are votes for the announced winner. Since $\tau_1(k,p_a,p_0,\ldots$...
+To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $1\%$ and risk limit $10\%$. 
+Suppose we would like to achieve a $90\%$ conditional stopping probability in each round of the audit.
+For \Providence, we can compute a new round size for each round based on the previous samples, but for \Minerva we need to commit to a round schedule in advance.
+So for \Minerva we choose a first round size which gives a $90\%$ probability of stopping, and then we set the subsequent marginal round sizes to be $1.5$ times each previous round size.
+This multiplier of $1.5$ is known to give a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
+Audit software ARLO, used by many states, commits to this round schedule by default.
+Both audits therefore begin with first round size of $17,272$ to give a $90\%$ probability of stopping, and both will stop in the first round if the sample contains at least $8725$ ballots for the winner.
+We now consider two cases for which the audit proceeds to a second round.
+
+In one case there are $8724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
+In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $95.4\%$.
+The \Providence audit samples more than $9000$ fewer ballots with a round size of $34,078$, achieving the desired $90\%$ probability of stopping.
+
+In a less lucky sample, the winner recieves $8637$ ballots, only a bit more than the loser recieves.
+In the \Minerva audit, we again have to use a second round size of $43180$, but now this round size only achieves a $72.7\%$ probability of stopping, significantly less than the desired $90\%$.
+Again, the \Providence audit can scale up the second round size according to the first sample and achieve the desired $90\%$ probability of stopping with $58,007$ ballots.
 

### 2022-10-12 providence/sections/prov.tex
broke down an equation

--- 2022-10-11 providence/sections/prov.tex
@@
-In this section we motivate and introduce the stopping condition of \Providence.
+In this section we introduce the stopping condition of \Providence and prove some properties.
 
-The \Minerva stopping condition tests the tail ratios of the probability distributions (given the null and alternate hypotheses) of winner votes $K_j$ in round $j$.
-Note that these distributions are not conditioned on $k_{j-1}$.
-Thus they may be viewed as a weighted average of various distributions, each conditioned on a possible value of $K_{j-1}$.
-The proof of \Minerva's risk-limiting property relies on the fact that, at every stopping point, the tail of the probability distribution function given $H_0$ is at most $\alpha$ times that given $H_a$.
-However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if the next round size is the same for all values of $K_{j-1}$.
-If not, this average pdf will not correctly represent the pdf of $K_j$.
-For example, if some values of $K_{j-1}$ would result in larger next round sizes, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller round sizes.
+Recall that the proof that the \Minerva audit is risk-limiting assumes that the round schedule of \Minerva is predetermined and that, in particular, an adversarial auditor cannot determine the next round size after drawing a sample.
+This presents difficulties because a non-adversarial election official might want to draw a small next round if the current sample comes close to satisfying the risk limit.
+Because the \Minerva round size is predetermined, however, the election official would be required to draw a larger round size than necessary for the sample.
+Conversely, if the current sample is not at all close to satisfying the risk limit, it would be advantageous to draw a larger round than the predetermined round size.
 
-For the \Providence audit we test the tail ratios of the distributions conditioned on $k_{j-1}$ and multiplied by the weighting factor.
-That is, we test this tail ratio separately for each weighted distribution corresponding to each possible value of $K_{j-1}$.
-Because of this, we no longer require that all values of $k_{j-1}$ should result in the same next round size.
-This results in a great deal of flexibility in multiple round audits.
+The \Providence audit is risk-limiting even if an adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.
 
@@
 
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence considerably more computationally efficient.
+Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence computationally considerably more efficient.
 
-\subsection{Proof of Risk-Limiting Property}
+\subsection{Risk-Limiting Property: Proof}
 \label{sec:proof}
@@
 It can be rewritten as
-$$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot \sum_{{k} = k_{min, j}}^{n_j} Pr[K_{j} = k \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
-$$
-Finally, because the total probability of stopping the audit under
-the alternative hypothesis is less than 1, we get
-$$
-\frac{Pr[\mathcal{A}=\text{Correct} \mid H_a]}{Pr[\mathcal{A}=\text{Correct} \mid H_0]}\ge \frac{1}{\alpha}
-$$
+\begin{equation*}
+\begin{aligned}
+\sum_{{k} = k_{min, j}}^{n_j} Pr[(K_{j} , K_{j-1}) = (k, k_{j-1}) \mid H_0, n_{j-1}, n_{j}] \leq \\
+\alpha \sum_{{k} = k_{min, j}}^{n_j} Pr[(K_{j} , K_{j-1}) = (k, k_{j-1}) \mid H_a, n_{j-1}, n_{j}]
+\end{aligned}
+\end{equation*}
+The left hand side above is the probability of stopping in the $j^{th}$ round and $K_{j-1} = k_{j-1}$, given the null hypothesis, which is smaller than $alpha$ times the same probability given the alternate hypothesis.
+Summing both sides over all values of $K_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).
+Note here that the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$.
+When both sides of the inequality are further summed over all rounds, we get:
+
 $$
@@
 \le
-Pr[\mathcal{A}=\text{Correct} \mid H_a] \cdot \alpha
-\le
+\alpha Pr[\mathcal{A}=\text{Correct} \mid H_a]
+$$
+Finally, because the total probability of stopping the audit under
+the alternative hypothesis is not greater than 1, we get
+$$
+Pr[\mathcal{A}=\text{Correct} \mid H_0] \le
 \alpha.

### 2022-10-12 providence/sections/adversary.tex
more

--- 2022-10-12 providence/sections/adversary.tex
@@
 
-% To see the practical implication of this result, we consider a toy example: an RLA of a two-candidate contest with margin $10\%$ and risk limit $10\%$. 
-% Suppose a first round of $100$ ballots is drawn and $60$ ballots in the sample are votes for the announced winner. Since $\tau_1(k,p_a,p_0,\ldots$...
+To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $1\%$ and risk limit $10\%$. 
+Suppose we would like to achieve a $0.9$ conditional stopping probability in each round of the audit.
+For \Providence, we can compute a new round size for each round based on the previous samples, but for \Minerva we need to commit to a round schedule in advance.
+So for \Minerva we choose a first round size which gives a $0.9$ probability of stopping, and then we set the subsequent marginal round sizes to be $1.5$ times each previous round size.
+This multiplier of $1.5$ is known to give a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
+Audit software ARLO, used by many states, commits to this round schedule by default.
+Both audits therefore begin with first round size of $17,272$ to give a $0.9$ probability of stopping, and both will stop in the first round if the sample contains at least $8,725$ ballots for the winner.
+We now consider two cases for which the audit proceeds to a second round.
+
+In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
+In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $95.4\%$.
+The \Providence audit samples more than $9,000$ fewer ballots with a round size of $34,078$, achieving the desired $0.9$ probability of stopping.
+
+In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
+In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $72.7\%$ probability of stopping, significantly less than the desired $0.9$.
+Again, the \Providence audit can scale up the second round size according to the first sample and achieve the desired $0.9$ probability of stopping with $58,007$ ballots.
 

### 2022-10-12 providence/sections/prov.tex
appendix proofs completed

@@
 $$
+where $k_{min,1} = k^{p_a, p_0, \alpha, 0}_{min, 1, 0, n_1}$ (see Lemma \ref{lemma:minerva2_kmin_exists} and Definition \ref{def:kmin}).
 
@@
 and Lemma \ref{lemma:minerva2_kmin_exists},
-$\mathcal{A}=\text{Correct}$ (the audit stops) only when
+$\mathcal{A}=\text{Correct}$ (the audit stops) if and only if
 \begin{equation*}
@@
 By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
-we see that there exists a $k_{min, j} = k_{min, j, n_{j-1}, n_j}^{p_a, p_0, k_{j-1}} \leq k_j$ 
+we see that there exists a $k_{min, j} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  \leq k_j$ 
 %$k_{min, j}\le k_j$ 
@@
 % Taking the sum over all possible audit histories, we get
-It can be rewritten as
+The above may be rewritten as
 \begin{equation*}

### 2022-10-12 providence/sections/adversary.tex
typo

@@
 To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $1\%$ and risk limit $10\%$. 
-Suppose we would like to achieve a $0.9$ conditional stopping probability in each round of the audit.
+Suppose we wish to achieve a conditional stopping probability $0.9$ in each round of the audit.
 For \Providence, we can compute a new round size for each round based on the previous samples, but for \Minerva we need to commit to a round schedule in advance.
@@
 In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
-In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $95.4\%$.
+In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $.954$.
 The \Providence audit samples more than $9,000$ fewer ballots with a round size of $34,078$, achieving the desired $0.9$ probability of stopping.
@@
 In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
-In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $72.7\%$ probability of stopping, significantly less than the desired $0.9$.
+In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $0.727$ probability of stopping, significantly less than the desired $0.9$.
 Again, the \Providence audit can scale up the second round size according to the first sample and achieve the desired $0.9$ probability of stopping with $58,007$ ballots.

### 2022-10-12 providence/sections/adversary.tex
toy example

@@
 Suppose we wish to achieve a conditional stopping probability $0.9$ in each round of the audit.
-For \Providence, we can compute a new round size for each round based on the previous samples, but for \Minerva we need to commit to a round schedule in advance.
-So for \Minerva we choose a first round size which gives a $0.9$ probability of stopping, and then we set the subsequent marginal round sizes to be $1.5$ times each previous round size.
-This multiplier of $1.5$ is known to give a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
-Audit software ARLO, used by many states, commits to this round schedule by default.
-Both audits therefore begin with first round size of $17,272$ to give a $0.9$ probability of stopping, and both will stop in the first round if the sample contains at least $8,725$ ballots for the winner.
+For \Providence, we can compute a new round size for each round based on the previous samples.
+For \Minerva, however, we would have a predetermined round schedule.
+We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by most states performing an RLA), which is is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
+This multiplier of $1.5$ is known to give, over a wide range of margins, a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
+
+Both the audits of our toy example therefore begin with a first round size of $17,272$ with a $0.9$ probability of stopping, and both will stop in the first round if the sample contains at least $8,725$ ballots for the winner.
 We now consider two cases for which the audit proceeds to a second round.
-
-In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
+\begin{description}
+\item In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
 In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $.954$.
 The \Providence audit samples more than $9,000$ fewer ballots with a round size of $34,078$, achieving the desired $0.9$ probability of stopping.
-
-In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
+\item In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
 In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $0.727$ probability of stopping, significantly less than the desired $0.9$.
 Again, the \Providence audit can scale up the second round size according to the first sample and achieve the desired $0.9$ probability of stopping with $58,007$ ballots.
+\end{description}
 
@@
 For any risk-limit $\alpha \in (0, 1)$, for any margin
-and for any round schedule $[n_1, \ldots, n_r]$, 
- \Providence RLA is more efficient than \R \BRAVO.
+and for any round schedule $[n_1, \ldots, n_j]$, 
+the \Providence RLA is more efficient than EoR \BRAVO.
 \end{lemma}
 \begin{proof}
- Let $[n_1, \ldots, n_r]$ be a round schedule, and  lets assume that \R \BRAVO audit stopped exactly at
- round $r$, 
- after observing $k_1, \ldots, k_r$ ballots for the announced winner in each round respectively. This means that
+ Let $[n_1, \ldots, n_j]$ be a round schedule, and lets assume that an EoR \BRAVO audit stopped exactly at
+ round $j$, 
+ after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively. This means that
  the following conditions hold:
  \begin{enumerate}
-  \item for all $j = 1, \ldots, r-1$: 
+  \item for all $i = 1, \ldots, j-1$: 
 %   $k_j < k^{p_a, p_0, \alpha}_{min, \BRAVO, n_j}$,
-  $\frac{P[{K_j = k_{j}} | H_a, n_j]}{P[K_j = {k_{j}} | H_0, n_j]} < \frac{1}{\alpha}$,
-  \item for $j = r$ it holds: % $k_r \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}.$
-  $\frac{P[{K_r = k_{r}} | H_a, n_r]}{P[K_r = {k_{r}} | H_0, n_r]} \geq \frac{1}{\alpha}$,
+  $\frac{P[{K_i = k_{i}} | H_a, n_i]}{P[K_i = {k_{i}} | H_0, n_i]} < \frac{1}{\alpha}$,
+  \item for $i = j$: % $k_r \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}.$
+  $\frac{P[{K_j = k_{j}} | H_a, n_j]}{P[K_j = {k_{j}} | H_0, n_j]} \geq \frac{1}{\alpha}$,
  \end{enumerate}
@@
  \[
- \tau_1(k, p_a, p_0, n_r) = 
-  \frac{P[K_r = {k} | H_a, n_r]}{P[K_r = {k} | H_0, n_r]}
+ \tau_1(k, p_a, p_0, n_j) = 
+  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
  \]
-is a stricly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
+is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
 % \textit{i.e.,}
@@
 
-If \BRAVO test is satisfied for $k_r$ then:
+If the \BRAVO test is satisfied for $k_j$ then:
 
@@
 \[
- \frac{1}{\alpha} \leq \sigma(k_{r}, p_a, p_0, n_{r}) =
+ \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) =
 \]
 \[
- = \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \sigma(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1})  \leq
+ = \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \sigma(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1})  
 \]
 \[
- \leq^{(*)} \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \tau_1(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1}) =
+ \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) =
 \]
 \[
- = \omega_r(k_r, k_{r-1}, p_a, p_0, n_r, n_{r-1}).
+ = \omega_r(k_j, k_{j-1}, p_a, p_0, n_j, n_{j-1}).
 \]

### 2022-10-12 providence/sections/appendix.tex
toy example

--- 2022-05-16 providence/sections/appendix.tex
@@
 \begin{proof}
-From $0<p_0< p_a< 1$, we get
-$$
-\frac{p_a}{p_0}>1
-$$
-and
-$$
-1-p_0>1-p_a\implies \frac{1-p_0}{1-p_a}>1,
-$$
-and thus
-$$
-\frac{p_a(1-p_0)}{p_0(1-p_a)}>1.
-$$
-Now simply observe that
-$$
-\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\sigma(k, p_a, p_0, n)
-=\frac{p_a(1-p_0)}{p_0(1-p_a)}\cdot\frac{p_a^{k} (1-p_a)^{n-k}}{p_a^{k} (1-p_a)^{n-k}}
-$$$$
-=\frac{p_a^{k+1} (1-p_a)^{n-(k+1)}}{p_a^{k+1} (1-p_a)^{n-(k+1)}}
-=\sigma(k+1, p_a, p_0, n).
-$$
+See \cite[Lemma 4]{usenix_minerva}. 
 \end{proof}
@@
 \begin{proof}
-    Note that $z_i$ is a weighted average of the values of $\frac{a_j}{b_j}$ for $j \geq i$:
-    $$z_i = \sum_{j=i}^n y_j \frac{a_j}{b_j}$$  ~for~ $$y_j = \frac{b_j}{\sum _{j=i}^n b_j} > 0.$$
-    Further, $\sum _{j=i}^n y_j = 1$ and hence  $y_j \leq 1$ ~and~ $y_j = 1 \Leftrightarrow i=j=n$.
-    Observe that, because $\frac{a_i}{b_i}$ is monotone increasing,
-    $z_i \geq \frac{a_i}{b_i}$
-    with equality if and only if $i=n$.
-    Suppose $i < n$. Then
-    $$z_{i+1} \geq \frac{a_{i+1}}{b_{i+1}} > \frac{a_i}{b_i}, $$
-    and
-    $$z_i = y_i \frac{a_i}{b_i} + (1-y_i) z_{i+1} < z_{i+1}.$$
-    Thus $z_i$ is also monotone increasing.
+See \cite[Lemma 2]{usenix_minerva}. 
 \end{proof}
@@
     \label{lemma:tau1_increasing}
-For $0<p_0< p_a< 1$ and $n>0$, the ratio $\tau_1(k,p_a,p_0,n)$ is strictly increasing as a function $k$ for $0\le k\le n$.
+For $0<p_0< p_a< 1$ and $n>0$, the ratio $\tau_1(k,p_a,p_0,n)$ is strictly increasing as a function of $k$ for $0\le k\le n$.
 \end{lemma}
@@
     \label{lemma:minerva2_kmin_exists}
-    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a $k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
+    For $\mathcal{A}=(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence, there exists\\ a 
+    $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  = 
+    k_{min,j}(\Providence, p_a, p_0, k_{j-1}, n_{j-1}, n_j)$ such that $$\mathcal{A}(X_j)=\text{Correct}\iff k_j\ge k_{min,j}(\Providence,  \bm{n_j}, p_a, p_0).$$
 \end{lemma}
@@
 \end{proof}
+
+\begin{definition} 
+\label{def:kmin}
+Let $[n_1, \ldots, n_j]$ be the round schedule of an audit that has not stopped by the round $j-1$. Let us define 
+\begin{small}
+\begin{equation}\label{eq:kMin}
+k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  =
+  \min\left\{k : \omega_j(k, k_{j-1},p_a,p_0,n_j, n_{j-1}) \geq \frac{1}{\alpha}  \right\}.
+%  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\}$
+\end{equation}
+\end{small}
+\end{definition}
+As we have seen in Lemma \label{lemma:minerva2_kmin_exists}, such a value of $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}$ exists and $k_j \geq k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $ if and only if the result of the audit is Correct, (\textit{i.e.,} the stopping condition in Definition~\ref{def:minervatwo} holds).
+
+The following lemma shows a Markov-like property of \Providence audit (\textit{i.e.,}
+for an audit that has not stopped in the first $j-1$ rounds, only cumulative results of the round $j-1$ matter: cumulative sample size $n_{j-1}$ and the number of ballots for the winner $k_{j-1}$).
+
+\begin{lemma}\label{lemma:markov}
+Let $[n_1, \ldots, n_{j-1}, n_j]$ be a round schedule for an execution of  \Providence audit that has not stopped
+in any of its first $j-1$ rounds (\textit{i.e.,} for every $i = 1, \ldots, j-1$:
+$k_i < k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $), then:
+
+\[ 
+k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j} .
+\]
+\end{lemma}
+% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
+\begin{proof}
+Let $k_{j-1}$ denote the number of ballots drawn for the declared winner up to the round $j-1$ (out of $n_{j-1}$ sampled ballots). The stopping decision for the round $j$ is made as follows:
+
+\[
+ k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
+\]
+\[
+  =  k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j}  
+\]
+
+\end{proof}
+
+That is, the stopping condition is equivalent to that of a two round audit with the same cumulative votes for the winner and cumulative round sizes: the first round is of size $n_{j-1}$ and has $k_{j-1}$ votes for the winner, and the second (cumulative) round size is $n_j$ with $k_j$ (cumulative) votes for the winner.
+Compare this to the similar property for the $\Bravo$ stopping condition.

### 2022-10-12 providence/sections/prov.tex
similarities Minerva + Providence

@@
 
-Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence computationally considerably more efficient.
+Notice that \Providence requires the computation of $\tau_j$ for $j=1$ and no other values of $j$.
+The value of $\tau_1$ is simply the ratio of the the tails of the binomial distributions for the two hypotheses and can be fairly efficiently computed.
+The computation of $\tau_j$ for $j \geq 2$, as required in \Minerva, relies on the convolution of two probability distribution functions and is hence computationally considerably more expensive.
+
+Notice also that \Providence and \Minerva are identical for $j=1$.
 

### 2022-10-12 providence/sections/adversary.tex

@@
 \begin{proof}
- Let $[n_1, \ldots, n_j]$ be a round schedule, and lets assume that an EoR \BRAVO audit stopped exactly at
- round $j$, 
- after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively. This means that
- the following conditions hold:
- \begin{enumerate}
-  \item for all $i = 1, \ldots, j-1$: 
-%   $k_j < k^{p_a, p_0, \alpha}_{min, \BRAVO, n_j}$,
-  $\frac{P[{K_i = k_{i}} | H_a, n_i]}{P[K_i = {k_{i}} | H_0, n_i]} < \frac{1}{\alpha}$,
-  \item for $i = j$: % $k_r \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}.$
-  $\frac{P[{K_j = k_{j}} | H_a, n_j]}{P[K_j = {k_{j}} | H_0, n_j]} \geq \frac{1}{\alpha}$,
- \end{enumerate}
+Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
+That is, the EoR \BRAVO stopping condition is true:
+$$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
+To see the the \Providence stopping condition is fulfilled, we rewrite as 
+%$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
 
- From Lemma~\ref{lemma:tau1_increasing} 
- \[
- \tau_1(k, p_a, p_0, n_j) = 
-  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
- \]
-is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
+%% From Lemma~\ref{lemma:tau1_increasing} 
+% \[
+% \tau_1(k, p_a, p_0, n_j) = 
+%  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
+% \]
+%is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
 % \textit{i.e.,}
@@
 
-If the \BRAVO test is satisfied for $k_j$ then:
+%If the \BRAVO test is satisfied for $k_j$ then:
 
@@
 \[
- \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) =
+ \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) 
 \]
@@
 \[
- \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) =
+ \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) 
 \]
@@
 
-Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
+Where inequality $(*)$ follows from \cite[Theorem 6]{athena}.
+%Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
 %the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then

### 2022-10-12 providence/sections/pilot.tex
edits

--- 2022-10-11 providence/sections/pilot.tex
@@
 The Rhode Island Board of Elections performed a pilot audit in Providence 
-in February 2022. The contest audited was a single yes-or-no question in the November 202 1 election: Portsmouth's
-Issue 1, "School Construction and Renovation Projects". The question had announced margin $25.67\%$ and the audit used risk-limit $10\%$.
+in February 2022. The contest audited was a single yes-or-no question in the November 2021 election: Portsmouth's
+Issue 1, "School Construction and Renovation Projects". The question had a reported margin of $0.2567$ and the audit used a risk-limit of $0.10$.
 
-A first round size of $140$ ballots with large probability of stopping ($95\%$) was selected
-in order to give the potential for more interesting analysis afterwards; the large margin 
-made a first round with large stopping probability practical.
+A first round size of $140$ ballots with large probability of stopping ($0.95$) was selected.
 Selection order was tracked for the sake of analysis.
-As expected, the audit concluded in the first round with a \Providence risk of $4.18\%$.
-Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Minerva and \BRAVO (both EoR and SO).
+As expected, the audit concluded in the first round.
+The \Providence risk was $0.0418$.
+Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO).
 
@@
 \hline
-140 & \bf{4.18\%} & \bf{4.18\%} & \bf{5.41\%} & 36.6\% \\
+140 & \bf{0.0418} & \bf{0.0418} & \bf{0.0541} & 0.366 \\
 \hline
@@
 
-Note that the risk measures shown in Table~\ref{tab:pilot-risk} imply that, for the sample obtained in the pilot audit, an EoR \BRAVO audit would not have been able to stop in the first round, despite the large round size.
+Note that the risk measures shown in Table~\ref{tab:pilot-risks} imply that, for the sample obtained in the pilot audit, an EoR \BRAVO audit would not have stopped in the first round, despite the large round size.
 Further, if the risk limit had been $0.05$ instead of $0.10$, SO \BRAVO also would have required moving on to a second round.
@@
 Each round size is chosen to give a probability of stopping $p$ assuming the announced tally and given the results of previous rounds.
-We use the same $10\%$ risk limit and margin of $25.57\%$.
+We use the same $0.1$ risk limit and margin of $0.2557$.
 
 Figure~\ref{fig:pilot_sims} shows the average number of ballots sampled for each value of $p$ in the simulations.
-The vertical line denotes the stopping probability of the first round size chosen in the pilot of $140$ ballots.
-The average number of ballots for such a high stopping probability round schedule is high.
-blah blah based on plot and some numbers illustrating that point...
+The vertical line denotes the stopping probability of the first round size actually chosen in the pilot ($140$ ballots).
+The large value of $p$ corresponds to a large first round size and a corresponding large value of average number of ballots.
+In later sections we show why average number of ballots is not the only metric to optimize, and how large round sizes can be beneficial from the perspective of other important metrics.
+%The average number of ballots for such a high stopping probability round schedule is high.
+For the first round size used in the pilot there was an approximately $0.95$ probability of stopping, and consequently, in Figure~\ref{fig:pilot_sims}, this corresponds to a higher average total ballots sampled than lower values of $p$.
 Of course, average number of ballots is not the only metric that matters when planning an audit.
@@
 \includegraphics[width=.5\textwidth]{pilot_sims.png}
-\caption{The total number of ballots sampled on average in our simulations of audits using the same contest parameters and risk limit as the pilot audit.
-The average total ballots are shown for various round schedules for which we ran simulations, parameterized by $p$ the conditional stopping probability used to select each round size.}
+\caption{The total number of ballots sampled on average in our audit simulations using the same contest parameters and risk limit as the Rhode Island pilot.
+The average total ballots are shown for various round schedules, parameterized by $p$, the conditional stopping probability used to select each round size.}
 \label{fig:pilot_sims}
@@
 
-For this pilot audit, extensive planning of the round schedule was not necessary because the margin was high enough the relatively few ballots were needed even to achieve the high probability of stopping.
-In Section~\ref{sec:workload} we consider a larger, state-wide contest in Virginia where selecting the round schedule has more significant implications.
-Virginia also currently uses ballot polling RLAs whereas Rhode Island primarily uses batch comparison RLAs.
-Some of the ideas introduced in Section~\ref{sec:workload} we do apply to this pilot case as well.
+For this pilot audit, extensive planning of the round schedule was not necessary because the margin was large enough that relatively few ballots were needed to achieve the high probability of stopping.
+In Section~\ref{sec:workload} we consider a larger state-wide contest in Virginia, where selecting the round schedule has more significant implications.
+Virginia also currently uses ballot polling RLAs, whereas Rhode Island primarily uses batch comparison RLAs.
+Some of the ideas introduced in Section~\ref{sec:workload} provide a context for this pilot case as well.
 
@@
 We observe two interesting phenomena in this particular sample's selection order.
-First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
-It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this where the condition is met early in the sample but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
-Secondly, after the first $11$ ballots were drawn, only $5$ of them were for the announced winner, giving misleading evidence due to a too-small sample size.
+\begin{description}
+\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
+It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
+\item Second, only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have provided misleading evidence due to a too-small sample size. 
+\end{description}
 Both these ideas are addressed more thoroughly in Section~\ref{sec:workload}.

### 2022-10-12 providence/sections/sims.tex

@@
 % simulations
-We use simulations to provide additional evidence for theoretical claims and gain insight into audit behavior.
-As in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins between the leading two candidates of 5\% or more.
-Narrower margins are computationally expensive, especially for the simulations with an underyling tie, which, by design, have a low probability of stopping and hence require all rounds and quickly increase in sample size.
-We use the simulator in the R2B2 software library\cite{r2b2}.
+We use simulations to provide additional evidence for theoretical claims and to gain insight into audit behavior.
+As done in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins of 5\% or more between the two leading candidates.
+Narrower margins are computationally expensive, especially for the simulations of tied elections, which, by design, have a low probability of stopping and hence quickly increase in sample size.
+We use the simulator in the R2B2 software library\cite{r2b2_anon}.
 For each margin, for each hypothesis, we perform $10^4$ trials.
@@
 
-Figure~\ref{fig:prov-asn} plots the probability of stopping as a function of the number of ballots sampled.
+We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single contest as an example, Texas with margin $5.7\%$.
+We run an additional $10^4$ simulations for each of the three other audits assuming the announced margin for comparison.
+Both \BRAVO implementations use a conditional stopping probability of $0.9$ for each round, while \Minerva uses a first round size with stopping probability $0.9$ and a multiplier of $1.5$ to obtain subsequent round sizes.
+Figure~\ref{fig:prov-asn} shows the probability of stopping as a function of the number of ballots sampled, a plot similar to those presented in \cite{simulations}.
 Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
 As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work before a certification deadline.
 % Section~\ref{sec:workload} discusses workload in more depth.

### 2022-10-12 providence/sections/adversary.tex
Update adversary.tex

@@
 
-To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $1\%$ and risk limit $10\%$. 
+To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $0.01$ and risk limit $0.1$. 
 Suppose we wish to achieve a conditional stopping probability $0.9$ in each round of the audit.
@@
 $$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
-To see the the \Providence stopping condition is fulfilled, we rewrite as 
+To see the \Providence stopping condition is fulfilled, we rewrite as 
 %$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
@@
 
-Where inequality $(*)$ follows from \cite[Theorem 6]{athena}.
+Where inequality $(*)$ follows from \cite[Theorem 6]{athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
 %Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.

### 2022-10-12 providence/sections/intro.tex
,

@@
 On the other hand, they require the use of special election technology and are not always feasible.
-Ballot-polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system.
+Ballot polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system.
 What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).
@@
 
-A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins, because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
+\subsection{Open Questions in the Literature}
+A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
 \BRAVO, on the other hand, is not limited in this manner.
@@
 Election officials, on the other hand, greatly prefer drawing many ballots at once.
-This preference is likely because each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
-Further, there is an overhead to finding a storage box and unsealing it.
+This preference is likely due to the following.
+\begin{description}
+\item Firstly, each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
+\item Secondly, there is an overhead to finding a storage box and unsealing it.
 For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each).
 For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
-Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large.
+\item Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large.
+\end{description}
 Thus the workload of an audit is not simply a linear (or affine) function of the number of ballots drawn.

### 2022-10-12 providence/sections/sims.tex
this and that

@@
 % simulations
-We use simulations to provide additional evidence for theoretical claims and to gain insight into audit behavior.
-As done in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins of 5\% or more between the two leading candidates.
+We use simulations to provide additional evidence for our theoretical claims regarding \Providence and to gain insight into audit behavior.
+As done in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins of $0.05$ or larger between the two leading candidates.
 Narrower margins are computationally expensive, especially for the simulations of tied elections, which, by design, have a low probability of stopping and hence quickly increase in sample size.
 We use the simulator in the R2B2 software library\cite{r2b2_anon}.
-For each margin, for each hypothesis, we perform $10^4$ trials.
-All trials use a $10\%$ risk limit, a maximum of $5$ rounds, and a conditional stopping probability of $0.90$ in each round.
-That is, each subsequent round size is selected to be large enough, assuming the announced tally is correct and given the tally of previous rounds, to give a $0.90$ probability of stopping in the current round.
+For each margin, we perform $10^4$ \Providence audit trials each on a tied election (hypothesis $H_0$, the null hypothesis) and the election as reported (hypothesis $H_a$, the alternate hypothesis).
+All trials have risk limit $\alpha = 0.1$, a maximum of $5$ rounds, and a conditional stopping probability of $0.90$ in each round.
+That is, each next round size is selected to be large enough to give a $0.90$ conditional probability of stopping in that round, assuming the announced tally is correct and given the tally of previous rounds.
+We use a maximum of five rounds because virtually no audits would progress beyond five rounds given the large conditional probability of stopping.
 
@@
 \includegraphics[width=.5\textwidth]{prov_risk.png}
-\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds at a $10\%$ risk limit).
+\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds at a risk limit of $0.1$) as a function of contest margin.
 This value is an estimate of the maximum risk of the \Providence audit.}
@@
 \includegraphics[width=.5\textwidth]{prov_sprob.png}
-\caption{The fraction of simulated \Providence audits of the election as announced that stopped for each round.
+\caption{The fraction of simulated \Providence audits of the election as reported that stopped for each round as a function of margin.
 This value is an estimate of the stopping probability conditioned on the sample of the previous round.
-The average fraction for rounds 1, 2, and 3 is $89.96\%$, $90.52\%$, and $90.98\%$ respectively.
-We show only the first three rounds since so few audits make it to rounds 4 and 5.}
+The average fraction for rounds 1, 2, and 3 is $0.8996$, $0.9052$, and $0.9098$ respectively.
+We show only the first three rounds since so few audits make it to rounds 4 and 5 (of the order of $10^4 \times (0.1)^3$ and $10^4 \times (0.1)^4$ respectively).}
 \label{fig:prov-sprob}
@@
 
-Simulations of audits of the election as announced provide insight into stopping probability and number of ballots drawn when the election is as announced.
-We wish the stopping probability to be as predicted, and the number of ballots drawn to be small.
-Figure~\ref{fig:prov-sprob} shows that the stopping probabilities over the first rounds are near and slightly above $90\%$ as expected since our software chose round sizes to give at least a $90\%$ conditional stopping probability.
+Simulations of audits of the election as reported provide insight into stopping probability and number of ballots drawn when the election is as reported.
+Figure~\ref{fig:prov-sprob} shows that the stopping probabilities over the first rounds are near and slightly above $0.9$ as expected, since our software chose round sizes to give at least a $0.9$ conditional stopping probability.
+The values are not as tight around $0.9$ for later rounds because fewer audit trials make it to later rounds, and our experimental probability estimates are not as accurate.
 
-We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single contest as an example, Texas with margin $5.7\%$.
-We run an additional $10^4$ simulations for each of the three other audits assuming the announced margin for comparison.
+We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single margin as an example: the 2020 US Presidential election in the state of Texas, with margin $0.057$.
+We run an additional $10^4$ simulations for each of the three other audits on the same underlying election and on a tied election.
 Both \BRAVO implementations use a conditional stopping probability of $0.9$ for each round, while \Minerva uses a first round size with stopping probability $0.9$ and a multiplier of $1.5$ to obtain subsequent round sizes.
@@
 As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work before a certification deadline.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than $0.03$) the difference in number of ballots sampled could correspond to many days of work which would need to be completed before a certification deadline.
 % Section~\ref{sec:workload} discusses workload in more depth.
@@
 \includegraphics[width=.5\textwidth]{prov_asn.png}
-\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability for average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
+\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability as a function of the average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
 The average sample number (ASN) for \B \BRAVO is included for context.}

### 2022-10-12 providence/sections/sims.tex
newpage in sims

@@
 \includegraphics[width=.5\textwidth]{prov_asn.png}
-\caption{For the entire audit, consisting of all five rounds, the estimated stopping probability as a function of the average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
+\caption{For the entire audit, consisting of all five rounds, the fraction of simulated audits that stopped as a function of the average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
 The average sample number (ASN) for \B \BRAVO is included for context.}
@@
 \end{figure}
+
+\newpage

### 2022-10-12 providence/sections/pilot.tex
small changes

@@
 
-\begin{table}
+\begin{table}[h!]
 \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{pilot_sims.png}
-\caption{The total number of ballots sampled on average in our audit simulations using the same contest parameters and risk limit as the Rhode Island pilot.
-The average total ballots are shown for various round schedules, parameterized by $p$, the conditional stopping probability used to select each round size.}
+\caption{The total number of ballots sampled on average as a function of $p$, the conditional stopping probability used to select each round size.
+We use the same contest parameters and risk limit as the Rhode Island pilot.}
 \label{fig:pilot_sims}
@@
 It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
-\item Second, only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have provided misleading evidence due to a too-small sample size. 
+\item Second, only $5$ of the first $11$ ballots were for the announced winner.
+A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size.
 \end{description}

### 2022-10-12 providence/sections/prov.tex

@@
     \label{def:minervatwo}
-    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
+    For cumulative round size $n_j$ for round $j$ and a cumulative $k_j$ ballots for the reported winner found in round $j$, the \R \Providence stopping rule for the $j^{th}$ round is:
 $$

### 2022-10-12 providence/sections/intro.tex

@@
 We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections.
-Our implementation of \Providence is likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
+Our open-source implementations of \Providence and our audit planning tools are likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
 

### 2022-10-12 providence/sections/workload.tex

@@
 Thus the time and personhours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
-Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading: the true winner has fewer votes in the audit sample than some other candidate.
+Secondly, smaller first round sizes are not large enough to accurately capture the distribution of votes.
+There is a higher probability that the true winner has fewer votes in the audit sample than some other candidate.
 On the other hand, a one-and-done audit may draw more ballots than are necessary; a more efficient round schedule could require less effort and time pre-certification.
 To evaluate the quality of various round schedules, we construct a simple workload model.
-Under this model we show how optimal round schedules can be chosen.
+Using this model we show how optimal round schedules can be chosen.
 We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like maximum allowed probability of a misleading audit sample.
 
-As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $5.3\%$ between the two candidates with the most votes.
+As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $0.053$ between the two candidates with the most votes.
 Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
@@
 %We will use the more standard approach of simulations to give an example here.
-Therefore we use the typical approach of simulations, again with risk limit $10\%$.
-We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website including irrelevant ballots.
+Therefore we use the typical approach of simulations, again with risk limit $0.1$.
+We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website, including irrelevant ballots.
 We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$.
 That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again have a probability of stopping $p$ in the second round.
-Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each sub-audit between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that sub-audit.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the announced results are correct.
+Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each pairwise contest between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that pairwise contest.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the reported results are correct.
+
+Note that simulations of audits of tied elections are not necessary, as all the audits we are considering are risk-limiting and hence we already know the performance to expect when auditing a tied election, even when it is not reported as such.
 

### 2022-10-12 providence/sections/Motivation_old.tex

@@
 This results in a great deal of flexibility in multiple round audits.
+
+\begin{definition}
+Let $d \in \{0, a\}$ stand for the null or the alternative hypothesis respectively.
+Let $s_{i, d, [n_1, \ldots, n_r]}$ be the probability of observing $i$ ballots for the winner after drawing the $r^{th}$ round of an audit with round schedule $[n_1, \ldots, n_r]$ conditioned on hypothesis $H_d$ \textit{i.e.,}
+\[
+s_{i, d, [n_1, \ldots, n_r]} = P\left[K_r = i | (n_1, \ldots, n_r), H_d\right]. 
+%  = s_{i, d, [n_r]},
+\]
+For $r = 1$ 
+\[
+ s_{i, d, [n_r]} = \mathsf{Bin}(n_r, p_d, i).
+\]
+\end{definition}
+
+\begin{example}[Adaptive round size selection] 
+Let us assume a malicious auditor with the following strategy for \Minerva audit:
+\begin{itemize}
+ \item select second round size $n_2$ if $k_{1} = k^* = k^{p_a, p_0, \alpha}_{min, 1} - 1$,
+ \item select round size $n'_2$ for $k_1 \leq k^{p_a, p_0, \alpha}_{min, 1} - 2$.
+\end{itemize}
+This leads to the situation where the true value of $s_{i, d, [n_1, n_2]}$ is missing any contributions from $K_1=k^*$.
+
+\[
+ s^{k^*, up}_{i, d, [n_1, n_{2}]} = \sum_{j = x}^{\min\{i, k^{p_a, p_0, \alpha}_{min, 1}-1\}} Bin(n_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
+\]
+
+\[
+ s^{k^*, down}_{i, d, [n_1, n'_{2}]} = \sum_{j = \max\{0, i - (n'_2 - n_1)\}}^{k^*-1} Bin(n'_{2} - n_1, p_d, i - j) \cdot s_{j, d, [n_1]},
+\]
+
+One would need to to prove that for $n_2, n'_2$ and for any choice of $k^*$ the following holds:
+\[
+\sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*, up}_{i, 0, [n_1, n_2]} \le \alpha \sum_{i = k^{p_a, p_0, \alpha}_{min, 2, [n_1, n_2]}}^{n_r} s^{k^*}_{i, A, [n_1, n_2]}.
+\]
+\end{example}

### 2022-10-12 providence/sections/adversary.tex

@@
 
-To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $0.01$ and risk limit $0.1$. 
+To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $1\%$ and risk limit $10\%$. 
 Suppose we wish to achieve a conditional stopping probability $0.9$ in each round of the audit.

### 2022-10-12 providence/sections/appendix-copy.tex

--- 2022-10-11 providence/sections/appendix.tex
@@
 
-\fpo{if we want to have lemmas in appendix formulated, we would need to set the lemma counter by hand}
-\setcounter{lemma}{0}
-\begin{lemma}
+\begin{lemma}\label{lemma:markov}
 Let $[n_1, \ldots, n_{r-1}, n_r]$ be a round schedule for an execution of  \Providence audit that has not stopped

### 2022-10-12 providence/sections/intro.tex

--- 2022-10-12 providence/sections/intro.tex
@@
 \subsection{Open Questions in the Literature}
-A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
+A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins, because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
 \BRAVO, on the other hand, is not limited in this manner.

### 2022-10-12 providence/sections/sims.tex

--- 2022-10-12 providence/sections/sims.tex
@@
 We use simulations to provide additional evidence for theoretical claims and to gain insight into audit behavior.
-As done in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins of 5\% or more between the two leading candidates.
+As in \cite{simulations}, we use margins from the 2020 US Presidential election---state-wide pairwise margins of 5\% or more between the two leading candidates.
 Narrower margins are computationally expensive, especially for the simulations of tied elections, which, by design, have a low probability of stopping and hence quickly increase in sample size.
@@
 
-We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single contest as an example, Texas with margin $5.7\%$.
-We run an additional $10^4$ simulations for each of the three other audits assuming the announced margin for comparison.
-Both \BRAVO implementations use a conditional stopping probability of $0.9$ for each round, while \Minerva uses a first round size with stopping probability $0.9$ and a multiplier of $1.5$ to obtain subsequent round sizes.
-Figure~\ref{fig:prov-asn} shows the probability of stopping as a function of the number of ballots sampled, a plot similar to those presented in \cite{simulations}.
+Figure~\ref{fig:prov-asn} plots the probability of stopping as a function of the number of ballots sampled.
 Points above (higher probability of stopping) and to the left (fewer ballots) represent more efficient audits.
 As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work before a certification deadline.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins less than $3\%$) the difference in number of ballots sampled could correspond to many days of work.
 % Section~\ref{sec:workload} discusses workload in more depth.

### 2022-10-12 providence/sections/adversary.tex

--- 2022-10-12 providence/sections/adversary.tex
@@
 For \Minerva, however, we would have a predetermined round schedule.
-We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by most states performing an RLA), which is is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
+We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by many states performing an RLA), which is is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
 This multiplier of $1.5$ is known to give, over a wide range of margins, a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.

### 2022-10-12 providence/sections/appendix.tex

@@
 \end{definition}
-As we have seen in Lemma \label{lemma:minerva2_kmin_exists}, such a value of $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}$ exists and $k_j \geq k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $ if and only if the result of the audit is Correct, (\textit{i.e.,} the stopping condition in Definition~\ref{def:minervatwo} holds).
+As we have seen in Lemma \ref{lemma:minerva2_kmin_exists}, such a value of $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}$ exists and $k_j \geq k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $ if and only if the result of the audit is Correct, (\textit{i.e.,} the stopping condition in Definition~\ref{def:minervatwo} holds).
 

### 2022-10-12 providence/sections/workload.tex
some changes

@@
 Firstly, rounds have some overhead in both time and effort.
-Thus the time and personhours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
+Thus the time and person-hours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
 Secondly, smaller first round sizes are not large enough to accurately capture the distribution of votes.
@@
 Therefore we use the typical approach of simulations, again with risk limit $0.1$.
+
 We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website, including irrelevant ballots.
@@
 Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each pairwise contest between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that pairwise contest.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the reported results are correct.
+For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
+We run $10^4$ trial audits for each value of $p$, assuming the reported results are correct\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.}.
 
-Note that simulations of audits of tied elections are not necessary, as all the audits we are considering are risk-limiting and hence we already know the performance to expect when auditing a tied election, even when it is not reported as such.
+Note that simulations of audits of tied elections are not necessary, as all the audits we are considering are risk-limiting and hence we already know the performance to expect when auditing a tied election, even one not reported as such.
 
@@
 
-\noindent\textbf{Average total ballots.} 
-The simplest workload models are a function of just the total number of a ballots sampled.\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is insignificant\cite{athena}.
-Modifying the model we discuss here to account for total distinct ballots is easy.} Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
-Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the \Providence values.
-It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test in the case where each round is a single ballot.
+\subsubsection{Average total ballots.} 
+The simplest workload models are a function of just the total number of a ballots sampled.\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is insignificant\cite{arxiv_athena}.
+It is straightforward to modify the model we discuss here to account for total distinct ballots.} Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled as a function of $p$.
+\begin{figure}[h!]
+\includegraphics[width=.5\textwidth]{avg_bals.png}
+\caption{The total number of ballots sampled on average, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
+\label{fig:avg_bals}
+\end{figure}
+
+Figure~\ref{fig:avg_bals_ratio} provides the same number as a fraction of the \Providence values.
+It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test when each round corresponds to a single ballot.
 Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e.
@@
 In particular, the savings of \Providence become larger as $p$ increases; for $p=.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{avg_bals.png}
-\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
-\label{fig:avg_bals}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
-\caption{The total number of ballots sampled on average in our simulations given as a fraction of those sampled by \Providence, for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
+\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
 \label{fig:avg_bals_ratio}
@@
 
-\noindent\textbf{Round overhead.} 
+\subsubsection{Round overhead.} 
 It is clear that average number of ballots alone is an inadequate workload measure. 
@@
 
-In a US state-wide RLA, the state organizes the audit by determining the random sample and communicating with the counties, but election officials at the county level physically sample and inspect the ballots from their precincts.
+In a US state-wide RLA, the state organizes the audit by determining the random sample and communicating with the counties, but election officials at the county level physically sample and inspect the ballots after drawing them from secure storage boxes stored in county locations.
 Therefore each audit round requires some number of person-hours for set up and communication between state and county.
@@
 
-Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials for their estimates.
-For simplicity, (and without loss of generality), we assume the per ballot workload is one, $w_b=1$.
-Then the per round workload $w_r$ tells us the workload of a round as a number of ballots.
-We begin with $w_r=1000$ as a conservative example.
+Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials to represent, for example, the effort of constructing a ballot manifest.
+For simplicity, (and without loss of generality), we measure in multiples of the per ballot workload; that is, we assume it is one unit, $w_b=1$.
+A per round workload of $w_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload.
+We use $w_r=1000$ as a conservative example.
 That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
 Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload.
-This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
+This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
 As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.
@@
 \caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
-\label{fig:avg_bals}
+\label{fig:misleading}
 \end{figure}

### 2022-10-12 providence/sections/workload.tex

@@
 This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
-As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.
 
-\begin{figure}
+As shown in Figure~\ref{fig:with_round_workload}, average workloads first reduce as stopping probability increases; this is likely due to a decrease in the number of rounds.
+After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not increase much and the cost changes because of number of ballots drawn, which increases with round size.
+\Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $w_r=1000$.
+
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{with_round_workload.png}
 \caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$.
-Expected workload is found using Equation~\ref{eq:workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
+Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
 \label{fig:with_round_workload}
@@
 
-Importantly, this gives us a way to estimate the expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload.
+Importantly, this gives us a way to estimate the minimum expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload.
 For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.
+
 Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads.
-For very low round workloads, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less.
-On the other hand, for extremely larger values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values.
+For very low round workloads, the workload function approaches just the total number of ballots, and so workload is minimized by minimizing the number of ballots drawn, which corresponds to small round sizes, and we would expect all three audits to behave similarly, as ballot-by-ballot audits, with the smallest workload.
+On the other hand, for extremely large values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values, all corresponding to large round sizes in order to minimize the number of rounds.
+We know that there is variation in the number of ballots used by each type of audit for large round sizes (a factor of two for $p=0.9$), but these values would be small in comparison to $w_r$.
+We observe this behaviour in Figure~\ref{fig:optimal_workloads} for extremely small and large workload values.
 For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
+\begin{figure}
+\includegraphics[width=.5\textwidth]{optimal_workloads.png}
+\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.}
+\label{fig:optimal_workloads}
+\end{figure}
+
 Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
-As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $w_r$.
+As expected, an overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audits $p$ increases as a function of $w_r$.
 Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO.
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{optimal_workloads.png}
-\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.
-(We show the value for varying $w_r$ since the value of $w_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $w_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
-\label{fig:optimal_workloads}
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=.5\textwidth]{optimal_ps.png}
 \caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$.
-With $w_b=1$, the varying value of $w_r$ can equivalently be thought of as the ratio of the workload of a round to the workload of a ballot.
 (Note that the steps in this function are a consequence of our subsampling the workload function.
-That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.5$.)}
+That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.05$.)}
 \label{fig:optimal_ps}
@@
 
-\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
-The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box.
+\subsubsection{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
+If.
+around requires multiple ballots from a single container, the container need only be unsealed once.
 Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container.
 Typically available election results give per-precinct granularity of vote tallies, rather than individual container information.
-In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant workload, $w_p$.
+In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct workload, $w_p$.
 In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round.
@@
 \subsection{Real time}
-Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
+Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
 Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
@@
 Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
-Figure~\ref{fig:real_time} shows how the esimated real time for these values differs as a function of $p$.
-It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retreiving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors.
+Figure~\ref{fig:real_time} shows how the estimated real time for these values differs as a function of $p$.
+It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retrieving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors.
 We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function.

### 2022-10-12 providence/sections/adversary.tex
move efficiency proof

@@
 \subsection{Efficiency}
-\begin{lemma}
+\newtheorem{lem}{Lemma}
+\begin{restatable}{lem}{efficiencylemma}
+\label{lem:efficiency}
 % For any round schedule %$[n_1, \ldots, n_r]$
@@
 the \Providence RLA is more efficient than EoR \BRAVO.
-\end{lemma}
-\begin{proof}
-Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
-That is, the EoR \BRAVO stopping condition is true:
-$$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
-To see the \Providence stopping condition is fulfilled, we rewrite as 
-%$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
-
-%% From Lemma~\ref{lemma:tau1_increasing} 
-% \[
-% \tau_1(k, p_a, p_0, n_j) = 
-%  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
-% \]
-%is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
-% \textit{i.e.,}
-% \begin{equation}\label{eq_incr}
-% % \frac{1}{\alpha} \leq
-%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}.
-% \end{equation}
-
-%If the \BRAVO test is satisfied for $k_j$ then:
-
-% From the properties of binomial distribution we obtain:
-% \[
-%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
-%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
-%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
-% \]
-\[
- \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) 
-\]
-\[
- = \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \sigma(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1})  
-\]
-\[
- \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) 
-\]
-\[
- = \omega_r(k_j, k_{j-1}, p_a, p_0, n_j, n_{j-1}).
-\]
-
-Where inequality $(*)$ follows from \cite[Theorem 6]{athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
-%Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
-%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
-%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.
-
-% 
-% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
-% 
-% 
-%  \[
-%  k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_r]}  
-%   = \min\left\{k :
-% \frac{
-% s_{k_{r-1}, a, [n_{r-1}]}
-% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
-% }{
-% s_{k_{r-1}, 0, [n_{r-1}]}
-% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
-% }\geq \frac{1}{\alpha}  \right\} =
-% \]
-% \[
-%   = \min\left\{k :
-% \frac{
-% s_{k_{r-1}, a, [n_{r-1}]}
-% \left(
-% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
-% + \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
-% \right)
-% }{
-% s_{k_{r-1}, 0, [n_{r-1}]}
-% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
-% }\geq \frac{1}{\alpha}  \right\} =
-% \]
-
-%  \[
-%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
-%         = \min\left\{k :
-% \frac{
-% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
-% }{
-% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
-% }\geq \frac{1}{\alpha}  \right\} =
-% \]
-
-% 
-%  
-%  Now, let us consider $j$-{th} round for $j = 1, \ldots, r$. We are going to show that
-%  $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.
-%  
-%  From Lemma~\ref{eq:kMin} we have:
-%  \[
-%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
-%   = \min\left\{k :
-% \frac{
-% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
-% }{
-% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
-% }\geq \frac{1}{\alpha}  \right\} =
-% \]
-% 
-% From the definition of $k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ we have that for every $k \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ it holds: $s_{k, a, [n_j]}/s_{k, 0, [n_j]} \geq 1/\alpha$ which means that:
-% \[
-% \frac{
-% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, a, [n_j]}
-% }{
-% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, 0, [n_j]}
-% }\geq \frac{1}{\alpha}
-% \]
-% which means that $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.
-
-\end{proof}
+\end{restatable}
+\begin{proof} Appendix~\ref{pf:efficiency_proof}\end{proof}

### 2022-10-12 providence/sections/appendix.tex
move efficiency proof

@@
 Compare this to the similar property for the $\Bravo$ stopping condition.
+
+\efficiencylemma
+
+\begin{proof}
+\label{pf:efficiency_proof}
+Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
+That is, the EoR \BRAVO stopping condition is true:
+$$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
+To see the \Providence stopping condition is fulfilled, we rewrite as 
+%$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
+
+%% From Lemma~\ref{lemma:tau1_increasing} 
+% \[
+% \tau_1(k, p_a, p_0, n_j) = 
+%  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
+% \]
+%is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
+% \textit{i.e.,}
+% \begin{equation}\label{eq_incr}
+% % \frac{1}{\alpha} \leq
+%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}.
+% \end{equation}
+
+%If the \BRAVO test is satisfied for $k_j$ then:
+
+% From the properties of binomial distribution we obtain:
+% \[
+%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
+%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
+%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
+% \]
+\[
+ \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) 
+\]
+\[
+ = \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \sigma(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1})  
+\]
+\[
+ \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) 
+\]
+\[
+ = \omega_r(k_j, k_{j-1}, p_a, p_0, n_j, n_{j-1}).
+\]
+
+Where inequality $(*)$ follows from \cite[Theorem 6]{athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
+%Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
+%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
+%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.
+
+% 
+% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
+% 
+% 
+\end{proof}

### 2022-10-12 providence/sections/workload.tex
bits and pieces

@@
 For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
-\begin{figure}
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{optimal_workloads.png}
@@
 This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.
-
-\begin{figure}
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{optimal_ps.png}
@@
 \includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
-\caption{The optimal average workload of the audits we simulated using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence.
-Similar to Figure~\ref{optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case.
+\caption{Optimal average workload using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence.
+Similar to Figure~\ref{fig:optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case.
 If the time for a single ballot is $75$ seconds, then $c_p=50$ corresponds to over an hour of extra time to sample a ballot from a new container.}
@@
 
-Unfortunately, efficiency alone is not sufficient for planning audits. In the US today, election officials need to make legitimate safety considerations.
-When drawing a random sample of the ballots, it is always possible that the tally of the sample provides misleading information.
-In a random sample, a true loser may recieve more votes than the true winner.
+Unfortunately, efficiency alone is not sufficient for planning audits. In the US today, election officials have a legitimate need to include personal safety as a consideration.
+In a random sample, a true loser may receive more votes than the true winner.
 This happens more often when the sample sizes are small, like for a hypothetical first round size of $11$ in the pilot audit, as seen in Figure~\ref{fig:pilot_sequence}.
-In these RLAs, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
+In the abstract, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
 
-Imagine that Alice beats Bob in an election contest both truly and by the announced results, but Bob's supporters are insistent he really won.
-When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner.
+Imagine that Alice beats Bob in an election contest both truly and in the reported results, but Bob's supporters are insistent he really won.
+When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner (and drawing fewer ballots on average).
 After the first round, by chance, there are more votes for Bob than for Alice in the sample.
@@
 After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
-This is an undesirable situation.
+This is an undesirable situation, as it can appear to Alice's supporters that election officials are simply drawing ballots till a chosen outcome is obtained.
 
@@
 To achieve a probability of a misleading sample of at most $0.1$, a round schedule with at least roughly $p=.3$ is needed.
-To achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
+To achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is needed, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
 It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US and the associated threats of violence.
@@
 \caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
-\label{fig:misleading}
+\label{fig:misleading_2}
 \end{figure}

### 2022-10-12 providence/sections/workload.tex
shrink table
table

@@
 
-\begin{table}
-\center
-\begin{tabular}{ |l|l|r|c|c|c| }
-\hline
-$M$ & margin & $n$ & Prov & SO & EoR \\
-\hline
-0.1&0.25&25&0.221&0.152&0.115\\
-&0.2&41&0.178&0.169&0.105\\
-&0.15&73&0.202&0.186&0.141\\
-&0.1&163&0.222&0.182&0.107\\
-&0.05&657&0.227&0.192&0.127\\
-&0.04&1027&0.237&0.193&0.124\\
-&0.03&1825&0.246&0.194&0.124\\
-&0.02&4105&0.246&0.195&0.124\\
-&0.01&16423&0.246&0.196&0.124\\
-\hline
-0.01&0.25&85&0.792&0.707&0.559\\
-&0.2&133&0.826&0.71&0.593\\
-&0.15&239&0.817&0.712&0.549\\
-&0.1&539&0.805&0.717&0.567\\
-&0.05&2163&0.817&0.721&0.569\\
-&0.04&3381&0.82&0.722&0.563\\
-&0.03&6011&0.824&0.723&0.573\\
-&0.02&13527&0.824&0.723&0.57\\
-&0.01&54117&0.824&0.724&0.57\\
-\hline
-0.001&0.25&149&0.962&0.889&0.783\\
-&0.2&235&0.963&0.89&0.768\\
-&0.15&421&0.958&0.894&0.801\\
-&0.1&951&0.958&0.894&0.793\\
-&0.05&3815&0.96&0.896&0.785\\
-&0.04&5965&0.961&0.896&0.791\\
-&0.03&10607&0.961&0.897&0.787\\
-&0.02&23869&0.962&0.897&0.787\\
-&0.01&95491&0.962&0.897&0.787\\
-\hline
-\end{tabular}
-\caption{For various margins, this table gives the minimum first round size $n$ to achieve at most a probability $M$ of a \emph{misleading sample} in the first round.
-The corresponding stopping probabilities of \Providence, SO \BRAVO, and EoR \BRAVO are given for each value of $n$.}
-\label{tab:misleading}
-\end{table}
-
-\textbf{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
+\subsubsection{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
 
@@
 \includegraphics[width=.5\textwidth]{so_misleading.png}
-\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
-\label{fig:misleading_2}
+\caption{The proportion of sequences that are misleading sequences in the SO \BRAVO audit as a function of $p$.}
+\label{so_misleading}
 \end{figure}

### 2022-10-12 providence/sections/workload.tex
more

@@
 Recall from Section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}.
-In particular, the audit passed with an SO \BRAVO risk measure of $5.41\%$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $36.6\%$.
+In particular, the audit passed with an SO \BRAVO risk measure of $0.0541$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $0.366$.
 

### 2022-10-12 providence/sections/conc.tex

@@
 A rigorous tabulation audit is an important part of a secure election.
-Ballot polling RLAs are commonly used and simple, not relying on special election equipment like comparison RLAs.
-We present \Providence which is the most efficient and secure ballot polling RLA, as efficient as \Minerva  and flexible as \BRAVO.
-We present proofs and simulation results to verify the claimed properties of \Providence, and we provide an open source implementation of the stopping condition and useful related functionality.
+We present \Providence and demonstrate that it is as efficient as \Minerva and as flexible as \BRAVO.
+We present proofs and simulation results to verify the claimed properties of \Providence, and we provide an open source implementation of the stopping condition and useful related functionality for planning audits.
+We define the constraint of an acceptable probability of a misleading audit sample, and describe its importance to the planning process.

### 2022-10-12 providence/sections/appendix.tex

@@
 \begin{proof}
-\label{pf:efficiency_proof}
 Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
@@
 
-Where inequality $(*)$ follows from \cite[Theorem 6]{athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
+Where inequality $(*)$ follows from \cite[Theorem 6]{arxiv_athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
 %Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.

### 2023-02-07 providence/sections/adversary.tex
.
fix lemma numbering

@@
 For \Minerva, however, we would have a predetermined round schedule.
-We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by many states performing an RLA), which is is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
+We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by many states performing an RLA), which is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
 This multiplier of $1.5$ is known to give, over a wide range of margins, a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
@@
 \subsection{Efficiency}
-\newtheorem{lem}{Lemma}
-\begin{restatable}{lem}{efficiencylemma}
+\begin{lemma}
 \label{lem:efficiency}
@@
 and for any round schedule $[n_1, \ldots, n_j]$, 
-the \Providence RLA is more efficient than EoR \BRAVO.
-\end{restatable}
-\begin{proof} Appendix~\ref{pf:efficiency_proof}\end{proof}
+the \Providence RLA stops before or at the same time as EoR \BRAVO.
+\end{lemma}
+\begin{proof} Appendix~\ref{sec:proofs}\end{proof}

### 2023-02-07 providence/sections/prov.tex
.

@@
 The computation of $\tau_j$ for $j \geq 2$, as required in \Minerva, relies on the convolution of two probability distribution functions and is hence computationally considerably more expensive.
+Lesser computational complexity makes audit planning and analysis using simulations as in Section~\ref{sec:workload} more feasible.
 
@@
 \label{sec:proof}
-We now prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.
+We now prove that \Providence is risk-limiting using lemmas from basic algebra which are given in Appendix~\ref{sec:proofs}.
 
@@
 $$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge \frac{1}{\alpha}.
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}] Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}] Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}
+$$$$\ge \frac{1}{\alpha}.
 $$
@@
 $$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}] Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}] Pr[K_{j} \ge k_{j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}\ge
 $$
 $$
-\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}]\cdot Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]} \ge
+\frac{\Pr[K_{j-1} = {k_{j-1}} \mid H_a, n_{j-1}] Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_a, n_{j-1}, n_{j}]}{\Pr[K_{j-1} = {k_{j-1}} \mid H_0, n_{j-1}] Pr[K_{j} \ge k_{min, j} \mid {k_{j-1}}, H_0, n_{j-1}, n_{j}]}
+$$$$
+\ge 
 \frac{1}{\alpha}

### 2023-02-07 providence/sections/related.tex
shortening
.

@@
 % is drawn (a \B \BRAVO audit), 
-It is easy to see that the ratio $\sigma$ is the 
-likelihood ratio:
+If $K$ is the random variable indicating the number of ballots in the sample that contain a vote for the reported winner, it is easy to see that the ratio $\sigma$ is the likelihood ratio:
 $$
@@
 
-It now becomes useful to have shorthand for a sequence of cumulative round sizes and the corresponding sequence
+Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
+Now it becomes useful to have shorthand for a sequence of cumulative round sizes and the corresponding sequence
 of cumulative winner ballot tallies.
 We use:
-$$\bm{k_j}\triangleq(k_1,k_2,\ldots,k_j)$$
-$$\bm{n_j}\triangleq(n_1,n_2,\ldots,n_j)$$
-
-Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
+$$\bm{n_j}\triangleq(n_1,n_2,\ldots,n_j) \quad\text{and}\quad \bm{k_j}\triangleq(k_1,k_2,\ldots,k_j)$$
+Also, let $K_j$ be the random variable indicating the cumulative number of ballots in the sample after the $j$th round is drawn.
 

### 2023-02-07 providence/sections/workload.tex
.

@@
 larger rounds), \Providence requires fewer ballots on average.
-In particular, the savings of \Providence become larger as $p$ increases; for $p=.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
+In particular, the savings of \Providence become larger as $p$ increases; for $p=0.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
 \begin{figure}[h!]
@@
 \includegraphics[width=.5\textwidth]{with_round_workload.png}
-\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$.
+\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
 Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
@@
 \subsubsection{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
-If.
-around requires multiple ballots from a single container, the container need only be unsealed once.
+If a round requires multiple ballots from a single container, the container need only be unsealed once.
 Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container.
@@
 Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
-For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
+For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_p$.
 Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
@@
 Figure~\ref{fig:pilot_misleading} shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
-Despite the large difference in margin ($\sim 5\%$ in Virginia and $\sim 25\%$ in the pilot) we still observe that a \emph{misleading limit} of $0.01$ is first achieved at roughly $p=0.8$ and $0.001$ at $p=0.95$.
+Despite the large difference in margin ($\sim 0.05$ in Virginia and $\sim 0.25$ in the pilot) we still observe that a \emph{misleading limit} of $0.01$ is first achieved at roughly $p=0.8$ and $0.001$ at $p=0.95$.
 

### 2023-02-07 providence/sections/intro.tex
.

--- 2022-10-12 providence/sections/intro.tex
@@
 One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
-Many US states have had pilot RLA programs.
-Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
 
-We propose the \Providence audit, a new approach to the ballot polling RLA, and propose a new model for the work load of an election.
+RLAs guarantee with known, high probability that the reported election outcome is correct but only require the effort of manually interpretting a small fraction of the total ballots.
+Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
+
+%We propose the \Providence audit, a new approach to the ballot polling RLA, and propose a new model for the work load of an election.
 We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections.
-Our implementation of \Providence is likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
+%Our open-source implementations of \Providence and our audit planning tools are likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
 
 \subsection{Background}
-Ballot comparison RLAs require the fewest ballots of all known RLA approaches.
+In a ballot comparison RLA, the manual and machine interpretatations of sampled ballots are compared.
+Comparison RLAs require the fewest ballots of all known RLA approaches.
 On the other hand, they require the use of special election technology and are not always feasible.
+In a ballot polling RLA, the manual interpretations of ballots are simply tallied.
 Ballot polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system.
@@
 A risk measure is then computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn.
+It is possible to design audits which output a third determination, (c) the audit proceeds to a full manual hand count.
+To allow maximum flexibility for election officials to make this logistically involved decision, this third output is generally not included.
 Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
@@
 \item Results and analysis from the use of \Providence in a pilot audit in Rhode Island.
-\item Open source implementation of \Providence. 
 \item A model of workload that includes the overhead effort of each round and the overhead effort of retrieving a storage unit of ballots; simulations that illustrate the use of this model to compare the different types of ballot polling audits and to plan an audit with minimal workload.
 \item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
+\item Open source implementation of \Providence and audit planning tools. 
 %including the novel metric Probability of Misleading(name?)
@@
 
-Our results demonstrate the superiority of \Providence over the other audits. Our work may be used by election officials to plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
+%Our results demonstrate the superiority of \Providence over the other audits. Our work may be used by election officials to plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
 

### 2023-02-07 providence/sections/related.tex
.

@@
 
+Note that \BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT).
+In the more general SPRT, the test can also reject the alternative hypothesis, and there is an additional parameter $\beta$, the probability of incorrectly rejecting the alternative hypothesis.
+In an RLA, this corresponds to the audit having a third output: \emph{proceed to a full manual count of the ballots}.
+In the existing literature, ballot polling audits do not include this possibility (i.e.
+they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
+
 Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
@@
 \end{comment}
+
+In this work, we consider ballot polling RLAs only and thus compare \Providence with \BRAVO and \Minerva.

### 2023-02-07 providence/sections/prov.tex
.

@@
 
+Before defining \Providence, we give some intuition for how it is designed to avoid the problem of \Minerva.
+In round $j$, for every possible value of $K_j$ for which the audit stops, a risk is incurred.
+To obtain the total risk for the round, one adds the risks corresponding to each value of $K_j$ for which the audit can stop, weighted by the probability of drawing that value of $K_j$.
+The stopping condition provides relationships among various quantities.
+
+In \Minerva, the stopping condition relates the weighted average of the risks to the weighted average of the stopping probabilities over all values of $K_j$, for a given round size.
+Separate relationships between risk and stopping probability are not available for individual values of $K_j$.
+If the next round size depends on $K_j$, we do not have expressions relating the risks, and so \Minerva may be vulnerable to an adversary choosing round sizes.
+
+In \Providence, we choose a stopping condition that applies separately to the risk and stopping probabilities for each value of $K_j$, avoiding the problem of \Minerva, and allowing for optimal round size choices, which depend on the drawn sample.
 The \Providence audit is risk-limiting even if an adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.

### 2023-02-07 providence/sections/adversary.tex
.

@@
 and for any round schedule $[n_1, \ldots, n_j]$, 
-the \Providence RLA stops before or at the same time as EoR \BRAVO.
+the \Providence RLA stops before or in the same round as EoR \BRAVO.
 \end{lemma}

### 2023-02-07 providence/sections/intro.tex
.

@@
 
-A ballot polling audit begins when a {\em round} \cite{usenix_minerva} of multiple randomly-chosen ballots is drawn.
-A risk measure is then computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn.
+A ballot polling audit begins with choosing a first {\em round} \cite{usenix_minerva} size, the number of ballots sampled before first checking the stopping condition.
+Ballots on the ballot manifest are then sampled uniformly at random using a pseudorandom number generator typically seeded by a natural source of randomness like rolling dice.
+The physical ballots are then found and manually interpreted, recording the manual interpretations.
+Based on the manual interpretations, a risk measure is computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn.
 It is possible to design audits which output a third determination, (c) the audit proceeds to a full manual hand count.
@@
 Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
+
 Ballot polling audits have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.

### 2023-02-07 providence/sections/prov.tex
typo

@@
 \end{equation*}
-The left hand side above is the probability of stopping in the $j^{th}$ round and $K_{j-1} = k_{j-1}$, given the null hypothesis, which is smaller than $alpha$ times the same probability given the alternate hypothesis.
+The left hand side above is the probability of stopping in the $j^{th}$ round and $K_{j-1} = k_{j-1}$, given the null hypothesis, which is smaller than $\alpha$ times the same probability given the alternate hypothesis.
 Summing both sides over all values of $K_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).

### 2023-02-07 providence/sections/prov.tex
.

@@
 Separate relationships between risk and stopping probability are not available for individual values of $K_j$.
-If the next round size depends on $K_j$, we do not have expressions relating the risks, and so \Minerva may be vulnerable to an adversary choosing round sizes.
+If the next round size depends on $K_j$, we do not have expressions relating the risks, and so \Minerva may be vulnerable to an adversary choosing round sizes \cite{usenix_minerva, arxiv_athena}.
 

### 2023-02-12 providence/sections/intro.tex
address why minerva is not in workload analysis directly
Intro

--- 2022-10-12 providence/sections/intro.tex
@@
 One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
-Many US states have had pilot RLA programs.
-Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
+A {\em risk-limiting audit} guarantees a minimum probability of a full hand count if the election outcome is incorrect.
+Conversely, it guarantees a maximum probability of audit error, termed the {\em risk limit}, which is the maximum probability with which the audit would declare an incorrect election outcome as being correct.
 
-We propose the \Providence audit, a new approach to the ballot polling RLA, and propose a new model for the work load of an election.
+Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
+
+%We propose the \Providence audit, a new approach to a particular type of RLA---the ballot polling RLA described below---and propose a new model for the work load of an election.
 We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections.
-Our implementation of \Providence is likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
+\Providence was recently integrated as an option in Arlo \cite{arlo}, the most popular election audit software.
+%Our open-source implementations of \Providence and our audit planning tools are likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.
 
-\subsection{Background}
-Ballot comparison RLAs require the fewest ballots of all known RLA approaches.
-On the other hand, they require the use of special election technology and are not always feasible.
-Ballot polling RLAs require a larger number of ballots, but are more feasible because they do not require any additional functionality of the voting system.
+\subsection{Background on RLAs}
+We provide here the background necessary to evaluate our contributions.
+
+All RLAs sample one or more ballots at random; we will refer to each such set of ballots as a {\em round}.
+The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination).
+In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count.
+However, a manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary, hence (c) is generally not incorporated.
+Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
+They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
+
+\subsubsection{Types of RLAs}
+In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.
+A typical approach is to use a ballot serial number on both paper ballot and CVR.
+When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered.
+Additionally, some voting systems do not record a CVR.
+One may perform a transitive audit by rescanning unnumbered voted ballots with special scanners which produce CVRs and also print numbers on the ballots as they are scanned.
+This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible.
+
+In a ballot polling RLA \cite{RLA}, the manual interpretations of the sampled ballots are simply tallied.
+Ballot polling RLAs require a much larger number of ballots than ballot comparison RLAs, but are more feasible because they do not require any additional functionality of the voting system, and, in particular, do not need CVRs.
 What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).
 
-A ballot polling audit begins when a {\em round} \cite{usenix_minerva} of multiple randomly-chosen ballots is drawn.
-A risk measure is then computed to determine whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn.
-Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
-Ballot polling audits have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
+Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch.
+This approach typically requires the sampling of a very large number of ballots.
+However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.
+(Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is.
+A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
 
-A {\em round-by-round (R2)} audit is one where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections.
+This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+
+\subsubsection{Ballot Polling Audits}
+Ballot polling audits proceed as follows. 
+\begin{enumerate}
+\item A first round \cite{usenix_minerva} size---the number of ballots sampled before first checking the stopping condition---is chosen. 
+\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator typically seeded by a natural source of randomness like rolling dice.
+\item The physical ballots are found and manually interpreted, recording the manual interpretations. 
+\item Based on the manual interpretations, the stopping condition is computed. 
+\item If more ballots are to be drawn, the next round size is chosen.
+Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
+For RLAs required by statute or law, certification deadlines would play a large role, as the audit would need to be completed before the deadline.
+\end{enumerate}
+
+A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections.
 A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
 The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
-Election officials typically draw ballots in large round sizes, and \BRAVO needs to be modified for use in this manner.
+
+Election officials typically draw ballots in large round sizes:  see for example \cite{va-2022,RI-report}, and note that, in addition to allowing users to directly enter a round size, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$, and the expected number of ballots required by \BRAVO.
+Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
+\BRAVO hence cannot be used as a B2 audit in these scenarios.
+
 For use as an R2 audit, the \BRAVO stopping condition can be applied once at the end of each round (End-of Round (EoR)), or retroactively after each ballot drawn if ballot order is retained (Selection-Ordered (SO)).
@@
 
+\subsubsection{Adversarial Model for RLAs}
+Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
+For our purposes, we will assume that the best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes were verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
+We will further assume that all software used in the RLA is open source and well-defined, so its output may be reproduced and thus verified by an observer wishing to do so with their own software.
+
+Referring to the ballot polling audit steps described above, we further assume that a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
+Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed from knowledge of the seed and the drawn ballots.
+Thus the only variable is round size.
+We define a {\em weak adversary} as one who can choose the first round size and a {\em strong adversary} as one who can choose any round size.
+
+\subsection{The Literature and Open Questions}
 Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which does not need ballot order and relies only on sample and round tallies.
-They prove that it requires fewer ballots than EoR \BRAVO when both audits have the same pre-determined round schedule and the true tally is as announced.
-%\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand.
-In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn.
+They prove that it is risk-limiting when the number of relevant ballots drawn in each round is pre-determined before any ballots are examined; that is, for a weak adversary.
+They do not address the case of a strong adversary (such as an audit insider) who can determine the size of the next round after knowing what votes are on the ballots sampled thus far.
+In particular, an open question about \Minerva is whether the computation of a risk limit assuming a weak adversary applies to an attack by a strong adversary, or is the risk limit computation incorrect when the adversary is strong?
+Can the strong adversary increase the audit's error probability beyond its declared risk limit?
+Or is there no probabilistic adversarial advantage to being able to compute next round sizes after knowing the drawn sample?
+We do not answer this question, and to our knowledge, it remains open.
+
+Until \Minerva is proven to be risk-limiting to a given risk limit for the strong adversary, it may not be used in audits whose round sizes are not pre-determined.
+This presents a major limitation, because the stopping probability of the next round is better estimated using information of the sample drawn thus far, but this would not be allowed for \Minerva.
+The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter.
+Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
+However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that a fixed draw size is sufficient to limit adversaries to weak ones, though this is not explicitly proven in \cite{usenix_minerva}.
+
+% In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn.
 Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general.
-(We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
-They also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced. 
+(We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.
+
+Zag\'{o}rski {\em et al.}  also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
 Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability \cite{simulations}, though the improvement from using \Minerva over either version of \BRAVO decreases with round size.
 
-\subsection{Open Questions in the Literature}
-A major limitation of \Minerva is that one needs to determine the round schedule before the audit begins, because \Minerva has not been shown to be risk-limiting if an adversary can choose subsequent round sizes after knowing the sample drawn.
-\BRAVO, on the other hand, is not limited in this manner.
+The risk limit for B2, EoR and SO \BRAVO is fixed whether the adversary is strong or weak.
 This allows \BRAVO audits the flexibility of choosing smaller subsequent round sizes if the sample drawn so far is a ``good'' sample.
@@
 If this were a correct measure of the workload of an audit, we would want to use B2 audits (round size is one) and make decisions about stopping the audit after drawing each ballot, because this leads to the smallest expected number of ballots.
-Election officials, on the other hand, greatly prefer drawing many ballots at once.
+As described above, election officials, on the other hand, greatly prefer drawing many ballots at once.
 This preference is likely due to the following.
@@
 \subsection{Our Contributions}
-We present \Providence, and provide the following:
+Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to a strong adversary.
+The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
+We are able to derive a new stopping condition for which a far simpler proof of the risk-limiting property is possible.
+In particular, this proof does not require an assumption about round sizes.
+We provide the following:
 \begin{enumerate}
-\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples.
+\item Proof that \Providence is an RLA and resistant to a strong adversary.
 \item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence uses number of ballots similar to those of \Minerva, both fewer than either version of \BRAVO.
 \item Results and analysis from the use of \Providence in a pilot audit in Rhode Island.
-\item Open source implementation of \Providence. 
 \item A model of workload that includes the overhead effort of each round and the overhead effort of retrieving a storage unit of ballots; simulations that illustrate the use of this model to compare the different types of ballot polling audits and to plan an audit with minimal workload.
 \item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
+\item Open source implementation of \Providence and audit planning tools. 
 %including the novel metric Probability of Misleading(name?)

### 2023-02-12 oliver_copy_providence/ForArchive/sections/sims.tex
minor edits

--- 2022-10-12 providence/sections/sims.tex
@@
 Narrower margins are computationally expensive, especially for the simulations of tied elections, which, by design, have a low probability of stopping and hence quickly increase in sample size.
-We use the simulator in the R2B2 software library\cite{r2b2_anon}.
+We use the simulator in the R2B2 software library\cite{r2b2}.
 For each margin, we perform $10^4$ \Providence audit trials each on a tied election (hypothesis $H_0$, the null hypothesis) and the election as reported (hypothesis $H_a$, the alternate hypothesis).
@@
 
-\begin{figure}
+In the simulations of \Providence audits of a tied election, the fraction of audits that stop, as shown in Figure~\ref{fig:prov-risk}, is an estimate of maximum risk.
+For all margins, this estimated maximum risk is less than the risk limit, supporting the claim that \Providence is risk-limiting.
+
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{prov_risk.png}
@@
 
-In the simulations of \Providence audits of a tied election, the fraction of audits that stop, as shown in Figure~\ref{fig:prov-risk}, is an estimate of maximum risk.
-For all margins, this estimated maximum risk is less than the risk limit, supporting the claim that \Providence is risk-limiting.
+Simulations of audits of the election as reported provide insight into stopping probability and number of ballots drawn when the election is as reported.
+Figure~\ref{fig:prov-sprob} shows that the stopping probabilities over the first rounds are near and slightly above $0.9$ as expected, since our software chose round sizes to give at least a $0.9$ conditional stopping probability.
+The values are not as tight around $0.9$ for later rounds because fewer audit trials make it to later rounds, and our experimental probability estimates are not as accurate.
 
-\begin{figure}
+\begin{figure}[h!]
 \includegraphics[width=.5\textwidth]{prov_sprob.png}
@@
 
-Simulations of audits of the election as reported provide insight into stopping probability and number of ballots drawn when the election is as reported.
-Figure~\ref{fig:prov-sprob} shows that the stopping probabilities over the first rounds are near and slightly above $0.9$ as expected, since our software chose round sizes to give at least a $0.9$ conditional stopping probability.
-The values are not as tight around $0.9$ for later rounds because fewer audit trials make it to later rounds, and our experimental probability estimates are not as accurate.
-
 We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single margin as an example: the 2020 US Presidential election in the state of Texas, with margin $0.057$.
@@
 Both \BRAVO implementations use a conditional stopping probability of $0.9$ for each round, while \Minerva uses a first round size with stopping probability $0.9$ and a multiplier of $1.5$ to obtain subsequent round sizes.
+
+\begin{figure}[h!]
+\includegraphics[width=.5\textwidth]{prov_asn.png}
+\caption{For the entire audit, consisting of all five rounds, the fraction of simulated audits that stopped as a function of the average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
+The average sample number (ASN) for \B \BRAVO is included for context.}
+\label{fig:prov-asn}
+\end{figure}
 Figure~\ref{fig:prov-asn} shows the probability of stopping as a function of the number of ballots sampled, a plot similar to those presented in \cite{simulations}.
@@
 % Section~\ref{sec:workload} discusses workload in more depth.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{prov_asn.png}
-\caption{For the entire audit, consisting of all five rounds, the fraction of simulated audits that stopped as a function of the average number of ballots drawn for \Providence, \Minerva, EoR \BRAVO, and SO \BRAVO.
-The average sample number (ASN) for \B \BRAVO is included for context.}
-\label{fig:prov-asn}
-\end{figure}
-
-\newpage

### 2023-02-12 oliver_copy_providence/sections/intro copy.tex
minor edits

--- 2022-01-09 MultipleRoundSims/sections/01-intro.tex
@@
-%Intro
+%Older intro file, which is mostly a repeat of the Voting paper intro 
 The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances.
@@
 
-This paper provides insight into the main approaches to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results.
-While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available.
-This paper examines the number of ballots drawn over multiple rounds of both audits, the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.
-
-\subsection{Background}
-This paper focuses on ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
-Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Goergia, Indiana, Michigan, Ohio, Pennsylvania and elsewhere).
+This paper concerns ballot-polling RLAs, which require a large number of ballots relative to comparison RLAs but do not rely on any special features of the election technology.
+Since comparison RLAs are not always feasible, ballot-polling audits remain an important resource and have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere).
 In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}.
 A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit.
-Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
+If so, the audit stops; else another round is drawn.
+Election officials would typically decide to do a full manual hand count if the audit does not stop in spite of drawing a large number of ballots.
+Because the decision to stop or draw more ballots is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit.
 The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.
 
-The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced, and stopping decisions are made after each ballot draw.
+The \BRAVO audit is designed for use as a B2 audit: it requires the smallest expected number of ballots when the true tally of the underlying election is as announced and stopping decisions are made after each ballot draw.
 In practice, election officials draw many ballots at once, and the \BRAVO stopping rule needs to be modified for use in an R2 audit that is not B2.
@@
 Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced.
+%\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand.
+In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn.
+Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general.
+(We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.}
 They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.
+Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability.
 
-There are no results, either theoretical or based on simulations, regarding the number of ballots drawn over multiple rounds in a \Minerva audit with a pre-determined schedule.
-Because \BRAVO does not need to work on a pre-determined round schedule, it can optimize the size of the next round based on the sample drawn so far.
-Thus an open question is whether the constraint of a predetermined round schedule limits the efficacy of \Minerva in future rounds, and there is no literature comparing the number of ballots drawn by \Minerva and SO \BRAVO over multiple rounds.
-Note that the Average Sample Number (ASN) computations for \BRAVO \cite{bravo} apply only for B2 audits and are especially misleading as estimates of the number of ballots drawn over multiple rounds when first round sizes are large.
+While more efficient than \BRAVO, \Minerva requires that the round schedule is fixed in advance of the audit.
+This may lead to significant unnecessary work.
+For example, suppose the fixed round schedule for \Minerva is $10,000$ ballots per round (a reasonable number in a state-wide contest with a narrow margin), and the first round sample finds a number of ballots for the winner just short of that necessary for the audit to stop.
+It then may be sufficient to draw a very small second round and still stop with high probability in the second round, but the additional $10,000$ ballots must be drawn.
+In contrast, subsequent \BRAVO round sizes can be chosen based on preceding samples.
 
-Both \BRAVO and \Minerva have been integrated into election audit software {\em Arlo} \cite{arlo}, and, as such, are available for use in real election audits.
-Both have been used in real election audits\cite{usenix_minerva, virginia_rla}.
-For this reason, it is important to understand their properties over multiple rounds.
+An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.
 
 \subsection{Our Contributions}
-For a two candidate plurality contest with a risk limit of $10\%$, we show the following
+We present \Providence, and provide the following:
 \begin{enumerate}
-\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO.
-However, the improvement is considerably smaller than that when the stopping probability is $0.9$.
-\begin{itemize}
-\item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva.
-On the other hand, for a first round stopping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.
-\item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva.
-On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.
-\end{itemize}
-\item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three.
-When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively.
-Both our simulator and the code estimating probabilities and round sizes are flexible enough to enable the study of various predetermined round schedules.
+\item Proof that \Providence is an RLA and resistant to an adversary who can choose subsequent round sizes with knowledge of previous samples
+\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence has similar efficiency to \Minerva, both greater than either implementation of \BRAVO
+\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island
+\item Open source implementation of \Providence 
+%including the novel metric Probability of Misleading(name?)
 \end{enumerate}
 
-\subsection{Organization} Section \ref{sec:related} describes related work.
-The experiments we performed are described in section \ref{sec:expts} and sections \ref{sec:results1} and \ref{sec:results2} present our results.
+%\subsection{Organization} 
+%Section \ref{sec:related} describes related work.
+The experiments we performed are described in section \ref{sec:expts}, and sections \ref{sec:results1} and \ref{sec:results2} present our results.
 Section \ref{sec:conc} has our conclusions.

### 2023-02-12 oliver_copy_providence/sections/prov.tex
minor edits

--- 2023-02-07 providence/sections/prov.tex
@@
 Conversely, if the current sample is not at all close to satisfying the risk limit, it would be advantageous to draw a larger round than the predetermined round size.
+
+We now formalize these notions of weak and strong adversaries.
+
+\begin{definition}[Weak Adversary]
+A {\emph weak adversary} may choose the first round size as a function of audit parameters. That is, the first round size is a function $$n_1(\alpha, p_a, p_0).$$
+\end{definition}
+
+\begin{definition}[Strong Adversary]
+A {\emph strong adversary} may choose any round size as a function of audit parameters and all precreding samples. That is, the first round size is a function 
+$$n_1(\alpha, p_a, p_0),$$
+and for all rounds $j\ge 2$, the round size is a function
+$$n_{j+1}(\alpha, p_a, p_0, \bm{k_{j-1}}, \bm{n_{j-1}}).$$
+\end{definition}
 
@@
 \label{sec:proof}
-We now prove that \Providence is risk-limiting using lemmas from basic algebra which are given in Appendix~\ref{sec:proofs}.
+We now prove that \Providence is risk-limiting against a strong adversary using lemmas from basic algebra which are given in Appendix~\ref{sec:proofs}.
 
@@
 $$
-By Lemma \ref{lemma:minerva2_kmin_exists}, we see that this
-is equivalent to the following:
+By Lemma \ref{lemma:minerva2_kmin_exists} and Definition \ref{def:kmin}, there is 
+a value $k_{min,1} = k^{p_a, p_0, \alpha, 0}_{min, 1, 0, n_1}$ such that 
 $$
+\frac{Pr[K_{1} \geq k_{1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{1} \mid H_0, n_1]}
+\ge
 \frac{Pr[K_{1} \geq k_{min,1} \mid H_a, n_1]}{Pr[K_{1} \geq k_{min, 1} \mid H_0, n_1]}
-\ge \frac{1}{\alpha}.
+\ge 
+\frac{1}{\alpha}.
 $$
-where $k_{min,1} = k^{p_a, p_0, \alpha, 0}_{min, 1, 0, n_1}$ (see Lemma \ref{lemma:minerva2_kmin_exists} and Definition \ref{def:kmin}).
 

### 2023-02-12 oliver_copy_providence/sections/workload.tex
minor edits

--- 2023-02-07 providence/sections/workload.tex
@@
 Note that simulations of audits of tied elections are not necessary, as all the audits we are considering are risk-limiting and hence we already know the performance to expect when auditing a tied election, even one not reported as such.
+
+Importantly, note that \Minerva does not appear in the analysis in this section. 
+Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in Section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples.
+\Minerva is not known to be risk-limiting in this setting, and thus cannot be used for RLAs that proceed in this way.
 

### 2023-02-12 providence/sections/adversary.tex
.

@@
 
-\Minerva was proven to be a risk-limiting audit for a predetermined round schedule.
-As explained earlier, it is not clear that \Minerva is risk-limiting if an adversary can 
-adaptively select the round schedule as the audit proceeds.
+%\Minerva was proven to be a risk-limiting audit for a predetermined round schedule.As explained earlier, it is not clear that \Minerva is risk-limiting if an adversary can adaptively select the round schedule as the audit proceeds.
 In this section we prove that \Providence does not have this problem, and is risk-limiting even when the adversary can choose next round sizes based on knowledge of the current sample.
 
-\begin{definition}[Strategy-Proof RLA]
- An audit $\mathcal{A}$ is a Strategy-Proof Risk Limiting Audit with risk limit $\alpha$ iff for
- all strategies of selecting round schedule and for sample $X$ 
- \[
-  \Pr\left[\mathcal{A}(X) = \text{Correct} | H_0\right] \leq \alpha.
- \]
+%\begin{definition}[Strategy-Proof RLA]
+% An audit $\mathcal{A}$ is a Strategy-Proof Risk Limiting Audit with risk limit $\alpha$ iff for
+% all strategies of selecting round schedule and for sample $X$ 
+% \[
+%  \Pr\left[\mathcal{A}(X) = \text{Correct} | H_0\right] \leq \alpha.
+% \]
+%\end{definition}
 
-\end{definition}
-
-\begin{lemma}
-\Providence is a Strategy-Proof RLA. 
-\end{lemma}
-\begin{proof}
-This property follows from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
+%\begin{lemma}
+%\Providence is a Strategy-Proof RLA. 
+%\end{lemma}
+%\begin{proof}
+%This property follows from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}.
 Note that, as described in section \ref{sec:proof}, the proof of the risk-limiting nature of the audit does not rely on round sizes $n_j$ being identical for all values of $k_{j-1}$.
-\end{proof}
-
+%\end{proof}
+%
 To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $0.01$ and risk limit $0.1$. 

### 2023-02-12 providence/sections/prov.tex
.

@@
 Summing both sides over all values of $K_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).
-Note here that the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$.
+
+Note here that the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.
+
 When both sides of the inequality are further summed over all rounds, we get:
@@
 
-\subsection{Resistance to an adversary choosing round sizes}
+\subsection{Consequences of resistance to an adversary choosing round size}
 \label{sec:adversary}
-% Filip TBD
 \input{sections/adversary}

### 2023-02-12 providence/sections/prov.tex
.

@@
 The left hand side above is the probability of stopping in the $j^{th}$ round and $K_{j-1} = k_{j-1}$, given the null hypothesis, which is smaller than $\alpha$ times the same probability given the alternate hypothesis.
-Summing both sides over all values of $K_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).
+For different possible values of $k_{j-1}$, different round sizes $n_j$ can be used, and this same relationship will hold. 
+That is, the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather for each individual value of $K_{j-1}$.
+By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strong adversary.}
 
-Note here that the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.
-
+Summing both sides over all values of $k_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).
 When both sides of the inequality are further summed over all rounds, we get:

### 2023-02-12 providence/sections/intro.tex
undetected ---> undetectable
In definition of SI in Intro

@@
 Errors and/or manipulation may not always change an election outcome, but we would want to know when they do.
-{\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetected change in the election outcome.
+{\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome.
 {\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; incorrect outcomes are detected with high probability when the evidence is examined.

### 2023-02-12 providence/sections/intro.tex
some more stuff on rlas

@@
 {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome.
-{\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; incorrect outcomes are detected with high probability when the evidence is examined.
+{\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; the evidence is examined so that incorrect outcomes may be detected with high probability.
 One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
-A {\em risk-limiting audit} guarantees a minimum probability of a full hand count if the election outcome is incorrect.
-Conversely, it guarantees a maximum probability of audit error, termed the {\em risk limit}, which is the maximum probability with which the audit would declare an incorrect election outcome as being correct.
+
+An audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count.
+It may be evaluated based on how it deals with incorrect election outcomes, as the purpose is to detect these.
+A {\em risk-limiting audit (RLA)} guarantees a minimum probability of a full hand count if the election outcome is incorrect; this is the probability of correct functioning of the audit.
+Equivalently, an RLA guarantees a maximum probability of audit error if the election outcome is incorrect.
+This error measure is termed the {\em risk limit}, more rigorously defined as the maximum probability with which the audit would declare an incorrect election outcome as being correct.
 

### 2023-02-12 providence/sections/intro.tex
typo

@@
 Section \ref{sec:related} describes related work.
-Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
+Section \ref{sec:prov} describes the \Providence audit, Section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
 Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2020 US Presidential election in the state of Virginia.

### 2023-02-12 providence/sections/prov.tex
typo

@@
 For different possible values of $k_{j-1}$, different round sizes $n_j$ can be used, and this same relationship will hold. 
-That is, the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather for each individual value of $K_{j-1}$.
+That is, the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather than for each individual value of $K_{j-1}$.
 By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strong adversary.}

### 2023-02-12 providence/sections/intro.tex
added model. worked more on intro.

@@
 
-An audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count.
-It may be evaluated based on how it deals with incorrect election outcomes, as the purpose is to detect these.
-A {\em risk-limiting audit (RLA)} guarantees a minimum probability of a full hand count if the election outcome is incorrect; this is the probability of correct functioning of the audit.
-Equivalently, an RLA guarantees a maximum probability of audit error if the election outcome is incorrect.
-This error measure is termed the {\em risk limit}, more rigorously defined as the maximum probability with which the audit would declare an incorrect election outcome as being correct.
+A tabulation audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count.
+The integrity of an audit may be judged by how it deals with incorrect outcomes.
+A {\em risk-limiting audit (RLA)} guarantees a minimum probability that it will perform as it is supposed to (escalate to a full hand count) if the outcome is incorrect.
+Equivalently, this is also a guarantee limiting the probability of audit error when the outcome is incorrect.
+The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct (and not proceed to a full hand count).
+Lower risk limits are better.
 
@@
 \subsection{Background on RLAs}
-We provide here the background necessary to evaluate our contributions.
-
 All RLAs sample one or more ballots at random; we will refer to each such set of ballots as a {\em round}.
@@
 
-\subsubsection{Types of RLAs}
-In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
-Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.
-A typical approach is to use a ballot serial number on both paper ballot and CVR.
-When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered.
-Additionally, some voting systems do not record a CVR.
-One may perform a transitive audit by rescanning unnumbered voted ballots with special scanners which produce CVRs and also print numbers on the ballots as they are scanned.
-This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible.
+There are three types of RLAs: ballot comparison RLAs, batch comparison RLAs and ballot polling RLAs.
+This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+Ballot and batch comparison RLAs are described in \ref{sec:related}.
 
-In a ballot polling RLA \cite{RLA}, the manual interpretations of the sampled ballots are simply tallied.
-Ballot polling RLAs require a much larger number of ballots than ballot comparison RLAs, but are more feasible because they do not require any additional functionality of the voting system, and, in particular, do not need CVRs.
+\subsection{Ballot Polling RLAs}
+\label{sec:polling}
+Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system.
+For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots.
+A ballot polling RLA \cite{RLA}, on the other hand, simply tallies the sampled ballots.
+It requires a much larger number of ballots than a ballot comparison RLA, but tends to be more feasible because it does not require any additional functionality of the voting system; in particular it does not require the system to generate or provide CVRs.
 What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).
 
-A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
-Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch.
-This approach typically requires the sampling of a very large number of ballots.
-However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.
-(Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is.
-A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
-
-This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
-
-\subsubsection{Ballot Polling Audits}
 Ballot polling audits proceed as follows. 
@@
 
-Election officials typically draw ballots in large round sizes:  see for example \cite{va-2022,RI-report}, and note that, in addition to allowing users to directly enter a round size, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$, and the expected number of ballots required by \BRAVO.
+Election officials typically draw ballots in large round sizes:  see for example \cite{va-2022,RI-report}.
+Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping).
 Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
@@
 
-\subsubsection{Adversarial Model for RLAs}
+\subsection{Adversarial Model for RLAs}
 Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
-For our purposes, we will assume that the best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes were verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
+For our purposes, we will assume that the best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
 We will further assume that all software used in the RLA is open source and well-defined, so its output may be reproduced and thus verified by an observer wishing to do so with their own software.
 
-Referring to the ballot polling audit steps described above, we further assume that a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
-Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed from knowledge of the seed and the drawn ballots.
+Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
+Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed, because it can be checked by the public through knowledge of the seed and the drawn ballots.
 Thus the only variable is round size.
-We define a {\em weak adversary} as one who can choose the first round size and a {\em strong adversary} as one who can choose any round size.
 
-\subsection{The Literature and Open Questions}
+We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time before the round begins.
+
+\subsection{Open Questions}
 Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which does not need ballot order and relies only on sample and round tallies.
@@
 As described above, election officials, on the other hand, greatly prefer drawing many ballots at once.
-This preference is likely due to the following.
+From conversations with election officials and staff members of Verified Voting, Brennan Center and Common Cause who have been training election officials to perform RLAs, we estimate that this preference is likely due to the following.
 \begin{description}
@@
 For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
-\item Finally, in the current environment of misinformation, election officials would want to ensure that the probability of a misleading audit sample (falsely indicating that the loser won) is very small, which implies that round sizes should be large.
+\item Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
 \end{description}
@@
 In particular, this proof does not require an assumption about round sizes.
+
 We provide the following:
@@
 \subsection{Organization} 
-Section \ref{sec:related} describes related work.
+Sections \ref{sec:model} and \ref{sec:related} provide background, describing the audit model and related work respectively.
 Section \ref{sec:prov} describes the \Providence audit, Section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.

### 2023-02-12 providence/sections/model.tex
added model. worked more on intro.

--- 2022-10-11 providence/sections/related.tex
@@
-%Related Work
-Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
-
-The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn.
-
-The \Minerva audit \cite{usenix_minerva,arxiv_athena} was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
-First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins; simulations \cite{arxiv_athena} support these observations.
-Additional simulations \cite{simulations} have shown that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
-As expected, the advantage of \Minerva decreases for smaller stopping probability (smaller round sizes) as such round schedules approach the B2 round schedule (1,1,1,\ldots) for which \BRAVO is known to be most efficient.
-
-Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and to understand audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
-There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
-Simple workload estimates may be obtained by using the number of ballots drawn \cite{bernoulli-ballot-polling}, a more thorough workload estimation model includes the time taken to access individual ballots\cite{bernhard-diss}.
+% Model
 
@@
 \end{definition}
-
-The stopping conditions of \BRAVO and \Minerva rely on the following ratios.
-
-\begin{definition}[\BRAVO Ratio] \label{def:bravo-ratio} The \BRAVO audit uses the ratio $\sigma$.
-Consider a sample size of $n$ ballots with $k$ for the reported winner.
-The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
-\begin{equation}
-    \sigma(k, p_a, p_0, n) \triangleq \frac{p_a^{k} (1-p_a)^{n-k}}{p_0^{k} (1-p_0)^{n-k}} 
-    \label{eqn:bravoratio}
-\end{equation}
-\end{definition}
-
-In \BRAVO, $p_0=\frac{1}{2}$. A \BRAVO audit outputs correct if and only if
-$$\sigma(k,p_a,\frac{1}{2},n)\ge \frac{1}{\alpha}.$$
-
-%If testing the \BRAVO stopping condition after each individual ballot
-% is drawn (a \B \BRAVO audit), 
-It is easy to see that the ratio $\sigma$ is the 
-likelihood ratio:
-$$
-\frac{Pr[K=k|H_a,n]}{Pr[K=k|H_0,n]}= \frac{\binom{n}{k}p_a^{k} (1-p_a)^{n-k}}{\binom{n}{k}(\frac{1}{2})^n} =\sigma(k, p_a, \frac{1}{2}, n)
-$$
-
-\begin{comment}
-\begin{definition}[$(\alpha,p)$-\BRAVO ]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\BRAVO audit iff the following stopping condition is tested at each ballot draw.
-If the sample $X$ is of size $n$ and has $k$ ballots for the winner,
-\begin{equation}
-    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \text{Correct} & ~\sigma(k, p, \frac{1}{2}, n) 
-         %\triangleq \frac{p^{k} (1-p)^{n-k}}{(\frac{1}{2})^n} 
-        \geq \frac{1}{\alpha}\\
-        Undetermined & ~else 
-    \end{array}
-    \right .
-    \label{eqn:bravo}
-\end{equation}
-\end{definition}
-\end{comment}
-
-It now becomes useful to have shorthand for a sequence of cumulative round sizes and the corresponding sequence
-of cumulative winner ballot tallies.
-We use:
-$$\bm{k_j}\triangleq(k_1,k_2,\ldots,k_j)$$
-$$\bm{n_j}\triangleq(n_1,n_2,\ldots,n_j)$$
-
-Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
-
-\begin{definition}[\Minerva Ratio] \label{def:minerva_ratio} The \R \Minerva audit uses the ratio $\tau_j$.
-We use cumulative round sizes $\bm{n_j}$, with corresponding $\bm{k_j}$ ballots for the reported winner in each round.
-The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
-         \begin{equation}
-         \begin{aligned}
-             \label{eqn:tau}
-                 \tau_{j}(k_{j}, p_a,p_0, \bm{n_j}, \alpha )  \triangleq\\
-                 \frac{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_a, \bm{n_j}]}{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_0, \bm{n_j}]}
-         \end{aligned}
-         \end{equation}
-\end{definition}
-
-\begin{comment}
-\begin{definition}[$ (\alpha, p, \bm{n_j} ) $-\Minerva]
-     \label{def:minerva}
-     Given \B $(\alpha, p)$-\BRAVO and cumulative round sizes\\ $\bm{n_j}$, the corresponding \R \Minerva stopping rule for the $j^{th}$ round is:
- \begin{equation}
-     \mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \tau_{j}(k_{j}, p_a, \frac{1}{2}, \bm{n_j}, \alpha ) \geq \frac{1}{\alpha}\\
-             % & \\
-             % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
-             % & \\
-             Undetermined ~~else \\
-         \end{array}
-         \right .
-         \label{eqn:minerva-test}
- \end{equation}
-\end{definition}
-\end{comment}

### 2023-02-13 providence/sections/pilot.tex
.
tighter -> sharper

@@
 % pilot
-The Rhode Island Board of Elections performed a pilot audit in Providence 
+The Rhode Island Board of Elections performed a pilot audit in the city of Providence 
 in February 2022. The contest audited was a single yes-or-no question in the November 2021 election: Portsmouth's
@@
 \begin{description}
-\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a tighter test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
+\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
 It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.

### 2023-02-13 providence/sections/related.tex

@@
 
-We now summarize the model drawing largely from the notation and terminology of \cite{usenix_minerva,arxiv_athena,simulations,bravo}.
-The model is related work and not claimed to be original to this work.
+In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.
+A typical approach is to use a ballot serial number on both paper ballot and CVR.
+When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered.
+Additionally, some voting systems do not record a CVR.
+One may perform a transitive audit by rescanning unnumbered voted ballots with special scanners which produce CVRs and also print numbers on the ballots as they are scanned.
+This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible.
 
-An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
-\BRAVO and \Minerva are modeled as binary hypothesis tests where the null hypothesis $H_0$ corresponds to a tied election and the alternative hypothesis $H_a$ to an election tally as announced. 
-(When the number of ballots is odd, $H_0$ corresponds to the announced loser winning by one ballot.)
-Thus the null hypothesis is the outcome distinct from the announced one which is most difficult to detect; the probability of failing to detect it, given that the null hypothesis is true, is the worst case such probability and should be below the risk limit \cite{Bayesian-RLA}.
-
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff for sample $X$
-$$
-Pr[\mathcal{A}(X) 
-= \text{Correct} \,|\, H_0]\le \alpha
-$$
-\end{definition}
+A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
+Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch.
+This approach typically requires the sampling of a very large number of ballots.
+However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.
+(Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is.
+A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
 
@@
 
-Note that \BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT).
+\BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT) \cite{wald}.
 In the more general SPRT, the test can also reject the alternative hypothesis, and there is an additional parameter $\beta$, the probability of incorrectly rejecting the alternative hypothesis.

### 2023-02-13 providence/sections/adversary.tex
theoretical properties section

@@
 
-\subsection{Efficiency}
+\subsection{Theoretical Properties}
+Here we briefly give two interesting theoretical properties of \Providence.
+
+\subsubsection{Efficiency}
+In terms of expected ballots sampled, it is easy to see that \Providence dominates EoR \BRAVO in the sense that for any sample for which the EoR \BRAVO stopping condition is met, the \Providence stopping condition is met also.
 \begin{lemma}
@@
 \begin{proof} Appendix~\ref{sec:proofs}\end{proof}
+
+\subsubsection{Markov-like Stopping Condition}
+After $j-1$ rounds, having drawn $n_{j-1}$ cumulative ballots with $k_{j-1}$ cumulative winner ballots, the \Providence stopping condition in round $j$ with size $n_j$ is equivalent to the \Providence stopping condition for a second round where the first round size is $n_1=n_{j-1}$ with $k_1=k_{j-1}$ winner ballots and a second round size $n_2=n_j$.
+The particular sequence of round sizes and winner ballots found does not affect the \Providence stopping condition; only the previous cumulative round size and number of winner ballots matters.
+\Minerva does not have this property, but \BRAVO also does.
+
+\begin{lemma}\label{lemma:markov}
+Let $[n_1, \ldots, n_{j-1}, n_j]$ be a round schedule for an execution of \Providence audit that has not stopped
+in any of its first $j-1$ rounds (\textit{i.e.,} for every $i = 1, \ldots, j-1$:
+$k_i < k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $), then:
+
+\[ 
+k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j} .
+\]
+\end{lemma}
+\begin{proof}
+Appendix~\ref{sec:proofs}
+\end{proof}

### 2023-02-13 providence/sections/appendix.tex
theoretical properties section

@@
+\textbf{Lemma \ref{lem:efficiency}.}
+For any risk-limit $\alpha \in (0, 1)$, for any margin
+and for any round schedule $[n_1, \ldots, n_j]$, 
+the \Providence RLA is more efficient than EoR \BRAVO.
+
+\begin{proof}
+Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
+That is, the EoR \BRAVO stopping condition is true:
+$$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
+To see the \Providence stopping condition is fulfilled, we rewrite as 
+%$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
+
+%% From Lemma~\ref{lemma:tau1_increasing} 
+% \[
+% \tau_1(k, p_a, p_0, n_j) = 
+%  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
+% \]
+%is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
+% \textit{i.e.,}
+% \begin{equation}\label{eq_incr}
+% % \frac{1}{\alpha} \leq
+%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}.
+% \end{equation}
+
+%If the \BRAVO test is satisfied for $k_j$ then:
+
+% From the properties of binomial distribution we obtain:
+% \[
+%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
+%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
+%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
+% \]
+\[
+ \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) 
+\]
+\[
+ = \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \sigma(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1})  
+\]
+\[
+ \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) 
+\]
+\[
+ = \omega_r(k_j, k_{j-1}, p_a, p_0, n_j, n_{j-1}).
+\]
+
+Where inequality $(*)$ follows from \cite[Theorem 6]{arxiv_athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
+%Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
+%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
+%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.
+
+% 
+% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
+% 
+% 
+\end{proof}
 \begin{lemma}
@@
 
-\begin{lemma}\label{lemma:markov}
+\textbf{Lemma \ref{lemma:markov}.}
 Let $[n_1, \ldots, n_{j-1}, n_j]$ be a round schedule for an execution of  \Providence audit that has not stopped
@@
 \]
-\end{lemma}
 % \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
@@
 Compare this to the similar property for the $\Bravo$ stopping condition.
-
-\efficiencylemma
-
-\begin{proof}
-Let $[n_1, \ldots, n_j]$ be a round schedule, and assume that an EoR \BRAVO audit stops in round $j$, after observing $k_1, \ldots, k_j$ ballots for the announced winner in each round respectively.
-That is, the EoR \BRAVO stopping condition is true:
-$$\sigma(k_j,p_a,p_0,n_j) \ge \frac{1}{\alpha}.$$
-To see the \Providence stopping condition is fulfilled, we rewrite as 
-%$$\frac{1}{\alpha} \le \sigma(k_j,p_a,p_0,n_j)=\sigma(k_j,p_a,p_0,n_j)$$
-
-%% From Lemma~\ref{lemma:tau1_increasing} 
-% \[
-% \tau_1(k, p_a, p_0, n_j) = 
-%  \frac{P[K_r \geq {k} | H_a, n_j]}{P[K_r \geq {k} | H_0, n_j]}
-% \]
-%is a strictly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
-% \textit{i.e.,}
-% \begin{equation}\label{eq_incr}
-% % \frac{1}{\alpha} \leq
-%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}.
-% \end{equation}
-
-%If the \BRAVO test is satisfied for $k_j$ then:
-
-% From the properties of binomial distribution we obtain:
-% \[
-%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
-%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
-%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
-% \]
-\[
- \frac{1}{\alpha} \leq \sigma(k_{j}, p_a, p_0, n_{j}) 
-\]
-\[
- = \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \sigma(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1})  
-\]
-\[
- \leq^{(*)} \sigma(k_{j-1}, p_a, p_0, n_{j-1}) \cdot \tau_1(k_j - k_{j-1}, p_a, p_0, n_j - n_{j-1}) 
-\]
-\[
- = \omega_r(k_j, k_{j-1}, p_a, p_0, n_j, n_{j-1}).
-\]
-
-Where inequality $(*)$ follows from \cite[Theorem 6]{arxiv_athena}. Note that we apply this result on $\tau_j$ for just $j=1$.
-%Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
-%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
-%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.
-
-% 
-% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
-% 
-% 
-\end{proof}

### 2023-02-13 providence/sections/model.tex
incorporated wald; small changes

--- 2023-02-07 providence/sections/related.tex
@@
 An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
+As described in the introduction, we ignore a third option of moving to a hand count, leaving that decision to election officials.
+We provide more context on this in section \ref{sec:related}, on related work.
+
 \BRAVO and \Minerva are modeled as binary hypothesis tests where the null hypothesis $H_0$ corresponds to a tied election and the alternative hypothesis $H_a$ to an election tally as announced. 

### 2023-02-13 providence/sections/related.tex
incorporated wald; small changes

@@
 
+\BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT) \cite{wald}.
+Applying the general SPRT to RLAs, there would be a third output, {\em Full Manual Hand Count}, in addition to {\em Correct} and {\em Undetermined}.
+The test requires an additional error parameter $\beta$ (of audit error when the outcome is correct): the probability of requiring a (unnecessary) hand count when the outcome is correct.
+The test is:
+\begin{equation}
+    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \emph{Correct} & ~\sigma(k, p_a, p_0, n) 
+        \geq \frac{1-\beta}{\alpha}\\ \\ \\
+       \emph{Hand~Count} & ~\sigma(k, p_a, p_0, n) 
+        < \frac{\beta}{1-\alpha}\\ \\ \\
+        \emph{Undetermined} & ~else 
+    \end{array}
+    \right .
+    \label{eqn:wald}
+\end{equation}
+
+In the existing literature, ballot polling audits do not include this possibility (i.e.
+they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
+
 Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and to understand audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
@@
 
-In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+For completeness, we describe here other types of RLAs. 
+\begin{description}
+\item In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
 Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.
@@
 This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible.
-
-A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
+\item A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
 Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch.
@@
 A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
-
-The stopping conditions of \BRAVO and \Minerva rely on the following ratios.
-
-\begin{definition}[\BRAVO Ratio] \label{def:bravo-ratio} The \BRAVO audit uses the ratio $\sigma$.
-Consider a sample size of $n$ ballots with $k$ for the reported winner.
-The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
-\begin{equation}
-    \sigma(k, p_a, p_0, n) \triangleq \frac{p_a^{k} (1-p_a)^{n-k}}{p_0^{k} (1-p_0)^{n-k}} 
-    \label{eqn:bravoratio}
-\end{equation}
-\end{definition}
-
-In \BRAVO, $p_0=\frac{1}{2}$. A \BRAVO audit outputs correct if and only if
-$$\sigma(k,p_a,\frac{1}{2},n)\ge \frac{1}{\alpha}.$$
-
-%If testing the \BRAVO stopping condition after each individual ballot
-% is drawn (a \B \BRAVO audit), 
-If $K$ is the random variable indicating the number of ballots in the sample that contain a vote for the reported winner, it is easy to see that the ratio $\sigma$ is the likelihood ratio:
-$$
-\frac{Pr[K=k|H_a,n]}{Pr[K=k|H_0,n]}= \frac{\binom{n}{k}p_a^{k} (1-p_a)^{n-k}}{\binom{n}{k}(\frac{1}{2})^n} =\sigma(k, p_a, \frac{1}{2}, n)
-$$
-
-\begin{comment}
-\begin{definition}[$(\alpha,p)$-\BRAVO ]\label{def:bravo}  An audit $\mathcal{A}$ is the \B~$(\alpha, p)$-\BRAVO audit iff the following stopping condition is tested at each ballot draw.
-If the sample $X$ is of size $n$ and has $k$ ballots for the winner,
-\begin{equation}
-    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \text{Correct} & ~\sigma(k, p, \frac{1}{2}, n) 
-         %\triangleq \frac{p^{k} (1-p)^{n-k}}{(\frac{1}{2})^n} 
-        \geq \frac{1}{\alpha}\\
-        Undetermined & ~else 
-    \end{array}
-    \right .
-    \label{eqn:bravo}
-\end{equation}
-\end{definition}
-\end{comment}
-
-\BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT) \cite{wald}.
-In the more general SPRT, the test can also reject the alternative hypothesis, and there is an additional parameter $\beta$, the probability of incorrectly rejecting the alternative hypothesis.
-In an RLA, this corresponds to the audit having a third output: \emph{proceed to a full manual count of the ballots}.
-In the existing literature, ballot polling audits do not include this possibility (i.e.
-they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
-
-Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
-Now it becomes useful to have shorthand for a sequence of cumulative round sizes and the corresponding sequence
-of cumulative winner ballot tallies.
-We use:
-$$\bm{n_j}\triangleq(n_1,n_2,\ldots,n_j) \quad\text{and}\quad \bm{k_j}\triangleq(k_1,k_2,\ldots,k_j)$$
-Also, let $K_j$ be the random variable indicating the cumulative number of ballots in the sample after the $j$th round is drawn.
-
-\begin{definition}[\Minerva Ratio] \label{def:minerva_ratio} The \R \Minerva audit uses the ratio $\tau_j$.
-We use cumulative round sizes $\bm{n_j}$, with corresponding $\bm{k_j}$ ballots for the reported winner in each round.
-The proportion of ballots for the reported winner under the alternative hypothesis and null hypothesis are $p_a$ and $p_0$ respectively.
-         \begin{equation}
-         \begin{aligned}
-             \label{eqn:tau}
-                 \tau_{j}(k_{j}, p_a,p_0, \bm{n_j}, \alpha )  \triangleq\\
-                 \frac{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_a, \bm{n_j}]}{Pr[K_{j} \geq k_{j} \wedge \forall_{i < j} ({\mathcal{A}}(X_i) ~\neq \text{Correct}) \mid H_0, \bm{n_j}]}
-         \end{aligned}
-         \end{equation}
-\end{definition}
-
-\begin{comment}
-\begin{definition}[$ (\alpha, p, \bm{n_j} ) $-\Minerva]
-     \label{def:minerva}
-     Given \B $(\alpha, p)$-\BRAVO and cumulative round sizes\\ $\bm{n_j}$, the corresponding \R \Minerva stopping rule for the $j^{th}$ round is:
- \begin{equation}
-     \mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \tau_{j}(k_{j}, p_a, \frac{1}{2}, \bm{n_j}, \alpha ) \geq \frac{1}{\alpha}\\
-             % & \\
-             % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
-             % & \\
-             Undetermined ~~else \\
-         \end{array}
-         \right .
-         \label{eqn:minerva-test}
- \end{equation}
-\end{definition}
-\end{comment}
+\end{description}
 

### 2023-02-13 providence/sections/intro.tex
added margin and bravo references to beta

@@
 One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
+To have any value, the audits should be performed before the election is certified.
 
@@
 The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination).
-In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count.
-However, a manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary, hence (c) is generally not incorporated.
+In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count, which is generally not incorporated \cite{bravo}.
+A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary.
 Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
@@
 \item Based on the manual interpretations, the stopping condition is computed. 
-\item If more ballots are to be drawn, the next round size is chosen.
+\item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
+\end{enumerate}
+
 Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
 For RLAs required by statute or law, certification deadlines would play a large role, as the audit would need to be completed before the deadline.
-\end{enumerate}
 
-A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections.
+A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
 A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
@@
 
-Election officials typically draw ballots in large round sizes:  see for example \cite{va-2022,RI-report}.
-Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping).
+Election officials typically draw ballots in large round sizes, see for example \cite{va-2022,RI-report}.
+Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common).
 Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).

### 2023-02-13 providence/sections/model.tex
added margin and bravo references to beta

@@
 As described in the introduction, we ignore a third option of moving to a hand count, leaving that decision to election officials.
-We provide more context on this in section \ref{sec:related}, on related work.
+This is standard in the literature, see, for example, \cite{bravo}.
+We provide more detail on the use of this option in section \ref{sec:related}, on related work.
 
@@
 
+Notice that the ratios for \BRAVO and \Minerva are for two candidates.
+They may both be extended to multi-candidate plurality elections through the computation of $\ell -1$ audits for $\ell$ candidates by pairing the announced winner with all other candidates.
+
+The {\em margin}, given a contest and two candidates, is the difference between the votes received by each as a fraction of the total number of votes.
+
 \begin{comment}

### 2023-02-13 providence/sections/related.tex
added margin and bravo references to beta

@@
 
-In the existing literature, ballot polling audits do not include this possibility (i.e.
+An example of the above use is \cite{RLA}.
+In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e.
 they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
@@
 
-For completeness, we describe here other types of RLAs. 
+For completeness, we describe other types of RLAs below. 
 \begin{description}

### 2023-02-13 providence/sections/appendix.tex
mostly proofs

@@
     Given a monotone increasing sequence: $\frac{a_1}{b_1}, \frac{a_2}{b_2}, \ldots, \frac{a_n}{b_n}$, for $a_i, b_i > 0$, the sequence:
-    $$z_i = \frac{\sum_{j=i}^n a_j}{\sum_{j=i}^n b_j}$$
+    $z_i = \frac{\sum_{j=i}^n a_j}{\sum_{j=i}^n b_j}$
     is also monotone increasing.
@@
 Then this is simply equal to
- $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)}
- $$$$
- =\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
- \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
+ $$\sigma(k_m, p_a, p_0, n_m)\cdot\frac{\text{Bin}(k_{m+1}',n_{m+1}',p_a)}{\text{Bin}(k_{m+1}',n_{m+1}',p_0)} = \sigma(k_{m+1}, p_a, p_0, n_{m+1})
  $$
- $$
- =\sigma(k_{m+1}, p_a, p_0, n_{m+1})
- $$
+ %$$
+ %=\frac{p_a^{k_m} (1-p_a)^{n_m-k_m}}{p_0^{k_m} (1-p_0)^{n_m-k_m}} \cdot
+% \frac{p_a^{k_{m+1}'} (1-p_a)^{n_{m+1}'-k_{m+1}'}}{p_0^{k_{m+1}'} (1-p_0)^{n_{m+1}'-k_{m+1}'}}
+ %$$
+ %$$
+ %=\sigma(k_{m+1}, p_a, p_0, n_{m+1})
+ %$$
 \end{proof}
@@
 \end{definition}
-As we have seen in Lemma \ref{lemma:minerva2_kmin_exists}, such a value of $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}$ exists and $k_j \geq k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $ if and only if the result of the audit is Correct, (\textit{i.e.,} the stopping condition in Definition~\ref{def:minervatwo} holds).
+As we have seen in Lemma \ref{lemma:minerva2_kmin_exists}, such a value of $k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}$ exists and $k_j \geq k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $ if and only if the result of the audit is Correct.
+% (\textit{i.e.,} the stopping condition in Definition~\ref{def:minervatwo} holds).
 
-The following lemma shows a Markov-like property of \Providence audit (\textit{i.e.,}
-for an audit that has not stopped in the first $j-1$ rounds, only cumulative results of the round $j-1$ matter: cumulative sample size $n_{j-1}$ and the number of ballots for the winner $k_{j-1}$).
+%The following lemma shows a Markov-like property of \Providence audit (\textit{i.e.,}
+%for an audit that has not stopped in the first $j-1$ rounds, only cumulative results of the round $j-1$ matter: cumulative sample size $n_{j-1}$ and the number of ballots for the winner $k_{j-1}$).
 
-\textbf{Lemma \ref{lemma:markov}.}
-Let $[n_1, \ldots, n_{j-1}, n_j]$ be a round schedule for an execution of  \Providence audit that has not stopped
-in any of its first $j-1$ rounds (\textit{i.e.,} for every $i = 1, \ldots, j-1$:
-$k_i < k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $), then:
+%\textbf{Lemma \ref{lemma:markov}.}
+%Let $[n_1, \ldots, n_{j-1}, n_j]$ be a round schedule for an execution of  \Providence audit that has not stopped
+%in any of its first $j-1$ rounds (\textit{i.e.,} for every $i = 1, \ldots, j-1$:
+%$k_i < k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} $), then: 
+%\[ 
+%$k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j}$.
+%\]
+% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
+%\begin{proof}
+%Let $k_{j-1}$ denote the number of ballots drawn for the declared winner up to the round $j-1$ (out of $n_{j-1}$ sampled ballots). The stopping decision for the round $j$ is made as follows:
+%
+%\[
+% k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
+%\]
+%\[
+%  =  k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j}  
+%\]
+%This is easily observed. 
+%\end{proof}
 
-\[ 
-k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j} = k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j} .
-\]
-% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
-\begin{proof}
-Let $k_{j-1}$ denote the number of ballots drawn for the declared winner up to the round $j-1$ (out of $n_{j-1}$ sampled ballots). The stopping decision for the round $j$ is made as follows:
-
-\[
- k^{p_a, p_0, \alpha, k_{j-1}}_{min, j, n_{j-1}, n_j}  = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
-\]
-\[
-  =  k^{p_a, p_0, \alpha, k_{j-1}}_{min, 2, n_{j-1}, n_j}  
-\]
-
-\end{proof}
-
-That is, the stopping condition is equivalent to that of a two round audit with the same cumulative votes for the winner and cumulative round sizes: the first round is of size $n_{j-1}$ and has $k_{j-1}$ votes for the winner, and the second (cumulative) round size is $n_j$ with $k_j$ (cumulative) votes for the winner.
+%That is, the stopping condition is equivalent to that of a two round audit with the same cumulative votes for the winner and cumulative round sizes: the first round is of size $n_{j-1}$ and has $k_{j-1}$ votes for the winner, and the second (cumulative) round size is $n_j$ with $k_j$ (cumulative) votes for the winner.
 Compare this to the similar property for the $\Bravo$ stopping condition.

### 2023-02-13 providence/sections/intro.tex
mostly proofs

@@
 \subsection{Adversarial Model for RLAs}
+\label{sec:adv}
 Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
@@
 
-We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time before the round begins.
+We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time (before, of course, that particular round begins).
+In particular, a strong adversary may use information about the sample drawn thus far to decide the next round size, but a weak adversary may not.
 

### 2023-02-13 providence/sections/prov.tex
mostly proofs

@@
 
-Recall that the proof that the \Minerva audit is risk-limiting assumes that the round schedule of \Minerva is predetermined and that, in particular, an adversarial auditor cannot determine the next round size after drawing a sample.
+Recall that the proof that the \Minerva audit is risk-limiting assumes that the round schedule of \Minerva is predetermined and that, in particular, an adversarial auditor cannot determine the next round size after drawing a sample (we termed such an adversary a weak adversary in section \ref{sec:adv}).
 This presents difficulties because a non-adversarial election official might want to draw a small next round if the current sample comes close to satisfying the risk limit.
@@
 Conversely, if the current sample is not at all close to satisfying the risk limit, it would be advantageous to draw a larger round than the predetermined round size.
+Thus it would be advantageous to have an RLA with \Minerva's efficiency advantages in the face of an adversarial auditor who could choose next round sizes after drawing the sample (we termed such an adversary a strong adversary).
+This would enable honest participants to better plan audits.
 
-We now formalize these notions of weak and strong adversaries.
+Before proceeding to define the audit \Providence which is an RLA in the presence of the stronger adversary, we formalize the notions of weak and strong adversaries introduced in section \ref{sec:adv}.
 
 \begin{definition}[Weak Adversary]
-A {\emph weak adversary} may choose the first round size as a function of audit parameters. That is, the first round size is a function $$n_1(\alpha, p_a, p_0).$$
+A {\emph weak adversary} may choose the first and consequent round sizes as a pre-determined function of audit parameters.
+That is, the $j^{th}$ round size is a function $$n_j(\alpha, p_a, p_0)$$ determined before the audit begins.
 \end{definition}
@@
 \begin{definition}[Strong Adversary]
-A {\emph strong adversary} may choose any round size as a function of audit parameters and all precreding samples. That is, the first round size is a function 
+A {\emph strong adversary} may choose any round size as any function of audit parameters and all preceding samples. That is, the first round size is a function 
 $$n_1(\alpha, p_a, p_0),$$
 and for all rounds $j\ge 2$, the round size is a function
-$$n_{j+1}(\alpha, p_a, p_0, \bm{k_{j-1}}, \bm{n_{j-1}}).$$
+$$n_{j}(\alpha, p_a, p_0, \bm{k_{j-1}}, \bm{n_{j-1}})$$ The functions $n_j$, $j \geq 1$ may be chosen at any time before the $j^{th}$ round begins. 
 \end{definition}
@@
 Before defining \Providence, we give some intuition for how it is designed to avoid the problem of \Minerva.
-In round $j$, for every possible value of $K_j$ for which the audit stops, a risk is incurred.
+
+For any ballot polling audit, in round $j$, designed to stop for some value of $K_j$, however large $K_j$ is, there is, in general, a probability, however small, that the value of $K_j$ has resulted from an underlying incorrect election outcome.
+Thus, corresponding to each such value of $K_j$, a risk---probability that the audit will stop given that the outcome is incorrect---is incurred.
 To obtain the total risk for the round, one adds the risks corresponding to each value of $K_j$ for which the audit can stop, weighted by the probability of drawing that value of $K_j$.
-The stopping condition provides relationships among various quantities.
+The stopping condition may provide relationships among the various quantities.
 
@@
 Separate relationships between risk and stopping probability are not available for individual values of $K_j$.
-If the next round size depends on $K_j$, we do not have expressions relating the risks, and so \Minerva may be vulnerable to an adversary choosing round sizes \cite{usenix_minerva, arxiv_athena}.
+If the next round size depends on $K_j$, expressions relating the risks are not available, and so we are not able to say if \Minerva is an RLA in this case, or if it is vulnerable to an adversary choosing round sizes \cite{usenix_minerva, arxiv_athena}.
 
 In \Providence, we choose a stopping condition that applies separately to the risk and stopping probabilities for each value of $K_j$, avoiding the problem of \Minerva, and allowing for optimal round size choices, which depend on the drawn sample.
-The \Providence audit is risk-limiting even if an adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.
+The \Providence audit is risk-limiting even if a strong adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.
 
@@
 \label{thm:minerva2_is_rla_new}
-An $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence audit is an
+$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence is an
 $\alpha$-RLA.
@@
 
-Summing both sides over all values of $k_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively).
+Summing both sides over all values of $k_{j-1} < k_{min, j-1}$ gives us a similar relationship between the probabilities of stopping in round $j$ (given the null and alternate hypotheses respectively)\footnote{In fact, this is the relationship \Minerva enforces for its stopping condition, additionally requiring that $n_j$ be fixed for all values of $k_{j-1}$.}.
 When both sides of the inequality are further summed over all rounds, we get:

### 2023-02-13 providence/sections/intro.tex
not clear enough
some changes to intro

@@
 Equivalently, this is also a guarantee limiting the probability of audit error when the outcome is incorrect.
-The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct (and not proceed to a full hand count).
+The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct.
 Lower risk limits are better.
@@
 The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination).
-In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count, which is generally not incorporated \cite{bravo}.
+In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count.
+This option is generally not incorporated \cite{bravo}.
 A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary.
@@
 There are three types of RLAs: ballot comparison RLAs, batch comparison RLAs and ballot polling RLAs.
-This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere) and in real statewide audits (Georgia, Virgina) \cite{vv_audits} as well as in audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere), real statewide audits (Georgia, Virgina) \cite{vv_audits} and  audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
 Ballot and batch comparison RLAs are described in \ref{sec:related}.
@@
 \begin{enumerate}
-\item A first round \cite{usenix_minerva} size---the number of ballots sampled before first checking the stopping condition---is chosen. 
-\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator typically seeded by a natural source of randomness like rolling dice.
+\item The ballot manifest is published. 
+\item A first round size \cite{usenix_minerva} is chosen. %\fpo{the sentence is hard to understand}. pv: fixed
+\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator, which is typically seeded by a natural source of randomness like rolling dice.
 \item The physical ballots are found and manually interpreted, recording the manual interpretations. 
@@
 
+The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants.
+Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
+\BRAVO is also used for what are termed {\em stratified} audits [REF], where a single contest is audited using different approaches (ballot polling, ballot comparison, batch comparison) in different counties.
+When a ballot polling audit is used for the entire set of ballots, the stopping condition assumes a tied outcome (the hardest incorrect outcome to detect) when measuring the risk limit.
+In a stratified audit, however, when testing individual units such as counties, one may need to assume a different outcome.
+This is because, for example, if a county with an announced large number of ballots for the winner were actually tied, the overall outcome would be incorrect.
+The proof of \Minerva's risk-limiting property assumes a tied outcome while measuring risk, and it is not immediately obvious how the proof may be modified to allow other possibilities.
+Thus, while \Minerva may be used for multi-candidate contests when all counties participate in the ballot polling audit, it cannot be used for stratified audits.
+\BRAVO, and \Providence, on the other hand, can.
+
 A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
@@
 Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common).
-Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
+For the two audits we attended, election officials chose stopping probabilities of at least $0.9$.
+We are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
 \BRAVO hence cannot be used as a B2 audit in these scenarios.
@@
 SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
-But it requires the additional effort of tracking the order of ballots.
+But it requires the additional effort of tracking the order of ballots, and should be expected to be inefficient because it does not use the information in the ballots drawn at the end.
 
@@
 
-Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
+Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that the ballot manifest is verified by the compliance audit; a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
 Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed, because it can be checked by the public through knowledge of the seed and the drawn ballots.
@@
 In particular, an open question about \Minerva is whether the computation of a risk limit assuming a weak adversary applies to an attack by a strong adversary, or is the risk limit computation incorrect when the adversary is strong?
-Can the strong adversary increase the audit's error probability beyond its declared risk limit?
+Can the strong adversary change the election outcome and then increase the audit's error probability beyond its declared risk limit?
 Or is there no probabilistic adversarial advantage to being able to compute next round sizes after knowing the drawn sample?
@@
 For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each).
-For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
+%For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
 \item Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
@@
 Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to a strong adversary.
+Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva.
+
 The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
-We are able to derive a new stopping condition for which a far simpler proof of the risk-limiting property is possible.
-In particular, this proof does not require an assumption about round sizes.
+Further, it assumes that the null hypothesis is a tie.
+A far simpler proof is possible for the risk-limiting property of \Providence.
+In particular, the proof does not require an assumption about round sizes and does not assume any particular fractional vote distribution for the null.
 
@@
 
+\Providence may be used in any audit, stratified or otherwise, where audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.
+
 %Our results demonstrate the superiority of \Providence over the other audits. Our work may be used by election officials to plan ballot polling audits, including in Georgia and Pennsylvania in 2022.
@@
 Sections \ref{sec:model} and \ref{sec:related} provide background, describing the audit model and related work respectively.
-Section \ref{sec:prov} describes the \Providence audit, Section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
+Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
 Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2020 US Presidential election in the state of Virginia.

### 2023-02-13 providence/sections/related.tex
some changes to intro

@@
 This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible.
+There is recent work on {\em adaptive comparison} RLAs, which relies on scanning only the batches necessary \cite{adaptive-comparison}.
+In either case, the workload of a transitive ballot comparison audit, measured in hours, would not be simply a linear function the number of ballots drawn, but would include a term which would depend on the total number of ballots rescanned.
 \item A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
-Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch.
-This approach typically requires the sampling of a very large number of ballots.
+Thus, while this type of audit does not need CVRs, it does need the publication of the tally of each batch.
+Batch comparison typically requires the sampling of a very large number of ballots, larger than polling audits except for small enough margins.
 However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.

### 2023-02-13 providence/sections/intro.tex

@@
 
-The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants.
+The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants and formulated as a binary hypothesis test.
 Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
+
 \BRAVO is also used for what are termed {\em stratified} audits [REF], where a single contest is audited using different approaches (ballot polling, ballot comparison, batch comparison) in different counties.
-When a ballot polling audit is used for the entire set of ballots, the stopping condition assumes a tied outcome (the hardest incorrect outcome to detect) when measuring the risk limit.
-In a stratified audit, however, when testing individual units such as counties, one may need to assume a different outcome.
+When a ballot polling audit is used for the entire set of ballots, the null hypothesis is a tied outcome (the hardest incorrect outcome to detect).
+In a stratified audit, however, when testing individual units such as counties, one may need to assume a different null.
 This is because, for example, if a county with an announced large number of ballots for the winner were actually tied, the overall outcome would be incorrect.
-The proof of \Minerva's risk-limiting property assumes a tied outcome while measuring risk, and it is not immediately obvious how the proof may be modified to allow other possibilities.
+The proof of \Minerva's risk-limiting property assumes a tied null, and it is not immediately obvious how the proof may be modified to allow other possibilities.
 Thus, while \Minerva may be used for multi-candidate contests when all counties participate in the ballot polling audit, it cannot be used for stratified audits.
@@
 This presents a major limitation, because the stopping probability of the next round is better estimated using information of the sample drawn thus far, but this would not be allowed for \Minerva.
-The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter.
-Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
-However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that a fixed draw size is sufficient to limit adversaries to weak ones, though this is not explicitly proven in \cite{usenix_minerva}.
+The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter\footnote{Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
+However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that a fixed draw size is sufficient to limit adversaries to weak ones, though this is not explicitly proven in \cite{usenix_minerva}.}.
+This could lead to the drawing of too few or too many ballots and greatly constrains audit planning.
 
@@
 From conversations with election officials and staff members of Verified Voting, Brennan Center and Common Cause who have been training election officials to perform RLAs, we estimate that this preference is likely due to the following.
-\begin{description}
-\item Firstly, each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
-\item Secondly, there is an overhead to finding a storage box and unsealing it.
+
+Firstly, each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored).
+Secondly, there is an overhead to finding a storage box and unsealing it.
 For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each).
 %For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
-\item Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
-\end{description}
+Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
+
 Thus the workload of an audit is not simply a linear (or affine) function of the number of ballots drawn.
@@
 Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva.
+\BRAVO is a very flexible audit enabling audit planning; with the introduction of \Providence, similar planning is possible with greater efficiency.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than 0:03) the difference in number of ballots sampled using \Providence over \BRAVO could correspond to many days of work which would need to be completed before a certification deadline.
 

### 2023-02-13 providence/sections/intro.tex
Update intro.tex

@@
 
-\BRAVO is also used for what are termed {\em stratified} audits [REF], where a single contest is audited using different approaches (ballot polling, ballot comparison, batch comparison) in different counties.
+%\BRAVO is also used for what are termed {\em stratified} audits [REF], where a single contest is audited using different approaches (ballot polling, ballot comparison, batch comparison) in different counties.
 When a ballot polling audit is used for the entire set of ballots, the null hypothesis is a tied outcome (the hardest incorrect outcome to detect).

### 2023-02-13 providence/sections/sims.tex
a

--- 2022-10-12 providence/sections/sims.tex
@@
 For each margin, we perform $10^4$ \Providence audit trials each on a tied election (hypothesis $H_0$, the null hypothesis) and the election as reported (hypothesis $H_a$, the alternate hypothesis).
-All trials have risk limit $\alpha = 0.1$, a maximum of $5$ rounds, and a conditional stopping probability of $0.90$ in each round.
+We use risk limit $\alpha = 0.1$, as is common in RLAs, see for example \cite{va-2022} and \cite{RI-report}.
+All trials have a maximum of $5$ rounds and a conditional stopping probability of $0.90$ in each round.
 That is, each next round size is selected to be large enough to give a $0.90$ conditional probability of stopping in that round, assuming the announced tally is correct and given the tally of previous rounds.

### 2023-02-13 providence/sections/intro.tex
Update intro.tex

@@
 Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to a strong adversary. 
-Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva.
+%Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva. 
 \BRAVO is a very flexible audit enabling audit planning; with the introduction of \Providence, similar planning is possible with greater efficiency.
@@
 The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
-Further, it assumes that the null hypothesis is a tie.
-A far simpler proof is possible for the risk-limiting property of \Providence.
-In particular, the proof does not require an assumption about round sizes and does not assume any particular fractional vote distribution for the null.
+%Further, it assumes that the null hypothesis is a tie. A far simpler proof is possible for the risk-limiting property of \Providence. 
+In particular, the proof does not require an assumption about round sizes.
 
@@
 
-\Providence may be used in any audit, stratified or otherwise, where audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.
+\Providence may be used in any audit where audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.
 

### 2023-02-13 providence/sections/related.tex
Update related.tex

@@
 Additional simulations \cite{simulations} have shown that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
-As expected, the advantage of \Minerva decreases for smaller stopping probability (smaller round sizes) as such round schedules approach the B2 round schedule (1,1,1,\ldots) for which \BRAVO is known to be most efficient.
+As expected, the advantage of \Minerva decreases for smaller stopping probability (smaller round sizes) as it approaches a B2 audit,
+%round schedule (1,1,1,\ldots) 
+for which \BRAVO is known to be most efficient.
 
@@
 
-Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and to understand audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
+Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and understanding audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
 There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
@@
 
-For completeness, we describe other types of RLAs below.  
-\begin{description}
-\item In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+For completeness, we describe other types of RLAs.
+In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
 Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.
-A typical approach is to use a ballot serial number on both paper ballot and CVR.
+Not all voting systems record CVRs and their use can present privacy challenges.
+%A typical approach is to use a ballot serial number on both paper ballot and CVR.
 When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered.
@@
 In either case, the workload of a transitive ballot comparison audit, measured in hours, would not be simply a linear function the number of ballots drawn, but would include a term which would depend on the total number of ballots rescanned.
-\item A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
-Thus, while this type of audit does not need CVRs, it does need the publication of the tally of each batch.
+A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch.
+%Thus, while this type of audit does not need CVRs, it does need the publication of the tally of each batch.
 Batch comparison typically requires the sampling of a very large number of ballots, larger than polling audits except for small enough margins. 
-However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.
+%However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied.
 (Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is.
 A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
-\end{description}
 

### 2023-02-13 providence/sections/workload.tex
.

@@
 Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
-For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_p$.
+Note that some other county may be slower, having fewer votes but also less auditing resources; but still, a slowest county exists. In this example, we take it to be Fairfax, the largest.
+For the slowest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_p$.
 Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
@@
 After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
-This is an undesirable situation, as it can appear to Alice's supporters that election officials are simply drawing ballots till a chosen outcome is obtained.
+This is an undesirable situation, as it can appear to Bob's supporters that election officials are simply drawing ballots till a chosen outcome is obtained.
 

### 2023-02-13 providence/sections/workload.tex
.

@@
 Bob's supporters celebrate their victory that the audit has in fact revealed that Bob really won, but the election officials have to explain that they are moving on to a second round.
-After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
+So there is a reason to structure round sizes so that they are unlikely to appear to support a true loser.
+
+%After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
 This is an undesirable situation, as it can appear to Bob's supporters that election officials are simply drawing ballots till a chosen outcome is obtained.

### 2023-02-13 providence/sections/model.tex
minor typo
various

@@
 This is standard in the literature, see, for example, \cite{bravo}.
-We provide more detail on the use of this option in section \ref{sec:related}, on related work.
+We provide more detail on the use of this option in Section \ref{sec:related}, on related work.
 
@@
 
+Here, if we were to condition the ratio on $\bm{k_{j-1}}$, we would get the binomial distributions and be running a new first round each time.
+
 Notice that the ratios for \BRAVO and \Minerva are for two candidates.

### 2023-02-13 providence/sections/prov.tex
add strong adversary to main theorem statement
various

@@
     \label{def:minervatwo}
-    For cumulative round size $n_j$ for round $j$ and a cumulative $k_j$ ballots for the reported winner found in round $j$, the \R \Providence stopping rule for the $j^{th}$ round is:
+    For cumulative round size $n_j$ for round $j$ and a cumulative $k_j$ ballots for the reported winner found in round $j$, where samples are drawn with replacement, the \R \Providence stopping rule for the $j^{th}$ round is:
 $$
@@
 Notice that \Providence requires the computation of $\tau_j$ for $j=1$ and no other values of $j$.
-The value of $\tau_1$ is simply the ratio of the the tails of the binomial distributions for the two hypotheses and can be fairly efficiently computed.
+The value of $\tau_1$ is simply the ratio of the the tails of the binomial distributions (the distributions are binomial because the auditing is with replacement) for the two hypotheses and can be fairly efficiently computed.
 The computation of $\tau_j$ for $j \geq 2$, as required in \Minerva, relies on the convolution of two probability distribution functions and is hence computationally considerably more expensive.
@@
 $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence is an
-$\alpha$-RLA.
+$\alpha$-RLA in the presence of a strong adversary.
 \end{theorem}

### 2023-02-13 providence/sections/pilot.tex
alpha and section not Section

@@
 For this pilot audit, extensive planning of the round schedule was not necessary because the margin was large enough that relatively few ballots were needed to achieve the high probability of stopping.
-In Section~\ref{sec:workload} we consider a larger state-wide contest in Virginia, where selecting the round schedule has more significant implications.
+In section~\ref{sec:workload} we consider a larger state-wide contest in Virginia, where selecting the round schedule has more significant implications.
 Virginia also currently uses ballot polling RLAs, whereas Rhode Island primarily uses batch comparison RLAs.
-Some of the ideas introduced in Section~\ref{sec:workload} provide a context for this pilot case as well.
+Some of the ideas introduced in section~\ref{sec:workload} provide a context for this pilot case as well.
 
@@
 \end{description}
-Both these ideas are addressed more thoroughly in Section~\ref{sec:workload}.
+Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.
 %first draft, needed to be shorter: ...stopping condition for the cumulative tally at the end of the round would see that the \BRAVO condition is not met (and would be correct that an EoR \BRAVO audit would require a second round for this sample).

### 2023-02-13 providence/sections/workload.tex
alpha and section not Section

@@
 Importantly, note that \Minerva does not appear in the analysis in this section. 
-Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in Section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples.
+Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples.
 \Minerva is not known to be risk-limiting in this setting, and thus cannot be used for RLAs that proceed in this way.
@@
 
-Recall from Section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}.
+Recall from section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}.
 In particular, the audit passed with an SO \BRAVO risk measure of $0.0541$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $0.366$.

### 2023-02-18 providence/sections/adversary.tex
mostly proofs
Lots of changes

@@
 Both the audits of our toy example therefore begin with a first round size of $17,272$ with a $0.9$ probability of stopping, and both will stop in the first round if the sample contains at least $8,725$ ballots for the winner.
+The audits are identical in the first round, and both numbers of ballots may be computed as described in \cite{usenix_minerva} and integrated into Arlo.
 We now consider two cases for which the audit proceeds to a second round.
-\begin{description}
-\item In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
+
+In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
 In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $.954$.
 The \Providence audit samples more than $9,000$ fewer ballots with a round size of $34,078$, achieving the desired $0.9$ probability of stopping.
-\item In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
+
+In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
 In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $0.727$ probability of stopping, significantly less than the desired $0.9$.
 Again, the \Providence audit can scale up the second round size according to the first sample and achieve the desired $0.9$ probability of stopping with $58,007$ ballots.
-\end{description}
 
@@
 \begin{proof}
-Appendix~\ref{sec:proofs}
+This is easily observed. 
+%Appendix~\ref{sec:proofs}
 \end{proof}

### 2023-02-18 providence/sections/intro.tex
various
Lots of changes

@@
 {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome.
-{\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; the evidence is examined so that incorrect outcomes may be detected with high probability.
-One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
-To have any value, the audits should be performed before the election is certified.
-
-A tabulation audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count.
-The integrity of an audit may be judged by how it deals with incorrect outcomes.
-A {\em risk-limiting audit (RLA)} guarantees a minimum probability that it will perform as it is supposed to (escalate to a full hand count) if the outcome is incorrect.
-Equivalently, this is also a guarantee limiting the probability of audit error when the outcome is incorrect.
-The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct.
-Lower risk limits are better.
-
-Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified.
-
+The {\em evidence-based elections} \cite{evidence-based} approach requires that election outcomes be supported by strong affirmative evidence.
+The evidence is generated during the election and publicly examined at the end, so that citizens may determine whether it provides strong support of the election outcome or is not sufficiently convincing.
+%by requiring not merely the use software independent systems to produce trustworthy evidence of outcome correctness; but also the examination of the evidence so that incorrect outcomes may be detected with high probability.
+One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely and procedures were followed; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.
+%Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified. 
 %We propose the \Providence audit, a new approach to a particular type of RLA---the ballot polling RLA described below---and propose a new model for the work load of an election.
@@
 
+As a formalized approach to the examination of the evidence supporting vote tabulation, an RLA is an important part of an evidence-based election.
+When correctly used, it serves to greatly improve the trustworthiness of an election.
+
+Significant efforts by nationwide organizations (such as Verified Voting, Brennan Center for Justice, Common Cause and Democracy Fund), local organizations, citizen advocates and experts have led to great progress towards the use of RLAs.
+Nonprofit VotingWorks has developed open-source election audit software, Arlo \cite{arlo}, and provides training in its use.
+Six states (Colorado, Georgia, Nevada, Pennsylvania, Rhode Island, Virginia) now require RLAs; three have statutory pilot programs (Indiana, Kentucky, Texas); four allow RLAs to satisfy a more general audit requirement (California, Ohio, Oregon, Washington); and two have an administrative pilot program (Michigan, New Jersey) \cite{vv_audits}.
+The effort to broaden the reach of RLAs continues, as election officials appear very keen to improve the trustworthiness of the elections they administer.
+
 \subsection{Background on RLAs}
-All RLAs sample one or more ballots at random; we will refer to each such set of ballots as a {\em round}.
-The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination).
-In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count.
-This option is generally not incorporated \cite{bravo}.
-A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary.
-Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
-They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
+A tabulation audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count.
+The integrity of an audit may be judged by how it deals with incorrect outcomes.
+A {\em risk-limiting audit (RLA)} guarantees a minimum probability that it will perform as it is supposed to (escalate to a full hand count) if the outcome is incorrect.
+Equivalently, this is also a guarantee of a maximum probability with which it will perform erroneously (declare the audit correct) when the outcome is incorrect.
+The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct.
+Lower risk limits are better.
 
 There are three types of RLAs: ballot comparison RLAs, batch comparison RLAs and ballot polling RLAs.
-This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere), real statewide audits (Georgia, Virgina) \cite{vv_audits} and  audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
-Ballot and batch comparison RLAs are described in \ref{sec:related}.
+This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere), real statewide audits (Georgia, Virginia) \cite{vv_audits} and  audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}.
+Ballot and batch comparison RLAs are described in section \ref{sec:related}.
 
-\subsection{Ballot Polling RLAs}
+\subsubsection{The Workflow of a Ballot Polling RLA}
 \label{sec:polling}
-Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system.
+%Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system.
 For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots.
-A ballot polling RLA \cite{RLA}, on the other hand, simply tallies the sampled ballots.
+A ballot polling RLA \cite{RLA} is based on the manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
+%, on the other hand, simply tallies the sampled ballots.
 It requires a much larger number of ballots than a ballot comparison RLA, but tends to be more feasible because it does not require any additional functionality of the voting system; in particular it does not require the system to generate or provide CVRs.
-What is needed is a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example).
+More detail about the storage of ballots is required, however: a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example) to enable the sampling of specific ballots from the list.
+
+All RLAs draw one or more ballots at a time; each such set of ballots is referred to as a {\em round}.
+We use notation and terminology from \cite{usenix_minerva,arxiv_athena,simulations,bravo} and also assume ballots are drawn with replacement.
 
@@
 \item A first round size \cite{usenix_minerva} is chosen. %\fpo{the sentence is hard to understand}. pv: fixed
-\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator, which is typically seeded by a natural source of randomness like rolling dice.
+\item Ballots on the ballot manifest are sampled uniformly at random, with replacement, using a pseudorandom number generator, which is typically seeded by a natural source of randomness like rolling dice.
 \item The physical ballots are found and manually interpreted, recording the manual interpretations. 
-\item Based on the manual interpretations, the stopping condition is computed. 
+\item The stopping condition $\mathcal{A}$, a function of the manual interpretations of the current cumulative sample of ballots $X$, is computed. It outputs: 
+\begin{enumerate}
+\item \emph{Correct: stop the audit} or 
+\item \emph{Undetermined: sample more ballots}. 
+\end{enumerate}
+Election officials may choose to abort this procedure and go to a full hand count at any time; we discuss this in more detail in section \ref{sec:related}.  
 \item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
@@
 Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
-For RLAs required by statute or law, certification deadlines would play a large role, as the audit would need to be completed before the deadline.
+For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there might be no legal means of changing the outcome and this would significantly impact citizen confidence.
+For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome.
+This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size deternination.
 
-The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants and formulated as a binary hypothesis test.
+%In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count.
+This option is generally not incorporated \cite{bravo}.
+A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary.
+% They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
+%This value is the smallest risk limit of the RLA that would stop at the end of the round. 
+%The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants and formulated as a binary hypothesis test.
 Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
+The \BRAVO and \Minerva stopping conditions are provided in section \ref{sec:model}.
 
@@
 
-A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
-A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
-The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
+%Using notation and terminology from , an audit $\mathcal{A}$ is a function that takes as input the current sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
+As mentioned earlier, we ignore a third option of a full hand count, leaving that decision to election officials, see section \ref{sec:related} for more detail.
 
-Election officials typically draw ballots in large round sizes, see for example \cite{va-2022,RI-report}.
-Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common).
-For the two audits we attended, election officials chose stopping probabilities of at least $0.9$.
-We are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
-\BRAVO hence cannot be used as a B2 audit in these scenarios.
+\subsubsection{The Audit Model}
+An audit is typically defined as a binary hypothesis test.
+If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}), we have the following definition of an RLA.
 
-For use as an R2 audit, the \BRAVO stopping condition can be applied once at the end of each round (End-of Round (EoR)), or retroactively after each ballot drawn if ballot order is retained (Selection-Ordered (SO)).
-SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
-But it requires the additional effort of tracking the order of ballots, and should be expected to be inefficient because it does not use the information in the ballots drawn at the end.
+\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+risk limit $\alpha$ iff for sample $X$
+$$
+Pr[\mathcal{A}(X) 
+= \text{Correct} \,|\, H_0]\le \alpha
+$$
+\end{definition}
 
-\subsection{Adversarial Model for RLAs}
-\label{sec:adv}
-Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
+The definition above is valid at the end of the RLA, and not at the end of each round. 
+%When $X$ is used to represent a sample while the RLA is ongoing, it is understood to be the cumulative sample, representing all the information obtained thus far about the underlying vote collection.
+
+The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, have the audit declare an incorrect outcome as correct more often than guaranteed.
+An audit is an RLA only if there is a proof of its risk-limiting property.
+Thus, the adversary would need to, at the very least, invalidate an assumption of the proof to obtain:
+$$
+Pr[\mathcal{A}(X) 
+= \text{Correct} \,|\, H_0] > \alpha
+$$
+
+\subsection{Related Work}
+\label{sec:related}
+\input{sections/related}
+%\subsection{Adversarial Model for RLAs}
+%\label{sec:adv}
+%Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
 For our purposes, we will assume that the best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
@@
 
-Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that the ballot manifest is verified by the compliance audit; a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
+%Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that the ballot manifest is verified by the compliance audit; a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
 Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed, because it can be checked by the public through knowledge of the seed and the drawn ballots.
@@
 
-We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time (before, of course, that particular round begins).
+%We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time (before, of course, that particular round begins).
 In particular, a strong adversary may use information about the sample drawn thus far to decide the next round size, but a weak adversary may not.
 
-\subsection{Open Questions}
-Zag\'{o}rski {\em et al.} propose ballot polling RLA \Minerva \cite{usenix_minerva}, which does not need ballot order and relies only on sample and round tallies.
-They prove that it is risk-limiting when the number of relevant ballots drawn in each round is pre-determined before any ballots are examined; that is, for a weak adversary.
-They do not address the case of a strong adversary (such as an audit insider) who can determine the size of the next round after knowing what votes are on the ballots sampled thus far.
-In particular, an open question about \Minerva is whether the computation of a risk limit assuming a weak adversary applies to an attack by a strong adversary, or is the risk limit computation incorrect when the adversary is strong?
-Can the strong adversary change the election outcome and then increase the audit's error probability beyond its declared risk limit?
+\subsection{Gaps in Prior Work and Our Contributions}
+Having provided context above, we now describe gaps in our knowledge of audits and describe our contributions. 
+\subsubsection{Limitations of \Minerva}
+Zag\'{o}rski {\em et al.}  prove that \Minerva \cite{usenix_minerva} is risk-limiting when the number of relevant ballots drawn in each round is pre-determined before any ballots are examined.
+They do not address the case of a stronger adversary (such as an audit insider) who can determine the size of the next round after knowing what votes are on the ballots sampled thus far.
+In particular, an open question about \Minerva is whether its RLA proof holds in this case.
+Can the audit insider increase the audit's error probability beyond its declared risk limit?
 Or is there no probabilistic adversarial advantage to being able to compute next round sizes after knowing the drawn sample?
@@
 
-Until \Minerva is proven to be risk-limiting to a given risk limit for the strong adversary, it may not be used in audits whose round sizes are not pre-determined.
+Until \Minerva is proven to be risk-limiting to a given risk limit for the adversary who can determine next round size after examining the current sample, it may not be used in audits whose round sizes are not pre-determined.
 This presents a major limitation, because the stopping probability of the next round is better estimated using information of the sample drawn thus far, but this would not be allowed for \Minerva.
 The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter\footnote{Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
-However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that a fixed draw size is sufficient to limit adversaries to weak ones, though this is not explicitly proven in \cite{usenix_minerva}.}.
+However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that differences in the number of ballots average out, and that a fixed draw size is sufficient, though this is not explicitly proven in \cite{usenix_minerva}.}.
 This could lead to the drawing of too few or too many ballots and greatly constrains audit planning.
@@
 
-Zag\'{o}rski {\em et al.}  also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
-Broadrick {\em et al.} provide further simulations that show \Minerva requires fewer ballots over multiple rounds and for lower stopping probability \cite{simulations}, though the improvement from using \Minerva over either version of \BRAVO decreases with round size.
-
-The risk limit for B2, EoR and SO \BRAVO is fixed whether the adversary is strong or weak.
+The risk limit for B2, EoR and SO \BRAVO is fixed independent of whether next round sizes are determined with or without knowledge of the current sample.
 This allows \BRAVO audits the flexibility of choosing smaller subsequent round sizes if the sample drawn so far is a ``good'' sample.
@@
 
+\subsubsection{Limitations in Existing Workload Measures}
 A major limitation of our understanding of the ballot polling problem as a community is that we use the number of ballots drawn or values proportional to this number \cite{mclaughlin_thesis,bernhard-diss,RI-report} as measures of the workload of an audit.
@@
 
-\subsection{Our Contributions}
-Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to a strong adversary. 
+\subsubsection{Our Contributions}
+Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to an adversary who can choose the next round size after knowing the current sample. 
 %Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva. 
 \BRAVO is a very flexible audit enabling audit planning; with the introduction of \Providence, similar planning is possible with greater efficiency.
-In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than 0:03) the difference in number of ballots sampled using \Providence over \BRAVO could correspond to many days of work which would need to be completed before a certification deadline.
+In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than 0.03) the difference in number of ballots sampled using \Providence over \BRAVO could correspond to many days of work which would need to be completed before a certification deadline.
 
 The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
-%Further, it assumes that the null hypothesis is a tie. A far simpler proof is possible for the risk-limiting property of \Providence. 
+%Further, it assumes that the null hypothesis is a tie. 
+A far simpler proof is possible for the risk-limiting property of \Providence. 
 In particular, the proof does not require an assumption about round sizes.
@@
 \begin{enumerate}
-\item Proof that \Providence is an RLA and resistant to a strong adversary.
+\item Proof that \Providence is an RLA and resistant to an adversary who can choose next round sizes after knowing the current sample.
 \item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence uses number of ballots similar to those of \Minerva, both fewer than either version of \BRAVO.
@@
 
-\Providence may be used in any audit where audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.
+\Providence may be used in any audit where sampling is with replacement and audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.
 
@@
 
-\subsection{Organization} 
-Sections \ref{sec:model} and \ref{sec:related} provide background, describing the audit model and related work respectively.
+\subsubsection{Organization} 
+Sections \ref{sec:model} provides preliminaries on the \BRAVO and \Minerva audits.
 Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.

### 2023-02-18 providence/sections/model.tex
alpha and section not Section
Lots of changes

--- 2023-02-13 providence/sections/model.tex
@@
 
-We now summarize the model drawing largely from the notation and terminology of \cite{usenix_minerva,arxiv_athena,simulations,bravo}.
+%We now summarize the model drawing largely from the notation and terminology of \cite{usenix_minerva,arxiv_athena,simulations,bravo}.
 The model is related work and not claimed to be original to this work.
 
-An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
+%An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}.
 As described in the introduction, we ignore a third option of moving to a hand count, leaving that decision to election officials.
-We provide more context on this in section \ref{sec:related}, on related work.
+This is standard in the literature, see, for example, \cite{bravo}.
+We provide more detail on the use of this option in section \ref{sec:related}, on related work.
 
@@
 (When the number of ballots is odd, $H_0$ corresponds to the announced loser winning by one ballot.)
-Thus the null hypothesis is the outcome distinct from the announced one which is most difficult to detect; the probability of failing to detect it, given that the null hypothesis is true, is the worst case such probability and should be below the risk limit \cite{Bayesian-RLA}.
+%Thus the null hypothesis is the outcome distinct from the announced one which is most difficult to detect; the probability of failing to detect it, given that the null hypothesis is true, is the worst case such probability and should be below the risk limit \cite{Bayesian-RLA}.
 
-\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
-An audit $\mathcal{A}$ is a Risk Limiting Audit with 
-risk limit $\alpha$ iff for sample $X$
-$$
-Pr[\mathcal{A}(X) 
-= \text{Correct} \,|\, H_0]\le \alpha
-$$
-\end{definition}
+%\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+%An audit $\mathcal{A}$ is a Risk Limiting Audit with 
+%risk limit $\alpha$ iff for sample $X$
+%$$
+%Pr[\mathcal{A}(X) 
+%= \text{Correct} \,|\, H_0]\le \alpha
+%$$
+%\end{definition}
 
@@
 
+\BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT) \cite{wald}.
+Applying the general SPRT to RLAs, there would be a third output, {\em Full Manual Hand Count}, in addition to {\em Correct} and {\em Undetermined}.
+The test requires an additional error parameter $\beta$ (of audit error when the outcome is correct): the probability of requiring a (unnecessary) hand count when the outcome is correct.
+The test is:
+\begin{equation}
+    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \emph{Correct} & ~\sigma(k, p_a, p_0, n) 
+        \geq \frac{1-\beta}{\alpha}\\ \\ \\
+       \emph{Hand~Count} & ~\sigma(k, p_a, p_0, n) 
+        < \frac{\beta}{1-\alpha}\\ \\ \\
+        \emph{Undetermined} & ~else 
+    \end{array}
+    \right .
+    \label{eqn:wald}
+\end{equation}
+
+An example of the above use is \cite{RLA}.
+In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e.
+they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
+%In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. 
+% Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.
+
+A manual hand count presents significant logistical challenges.
+Left to election officials, the decision would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
+The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count.
+Audits going towards statutory or legal requirements would need to be completed by the certification deadline; this provides an incentive to go to a hand count.
+Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.
+A better workload model would enable a better determination of when election officials are better off stopping than drawing more ballots
+
 Where \BRAVO uses the ratio of the values of the probability distribution functions, \Minerva uses the ratio of their \emph{tails}.
@@
 
+Note that $\tau$ depends only on the most recent cumulative votes for the winner, $k_j$ and not on older values.
+Here, if we were to condition the ratio on the vector of cumulative winner votes $\bm{k_{j-1}}$, or, equivalently, on the most recent cumulative winner votes ${k_{j-1}}$, we would get the binomial distributions and be running a new first round each time.
+
+Notice that the ratios for \BRAVO and \Minerva are for two candidates.
+Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
+
+The {\em margin}, given a contest and two candidates, is the difference between the votes received by each as a fraction of the total number of votes.
+
 \begin{comment}

### 2023-02-18 providence/sections/prov.tex
alpha and section not Section
Lots of changes

@@
-In this section we introduce the stopping condition of \Providence and prove some properties.
+In this section we describe the adversarial model and introduce the stopping condition of \Providence, proving some of its properties.
 
-Recall that the proof that the \Minerva audit is risk-limiting assumes that the round schedule of \Minerva is predetermined and that, in particular, an adversarial auditor cannot determine the next round size after drawing a sample (we termed such an adversary a weak adversary in section \ref{sec:adv}).
+%Recall that the proof that the \Minerva audit is risk-limiting assumes that the round schedule of \Minerva is predetermined and that, in particular, an adversarial auditor cannot determine the next round size after drawing a sample (we termed such an adversary a weak adversary in section \ref{sec:adv}).
 This presents difficulties because a non-adversarial election official might want to draw a small next round if the current sample comes close to satisfying the risk limit.
@@
 
-Before proceeding to define the audit \Providence which is an RLA in the presence of the stronger adversary, we formalize the notions of weak and strong adversaries introduced in section \ref{sec:adv}.
+%Before proceeding to define the audit \Providence which is an RLA in the presence of the stronger adversary, we formalize the notions of weak and strong adversaries introduced in section \ref{sec:adv}.
 
-\begin{definition}[Weak Adversary]
-A {\emph weak adversary} may choose the first and consequent round sizes as a pre-determined function of audit parameters.
-That is, the $j^{th}$ round size is a function $$n_j(\alpha, p_a, p_0)$$ determined before the audit begins.
+\subsection{Adversarial Model for RLAs}
+\label{sec:adv}
+Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
+For our purposes, we will assume that best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
+We will further assume that all software used in the RLA is open source and well-defined, so its output may be reproduced and thus verified by an observer wishing to do so with their own software.
+
+Referring to the ballot polling audit steps described in section \ref{sec:polling}, we further assume that the ballot manifest is verified by the compliance audit; a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public.
+Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed, because it can be checked by the public through knowledge of the seed and the drawn ballots.
+Thus the only variable is round size.
+
+We define a {\em weakly round-choosing adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strongly round-choosing adversary} as one who can choose any round size at any time (before, of course, that particular round begins).
+In particular, a strongly round-choosing adversary may use information about the sample drawn thus far to decide the next round size, but a weakly round-choosing adversary may not.
+
+\begin{definition}[Weakly Round-Choosing Adversary]
+A {\emph weakly round-choosing adversary} may choose the first and consequent round sizes as a pre-determined function of audit parameters.
+That is, the $j^{th}$ round size is a function $$n_j(\alpha, p_a, p_0, ballot\_manifest)$$ determined before the audit begins.
 \end{definition}
 
-\begin{definition}[Strong Adversary]
-A {\emph strong adversary} may choose any round size as any function of audit parameters and all preceding samples. That is, the first round size is a function 
-$$n_1(\alpha, p_a, p_0),$$
+\begin{definition}[Strongly Round-Choosing Adversary]
+A {\emph strongly round-choosing adversary} may choose any round size as any function of audit parameters and all preceding samples. That is, the first round size is a function 
+$$n_1(\alpha, p_a, p_0, ballot\_manifest),$$
 and for all rounds $j\ge 2$, the round size is a function
-$$n_{j}(\alpha, p_a, p_0, \bm{k_{j-1}}, \bm{n_{j-1}})$$ The functions $n_j$, $j \geq 1$ may be chosen at any time before the $j^{th}$ round begins. 
+$$n_{j}(\alpha, p_a, p_0, ballot\_manifest, \bm{k_{j-1}}, \bm{n_{j-1}})$$ The functions $n_j$, $j \geq 1$ may be chosen at any time before the $j^{th}$ round begins. 
 \end{definition}
 
+\subsection{The Intuition Behind the Properties of \Providence}
 Before defining \Providence, we give some intuition for how it is designed to avoid the problem of \Minerva.
 
-For any ballot polling audit, in round $j$, designed to stop for some value of $K_j$, however large $K_j$ is, there is, in general, a probability, however small, that the value of $K_j$ has resulted from an underlying incorrect election outcome.
-Thus, corresponding to each such value of $K_j$, a risk---probability that the audit will stop given that the outcome is incorrect---is incurred.
+Consider any ballot polling audit in round $j$, designed to stop for some value of $K_j=k_j$ smaller than the round size.
+In general, if $k_j$ is large, the probability of this value of $K_j$ given a correct outcome, $Pr[K_j = k_j \mid H_1]$, is larger than the corresponding risk---$Pr[K_j = k_j \mid H_0]$, the probability of this value of $K_j$ given a tie.
+The risk is generally not zero, however.
+Thus, corresponding to each such value of $K_j$, a risk is incurred.
 To obtain the total risk for the round, one adds the risks corresponding to each value of $K_j$ for which the audit can stop, weighted by the probability of drawing that value of $K_j$.
 The stopping condition may provide relationships among the various quantities.
+To obtain the risk of the entire audit, one sums the risks of each round.
 
@@
 Separate relationships between risk and stopping probability are not available for individual values of $K_j$.
-If the next round size depends on $K_j$, expressions relating the risks are not available, and so we are not able to say if \Minerva is an RLA in this case, or if it is vulnerable to an adversary choosing round sizes \cite{usenix_minerva, arxiv_athena}.
+If the next round size depends on $K_j$, expressions relating the risks are not available, and we are not able to obtain an expression bounding the sum of the risks across rounds.
+Thus we are not able to determine if \Minerva is an RLA in this case, or if it is vulnerable to the strongly round-choosing adversary \cite{usenix_minerva, arxiv_athena, alpha}.
 
 In \Providence, we choose a stopping condition that applies separately to the risk and stopping probabilities for each value of $K_j$, avoiding the problem of \Minerva, and allowing for optimal round size choices, which depend on the drawn sample.
-The \Providence audit is risk-limiting even if a strong adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.
+The \Providence audit is risk-limiting even if a strongly round-choosing adversarial auditor determines round sizes after drawing the sample, and next round size computations may use knowledge of the current sample.
 
@@
 The computation of $\tau_j$ for $j \geq 2$, as required in \Minerva, relies on the convolution of two probability distribution functions and is hence computationally considerably more expensive.
-Lesser computational complexity makes audit planning and analysis using simulations as in Section~\ref{sec:workload} more feasible.
+Smaller computational complexity makes audit planning and analysis using simulations as in section~\ref{sec:workload} more feasible.
 
@@
 \label{sec:proof}
-We now prove that \Providence is risk-limiting against a strong adversary using lemmas from basic algebra which are given in Appendix~\ref{sec:proofs}.
+We now prove that \Providence is risk-limiting against a strongly round-choosing adversary using lemmas from basic algebra which are given in Appendix~\ref{sec:proofs}.
 
@@
 $(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence is an
-$\alpha$-RLA in the presence of a strong adversary.
+$\alpha$-RLA in the presence of a strongly round-choosing adversary.
 \end{theorem}
@@
 For different possible values of $k_{j-1}$, different round sizes $n_j$ can be used, and this same relationship will hold. 
-That is, the relationship holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather than for each individual value of $K_{j-1}$.
-By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strong adversary.}
+That is, the relationship\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather than for each individual value of $K_{j-1}$.
+By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strongly round-choosing adversary.} holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.
 

### 2023-02-18 providence/sections/related.tex
alpha and section not Section
Lots of changes

@@
 %Related Work
-Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
+In this section, we describe related work.
 
-The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn.
+{\bf Ballot Polling RLA Process:} Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
 
-The \Minerva audit \cite{usenix_minerva,arxiv_athena} was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
+Election officials typically draw ballots in large round sizes, see for example \cite{va-2022,RI-report}.
+Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common).
+For the two audits we attended, election officials chose stopping probabilities of at least $0.9$.
+We are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
+\BRAVO hence cannot be used as a B2 audit in these scenarios.
+
+{\bf R2 and B2 Audits and the Classical \BRAVO audit:} A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
+A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
+The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
+
+For use as an R2 audit, the \BRAVO stopping condition can be applied once at the end of each round (End-of Round (EoR)), or retroactively after each ballot drawn if ballot order is retained (Selection-Ordered (SO)).
+SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
+But it requires the additional effort of tracking the order of ballots, and should be expected to be inefficient because it does not use the information in the ballots drawn at the end.
+
+%The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn. 
+{\bf Newer Ballot Polling Audits:} The \Minerva audit \cite{usenix_minerva,arxiv_athena} does not need ballot order and relies only on sample and round tallies.
+It was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
 First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins; simulations \cite{arxiv_athena} support these observations.
-Additional simulations \cite{simulations} have shown that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
+
+The \ALPHA audit \cite{alpha} generalizes \BRAVO to gain efficiency in cases where the reported outcome is correct but the reported margin is erroneous.
+
+{\bf Simulations:} Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and understanding audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20,bravo}.
+There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
+Simple workload estimates may be obtained by using the number of ballots drawn \cite{bernoulli-ballot-polling}, a more thorough workload estimation model includes the time taken to access individual ballots\cite{bernhard-diss}.
+
+Zag\'{o}rski {\em et al.}  also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
+Broadrick {\em et al.} provide further simulations showing that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
 As expected, the advantage of \Minerva decreases for smaller stopping probability (smaller round sizes) as it approaches a B2 audit,
@@
 
-\BRAVO is an instance of Wald's Sequential Probability Ratio Test (SPRT) \cite{wald}.
-Applying the general SPRT to RLAs, there would be a third output, {\em Full Manual Hand Count}, in addition to {\em Correct} and {\em Undetermined}.
-The test requires an additional error parameter $\beta$ (of audit error when the outcome is correct): the probability of requiring a (unnecessary) hand count when the outcome is correct.
-The test is:
-\begin{equation}
-    \mathcal{A}(X) =  \left\{ \begin{array}{ll} \emph{Correct} & ~\sigma(k, p_a, p_0, n) 
-        \geq \frac{1-\beta}{\alpha}\\ \\ \\
-       \emph{Hand~Count} & ~\sigma(k, p_a, p_0, n) 
-        < \frac{\beta}{1-\alpha}\\ \\ \\
-        \emph{Undetermined} & ~else 
-    \end{array}
-    \right .
-    \label{eqn:wald}
-\end{equation}
-
-An example of the above use is \cite{RLA}.
-In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e.
-they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
-
-Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and understanding audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20}.
-There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
-Simple workload estimates may be obtained by using the number of ballots drawn \cite{bernoulli-ballot-polling}, a more thorough workload estimation model includes the time taken to access individual ballots\cite{bernhard-diss}.
-
-For completeness, we describe other types of RLAs.
-In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+{\bf Ballot and Batch Comparison Audits:} In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
 Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.

### 2023-02-18 providence/sections/pilot.tex
Update pilot.tex

@@
 \includegraphics[width=.5\textwidth]{pilot_sequence.png}
-\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner found in the sample is shown.}
+\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown, and the corresponding behavior is that of SO \BRAVO.}
 \label{fig:pilot_sequence}

### 2023-02-18 providence/sections/pilot.tex
Update pilot.tex

@@
 We observe two interesting phenomena in this particular sample's selection order.
-\begin{description}
-\item First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
+
+First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
 It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
-\item Second, only $5$ of the first $11$ ballots were for the announced winner.
+Second, only $5$ of the first $11$ ballots were for the announced winner.
 A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size.
-\end{description}
 Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.

### 2023-02-18 providence/sections/intro.tex
Various changes

@@
 \end{enumerate}
-Election officials may choose to abort this procedure and go to a full hand count at any time; we discuss this in more detail in section \ref{sec:related}.  
+Election officials may choose to abort this procedure and go to a full hand count at any time and should have a plan for how to decide whether to do so; we discuss this in more detail below and in section \ref{sec:related}.
 \item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
@@
 The definition above is valid at the end of the RLA, and not at the end of each round.
+
 %When $X$ is used to represent a sample while the RLA is ongoing, it is understood to be the cumulative sample, representing all the information obtained thus far about the underlying vote collection.
 
-The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, have the audit declare an incorrect outcome as correct more often than guaranteed.
+The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct more often than implied by the risk limit.
 An audit is an RLA only if there is a proof of its risk-limiting property.
@@
 
+It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, at least a $1-\alpha$ fraction of the audits of an incorrect outcome will never stop, and election officials will keep drawing new rounds of ballots.
+See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins).
+Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.
 \subsection{Related Work}
@@
 \item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
-\item Open source implementation of \Providence and audit planning tools. 
+\item Open source implementation of \Providence and audit planning tools. The implementation of \Providence has been integrated as an option in Arlo. 
 %including the novel metric Probability of Misleading(name?)

### 2023-02-18 providence/sections/model.tex
Various changes

@@
 Left to election officials, the decision would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
-The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count.
-Audits going towards statutory or legal requirements would need to be completed by the certification deadline; this provides an incentive to go to a hand count.
-Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.
-A better workload model would enable a better determination of when election officials are better off stopping than drawing more ballots
+A better workload model would better inform the decision.
+%The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count.
+Audits going towards statutory or legal requirements would need to be completed by the certification deadline\footnote{Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.}; this provides an incentive to go to a hand count.
+It is worth considering bring back a small value of $\beta$ into the stopping condition, or requiring election officials to declare beforehand when they would stop drawing ballots and go to a full hand count.
 

### 2023-02-18 providence/sections/workload.tex
add error bars to fig 6
avg bals ratio with error bars
errorbars
combined misleading plot

@@
 \begin{figure}[h!]
-\includegraphics[width=.5\textwidth]{avg_bals.png}
+%\includegraphics[width=.5\textwidth]{avg_bals.png}
+\includegraphics[width=.5\textwidth]{avg_bals_error_bars_every_three.png}
 \caption{The total number of ballots sampled on average, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
@@
 \begin{figure}[h!]
-\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
+%\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
+\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
 \caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
@@
 \begin{figure}[h!]
-\includegraphics[width=.5\textwidth]{with_round_workload.png}
+%\includegraphics[width=.5\textwidth]{with_round_workload.png}
+\includegraphics[width=.5\textwidth]{with_round_workload_errorbars.png}
 \caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
@@
 \begin{figure}[h!]
+%\includegraphics[width=.5\textwidth]{optimal_workloads.png}
 \includegraphics[width=.5\textwidth]{optimal_workloads.png}
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{misleading_limits.png}
-\caption{The proportion of simulated \Providence audits for the Virginia contest parameters that had a \emph{misleading sample} in any round.}
+%\includegraphics[width=.5\textwidth]{misleading_limits.png}
+\includegraphics[width=.5\textwidth]{combined_misleading.png}
+\caption{The proportion of simulated \Providence audits for the Virginia and pilot contest parameters that had a \emph{misleading sample} in any round.}
 \label{fig:misleading}
@@
 We observe a similar behavior in our simulations of audits on the contest from the pilot audit.
-Figure~\ref{fig:pilot_misleading} shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
+Figure~\ref{fig:misleading} also shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
 Despite the large difference in margin ($\sim 0.05$ in Virginia and $\sim 0.25$ in the pilot) we still observe that a \emph{misleading limit} of $0.01$ is first achieved at roughly $p=0.8$ and $0.001$ at $p=0.95$.
 
-\begin{figure}
-\includegraphics[width=.5\textwidth]{pilot_misleading_prop.png}
-\caption{The proportion of simulated \Providence audits for the pilot audit parameters that had a \emph{misleading sample} in any round.}
-\label{fig:pilot_misleading}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=.5\textwidth]{pilot_misleading_prop.png}
+%\caption{The proportion of simulated \Providence audits for the pilot audit parameters that had a \emph{misleading sample} in any round.}
+%\label{fig:pilot_misleading}
+%\end{figure}
 

### 2023-02-18 providence/sections/workload.tex
.
imgs now go 25,avg,75
error bars and caption

@@
 %\includegraphics[width=.5\textwidth]{avg_bals.png}
-\includegraphics[width=.5\textwidth]{avg_bals_error_bars_every_three.png}
-\caption{The total number of ballots sampled on average, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
+%\includegraphics[width=.5\textwidth]{avg_bals_error_bars_every_three.png}
+%\includegraphics[width=.5\textwidth]{avg_bals_errorbars.png}
+\includegraphics[width=.5\textwidth]{avg_bals_quantiles.png}
+\caption{The average and $0.25$ and $0.75$ quantiles for total number of ballots sampled, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the US State of Virginia.
+For sufficiently large $p$ ($p\ge 0.75$), the $0.25$ and $0.75$ quantiles are both equal to the first round size, and this shown by the downward arrows.}
 \label{fig:avg_bals}
@@
 %\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
-\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
+%\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
+%\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
+\includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
 \caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
@@
 %\includegraphics[width=.5\textwidth]{with_round_workload.png}
-\includegraphics[width=.5\textwidth]{with_round_workload_errorbars.png}
+%\includegraphics[width=.5\textwidth]{with_round_workload_errorbars.png}
+%\includegraphics[width=.5\textwidth]{round_workload_errorbars.png}
+\includegraphics[width=.5\textwidth]{round_workload_quantiles.png}
 \caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
@@
 
-\subsection{Misleading samples}
+\section{Misleading samples}
+\label{sec:misleading}
 

### 2023-02-18 providence/sections/workload.tex
mention quantiles in captions

@@
 \includegraphics[width=.5\textwidth]{avg_bals_quantiles.png}
-\caption{The average and $0.25$ and $0.75$ quantiles for total number of ballots sampled, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the US State of Virginia.
-For sufficiently large $p$ ($p\ge 0.75$), the $0.25$ and $0.75$ quantiles are both equal to the first round size, and this shown by the downward arrows.}
+\caption{The average total number of ballots sampled, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the US State of Virginia.
+Error bars show the $0.25$ and $0.75$ quantiles.
+For sufficiently large $p$ ($p\ge 0.75$), the $0.25$ and $0.75$ quantiles are both equal to the first round size, and this is shown by the downward arrows.}
 \label{fig:avg_bals}
@@
 \includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
-\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.}
+\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.
+The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}
 \label{fig:avg_bals_ratio}
@@
 \caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
-Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
+Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.
+The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}
 \label{fig:with_round_workload}

### 2023-02-18 providence/sections/intro.tex
rewrote full hand count in intro

--- 2023-02-18 providence/sections/intro.tex
@@
 \end{enumerate}
-Election officials may choose to abort this procedure and go to a full hand count at any time; we discuss this in more detail in section \ref{sec:related}.  
+Election officials may choose to abort this procedure and go to a full hand count at any time and should have a plan for how to decide whether to do so; we discuss this in more detail below and in section \ref{sec:model}.
 \item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
@@
 For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome.
-This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size deternination.
+This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size and hand count decisions.
 
@@
 An audit is typically defined as a binary hypothesis test.
-If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}), we have the following definition of an RLA.
+If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}, we have the following definition of an RLA.
 
 \begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
+\label{def:RLA}
 An audit $\mathcal{A}$ is a Risk Limiting Audit with 
@@
 
-The definition above is valid at the end of the RLA, and not at the end of each round. 
+Definition \ref{def:RLA} is valid at the end of the RLA, and not at the end of each round.
+Thus it is not the case that an incorrect election will pass the audit if a sufficient number of rounds is drawn.
+In fact, the larger a sample the more accurate the estimate of election correctness can be, though, of course, it is possible to declare an incorrect outcome correct even after drawing a large sample.
+
+We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit (though, technically, it is the maximum risk, corresponding to the election outcome hardest to detect).
+Every time an audit stops, it incurs a contribution to its risk.
+Audit design determines how the risk contributions are distributed across rounds, so long as the overall risk is limited by $\alpha$.
+
 %When $X$ is used to represent a sample while the RLA is ongoing, it is understood to be the cumulative sample, representing all the information obtained thus far about the underlying vote collection.
 
-The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, have the audit declare an incorrect outcome as correct more often than guaranteed.
+{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, at least a $1-\alpha$ fraction of the audits of an incorrect outcome will {\em never} stop, and election officials will keep drawing new rounds of ballots.
+See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins).
+Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.
+
+EXAMPLE
+
+To avoid a scenario where election officials fruitlessly draw round after round hoping the audit will stop when the election outcome is incorrect, it might be worthwhile for election officials to determine, before the audit begins, the latest date by which either the audit stops or a hand count must begin to be completed before certification.
+
+{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct more often than implied by the risk limit.
 An audit is an RLA only if there is a proof of its risk-limiting property.

### 2023-02-18 providence/sections/prov.tex
rewrote full hand count in intro

@@
 
-\subsection{The Intuition Behind the Properties of \Providence}
+\subsection{Intuition Behind the Properties of \Providence}
 Before defining \Providence, we give some intuition for how it is designed to avoid the problem of \Minerva.

### 2023-02-19 providence/sections/intro.tex
re: beta

@@
 
-As a formalized approach to the examination of the evidence supporting vote tabulation, an RLA is an important part of an evidence-based election.
-When correctly used, it serves to greatly improve the trustworthiness of an election.
+As a formalized approach to the examination of evidence supporting vote tabulation, an RLA is an important part of an evidence-based election.
+When correctly implemented, it serves to vastly improve the trustworthiness of the election.
 
@@
 For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots.
-A ballot polling RLA \cite{RLA} is based on the manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
+A ballot polling RLA \cite{RLA} is based on manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
 %, on the other hand, simply tallies the sampled ballots.
@@
 An audit is typically defined as a binary hypothesis test.
-If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}, we have the following definition of an RLA.
+If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}), we have the following definition of an RLA.
 
@@
 
-We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit (though, technically, it is the maximum risk, corresponding to the election outcome hardest to detect).
+We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}.
 Every time an audit stops, it incurs a contribution to its risk.
@@
 
-{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, at least a $1-\alpha$ fraction of the audits of an incorrect outcome will {\em never} stop, and election officials will keep drawing new rounds of ballots.
+{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and they will keep drawing new rounds of ballots.
 See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins).
@@
 
-To avoid a scenario where election officials fruitlessly draw round after round hoping the audit will stop when the election outcome is incorrect, it might be worthwhile for election officials to determine, before the audit begins, the latest date by which either the audit stops or a hand count must begin to be completed before certification.
+To avoid a scenario where election officials fruitlessly draw round after round hoping the audit will stop when the election outcome is incorrect, it might be worthwhile for election officials to determine, before the audit begins, the latest date by which either: the audit stops, or a hand count begins, so as to be completed before certification.
 
-{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct more often than implied by the risk limit.
+{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit.
 An audit is an RLA only if there is a proof of its risk-limiting property.
-Thus, the adversary would need to, at the very least, invalidate an assumption of the proof to obtain:
+Hence, at the very least, the adversary would need to invalidate an assumption of the proof to obtain:
 $$

### 2023-02-19 providence/sections/model.tex
re: beta

@@
 they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.
+For a given election, $\sigma$ is a function only of $k$, and increases with $k$.
+Any non-zero value of $\beta$ would reduce the value $\sigma$ is being compared with, and allow $\mathcal{A}$ to declare \emph{Correct} for lower values of $k$.
+Thus $\beta = 0$ is the most stringent of this form of audit.
+The choice of going to a full hand count when $\beta=0$ will only reduce the risk because it never results in an error (by definition, the true election outcome is one that would be obtained by a full manual hand count of the ballots).
 %In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. 
@@
 A manual hand count presents significant logistical challenges.
-Left to election officials, the decision would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
+The decision to move to one would be influenced by the certification deadline\footnote{As mentioned earlier, audits going towards statutory or legal requirements would need to be completed by the certification deadline.
+Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.}, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.
 A better workload model would better inform the decision.
 %The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count.
-Audits going towards statutory or legal requirements would need to be completed by the certification deadline\footnote{Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.}; this provides an incentive to go to a hand count.
-It is worth considering bring back a small value of $\beta$ into the stopping condition, or requiring election officials to declare beforehand when they would stop drawing ballots and go to a full hand count.
+Election officials should announce ahead of time their plans for when they would abort the audit procedure and go to a full hand count, justifying it with an eye towards completing certification requirements.
+This would provide a more transparent process and protect them from political pressures.
+See section \ref{sec:sims} for a discussion of the considerations informing the choice between drawing more rounds and performing a full hand count for the \Providence audit and large round size.
 

### 2023-02-19 providence/sections/intro.tex
VA 2020 ---> 2016

@@
 Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
-Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2020 US Presidential election in the state of Virginia.
+Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2016 US Presidential election in the state of Virginia.
 Our conclusions, the availability of an audit implementation and acknowledgements may be found in sections \ref{sec:conc}, \ref{sec:avail} and \ref{sec:ack} respectively.

### 2023-02-19 providence/sections/workload.tex
VA 2020 ---> 2016

@@
 \includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
-\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2020 US Presidential election in the state of Virginia.
+\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the state of Virginia.
 The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}

### 2023-02-19 providence/Extra.tex
Example ongoing

We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}. Loosely speaking, it is the sum of the probabilities of every sequence that stops the audit. The sequences that stop the audit depend on the audit, and not on the underlying vote distribution as the audit does not know what it is. The probabilities used for risk contributions are computed assuming the null hypothesis; those used for stopping probability computations assume the announced tally. Thus, every time an audit stops, it incurs a contribution to its risk and its stopping probability. Audit design determines, among other things, how the risk and stopping probability contributions are distributed across rounds, as long as the overall risk is limited by $\alpha$.

### 2023-02-19 providence/sections/intro.tex
Example ongoing

@@
 Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
-For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there might be no legal means of changing the outcome and this would significantly impact citizen confidence.
+For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence.
 For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome.
@@
 Thus it is not the case that an incorrect election will pass the audit if a sufficient number of rounds is drawn.
-In fact, the larger a sample the more accurate the estimate of election correctness can be, though, of course, it is possible to declare an incorrect outcome correct even after drawing a large sample.
-
-We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}.
-Every time an audit stops, it incurs a contribution to its risk.
-Audit design determines how the risk contributions are distributed across rounds, so long as the overall risk is limited by $\alpha$.
+In fact, the larger a sample, the more accurate the estimate of election correctness can be, and the more the audit will diverge from one that would stop (see the example below).
+Of course, though, it is possible to declare an incorrect outcome as correct even after drawing a large sample, but a good audit is less likely to make this error.
 
@@
 
-{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and they will keep drawing new rounds of ballots.
+\npara{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and new rounds will continue to be drawn.
 See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins).
@@
 
-EXAMPLE
+We present an example to illustrate the above idea.
+Consider the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with a total of 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.056614 in the pairwise contest (that is, ignoring the votes received by other candidates).
+The total number of votes cast in this contest, including invalid votes and votes for other candidates (all termed {\em irrelevant} votes for this pairwise contest), was 11, 315,056.
+A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of 0.1 for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
+The stopping condition of \Providence (see section \ref{sec:prov}) is a comparison of the \Providence ratio with the risk limit, stopping when the ratio is smaller.
+We ran two simulations using these parameters.
+
+Assuming the tally is correct, the probability that the audit stops in the first round is 0.9.
+Our simulation resulted in 1,138 votes for Trump, 1,054 for Biden and the rest were irrelevant votes.
+The \Providence ratio for this sample is 0.047, which is smaller than the desired risk limit, 0.1 and the audit stops.
+
+Assuming the null, a tied election, the probability that the audit stops over all possible rounds is 0.1, the risk limit. Our simulation of a five-round audit did not stop. 
+%The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09.
+Table \ref{tab:toy} lists the round sizes, votes for both candidates and the \Providence ratio after each draw.
+
+Our simulation draws 1,111 and 1,079 votes for Trump and Biden respectively, to obtain a \Providence ratio of 0.256 which is not small enough.
+The simulation then proceeds to draw another round of cumulative size 5,970 for a conditional stopping probability of 0.9, which corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second).
+Similarly, this round corresponds to a risk expenditure of approximately 0.009.
+Trump gets 2,940 votes; Biden 2,953; and the ratio is 4.251.
+The process continues, with cumulative round sizes of 16,685; 35,096; and 76,979; Trump votes of 8,281; 17,320; and 37,943; Biden votes of 8,171; 17,264; and 37,868; and \Providence ratios of 3,150.268; 380,220,376; and $2.5e^{+22}$.
 
@@
 
-{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit.
+\npara{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit.
 An audit is an RLA only if there is a proof of its risk-limiting property.

### 2023-02-19 providence/sections/intro.tex
Update intro.tex

@@
 The total number of votes cast in this contest, including invalid votes and votes for other candidates (all termed {\em irrelevant} votes for this pairwise contest), was 11, 315,056.
-A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of 0.1 for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
-The stopping condition of \Providence (see section \ref{sec:prov}) is a comparison of the \Providence ratio with the risk limit, stopping when the ratio is smaller.
+A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of $\alpha=0.1$ for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
+The audit stops when its p-value is smaller than $\alpha$.
 We ran two simulations using these parameters.
 
-Assuming the tally is correct, the probability that the audit stops in the first round is 0.9.
-Our simulation resulted in 1,138 votes for Trump, 1,054 for Biden and the rest were irrelevant votes.
-The \Providence ratio for this sample is 0.047, which is smaller than the desired risk limit, 0.1 and the audit stops.
+The first simulation assumed the tally is correct and resulted in 1,138 votes for Trump, 1,054 for Biden (the rest were irrelevant votes). The p-value was $0.047 < \alpha$ and the audit stops.
 
-Assuming the null, a tied election, the probability that the audit stops over all possible rounds is 0.1, the risk limit. Our simulation of a five-round audit did not stop. 
+The second simulation assumed a tied election and did not stop over five rounds. 
 %The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09.

### 2023-02-19 providence/sections/intro.tex
toy example done

@@
 The {\em evidence-based elections} \cite{evidence-based} approach requires that election outcomes be supported by strong affirmative evidence.
-The evidence is generated during the election and publicly examined at the end, so that citizens may determine whether it provides strong support of the election outcome or is not sufficiently convincing.
+The evidence is generated during the election and publicly examined at the end, enabling citizens to determine whether it provides strong support of the election outcome or is not sufficiently convincing.
 %by requiring not merely the use software independent systems to produce trustworthy evidence of outcome correctness; but also the examination of the evidence so that incorrect outcomes may be detected with high probability.
@@
 Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
-For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence.
+For RLAs required by statute or legislation, the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence.
 For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome.
@@
 Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.
+We present two example simulations from among the ones reported to illustrate this idea.
 
-We present an example to illustrate the above idea.
 Consider the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with a total of 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.056614 in the pairwise contest (that is, ignoring the votes received by other candidates).
@@
 A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of $\alpha=0.1$ for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
-The audit stops when its p-value is smaller than $\alpha$.
-We ran two simulations using these parameters.
+The audit stops when the inverse of the \Providence ratio (see section \ref{sec:prov}) is smaller than $\alpha$.
+We describe two audit simulations using these parameters.
 
-The first simulation assumed the tally is correct and resulted in 1,138 votes for Trump, 1,054 for Biden (the rest were irrelevant votes). The p-value was $0.047 < \alpha$ and the audit stops.
+The first simulation assumed that the tally is correct and resulted in 1,138 votes for Trump and 1,054 for Biden (the rest were irrelevant votes).
+The inverse of the \Providence ratio was $0.047 < \alpha$ and the audit stopped.
 
-The second simulation assumed a tied election and did not stop over five rounds. 
+The second simulation assumed a tied election and did not stop at the end of five rounds. 
 %The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09.
-Table \ref{tab:toy} lists the round sizes, votes for both candidates and the \Providence ratio after each draw.
-
-Our simulation draws 1,111 and 1,079 votes for Trump and Biden respectively, to obtain a \Providence ratio of 0.256 which is not small enough.
-The simulation then proceeds to draw another round of cumulative size 5,970 for a conditional stopping probability of 0.9, which corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second).
-Similarly, this round corresponds to a risk expenditure of approximately 0.009.
-Trump gets 2,940 votes; Biden 2,953; and the ratio is 4.251.
-The process continues, with cumulative round sizes of 16,685; 35,096; and 76,979; Trump votes of 8,281; 17,320; and 37,943; Biden votes of 8,171; 17,264; and 37,868; and \Providence ratios of 3,150.268; 380,220,376; and $2.5e^{+22}$.
+Table \ref{tab:toy} lists the round sizes, votes for both candidates and the inverse \Providence ratio after each draw.
+Each round size was chosen for a conditional stopping probability of 0.9, given that the audit did not stop so far.
+For example, this corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second), and so on.
+\begin{table}[h]
+\centering
+%\scriptsize
+\begin{tabular}{|crrrc|}
+\hline
+\hline
+Round& Cumulative& Trump & Biden & Inverse  \\
+ No. & Round Size &  & & Ratio \\
+\hline
+\hline
+1 & 2,217 & 1,111 & 1,079 & 0.256\\
+2 & 5,970 & 2,940 & 2,953 & 4.251\\
+3 & 16,685 & 8,281 & 8,171 & 3,150 \\
+4 & 35,096 & 17,320 & 17,264 & 380,220,376 \\
+5 & 76,979 & 37,943 & 37,868 & $2.5e^{+22}$ \\
+\hline
+\end{tabular}
+\caption{A simulation of a \Providence audit of a tied contest when the announced election has a margin of 0.056614, each round size is computed for a conditional stopping probability of $0.9$, and $\alpha=0.1$.}
+\label{tab:toy}
+\end{table}
 

### 2023-02-19 providence/sections/model.tex
small changes

@@
 Thus $\beta = 0$ is the most stringent of this form of audit.
-The choice of going to a full hand count when $\beta=0$ will only reduce the risk because it never results in an error (by definition, the true election outcome is one that would be obtained by a full manual hand count of the ballots).
+The choice of going to a full hand count when $\beta=0$ will only reduce the risk because it never results in an error (by definition, the true election outcome is one that would be obtained by a full manual hand count of the ballots) and prevents future errors.
 %In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. 
@@
 
-A manual hand count presents significant logistical challenges.
+A manual hand count presents numerous logistical challenges.
 The decision to move to one would be influenced by the certification deadline\footnote{As mentioned earlier, audits going towards statutory or legal requirements would need to be completed by the certification deadline.
@@
 This would provide a more transparent process and protect them from political pressures.
-See section \ref{sec:sims} for a discussion of the considerations informing the choice between drawing more rounds and performing a full hand count for the \Providence audit and large round size.
+%See section \ref{sec:sims} for a discussion of the considerations informing the choice between drawing more rounds and performing a full hand count for the \Providence audit and large round size.
 
@@
 
-Notice that the ratios for \BRAVO and \Minerva are for two candidates.
-Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
-
-The {\em margin}, given a contest and two candidates, is the difference between the votes received by each as a fraction of the total number of votes.
+The ratios for \BRAVO and \Minerva are for two candidates.
+As mentioned when describing the example in section \ref{sec:intro-model}, audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
 

### 2023-02-19 providence/sections/prov.tex
small changes

@@
 Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}.
-For our purposes, we will assume that best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public.
+For our purposes, we will assume that best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified; and
+%only legitimate voters cast no more than a single vote each; 
+the risk-limiting audit is public.
 We will further assume that all software used in the RLA is open source and well-defined, so its output may be reproduced and thus verified by an observer wishing to do so with their own software.

### 2023-02-19 providence/sections/prov.tex
added ballot manifest in one place

@@
 That is, the relationship\footnote{\Minerva enforces a similar relationship between risk and stopping probability but does so at the level of the round rather than for each individual value of $K_{j-1}$.
-By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strongly round-choosing adversary.} holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, \bm{k_{j-1}}, \bm{n_{j-1}})$.
+By enforcing this relationship for each value of $K_{j-1}$, \Providence is resistant to a strongly round-choosing adversary.} holds even if the values of $n_{j}$ depend on $k_{j-1}$, if $n_j$ is a function $n_j(\alpha, p_0, p_1, ballot\_manifest, \bm{k_{j-1}}, \bm{n_{j-1}})$.
 

### 2023-02-19 providence/sections/adversary.tex

@@
 To illustrate the practical implication of this property, we consider a toy example: an RLA of a two-candidate contest with margin $0.01$ and risk limit $0.1$. 
-Suppose we wish to achieve a conditional stopping probability $0.9$ in each round of the audit.
-For \Providence, we can compute a new round size for each round based on the previous samples.
-For \Minerva, however, we would have a predetermined round schedule.
-We use the default \Minerva round schedule of audit software Arlo \cite{arlo} (used by many states performing an RLA), which is $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one.
-This multiplier of $1.5$ is known to give, over a wide range of margins, a probability of stopping roughly $0.9$ in the second round if the first round size has probability of stopping $0.9$.
+For a conditional stopping probability of $0.9$ in each round of a \Providence audit, we can compute next round sizes based on the current sample.
+\Minerva, however, would have a predetermined round schedule.
+We use the default \Minerva round schedule of audit software Arlo: $[x, 2.5x, 6.25x, ...]$; that is, the next marginal round size is $1.5$ times the current one---equivalently, the next total round size is $2.5$ times the current one.
+This approach is known to give, over a wide range of margins, a conditional probability of stopping roughly $0.9$ in the second round if the first round size was determined for a probability of stopping $0.9$.
 
 Both the audits of our toy example therefore begin with a first round size of $17,272$ with a $0.9$ probability of stopping, and both will stop in the first round if the sample contains at least $8,725$ ballots for the winner.
-The audits are identical in the first round, and both numbers of ballots may be computed as described in \cite{usenix_minerva} and integrated into Arlo.
+The audits are identical in the first round, and both numbers of ballots may be computed as described in \cite{usenix_minerva} and used by Arlo.
 We now consider two cases for which the audit proceeds to a second round.
@@
 
-In a less lucky sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
+In a less lucky first round sample, the winner recieves $8,637$ ballots, few more than the loser recieves.
 In the \Minerva audit, we again have to use a second round size of $43,180$, but now this round size only achieves a $0.727$ probability of stopping, significantly less than the desired $0.9$.
@@
 \subsection{Theoretical Properties}
-Here we briefly give two interesting theoretical properties of \Providence.
+Here we present two interesting theoretical properties of \Providence.
 
 \subsubsection{Efficiency}
-In terms of expected ballots sampled, it is easy to see that \Providence dominates EoR \BRAVO in the sense that for any sample for which the EoR \BRAVO stopping condition is met, the \Providence stopping condition is met also.
+It is easy to see that, for any sample for which the EoR \BRAVO stopping condition is met, the \Providence stopping condition is also met.
+This implies that \Providence will never draw more ballots than EoR \BRAVO.
 \begin{lemma}
@@
 After $j-1$ rounds, having drawn $n_{j-1}$ cumulative ballots with $k_{j-1}$ cumulative winner ballots, the \Providence stopping condition in round $j$ with size $n_j$ is equivalent to the \Providence stopping condition for a second round where the first round size is $n_1=n_{j-1}$ with $k_1=k_{j-1}$ winner ballots and a second round size $n_2=n_j$.
-The particular sequence of round sizes and winner ballots found does not affect the \Providence stopping condition; only the previous cumulative round size and number of winner ballots matters.
+The particular sequence of round sizes and winner ballots does not affect the \Providence stopping condition; only the previous cumulative round size and number of winner ballots matters.
 \Minerva does not have this property, but \BRAVO also does.

### 2023-02-19 providence/sections/intro.tex

@@
 \subsubsection{The Audit Model}
+\label{sec:intro-model}
 An audit is typically defined as a binary hypothesis test.
@@
 In fact, the larger a sample, the more accurate the estimate of election correctness can be, and the more the audit will diverge from one that would stop (see the example below).
-Of course, though, it is possible to declare an incorrect outcome as correct even after drawing a large sample, but a good audit is less likely to make this error.
+It is possible to declare an incorrect outcome as correct even after drawing a large sample, but a good audit is less likely to make this error.
 
@@
 \npara{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and new rounds will continue to be drawn.
-See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins).
-Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.
-We present two example simulations from among the ones reported to illustrate this idea.
+See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins of 0.05 and larger in the statewide contests for US President in 2020), each simulation stopping after five rounds.
+Observe that, for each margin studied, more than 90\% of the audits do not stop after five rounds.
+We present two example simulations from among the ones reported for the state of Texas to illustrate this idea.
 
-Consider the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with a total of 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.056614 in the pairwise contest (that is, ignoring the votes received by other candidates).
-The total number of votes cast in this contest, including invalid votes and votes for other candidates (all termed {\em irrelevant} votes for this pairwise contest), was 11, 315,056.
+A total of 11, 315,056 votes were cast in the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.0566 in the pairwise contest.
+A pairwise contest between two candidates treats invalid votes and those received by other candidates as irrelevant to determining which of the two won the pairwise contest.
+Its {\em margin} is the difference between their vote counts as a fraction of the sum of their votes.
+Ballot polling audits used to audit government elections (such as \BRAVO, \Minerva and \Providence) audit a multi-candidate contest by conducting multiple audits of the pairwise contests between the winner and every other candidate.
+
 A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of $\alpha=0.1$ for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
@@
 %The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09.
-Table \ref{tab:toy} lists the round sizes, votes for both candidates and the inverse \Providence ratio after each draw.
+Table \ref{tab:toy} lists the cumulative round sizes, cumulative votes for both candidates and the inverse \Providence ratio after each draw.
 Each round size was chosen for a conditional stopping probability of 0.9, given that the audit did not stop so far.
@@
 \end{tabular}
-\caption{A simulation of a \Providence audit of a tied contest when the announced election has a margin of 0.056614, each round size is computed for a conditional stopping probability of $0.9$, and $\alpha=0.1$.}
+\caption{A simulation of a \Providence audit of a tied contest with an announced margin of 0.0566, and $\alpha=0.1$.
+Each round size is computed for a conditional stopping probability of $0.9$.
+The stopping condition is Inverse Ratio < $\alpha$}
 \label{tab:toy}
@@
 \subsection{Gaps in Prior Work and Our Contributions}
-Having provided context above, we now describe gaps in our knowledge of audits and describe our contributions. 
+We describe relevant gaps in our knowledge of audits and then describe our contributions. 
 \subsubsection{Limitations of \Minerva}
@@
 They do not address the case of a stronger adversary (such as an audit insider) who can determine the size of the next round after knowing what votes are on the ballots sampled thus far.
-In particular, an open question about \Minerva is whether its RLA proof holds in this case.
+An open question about \Minerva is whether its RLA proof holds in this case.
 Can the audit insider increase the audit's error probability beyond its declared risk limit?
@@
 This presents a major limitation, because the stopping probability of the next round is better estimated using information of the sample drawn thus far, but this would not be allowed for \Minerva.
-The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter\footnote{Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
+The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter\footnote{Note that every draw may contain irrelevant ballots, and thus the true number of relevant ballots can never be predetermined.
 However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that differences in the number of ballots average out, and that a fixed draw size is sufficient, though this is not explicitly proven in \cite{usenix_minerva}.}.
@@
 If this were a correct measure of the workload of an audit, we would want to use B2 audits (round size is one) and make decisions about stopping the audit after drawing each ballot, because this leads to the smallest expected number of ballots.
-As described above, election officials, on the other hand, greatly prefer drawing many ballots at once.
+As described in \ref{sec:related}, election officials, on the other hand, greatly prefer drawing many ballots at once.
 From conversations with election officials and staff members of Verified Voting, Brennan Center and Common Cause who have been training election officials to perform RLAs, we estimate that this preference is likely due to the following.
@@
 %For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample.
-Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
+Finally, in the current environment of misinformation, election officials fear a misleading audit sample (with more votes for an announced losing candidate than the winner), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.
 
@@
 Relatedly, an optimal round schedule is not completely determined by the expected number of ballots drawn.
-It depends on other variables as well.
-The consideration of all these variables is necessary while planning an audit.
+It depends on other variables as well, the consideration of which is necessary while planning an audit.
 
@@
 %Further, it assumes that the null hypothesis is a tie. 
-A far simpler proof is possible for the risk-limiting property of \Providence. 
-In particular, the proof does not require an assumption about round sizes.
+Attempting to simplify the proof of \Minerva's risk-limiting property, we were able to define a different audit \Providence.
+The RLA proof for \Providence does not make an assumption about round sizes.
 

### 2023-02-19 providence/sections/related.tex

@@
 
-{\bf Ballot Polling RLA Process:} Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
+\npara{\bf Ballot Polling RLA Process:} Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
 
 Election officials typically draw ballots in large round sizes, see for example \cite{va-2022,RI-report}.
-Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common).
+Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$.
 For the two audits we attended, election officials chose stopping probabilities of at least $0.9$.
+Estimates of round sizes with stopping probability 0.9 for each state in the 2020 US Presidential election may be found in \cite{usenix_minerva}.
+Thousands of ballots is quite a common estimate; many estimates are as large as tens and hundreds of thousands of ballots.
 We are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds).
-\BRAVO hence cannot be used as a B2 audit in these scenarios.
 
-{\bf R2 and B2 Audits and the Classical \BRAVO audit:} A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
+\npara{\bf R2 and B2 Audits and the Classical \BRAVO audit:} A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots.
 A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
 The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
+However, \BRAVO cannot be used as a B2 audit in the scenarios described in the previous paragraph.
 
@@
 SO \BRAVO is closer to the original B2 \BRAVO, and requires fewer ballots on average than EoR \BRAVO.
-But it requires the additional effort of tracking the order of ballots, and should be expected to be inefficient because it does not use the information in the ballots drawn at the end.
+But it requires the additional effort of tracking the order of ballots, and should be expected to be inefficient because it does not use the information in the ballots drawn towards the end of the round.
 
 %The \BRAVO audit \cite{bravo} is the most popular ballot polling audit. When ballots are sampled one at a time, it is the audit with the smallest expected number of ballots drawn. 
-{\bf Newer Ballot Polling Audits:} The \Minerva audit \cite{usenix_minerva,arxiv_athena} does not need ballot order and relies only on sample and round tallies.
+\npara{\bf Newer Ballot Polling Audits:} The \Minerva audit \cite{usenix_minerva,arxiv_athena} does not need ballot order and relies only on sample and round tallies.
 It was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
-First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins; simulations \cite{arxiv_athena} support these observations.
+First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins.
 
@@
 
-{\bf Simulations:} Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and understanding audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20,bravo}.
+\npara{\bf Simulations:} Ballot polling audit simulations provide a means of educating the public and election officials \cite{dice} and understanding audit properties \cite{mclaughlin_thesis,simulations_house, blom_IRV, DBLP:conf/evoteid/HuangRSTV20,bravo}.
 There is work measuring the amount of time taken to examine a single ballot \cite{RI-report}.
@@
 
-Zag\'{o}rski {\em et al.}  also present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
+Zag\'{o}rski {\em et al.}  present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
 Broadrick {\em et al.} provide further simulations showing that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.
@@
 
-{\bf Ballot and Batch Comparison Audits:} In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
+\npara{\bf Ballot and Batch Comparison Audits:} In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot.
 Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot.

### 2023-02-19 providence/sections/workload.tex
errorbars for real time plot
misleading plot back to just virginia
fig 12 caption

@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{real_time.png}
-\caption{The real time as estimated by Equation~\ref{eq:real_time} for varying $p$ with expected values as estimated by our simulations.}
+%\includegraphics[width=.5\textwidth]{real_time.png}
+\includegraphics[width=.5\textwidth]{real_time_errorbars.png}
+\caption{The real time as estimated by Equation~\ref{eq:real_time} for varying $p$ with expected values as estimated by our simulations.
+Error bars show the $0.25$ and $0.75$ quantiles.
+Unlike Figures \ref{fig:avg_bals}, \ref{fig:avg_bals_ratio}, and \ref{fig:with_round_workload}, the quantiles stil differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
 \label{fig:real_time}
@@
 \begin{figure}
-%\includegraphics[width=.5\textwidth]{misleading_limits.png}
-\includegraphics[width=.5\textwidth]{combined_misleading.png}
+\includegraphics[width=.5\textwidth]{misleading_limits.png}
+%\includegraphics[width=.5\textwidth]{combined_misleading.png}
 \caption{The proportion of simulated \Providence audits for the Virginia and pilot contest parameters that had a \emph{misleading sample} in any round.}

### 2023-02-19 providence/sections/sims.tex
.
Update sims.tex

@@
 We use risk limit $\alpha = 0.1$, as is common in RLAs, see for example \cite{va-2022} and \cite{RI-report}.
-All trials have a maximum of $5$ rounds and a conditional stopping probability of $0.90$ in each round.
+All trials have a maximum of five rounds and a conditional stopping probability of $0.90$ in each round.
 That is, each next round size is selected to be large enough to give a $0.90$ conditional probability of stopping in that round, assuming the announced tally is correct and given the tally of previous rounds.
-We use a maximum of five rounds because virtually no audits would progress beyond five rounds given the large conditional probability of stopping.
+We use a maximum of five rounds because, if the tally were correct, virtually no audits would progress beyond five rounds given the large conditional probability of stopping in each round.
 
+\subsection{Tied Elections}
 In the simulations of \Providence audits of a tied election, the fraction of audits that stop, as shown in Figure~\ref{fig:prov-risk}, is an estimate of maximum risk.
@@
 
+\subsection{Efficiency of \Providence}
 We now investigate the efficiency of \Providence compared to \Minerva, SO \BRAVO, and EoR \BRAVO by taking a single margin as an example: the 2020 US Presidential election in the state of Texas, with margin $0.057$.
@@
 As shown, \Providence has comparable efficiency to \Minerva, while both are significantly more efficient than either implementation of \BRAVO.
-In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than $0.03$) the difference in number of ballots sampled could correspond to many days of work which would need to be completed before a certification deadline.
 % Section~\ref{sec:workload} discusses workload in more depth.

### 2023-02-19 providence/sections/pilot.tex
Update pilot.tex

@@
 As expected, the audit concluded in the first round.
-The \Providence risk was $0.0418$.
-Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO).
+The \Providence risk measure was $0.0418$.
+This is the smallest risk for which the sample would have passed a \Providence RLA for the announced election---the p-value of the statistical test and the inverse of the \Providence ratio defined in section \ref{sec:prov}.
+Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO), all similarly defined.
 
@@
 \end{center}
-\caption{Risk measures for the drawn first round of $140$ ballots in the Providence, RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
+\caption{Risk measures for the drawn first round of $140$ ballots in the RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
 \label{tab:pilot-risks}
@@
 Figure~\ref{fig:pilot_sequence} shows the cumulative tally of winner ballots after each new ballot in the selection order is added to the sample.
+
 We observe two interesting phenomena in this particular sample's selection order. 
-
-First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample.\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.} EoR \BRAVO, however, does not stop.
+First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.}.
+EoR \BRAVO, however, does not stop at sample size $140$, the number of ballots drawn during the pilot.
 It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
@@
 \includegraphics[width=.5\textwidth]{pilot_sequence.png}
-\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown, and the corresponding behavior is that of SO \BRAVO.}
+\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown.
+Observe that SO \BRAVO would stop because the \BRAVO condition (blue line) is achieved early on (orange line) in the audit.
+However, because the \BRAVO condition is no longer satisfied at $140$ ballots, an EoR \BRAVO audit would not have stopped.
+}
 \label{fig:pilot_sequence}

### 2023-02-19 providence/sections/workload.tex
small changes

@@
 As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $0.053$ between the two candidates with the most votes.
-Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
+Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is not straightforward.
+%challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
 %A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.

### 2023-02-19 providence/sections/workload.tex
Update workload.tex

@@
 \subsubsection{Average total ballots.} 
-The simplest workload models are a function of just the total number of a ballots sampled.\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is insignificant\cite{arxiv_athena}.
-It is straightforward to modify the model we discuss here to account for total distinct ballots.} Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled as a function of $p$.
+The simplest workload models are a function of just the total number of a ballots sampled\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is very small\cite{arxiv_athena}.
+It is straightforward to modify the model we discuss here to account for total distinct ballots.}.
+Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled as a function of $p$.
 \begin{figure}[h!]
@@
 This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
+We do not consider $w_r < 1$ because it is not possible for the round overhead to be smaller than the workload corresponding to a single ballot.
 
 As shown in Figure~\ref{fig:with_round_workload}, average workloads first reduce as stopping probability increases; this is likely due to a decrease in the number of rounds.
-After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not increase much and the cost changes because of number of ballots drawn, which increases with round size.
+After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not decrease much and the cost changes because of number of ballots drawn, which increases with round size.
 \Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $w_r=1000$.
@@
 \subsection{Real time}
-Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
+Given tight certification deadlines,
+%\footnote{Virginia recently passed legislation requiring pre-certification RLAs.}, 
+the total real time to conduct the RLA is also an important factor to consider when planning audits.
 Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
@@
 Error bars show the $0.25$ and $0.75$ quantiles.
-Unlike Figures \ref{fig:avg_bals}, \ref{fig:avg_bals_ratio}, and \ref{fig:with_round_workload}, the quantiles stil differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
+Unlike Figures \ref{fig:avg_bals}, \ref{fig:avg_bals_ratio}, and \ref{fig:with_round_workload}, the quantiles still differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
 \label{fig:real_time}

### 2023-02-20 providence/sections/misleading-seq.tex
fits!

--- 2023-02-19 providence/sections/workload.tex
@@
-Some election audits have benefited from a one-and-done approach: draw a large sample with high probability of stopping in the first round and usually avoid a second round altogether.
-This is appealing for two reasons.
-Firstly, rounds have some overhead in both time and effort.
-Thus the time and person-hours of an audit grows not just with the number of ballots sampled but also with the number of rounds.
-Secondly, smaller first round sizes are not large enough to accurately capture the distribution of votes.
-There is a higher probability that the true winner has fewer votes in the audit sample than some other candidate.
-On the other hand, a one-and-done audit may draw more ballots than are necessary; a more efficient round schedule could require less effort and time pre-certification.
-To evaluate the quality of various round schedules, we construct a simple workload model.
-Using this model we show how optimal round schedules can be chosen.
-We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like maximum allowed probability of a misleading audit sample.
-
-As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $0.053$ between the two candidates with the most votes.
-Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds.
-%A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
-%We will use the more standard approach of simulations to give an example here.
-Therefore we use the typical approach of simulations, again with risk limit $0.1$.
-
-We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website, including irrelevant ballots.
-We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$.
-That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again have a probability of stopping $p$ in the second round.
-Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each pairwise contest between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that pairwise contest.
-For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.
-We run $10^4$ trial audits for each value of $p$, assuming the reported results are correct\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.}.
-
-Note that simulations of audits of tied elections are not necessary, as all the audits we are considering are risk-limiting and hence we already know the performance to expect when auditing a tied election, even one not reported as such.
-
-Importantly, note that \Minerva does not appear in the analysis in this section. 
-Questions about the efficiency of \Minerva for its necessarily fixed round schedules are addressed in section~\ref{sec:sims}, but in this section round sizes are chosen to have specified probabilities of stopping given previous samples.
-\Minerva is not known to be risk-limiting in this setting, and thus cannot be used for RLAs that proceed in this way.
-
-\subsection{Person-hours}
-
-\subsubsection{Average total ballots.} 
-The simplest workload models are a function of just the total number of a ballots sampled.\footnote{Sometimes total \emph{distinct} ballots sampled is used, but for the margins we use in our examples in this section, the difference between total distinct ballots and total ballots is insignificant\cite{arxiv_athena}.
-It is straightforward to modify the model we discuss here to account for total distinct ballots.} Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled as a function of $p$.
-\begin{figure}[h!]
-%\includegraphics[width=.5\textwidth]{avg_bals.png}
-%\includegraphics[width=.5\textwidth]{avg_bals_error_bars_every_three.png}
-%\includegraphics[width=.5\textwidth]{avg_bals_errorbars.png}
-\includegraphics[width=.5\textwidth]{avg_bals_quantiles.png}
-\caption{The average total number of ballots sampled, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the US State of Virginia.
-Error bars show the $0.25$ and $0.75$ quantiles.
-For sufficiently large $p$ ($p\ge 0.75$), the $0.25$ and $0.75$ quantiles are both equal to the first round size, and this is shown by the downward arrows.}
-\label{fig:avg_bals}
-\end{figure}
-
-Figure~\ref{fig:avg_bals_ratio} provides the same number as a fraction of the \Providence values.
-It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test when each round corresponds to a single ballot.
-Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e.
-larger rounds), \Providence requires fewer ballots on average.
-In particular, the savings of \Providence become larger as $p$ increases; for $p=0.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
-\begin{figure}[h!]
-%\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
-%\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
-%\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
-\includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
-\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the state of Virginia.
-The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}
-\label{fig:avg_bals_ratio}
-\end{figure}
-
-\subsubsection{Round overhead.} 
-It is clear that average number of ballots alone is an inadequate workload measure. 
-(Consider a state conducting its audit by selecting a single ballot at random, 
-notifying just the county where the ballot is located, and then waiting to hear back for the manual interpretation of the ballot before moving on to the next one. 
-This of course is inefficient and is why audits are actually performed in rounds.)
-
-In a US state-wide RLA, the state organizes the audit by determining the random sample and communicating with the counties, but election officials at the county level physically sample and inspect the ballots after drawing them from secure storage boxes stored in county locations.
-Therefore each audit round requires some number of person-hours for set up and communication between state and county.
-This overhead for a round includes choosing the round size, generating the random sample, and communicating that random sample to the counties, as well as the communication of the results back to the state afterwards.
-
-Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
-So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $W$ of the audit is
-\begin{equation}
-W(E_b,E_r) = E_b w_b + E_r w_r + C
-\label{eq:round_workload}
-\end{equation}
-
-Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials to represent, for example, the effort of constructing a ballot manifest.
-For simplicity, (and without loss of generality), we measure in multiples of the per ballot workload; that is, we assume it is one unit, $w_b=1$.
-A per round workload of $w_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload.
-We use $w_r=1000$ as a conservative example.
-That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
-Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload.
-This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
-
-As shown in Figure~\ref{fig:with_round_workload}, average workloads first reduce as stopping probability increases; this is likely due to a decrease in the number of rounds.
-After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not increase much and the cost changes because of number of ballots drawn, which increases with round size.
-\Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $w_r=1000$.
-
-\begin{figure}[h!]
-%\includegraphics[width=.5\textwidth]{with_round_workload.png}
-%\includegraphics[width=.5\textwidth]{with_round_workload_errorbars.png}
-%\includegraphics[width=.5\textwidth]{round_workload_errorbars.png}
-\includegraphics[width=.5\textwidth]{round_workload_quantiles.png}
-\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
-Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.
-The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}
-\label{fig:with_round_workload}
-\end{figure}
-
-Importantly, this gives us a way to estimate the minimum expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload.
-For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.
-
-Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads.
-For very low round workloads, the workload function approaches just the total number of ballots, and so workload is minimized by minimizing the number of ballots drawn, which corresponds to small round sizes, and we would expect all three audits to behave similarly, as ballot-by-ballot audits, with the smallest workload.
-On the other hand, for extremely large values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values, all corresponding to large round sizes in order to minimize the number of rounds.
-We know that there is variation in the number of ballots used by each type of audit for large round sizes (a factor of two for $p=0.9$), but these values would be small in comparison to $w_r$.
-We observe this behaviour in Figure~\ref{fig:optimal_workloads} for extremely small and large workload values.
-For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
-\begin{figure}[h!]
-%\includegraphics[width=.5\textwidth]{optimal_workloads.png}
-\includegraphics[width=.5\textwidth]{optimal_workloads.png}
-\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.}
-\label{fig:optimal_workloads}
-\end{figure}
-
-Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
-As expected, an overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audits $p$ increases as a function of $w_r$.
-Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO.
-This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.
-\begin{figure}[h!]
-\includegraphics[width=.5\textwidth]{optimal_ps.png}
-\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$.
-(Note that the steps in this function are a consequence of our subsampling the workload function.
-That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.05$.)}
-\label{fig:optimal_ps}
-\end{figure}
-
-\subsubsection{Precinct overhead.} For a more complete model, we can also introduce container-level workload.
-If a round requires multiple ballots from a single container, the container need only be unsealed once.
-Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container.
-Typically available election results give per-precinct granularity of vote tallies, rather than individual container information.
-In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct workload, $w_p$.
-In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round.
-That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$.
-Then the new model is
-\begin{equation}
-W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
-\label{eq:round_and_precinct_workload}
-\end{equation}
-
-We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_workload_precinct_workload_ratio}.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
-\caption{Optimal average workload using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence.
-Similar to Figure~\ref{fig:optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case.
-If the time for a single ballot is $75$ seconds, then $c_p=50$ corresponds to over an hour of extra time to sample a ballot from a new container.}
-\label{fig:optimal_workload_precinct_workload_ratio}
-\end{figure}
-
-\subsection{Real time}
-Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
-Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
-In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
-Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
-Note that some other county may be slower, having fewer votes but also less auditing resources; but still, a slowest county exists. In this example, we take it to be Fairfax, the largest.
-For the slowest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_p$.
-Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
-\begin{equation}
-T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p + C
-\label{eq:real_time}
-\end{equation}
-
-As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
-Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
-Figure~\ref{fig:real_time} shows how the estimated real time for these values differs as a function of $p$.
-It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retrieving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors.
-We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function.
-Use of this optimal scheduling tool would depend on parameter estimates tailored to each case.
-
-\begin{figure}
-%\includegraphics[width=.5\textwidth]{real_time.png}
-\includegraphics[width=.5\textwidth]{real_time_errorbars.png}
-\caption{The real time as estimated by Equation~\ref{eq:real_time} for varying $p$ with expected values as estimated by our simulations.
-Error bars show the $0.25$ and $0.75$ quantiles.
-Unlike Figures \ref{fig:avg_bals}, \ref{fig:avg_bals_ratio}, and \ref{fig:with_round_workload}, the quantiles stil differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
-\label{fig:real_time}
-\end{figure}
-
-\section{Misleading samples}
-\label{sec:misleading}
-
-Unfortunately, efficiency alone is not sufficient for planning audits. In the US today, election officials have a legitimate need to include personal safety as a consideration.
-In a random sample, a true loser may receive more votes than the true winner.
-This happens more often when the sample sizes are small, like for a hypothetical first round size of $11$ in the pilot audit, as seen in Figure~\ref{fig:pilot_sequence}.
-In the abstract, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
-
-Imagine that Alice beats Bob in an election contest both truly and in the reported results, but Bob's supporters are insistent he really won.
-When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner (and drawing fewer ballots on average).
-After the first round, by chance, there are more votes for Bob than for Alice in the sample.
-Bob's supporters celebrate their victory that the audit has in fact revealed that Bob really won, but the election officials have to explain that they are moving on to a second round.
-So there is a reason to structure round sizes so that they are unlikely to appear to support a true loser.
-
-%After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won.
-This is an undesirable situation, as it can appear to Bob's supporters that election officials are simply drawing ballots till a chosen outcome is obtained.
-
-We introduce the notion of a \emph{misleading sample}, any cumulative sample which, assuming the announced outcome is correct, contains more ballots for a loser than for the winner.
-We can again use our simulations to gain insight into the frequency of \emph{misleading samples}.
-For each stopping probability $p$, Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a \emph{misleading sample} at any point. 
-Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
-Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.
-
-In Figure~\ref{fig:misleading}, horizontal lines are included to show \emph{misleading limits} of $0.1$, $0.01$, and $0.001$.
-To achieve a probability of a misleading sample of at most $0.1$, a round schedule with at least roughly $p=.3$ is needed.
-To achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is needed, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
-It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US and the associated threats of violence.
-Consequently, the desired \emph{misleading limit} may be a deciding constraint in the choice of round schedule.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{misleading_limits.png}
-%\includegraphics[width=.5\textwidth]{combined_misleading.png}
-\caption{The proportion of simulated \Providence audits for the Virginia and pilot contest parameters that had a \emph{misleading sample} in any round.}
-\label{fig:misleading}
-\end{figure}
-
-We observe a similar behavior in our simulations of audits on the contest from the pilot audit.
-Figure~\ref{fig:misleading} also shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
-Despite the large difference in margin ($\sim 0.05$ in Virginia and $\sim 0.25$ in the pilot) we still observe that a \emph{misleading limit} of $0.01$ is first achieved at roughly $p=0.8$ and $0.001$ at $p=0.95$.
-
-%\begin{figure}
-%\includegraphics[width=.5\textwidth]{pilot_misleading_prop.png}
-%\caption{The proportion of simulated \Providence audits for the pilot audit parameters that had a \emph{misleading sample} in any round.}
-%\label{fig:pilot_misleading}
-%\end{figure}
-
-If election officials wish to enforce a \emph{misleading limit} for all the rounds, our simulation analysis could help.
-On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample.
-Table~\ref{tab:misleading} shows for various margins the minimum first round size $n$ that guarantees a probability of a \emph{misleading sample} at most $M\in\{0.1,0.01,0.001\}$.
-For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO.
-    As seen in the Table~\ref{tab:misleading}, to enforce $M=0.01$ requires minimum round sizes with at least roughly a $0.8$ probability of stopping in the first round.
-Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$ to choose the first round size, the election officials may opt to use this constraint on the probability of a \emph{misleading sample} as the deciding factor in planning their audits.
-
 \subsubsection{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.

### 2023-02-20 providence/sections/pilot.tex
small changes
fits!

@@
 Figure~\ref{fig:pilot_sims} shows the average number of ballots sampled for each value of $p$ in the simulations.
-The vertical line denotes the stopping probability of the first round size actually chosen in the pilot ($140$ ballots).
+The vertical line denotes the stopping probability of the first round size of relevant votes drawn in the pilot ($140$ ballots).
 The large value of $p$ corresponds to a large first round size and a corresponding large value of average number of ballots.
@@
 We observe two interesting phenomena in this particular sample's selection order. 
-First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the sample (orange line) surpasses the minimum number of winner ballots (blue line) earlier in the sample\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.}.
-EoR \BRAVO, however, does not stop at sample size $140$, the number of ballots drawn during the pilot.
-It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but poor evidence for the alternative hypothesis in the rest of the sample is ignored.
-Second, only $5$ of the first $11$ ballots were for the announced winner.
+First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the winner votes in the sample (orange line) surpasses the minimum number of winner ballots need to meet the \BRAVO stopping condition (blue line) earlier in the sample\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.}.
+EoR \BRAVO, however, does not stop at sample size $140$, the number of relevant ballots drawn during the pilot.
+It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but the rest of the sample is ignored.
+Second, the orange line is below the dotted green line, which represents half the sample size; only $5$ of the first $11$ ballots were for the announced winner.
 A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size.
-Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.
+%Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.
+
 %first draft, needed to be shorter: ...stopping condition for the cumulative tally at the end of the round would see that the \BRAVO condition is not met (and would be correct that an EoR \BRAVO audit would require a second round for this sample).
@@
 \caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown.
-Observe that SO \BRAVO would stop because the \BRAVO condition (blue line) is achieved early on (orange line) in the audit.
+Observe that SO \BRAVO would stop because the minimum number of ballots required to satisfy the \BRAVO stopping condition (blue line) is achieved early on (when the orange line crosses the blue one) in the audit.
 However, because the \BRAVO condition is no longer satisfied at $140$ ballots, an EoR \BRAVO audit would not have stopped.
-}
+Further, at a sample size of $11$, the orange line is not even above the dotted green lime representing half the sample size and the winner has fewer than half the relevant votes in the sample.}
 \label{fig:pilot_sequence}

### 2023-02-20 providence/sections/workload.tex
fits!

@@
 
-Figure~\ref{fig:avg_bals_ratio} provides the same number as a fraction of the \Providence values.
-It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test when each round corresponds to a single ballot.
-Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e.
+%Figure~\ref{fig:avg_bals_ratio} provides the same number as a fraction of the \Providence values.
+It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test when each round corresponds to a single ballot. Figures~\ref{fig:avg_bals} 
+%and \ref{fig:avg_bals_ratio} 
+shows that for larger stopping probabilities $p$ (i.e.
 larger rounds), \Providence requires fewer ballots on average.
 In particular, the savings of \Providence become larger as $p$ increases; for $p=0.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.
-\begin{figure}[h!]
+
+%\begin{figure}[h!]
 %\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
@@
 %\includegraphics[width=.5\textwidth]{avg_bals_ratio_errorbars.png}
-\includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
-\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the state of Virginia.
+%\includegraphics[width=.5\textwidth]{avg_bals_ratio_quantiles.png}
+%\caption{The total number of ballots sampled on average, as a fraction of those sampled by \Providence, as a function of $p$, the conditional stopping probability used to select each round size, for ballot polling audits of the 2016 US Presidential election in the state of Virginia.
 The $0.25$ and $0.75$ quantiles are shown as in Figure~\ref{fig:avg_bals}.}
-\label{fig:avg_bals_ratio}
-\end{figure}
+%\label{fig:avg_bals_ratio}
+%\end{figure}
 
@@
 As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
-Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
+Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order of magnitude for this variable.}.
+In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules.
 Figure~\ref{fig:real_time} shows how the estimated real time for these values differs as a function of $p$.
@@
 This happens more often when the sample sizes are small, like for a hypothetical first round size of $11$ in the pilot audit, as seen in Figure~\ref{fig:pilot_sequence}.
-In the abstract, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.
+In the abstract, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice it serves to create expectations or suspicions that then need to be managed by election officials.
+Hence there is reason to structure round sizes so that they are unlikely to misrepresent the true outcome.
 
-Imagine that Alice beats Bob in an election contest both truly and in the reported results, but Bob's supporters are insistent he really won.
+%, angry supporters of the losing candidate may be suspicious of election officials moving onto a next round, or, if the audit is declared correct after the second round, feel sufficiently disappointed to pose a threat.
+%but in practice the implications of this approach may be dangerous.
+
+%Imagine that Alice beats Bob in an election contest both truly and in the reported results, but Bob's supporters are insistent he really won.
 When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner (and drawing fewer ballots on average).
-After the first round, by chance, there are more votes for Bob than for Alice in the sample.
+% After the first round, by chance, there are more votes for Bob than for Alice in the sample.
 Bob's supporters celebrate their victory that the audit has in fact revealed that Bob really won, but the election officials have to explain that they are moving on to a second round.
-So there is a reason to structure round sizes so that they are unlikely to appear to support a true loser.
 
@@
 
-We observe a similar behavior in our simulations of audits on the contest from the pilot audit.
+%We observe a similar behavior in our simulations of audits on the contest from the pilot audit.
 Figure~\ref{fig:misleading} also shows the proportion of the pilot simulations which contained a \emph{misleading sample} in any round.
@@
 Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$ to choose the first round size, the election officials may opt to use this constraint on the probability of a \emph{misleading sample} as the deciding factor in planning their audits.
-
-\subsubsection{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.
-
-After drawing a cumulative $n>1$ ballots in a round, some number $k$ of them are votes for the announced winner.
-There are $\binom{n}{k}$ possible sequences of ballots which can lead to such a sample.
-Given a value of $k$, however, the particular sequence of the sample that led to that value of $k$ contains no additional information about whether the sample is more likely under the alternative or null hypotheses.
-That is to say, $\Pr[K=k|H_a]$ and $\Pr[K=k|H_0]$ have the same value regardless of the sequence.
-Despite this, the SO \BRAVO RLA stopping condition is not just a function of $n$ and $k$ but also a function of the sequence, the selection order.
-In particular, if the sequence of ballots is such that the standard \BRAVO stopping condition was met for some $n'<n$ and corresponding $k'<n$, the audit will stop, even if by the end of the sequence the values $k$ and $n$ no longer meet the \BRAVO condition.
-We refer to such sequences which stop under SO \BRAVO, but not under EoR \BRAVO, as \emph{misleading sequences}.
-To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result\cite{wald}.
-The misleading nature of such stoppages is the note we are making.
-This is another case where election officials might have difficulty explaining the misleading situation to the public.
-
-Recall from section~\ref{sec:pilot} that the pilot \Providence RLA performed in Providence, Rhode Island had an SO \BRAVO \emph{misleading sequence}.
-In particular, the audit passed with an SO \BRAVO risk measure of $0.0541$ but the final cumulative tally of the sample gives a \BRAVO risk measure of $0.366$.
-
-It is easy to use our simulations to see how often SO \BRAVO \emph{misleading sequences} occur by checking whether the final cumulative sample of each SO \BRAVO trial meets the EoR \BRAVO stopping condition and counting those which do not.
-Figure~\ref{so_misleading} shows the proportion of simulated SO \BRAVO audits that stopped with a \emph{misleading sequence}.
-Unlike the more general \emph{misleading sample} discussed so far, these \emph{misleading sequences} are unique to SO \BRAVO audits, and Figure~\ref{so_misleading} only shows the proportion of audits that stopped with a \emph{misleading sequence}; additional SO \BRAVO audits also contained \emph{misleading samples}.
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{so_misleading.png}
-\caption{The proportion of sequences that are misleading sequences in the SO \BRAVO audit as a function of $p$.}
-\label{so_misleading}
-\end{figure}

### 2023-02-20 providence/sections/workload.tex
Final version for second shepherd draft

@@
 Error bars show the $0.25$ and $0.75$ quantiles.
-Unlike Figures \ref{fig:avg_bals}, \ref{fig:avg_bals_ratio}, and \ref{fig:with_round_workload}, the quantiles still differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
+Unlike Figures \ref{fig:avg_bals} and \ref{fig:with_round_workload}, the quantiles still differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
 \label{fig:real_time}

### 2023-02-20 providence/sections/workload.tex
Fixed w_r, w_p and Fairfax caption

@@
 
-Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
+Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $c_r$.
 So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $W$ of the audit is
 \begin{equation}
-W(E_b,E_r) = E_b w_b + E_r w_r + C
+W(E_b,E_r) = E_b w_b + E_r c_r + C
 \label{eq:round_workload}
@@
 For simplicity, (and without loss of generality), we measure in multiples of the per ballot workload; that is, we assume it is one unit, $w_b=1$.
-A per round workload of $w_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload.
-We use $w_r=1000$ as a conservative example.
+A per round workload of $c_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload.
+We use $c_r=1000$ as a conservative example.
 That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots.
-Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload.
+Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload.
 This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.
-We do not consider $w_r < 1$ because it is not possible for the round overhead to be smaller than the workload corresponding to a single ballot.
+We do not consider $c_r < 1$ because it is not possible for the round overhead to be smaller than the workload corresponding to a single ballot.
 
@@
 After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not decrease much and the cost changes because of number of ballots drawn, which increases with round size.
-\Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $w_r=1000$.
+\Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $c_r=1000$.
 
@@
 \includegraphics[width=.5\textwidth]{round_workload_quantiles.png}
-\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various values of $p$.
+\caption{For workload parameters $w_b=1$ and $c_r=1000$, this plot shows the expected workload for various values of $p$.
 Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.
@@
 Importantly, this gives us a way to estimate the minimum expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload.
-For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.
+For each round workload $c_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.
 
@@
 On the other hand, for extremely large values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values, all corresponding to large round sizes in order to minimize the number of rounds.
-We know that there is variation in the number of ballots used by each type of audit for large round sizes (a factor of two for $p=0.9$), but these values would be small in comparison to $w_r$.
+We know that there is variation in the number of ballots used by each type of audit for large round sizes (a factor of two for $p=0.9$), but these values would be small in comparison to $c_r$.
 We observe this behaviour in Figure~\ref{fig:optimal_workloads} for extremely small and large workload values.
-For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
+For more reasonable values of the round workload $c_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
 \begin{figure}[h!]
@@
 \includegraphics[width=.5\textwidth]{optimal_workloads.png}
-\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.}
+\caption{For varying round workload $c_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.}
 \label{fig:optimal_workloads}
@@
 Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads.
-As expected, an overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audits $p$ increases as a function of $w_r$.
+As expected, an overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audits $p$ increases as a function of $c_r$.
 Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO.
@@
 \includegraphics[width=.5\textwidth]{optimal_ps.png}
-\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$.
+\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $c_r$.
 (Note that the steps in this function are a consequence of our subsampling the workload function.
-That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.05$.)}
+That is, the workload-minimizing value of $p$ for each $c_r$ is only allowed to take on values at increments of $0.05$.)}
 \label{fig:optimal_ps}
@@
 Typically available election results give per-precinct granularity of vote tallies, rather than individual container information.
-In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct workload, $w_p$.
-In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round.
+In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct workload, $c_p$.
+In this model, the workload estimate incurs an additional workload of $c_p$ every time a precinct is sampled from for the first time in a round.
 That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$.
@@
 \begin{equation}
-W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
+W(E_b, E_r, E_p) = E_b w_b + E_r c_r + E_p c_p + C
 \label{eq:round_and_precinct_workload}
@@
 \includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
-\caption{Optimal average workload using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence.
+\caption{Optimal average workload using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $c_p$, given as a fraction of the value for \Providence.
 Similar to Figure~\ref{fig:optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case.
@@
 Error bars show the $0.25$ and $0.75$ quantiles.
-Unlike Figures \ref{fig:avg_bals} and \ref{fig:with_round_workload}, the quantiles still differ for large $p$ because the first round size is no longer a constant; the number of ballots drawn in the first round in Fairfax County is variable.}
+Unlike Figures \ref{fig:avg_bals} and \ref{fig:with_round_workload}, the quantiles still differ for large $p$ because while the the number of ballots drawn in the first round in Virginia is constant, the number drawn in Fairfax County is variable.}
 \label{fig:real_time}

### 2023-02-27 providence/sections/intro.tex
.

@@
 
-A total of 11, 315,056 votes were cast in the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.0566 in the pairwise contest.
+A total of 11,315,056 votes were cast in the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.0566 in the pairwise contest.
 A pairwise contest between two candidates treats invalid votes and those received by other candidates as irrelevant to determining which of the two won the pairwise contest.
@@
 Each round size is computed for a conditional stopping probability of $0.9$.
-The stopping condition is Inverse Ratio < $\alpha$}
+The stopping condition is Inverse Ratio $\le \alpha$.}
 \label{tab:toy}

### 2023-02-27 providence/sections/intro.tex
made the two changes from email
minor edits

@@
 It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected.
-Errors and/or manipulation may not always change an election outcome, but we would want to know when they do.
+Errors and/or manipulation may not always change an election outcome, but we want to know when they do.
 {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome.
@@
 \subsection{Gaps in Prior Work and Our Contributions}
-We describe relevant gaps in our knowledge of audits and then describe our contributions. 
+We describe relevant gaps in current knowledge of audits and then describe our contributions. 
 \subsubsection{Limitations of \Minerva}

### 2023-02-28 providence/sections/intro.tex
Changes to intro and bib

@@
 \item A first round size \cite{usenix_minerva} is chosen. %\fpo{the sentence is hard to understand}. pv: fixed
-\item Ballots on the ballot manifest are sampled uniformly at random, with replacement, using a pseudorandom number generator, which is typically seeded by a natural source of randomness like rolling dice.
-\item The physical ballots are found and manually interpreted, recording the manual interpretations. 
+\item Ballots on the ballot manifest are sampled uniformly at random, with replacement, using a pseudorandom number generator---typically seeded by a natural source of randomness like rolling dice.
+\item The physical ballots are found and manually interpreted; the interpretations are recorded. 
 \item The stopping condition $\mathcal{A}$, a function of the manual interpretations of the current cumulative sample of ballots $X$, is computed. It outputs: 
@@
 Election officials may choose to abort this procedure and go to a full hand count at any time and should have a plan for how to decide whether to do so; we discuss this in more detail below and in section \ref{sec:model}.
-\item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
+\item If more ballots are to be drawn, the next round size is chosen, and the audit goes back to step 2. 
 \end{enumerate}
@@
 Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required.
-For RLAs required by statute or legislation, the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence.
+For RLAs required by statute or legislation, the successful completion of the RLA (or a full hand count confirming the certified outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence.
 For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome.
@@
 \npara{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and new rounds will continue to be drawn.
-See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins of 0.05 and larger in the statewide contests for US President in 2020), each simulation stopping after five rounds.
+See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins of 0.05 and larger in the statewide contests for US President in 2020).
 Observe that, for each margin studied, more than 90\% of the audits do not stop after five rounds.
-We present two example simulations from among the ones reported for the state of Texas to illustrate this idea.
+To illustrate this idea, we present two example simulations from among those used to compute the statistics reported in Figure \ref{fig:prov-risk}.
 
@@
 A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of $\alpha=0.1$ for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round).
-The audit stops when the inverse of the \Providence ratio (see section \ref{sec:prov}) is smaller than $\alpha$.
+If the \Providence ratio is denoted $\omega$, the audit stops when $\omega^{-1} \leq \alpha$, or equivalently (see Definition \ref{def:minervatwo}) if $\omega \geq \alpha^{-1}$.
 We describe two audit simulations using these parameters.
 
-The first simulation assumed that the tally is correct and resulted in 1,138 votes for Trump and 1,054 for Biden (the rest were irrelevant votes).
-The inverse of the \Providence ratio was $0.047 < \alpha$ and the audit stopped.
+The first simulation assumed that the tally is correct and resulted in 1,138 votes for Trump and 1,054 for Biden (the rest were irrelevant votes) and the audit stopped because: \[ \omega ^{-1} = 0.047 < \alpha \]
 
@@
 \hline
-Round& Cumulative& Trump & Biden & Inverse  \\
- No. & Round Size &  & & Ratio \\
+Round& Cumulative& Trump & Biden & $\omega^{-1}$  \\
+ No. & Round Size &  & &  \\
 \hline
@@
 Each round size is computed for a conditional stopping probability of $0.9$.
-The stopping condition is Inverse Ratio $\le \alpha$.}
+The stopping condition is $\omega^{-1} \le \alpha$.}
 \label{tab:toy}
@@
 
-\npara{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit.
+\npara{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, given an incorrect outcome, make the audit declare it as correct with a chance larger than the risk limit.
 An audit is an RLA only if there is a proof of its risk-limiting property.
@@
 
-The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
+The stopping condition for \Minerva does not take into account the sample obtained in previous rounds.
+The risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample.
 %Further, it assumes that the null hypothesis is a tie. 
@@
 \subsubsection{Organization} 
-Sections \ref{sec:model} provides preliminaries on the \BRAVO and \Minerva audits.
+Section \ref{sec:model} provides preliminaries on the \BRAVO and \Minerva audits.
 Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island.
-Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2016 US Presidential election in the state of Virginia.
+Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2016 US Presidential election in the state of Virginia, while section \ref{sec:misleading} introduces the notion of a misleading sample, illustrating it on the Virginia details as an example.
 Our conclusions, the availability of an audit implementation and acknowledgements may be found in sections \ref{sec:conc}, \ref{sec:avail} and \ref{sec:ack} respectively.

### 2023-02-28 providence/sections/related.tex
made the two changes from email
Changes to intro and bib

@@
 %Related Work
-In this section, we describe related work.
-
 \npara{\bf Ballot Polling RLA Process:} Bernhard provides a good description of the RLA and its assumptions, and also describes the process on the ground \cite{bernhard-sok}.
@@
 Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$.
-For the two audits we attended, election officials chose stopping probabilities of at least $0.9$.
+For the two audits we attended, election officials chose stopping probabilities of $0.9$ and $0.95$.
 Estimates of round sizes with stopping probability 0.9 for each state in the 2020 US Presidential election may be found in \cite{usenix_minerva}.
@@
 A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn.
-The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
+The popular \BRAVO audit \cite{bravo} requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).
 However, \BRAVO cannot be used as a B2 audit in the scenarios described in the previous paragraph.
@@
 It was developed for use with large first round sizes, and has been proven to be risk limiting when the round schedule for the audit is fixed before any ballots are drawn.
-First-round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins.
+First round sizes for a stopping probability of $0.9$ when the announced tally is correct have been shown to be smaller than those for EoR and SO \BRAVO for a wide range of margins.
 
@@
 
-Zag\'{o}rski {\em et al.}  present first-round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
+Zag\'{o}rski {\em et al.}  present first round simulations demonstrating that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first round sizes when the true tally is as announced.
 Broadrick {\em et al.} provide further simulations showing that \Minerva requires fewer ballots than EoR and SO \BRAVO over multiple rounds and for smaller stopping probability.

### 2023-02-28 providence/sections/sims.tex
Changes to intro and bib

@@
 \includegraphics[width=.5\textwidth]{prov_risk.png}
-\caption{The fraction of simulated \Providence audits on tied elections that stopped in any rounds (we performed five rounds at a risk limit of $0.1$) as a function of contest margin.
-This value is an estimate of the maximum risk of the \Providence audit.}
+\caption{The fraction of simulated \Providence audits on tied elections that stopped in any round (we performed five rounds at a risk limit of $0.1$) as a function of contest margin.
+This value is an estimate of the maximum risk of the \Providence audit.
+Observe that it is below the risk limit, as expected.}
 \label{fig:prov-risk}

### 2023-02-28 providence/sections/adversary.tex
typo

@@
 In one case there are $8,724$ votes for the winner in the sample, just one fewer than the minimum needed to meet the risk limit.
-In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $.954$.
+In the \Minerva audit, we are already committed to a second round size of $43,180$ which, given the nearly-passing sample of the first round is higher than necessary, achieving a stopping probability in the second round of $0.954$.
 The \Providence audit samples more than $9,000$ fewer ballots with a round size of $34,078$, achieving the desired $0.9$ probability of stopping.

### 2023-02-28 providence/sections/intro.tex
small changes

@@
 Observe that, for each margin studied, more than 90\% of the audits do not stop after five rounds.
-To illustrate this idea, we present two example simulations from among those used to compute the statistics reported in Figure \ref{fig:prov-risk}.
+To illustrate this idea, we present two example simulations from among those used to compute the statistics reported in section \ref{sec:sims}.
 
@@
 
-The first simulation assumed that the tally is correct and resulted in 1,138 votes for Trump and 1,054 for Biden (the rest were irrelevant votes) and the audit stopped because: \[ \omega ^{-1} = 0.047 < \alpha \]
+The first simulation assumed that the tally is correct and resulted in 1,138 votes for Trump and 1,054 for Biden (the rest were irrelevant votes).
+The audit stopped because: \[ \omega ^{-1} = 0.047 < \alpha \]
 

### 2023-02-28 providence/sections/model.tex
small changes

@@
 The ratios for \BRAVO and \Minerva are for two candidates.
-As mentioned when describing the example in section \ref{sec:intro-model}, audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates.
+As mentioned when describing the example in section \ref{sec:intro-model}, audits of multiple-candidate contests are treated as multiple audits of the pairwise contests between the announced winner and all other candidates.
 

### 2023-02-28 providence/sections/prov.tex
small changes

@@
 
-In \Minerva, the stopping condition relates the weighted average of the risks to the weighted average of the stopping probabilities over all values of $K_j$, for a given round size.
+In \Minerva, the stopping condition relates the weighted average of the risks to the weighted average of the stopping probabilities over all values of $K_j$ for which the audit stops, for a given round size.
 Separate relationships between risk and stopping probability are not available for individual values of $K_j$.
@@
 Notice that \Providence requires the computation of $\tau_j$ for $j=1$ and no other values of $j$.
-The value of $\tau_1$ is simply the ratio of the the tails of the binomial distributions (the distributions are binomial because the auditing is with replacement) for the two hypotheses and can be fairly efficiently computed.
+The value of $\tau_1$ is simply the ratio of the the tails of the binomial distributions (the distributions are binomial because the sampling is with replacement) for the two hypotheses and can be fairly efficiently computed.
 The computation of $\tau_j$ for $j \geq 2$, as required in \Minerva, relies on the convolution of two probability distribution functions and is hence computationally considerably more expensive.

### 2023-02-28 providence/sections/pilot.tex
I'm done

@@
 The \Providence risk measure was $0.0418$.
-This is the smallest risk for which the sample would have passed a \Providence RLA for the announced election---the p-value of the statistical test and the inverse of the \Providence ratio defined in section \ref{sec:prov}.
+This is the smallest risk for which the sample would have passed a \Providence RLA for the announced election---the p-value of the statistical test and the inverse of the \Providence ratio $\omega$ of Definition \ref{def:minervatwo}.
 Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO), all similarly defined.
@@
 \end{center}
-\caption{Risk measures for the drawn first round of $140$ ballots in the RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
+\caption{Risk measures for the drawn first round of $140$ ballots in the RI pilot audit. Risk measures in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
 \label{tab:pilot-risks}
@@
 It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but the rest of the sample is ignored.
-Second, the orange line is below the dotted green line, which represents half the sample size; only $5$ of the first $11$ ballots were for the announced winner.
+Second, the orange line is below the dotted green line, which represents half the sample size, at a sample size of $11$; only $5$ of the first $11$ ballots were for the announced winner.
 A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size.
@@
 However, because the \BRAVO condition is no longer satisfied at $140$ ballots, an EoR \BRAVO audit would not have stopped.
-Further, at a sample size of $11$, the orange line is not even above the dotted green lime representing half the sample size and the winner has fewer than half the relevant votes in the sample.}
+Further, at a sample size of $11$, the orange line is not even above the dotted green line representing half the sample size and the winner has fewer than half the relevant votes in the sample.}
 \label{fig:pilot_sequence}
